### 2020-06-11 article.tex
Initial commit

% SIAM Article Template
\documentclass[draft,review,onefignum,onetabnum]{siamart190516}

% Information that is shared between the article and the supplement
% (title and author information, macros, packages, etc.) goes into
% ex_shared.tex. If there is no supplement, this file can be included
% directly.

\input{shared}

% Optional PDF information
\ifpdf
\hypersetup{
  pdftitle={KSPHPDDM and PCHPDDM: A Class of Adaptive Krylov and Overlapping Schwarz Methods in PETSc},
  pdfauthor={P. Jolivet and S. Zampini}
}
\fi

% The next statement enables references to information in the
% supplement. See the xr-hyperref package for details.

% \externaldocument{supplement}

% FundRef data to be entered by SIAM
%<funding-group specific-use="FundRef">
%<award-group>
%<funding-source>
%<named-content content-type="funder-name"> 
%</named-content> 
%<named-content content-type="funder-identifier"> 
%</named-content>
%</funding-source>
%<award-id> </award-id>
%</award-group>
%</funding-group>
\newcommand\col{:}
\newcommand{\pk}[1]{{\color{gray!90!black} #1}}
\usepackage{stmaryrd}
\usepackage{pgfplotstable}
\pgfplotsset{compat=newest}
\pgfplotsset{colormap={paraview}{rgb(0cm)=(0.278431,0.278431,0.858824) rgb(0.1428571429cm)=(0,0,0.360784) rgb(0.2857142857cm)=(0,1,1) rgb(0.4285714286cm)=(0,0.501961,0) rgb(0.5714285714cm)=(1,1,0) rgb(0.7142857143cm)=(1,0.380392,0) rgb(0.8571428571cm)=(0.419608,0,0) rgb(1cm)=(0.878431,0.301961,0.301961)}}
\usepackage{fancyvrb}
\usepackage{subfig}
\usepackage{upquote,textcomp}
\newcommand*{\fvtextcolor}[2]{\textcolor{#1}{#2}}
\definecolor{last}{RGB}{44,123,182}
\definecolor{blueaccent}{RGB}{215,25,28}
\definecolor{myyellow}{RGB}{102,102,17}
\definecolor{myred}{RGB}{197,65,0}
\definecolor{myblue}{RGB}{0,0,235}
\definecolor{myaqua}{RGB}{9,156,156}
\definecolor{mygreen}{RGB}{63,126,28}
\definecolor{paraviewred}{rgb}{0.278431,0.278431,0.858824}
\definecolor{paraviewblue}{rgb}{0.878431,0.301961,0.301961}
\definecolor{darkgray}{RGB}{253,174,97}
\definecolor{mediumgray}{RGB}{255,255,191}
\usepackage{tikz}
\usepackage[binary-units=true]{siunitx}
\usetikzlibrary{decorations.text,arrows.meta,calc,plotmarks,backgrounds,shadows,spy,fixedpointarithmetic,patterns,intersections,arrows.meta,shapes,decorations.text,decorations.pathmorphing,backgrounds,fit,positioning,shapes.symbols,chains,3d,calc,external}
\long\def\ifnodedefined#1#2#3{%
    \@ifundefined{pgf@sh@ns@#1}{#3}{#2}%
}
\makeatother
\usepackage{amsmath,etoolbox}
\AtBeginEnvironment{pmatrix}{\setlength{\arraycolsep}{2pt}}
\pgfmathsetmacro{\dof}{4}
\pgfmathsetmacro{\overlap}{1}
\begin{document}

\maketitle

% REQUIRED
\begin{abstract}
  Contemporary applications in computational science and engineering often
    require the solution of linear systems which may be of different sizes,
    shapes, and structures. The goal of this paper is to explain how two
    libraries, PETSc and HPDDM, have been interfaced together in order to offer
    to end-users adaptive Krylov and overlapping Schwarz methods.  The
    flexibility of the implementation, which for example may deal with various
    matrix formats, are showcased and explained.  Diverse examples are
    displayed, ranging from fracture mechanics to linear stability analysis, and
    the sets of options are thoroughly documented, so that they may be readily
    tried on other applications.
\end{abstract}

% REQUIRED
\begin{keywords}
  Krylov methods, domain decomposition preconditioners, distributed computing
\end{keywords}

% REQUIRED
\begin{AMS}
  35Q68, 65Y05
\end{AMS}

\section{Introduction}
Implementing and maintaining robust and scalable applications is highly relevant
in the field of computational science and engineering. Some tools exist to help
end-users achieve this goal. One the one hand, PETSc \cite{balay1997cient}, the
Portable and Extensible Toolkit for Scientific computation, is a
well-established software from the community. It may be used to efficiently
discretize partial differential equations, or solve algebraic linear or
nonlinear systems of equations. It comes with many built-in features such as
advanced preconditioning or tailored matrix formats. Also, it can be used to
compile and interact with third-party libraries, such as
\emph{hypre}~\cite{falgout2002hypre} for linear solvers, TetGen~\cite{si2013tet}
for mesh generation, and so on. These appealing features convinced computational
scientists to use PETSc as one of the discretization and/or algebraic backend in
many different higher-level projects, such as finite element or finite volume
packages. On the other hand, HPDDM is a much smaller project purely focused on
domain decomposition preconditioners~\cite{jolivet2013scalable}, a small
subclass of preconditioners available in PETSc as well, and advanced iterative
methods~\cite{jolivet2016block}. Such methods deal efficiently with linear
systems when multiple right-hand sides are available simultaneously, and are
then referred to as block iterative methods, or when there is a recurrence of
varying coefficient matrices and right-hand sides, and are then referred to as
recycling iterative methods.
% Htool ?
In this paper, the integration of both aspects of HPDDM, its advanced iterative
methods and its domain decomposition preconditioners, into PETSc is explained.
The design principles of the interface are showcased, alongside minimalist and
easy-to-run examples. The paper is divided in two main parts:
\cref{sec:KSPHPDDM} which introduces \pk{KSPHPDDM}, and \cref{sec:PCHPDDM} which
introduces \pk{PCHPDDM}. Eventually, concluding remarks may be found
\cref{sec:conclusion}. All PETSc keywords are highlighted in \pk{gray}, these
are public and documented at \url{https://www.mcs.anl.gov/petsc/petsc-master/docs}.
\section{KSPHPDDM}\label{sec:KSPHPDDM}
  \subsection{Related work}
Krylov subspace methods are widely used in numerical linear algebra for solving
systems of equations. Some of their strengths are memory efficiency and the fact
that they are mostly format agnostic. Indeed, they mainly rely on matrix--vector
operations such as multiplications or transpose multiplications. For
challenging high-dimensional systems, finding appropriate preconditioners is
mandatory to make these methods robust, but this goes beyond the scope of this
section. In PETSc, Krylov methods inherit from the \pk{KSP} base class. Some of the
most used types are \pk{KSPGMRES}, \pk{KSPCG}, or \pk{KSPBCGS}, which respectively implement
the generalized minimal residual method~\cite{saad1986gmres}~(GMRES), conjugate
gradient~\cite{hestenes1952methods}~(CG), or biconjugate gradient stabilized method~\cite{van1992bi}~(BCGS). Some of these methods are also
available in other linear algebra backends such as~\emph{hypre} or DUNE~\cite{bastian2010generic}.

Because these methods may need long-term recurrences to mitigate round-off
errors during the generation of Krylov subspaces, it is common to introduce a
restart parameter in order to control their memory consumption and the amount of
global communication, e.g., when orthonormalizing a candidate basis vector. Such
restarts may hinder convergence of iterative methods, and even introduce
convergence plateaus. Recycling techniques have been introduced to attenuate
these effects. In PETSc, the Loose GMRES~\cite{baker2005technique}~(LGMRES) and the
Deflated GMRES~\cite{erhel1996restarted,wakam2013158}~(DGMRES), implemented as
\pk{KSPLGMRES} and \pk{KSPDGMRES}, are available. Neither handle variable
preconditioning, and the latter does not support complex arithmetic. Also note
that when solving a sequence of linear systems $A_i x_i = b_i, i = 1, 2,
\ldots$, it is not trivial to extend the theory of these methods to perform
recycling throughout both the restarts and the different systems. The
generalized conjugate residual method with inner orthogonalization and deflated
restarting~\cite{parks2006recycling}~(GCRODR) does not suffer from this
limitation.

Recycling put aside, another important aspect of Krylov methods is their ability
to deal with multiple right-hand sides simultaneously. Block Krylov methods~\cite{gutknecht2006block} are
designed for solving linear systems $A X = B$, where
$X$ and $B$ are tall-and-skinny matrices with $k \geq 1$ columns. These methods are
less frequently implemented in general purpose libraries because they require
somehow more involved kernels such as matrix--matrix multiplication, instead of
matrix--vector. Still, systems with multiple right-hand sides are ubiquitous,
e.g., in tomography~\cite{tournier2016micro} or data
analytics~\cite{KALANTZIS2018136}. Of course, it is possible to cast a system
with multiple right-hand sides into a sequence of systems with the same
coefficient matrix but changing right-hand sides $A x_i = b_i, i =
1,2,\ldots,k$. However, block Krylov methods, at the cost of higher arithmetic
intensities, generate larger subspaces and typically converge in fewer iterates.
There is none implemented in PETSc. It is nevertheless possible to solve
systems with multiple right-hand sides using a specialized type of matrices,
but with a very limited choice of preconditioners, cf.~\cref{sec:matkaij}.
In the context of optimization, it is also
frequent to solve a sequence of linear systems $A_i X_i = B_i, i = 1, 2,
\ldots$

Trilinos~\cite{heroux2005overview} is another well-known software for scientific
computing. Through its Belos package~\cite{bavier2012amesos2}, it provides most
of the advanced iterative methods aforementioned. Although having PETSc and
Trilinos interoperate is possible, there is currently no \pk{KSP} interface to Belos.
This limits its composability inside complete PETSc applications. Furthermore,
as displayed in~\cref{sec:interface-ksp}, some Krylov methods implemented in
HPDDM are not available in Belos. In the following section, the details of the
interface between HPDDM Krylov methods and PETSc matrix and preconditioner
classes will be explained. It is assumed that the right-hand sides, and
consequently the solutions, are always dense vectors or matrices.
  \subsection{Interfacing HPDDM Krylov methods in PETSc\label{sec:interface-ksp}}\
\pk{KSPHPDDM}, the interface between HPDDM Krylov methods and PETSc, is automatically
registered since PETSc version 3.12 when one configures PETSc with the extra
flag -{}-download-hpddm. It is then possible to select the corresponding \pk{KSP}
using the command line option -ksp\_type hpddm, or the routine \pk{KSPSetType}(ksp,
\pk{KSPHPDDM}) and its Fortran equivalent.
When \pk{KSPHPDDM} is being used, it is then possible to select between the following
Krylov methods:
\begin{itemize}
    \item pseudo-block GMRES or flexible GMRES~\cite{saad1993flexible};
    \item pseudo-block CG or flexible CG~\cite{notay2000flexible};
    \item pseudo-block GCRODR or flexible GCRODR~\cite{carvalho2011flexible};
    \item block GMRES or flexible GMRES~\cite{calandra2012flexible};
    \item block CG~\cite{o1980block};
    \item breakdown-free block CG~\cite{ji2017breakdown};
    \item block GCRODR or flexible GCRODR.
\end{itemize}
In pseudo-block variants, multiple similar operations like matrix--vector
products are fused to achieve higher arithmetic intensity, or to decrease the
number of global synchronizations, e.g., reductions for scalar products. But they are mathematically equivalent to
their ``standard'' counterpart. These methods may be selected using the additional
options -ksp\_hpddm\_type (gmres$|$cg$|$gcrodr$|$bgmres$|$bcg$|$bfbcg$|$bgcrodr$|$preonly) and
-ksp\_hpddm\_variant (left$|$right$|$flexible) to select whether preconditioning
is applied on the left, on the right, or variable. It is also possible to set
this option through the more common PETSc option -ksp\_pc\_side (left$|$right). 
Note that the value of the latter option is useless for BCG or BFBCG, and HPDDM CG only handles left and
variable preconditioning. According to the standard PETSc notation, ksp($A$, $P_B$)
defines a \pk{KSP} used to solve a system $Ax=b$ using a preconditioner $P_B$ built
using an operator $B$ which is usually $A$ but not necessarily. If the
type of ksp is \pk{KSPHPDDM}, the solution phase, invoked by \pk{KSPSolve}(ksp, $b$, $x$), will be carried out by HPDDM. To do
so, HPDDM will repeatedly call the following PETSc routines:
\begin{itemize}
    \item \pk{MatMult}($A$, $x$, $y$) for $y = Ax$;
    \item \pk{MatMatMult}($A$, $X$, {\scriptsize \pk{MAT\_REUSE\_MATRIX}}, {\scriptsize
        \pk{PETSC\_DEFAULT}}, $Y$)\footnote{or \pk{MatProductNumeric}($Y$) with PETSc 3.14.0 and above} for $Y =
        AX$;
    \item \pk{PCApply}($P_B$, $y$, $x$) for $x = P_By$.
\end{itemize}
Unfortunately, PETSc does not have a general routine for applying $P_B$ to a
tall-and-skinny dense matrix $Y$. In HPDDM, a specialized routine has been implemented
to compute $X = P_BY$ if the preconditioner is of one of the following type:
\begin{itemize}
    \item exact $LU$ or Cholesky factorization, i.e., $P_B = B^{-1}$, \pk{PCLU} or
        \pk{PCCHOLESKY};
    \item incomplete $LU$ or Cholesky factorization, \pk{PCILU} or \pk{PCICC};
    \item block Jacobi with a single subdomain per process and exact or
        incomplete subdomain solvers, \pk{PCBJACOBI};
    \item overlapping Schwarz methods with a single subdomain per process and
        exact or incomplete subdomain solvers, \pk{PCASM}.
\end{itemize}
Indeed, in these cases, it is possible to get access to a factored matrix $F$
and call \pk{MatMatSolve}($F$, $\tilde{Y}$, $\tilde{X}$). With the two first sets of
preconditioners, $\tilde{Y} = Y$ and $\tilde{X} = X$. With block Jacobi,
$\tilde{Y}$, resp. $\tilde{X}$, represents only the part of the matrix $Y$,
resp. $X$, local to a given process. With overlapping Schwarz methods, there are
furthermore overlapping ``ghost'' unknowns, interested readers are referred to
monographs on practical aspects of domain decomposition
methods~\cite{smith2004domain,dolean2014ddm}. For all other types of
preconditioners, HPDDM will loop over the columns of $Y$ and apply $P_B$
successively to get all columns of $X$.

The rest of the operations are then performed directly inside HPDDM. For
example, in all Krylov methods using the Arnoldi process, it is mandatory to orthonormalize
candidate basis vectors. It would be impractical to try to reuse PETSc routines
like \pk{KSPGMRESModifiedGramSchmidtOrthogonalization} for two reasons: 1) it does
not handle the orthonormalization of multiple vectors at once which is needed in
pseudo-block methods 2) HPDDM may be used without PETSc as a stand-alone
library. To give some more technical details to curious readers, in HPDDM, $QR$
factorizations of tall-and-skinny dense matrices, which are obviously needed for
block Krylov methods but also for recycling Krylov methods are computed using
the CholQR algorithm~\cite{stathopoulos2002block}, or the block modified or classical
Gram--Schmidt method. This may be adjusted with the option -ksp\_hpddm\_qr (cholqr$|$mgs$|$cgs).
For Hessenberg matrices generated by the Arnoldi
process, it is common to update their $QR$ decomposition using Givens
rotations~\cite{saad2003iterative}. For block Hessenberg matrices, Householder reflectors are
used instead~\cite{gutknecht2008updating}.

To monitor convergence of a \pk{KSP} object, calls to the PETSc \pk{KSPMonitor} routine
were added inside HPDDM Krylov methods. 
    \subsubsection{User-defined deflation space} Recycling Krylov methods try to
extract information at the end of each cycle or when convergence is reached by
solving small standard or generalized dense eigenproblems. Then, during a
subsequent solve, this information is reused appropriately. Two new routines were
added in PETSc to either get the information generated by such recycling
methods, or set the information in case users know what is a good initial
deflation subspace. These routines are respectively
\pk{KSPHPDDMGetDeflationSpace} and \pk{KSPHPDDMSetDeflationSpace}. The subspaces
are once again stored as distributed
dense tall-and-skinny matrices.
    \subsubsection{MatKAIJ and systems with multiple right-hand sides\label{sec:matkaij}}
There is a wide range of available matrix types in PETSc. One of them is
\pk{MatKAIJ}, for which it is assumed that an operator may be written as $I \otimes S
+ A \otimes T$, with $S$ and $T$ elements of $\mathbb{K}^{p\times q}$ and
$A$ a standard sparse matrix of dimension $n$. If $p=q=k$, $S = \lambda I$, and $T =
\mu I$, then solving $(I \otimes S + A \otimes T) x = b$ may be cast into a
system with multiple right-hand sides $(\lambda I + \mu A)X = B$ where $B$ is
the row-major representation of $b$, i.e., $b_{k \cdot i + j} = B_{i, j},
\forall (i, j) \in \llbracket 1;n\rrbracket \times \llbracket 1;k\rrbracket$.
On the one hand, due to the structure of the operator, it is challenging to
define appropriate preconditioners for $(I \otimes S + A \otimes T)$: only
\pk{PCPBJACOBI} is currently implemented for \pk{MatKAIJ}. On the other hand,
it may be simpler to efficiently approximate the inverse of $A$. In
\pk{KSPHPDDM}, the type of the operators is checked and if it is \pk{MatKAIJ},
the single right-hand side system is automatically cast into a system with
multiple right-hand sides instead. The standard function \pk{KSPSolve} takes as
input argument a \pk{KSP} and two \pk{Vec}. A new function
\pk{KSPMatSolve} was added in PETSc 3.14.0, which takes as input argument
a \pk{KSP} and two dense \pk{Mat}, for solving true
multiple right-hand side systems. If the input \pk{KSP} is not of type
\pk{KSPHPDDM}, the solution phase is done in a column by column fashion.
Otherwise, it is possible to use any \pk{KSPHPDDM} (pseudo-)block Krylov methods.
Furthermore, a function \pk{KSPSetMatSolveBlockSize} is also provided to decompose
a single large block of column vectors into multiple sub-blocks.
\pk{KSPMatSolve} is then called repeatedly until all sub-blocks are traversed.
This was inspired by MUMPS~\cite{amestoy2001fully} option
ICNTL(27)\footnote{\url{http://mumps.enseeiht.fr/doc/userguide_5.3.1.pdf}, section 6.1}.
  \subsection{Applications and numerical results}
    \subsubsection{Reproducibility of the results from Parks et
    al.~\cite{parks2006recycling}}
Alongside the paper introducing GCRODR, a MATLAB implementation was provided and
is since then
available\footnote{\url{https://www.sandia.gov/~mlparks/GCRODR.zip}}. It comes
with a sequence of ten ``linear systems from a finite element fracture mechanics
problem constructed by Philippe H.\ Geubelle and Spandan Maiti." The goal of
this paragraph is to explain how the results have been reproduced in PETSc.
Thanks to \pk{KSPHPDDM}, it is quite straightforward. First, the sequence of
MATLAB systems had to be converted to PETSc \pk{Mat} and \pk{Vec}. Then, it is
just a matter of calling \pk{KSPSolve} in sequence. The dataset and the source
code are available at
\url{https://gitlab.com/petsc/datafiles/-/tree/master/matrices/hpddm/GCRODR}
and
\url{https://www.mcs.anl.gov/petsc/petsc-master/src/ksp/ksp/tutorials/ex75.c.html}
respectively. In order to check the correctness of the \pk{KSPHPDDM} interface,
the following tests are performed:
\begin{itemize}
    \item in MATLAB, unpreconditioned GCRODR(40, 20), $ICC(0)$ left-preconditioned
        GCRODR(40, 20), and Jacobi right-preconditioned GCRODR(40, 20);
    \item in PETSc with \pk{KSPGMRES}, unpreconditioned unrestarted GMRES($\infty$);
    \item in PETSc with \pk{KSPHPDDM}, all of the above.
\end{itemize}
Except for the right-preconditioned GCRODR, these tests are the same as the
ones from the original GCRODR paper~\cite{parks2006recycling}. Furthermore,
\pk{PCICC} and ichol from MATLAB are not consistent. All PETSc tests are run on four MPI
processes. They can be launched using the following command line. \\[-4pt]
\begin{Verbatim}[fontsize=\footnotesize,frame=single,framerule=0.1mm,commandchars=&\[\]]
&fvtextcolor[mygreen][$] &fvtextcolor[myyellow][mpirun] -n 4 ./ex75 -ksp_converged_reason -pc_type none -ksp_rtol 1e-10             &fvtextcolor[myred][\]
  -ksp_type hpddm -ksp_hpddm_type gcrodr -ksp_gmres_restart 40 -ksp_hpddm_recycle 20 &fvtextcolor[myred][\] 
  -load_dir &fvtextcolor[myyellow][${DATAFILESPATH}]/matrices/hpddm/GCRODR
&fvtextcolor[mygreen][$] &fvtextcolor[myyellow][mpirun] -n 4 ./ex75 -ksp_converged_reason -sub_pc_type icc -ksp_rtol 1e-10          &fvtextcolor[myred][\]
  -ksp_type hpddm -ksp_hpddm_type gcrodr -ksp_gmres_restart 40 -ksp_hpddm_recycle 20 &fvtextcolor[myred][\] 
  -load_dir &fvtextcolor[myyellow][${DATAFILESPATH}]/matrices/hpddm/GCRODR
&fvtextcolor[mygreen][$] &fvtextcolor[myyellow][mpirun] -n 4 ./ex75 -ksp_converged_reason -pc_type jacobi -ksp_rtol 1e-10           &fvtextcolor[myred][\]
  -ksp_type hpddm -ksp_hpddm_type gcrodr -ksp_gmres_restart 40 -ksp_hpddm_recycle 20 &fvtextcolor[myred][\] 
  -ksp_pc_side right -load_dir &fvtextcolor[myyellow][${DATAFILESPATH}]/matrices/hpddm/GCRODR
&fvtextcolor[mygreen][$] &fvtextcolor[myyellow][mpirun] -n 4 ./ex75 -ksp_converged_reason -pc_type none -ksp_rtol 1e-10             &fvtextcolor[myred][\]
  -ksp_type hpddm -ksp_gmres_restart 500                                             &fvtextcolor[myred][\] 
  -load_dir &fvtextcolor[myyellow][${DATAFILESPATH}]/matrices/hpddm/GCRODR
\end{Verbatim}
The various number of iterations needed to reach the convergence criterion as
originally set to $10^{-10}$ are gathered in~\cref{fig:comparison}. Apart from
the tests \ref{pgfplots:r2} and \ref{pgfplots:r3} using $ICC(0)$ and \ref{pgfplots:r4} and \ref{pgfplots:r5}
using the Jacobi method, one may notice that \ref{pgfplots:r0} and
\ref{pgfplots:r1} (resp. \ref{pgfplots:r6} and \ref{pgfplots:r7}) are almost a
perfect match and reproduce partially Figure 4.2 from Parks et
al.~\cite{parks2006recycling}.
\pgfplotstableread{
sys none-matlab icc-matlab jacobi-matlab none-hpddm icc-hpddm jacobi-hpddm icc-gmres none-gmres icc-petsc none-petsc
401 498         293        457           497        131       460          127       438        127       438
402 228         125        187           231        53        184          127       449        127       449
403 209         113        182           206        47        177          127       450        127       450
404 199         112        175           198        47        174          127       450        127       450
405 198         110        174           198        47        172          127       451        127       451
406 198         111        174           199        47        174          127       452        127       452
407 206         114        180           206        49        179          126       458        126       458
408 207         115        181           208        49        182          128       458        135       458
409 207         110        181           207        49        181          128       458        128       458
410 206         110        179           206        49        179          128       457        128       457
}\loadedtable
    \begin{figure}
\centering \begin{tikzpicture}
    \begin{axis}[xlabel=System index,ylabel={\# of iterations},xmin=401,xmax=410,ymin=0,ymajorgrids,
        legend columns=1,enlarge x limits=0.05,
        legend style={font=\small,at={(1.02,0.25)}, legend cell align=left,
        anchor=south west}]
\addplot[very thick,myred]                       table[x expr=\thisrowno{0},y expr=\thisrowno{1}, col sep=space] {\loadedtable};\label{pgfplots:r0}
\addplot[very thick,mark=*,only marks,myred]     table[x expr=\thisrowno{0},y expr=\thisrowno{4}, col sep=space] {\loadedtable};\label{pgfplots:r1}
\addplot[very thick,myaqua]                      table[x expr=\thisrowno{0},y expr=\thisrowno{2}, col sep=space] {\loadedtable};\label{pgfplots:r2}
\addplot[very thick,mark=pentagon*,only marks,myaqua]    table[x expr=\thisrowno{0},y expr=\thisrowno{5}, col sep=space] {\loadedtable};\label{pgfplots:r3}
\addplot[very thick,myyellow]                    table[x expr=\thisrowno{0},y expr=\thisrowno{3}, col sep=space] {\loadedtable};\label{pgfplots:r4}
\addplot[very thick,mark=square*,only marks,myyellow]  table[x expr=\thisrowno{0},y expr=\thisrowno{6}, col sep=space] {\loadedtable};\label{pgfplots:r5}
\addplot[very thick,last]                      table[x expr=\thisrowno{0},y expr=\thisrowno{10}, col sep=space]{\loadedtable};\label{pgfplots:r6}
\addplot[very thick,mark=triangle*,only marks,last]    table[x expr=\thisrowno{0},y expr=\thisrowno{8}, col sep=space] {\loadedtable};\label{pgfplots:r7}
\legend{{MATLAB unpreconditioned GCRODR},{\pk{KSPHPDDM}},% unpreconditioned GCRODR(40, 20)},
        {MATLAB $ICC(0)$-preconditioned GCRODR},{\pk{KSPHPDDM}},% $ICC(0)$-preconditioned GCRODR(40, 20)},
        {MATLAB Jacobi-preconditioned GCRODR},{\pk{KSPHPDDM}},% Jacobi-preconditioned GCRODR(40, 20)},
        {Unpreconditioned \pk{KSPGMRES}},{\pk{KSPHPDDM}}}% unpreconditioned GMRES($\infty$)}}
\end{axis}
\end{tikzpicture}
        \caption{\# of iterations needed to converge for various
        configurations as originally tested by Parks et
        al.~\cite{parks2006recycling}. Note that \pk{PCICC} and
        \textnormal{ichol} from MATLAB are not consistent, hence the discrepancy
        between \ref{pgfplots:r2} and \ref{pgfplots:r3}. All other \pk{KPSHPDDM}
        numbers of iterations match with the respective MATLAB or PETSc
        reference implementation.\label{fig:comparison}}
    \end{figure}
    \subsubsection{Linear stability analysis}
SLEPc~\cite{hernandez2005ssf}, the Scalable Library for Eigenvalue Problem
computations, is built on top of PETSc. Thus, it is possible to reuse and
compose \pk{KSP} or preconditioners for solving eigenproblems efficiently, e.g.,
through spectral transformations. Recently, it has been observed that in the
field of computational fluid dynamics, the role of recycling can be quite
significant when performing linear stability analysis~\cite{moulin2018al}. 
The following generalized eigenvalue problem is considered next:
\begin{equation}\label{eq:lsa}
    J(q_b)x = \lambda \begin{bmatrix}M & 0 \\ 0 & 0 \end{bmatrix} x,
\end{equation}
where $J$ is the Jacobian of the incompressible steady-state Navier--Stokes equation
evaluated using a given base flow $q_b = \begin{bmatrix}u_b & p_b\end{bmatrix}^T$ and $M$ the discretization of the mass matrix
on the space of velocities. There are many ways to compute the eigenvalues
of~\cref{eq:lsa} near a complex-valued shift $\sigma$, but emphasis will be put
in this paragraph on Krylov--Schur methods~\cite{Stewart2002}. These rely on the successive
solution of linear systems such as:
\begin{equation}\label{eq:ks}
    \left(J(q_b) - \sigma \begin{bmatrix}M&0\\0&0\end{bmatrix}\right)x_i = b_i, i = 1,2,\ldots
\end{equation}
In SLEPc, these systems may be parameterized using the -st\_ prefix, e.g.,
-st\_pc\_type~lu. If they are solved iteratively, it is possible to
use a \pk{KSP} of type \pk{KSPHPDDM} to recycle information between
successive linear systems yielded by the convergence of the Krylov--Schur
algorithm. Using the FreeFEM~\cite{hecht2012new} mini-app from \url{https://github.com/prj-/moulin2019al}
which precondition~\cref{eq:ks} using a modified augmented Lagrangian
approach~\cite{Benzi2011a}, the effect of such recycling is shown next for
finding the 5 eigenpairs closest to the shift $\sigma = 10^{-6} + 0.6\mathfrak{i}$ for a
flow past a cylinder at Reynolds 100, interested readers are again referred to
the previously cited paper~\cite{moulin2018al} for more details and for larger
runs. Results can be reproduced using the following four commands, the first two
being merely for generating the base flow $q_b$ with \pk{SNESSolve} using a
continuation method on the Reynolds number. \\[-4pt]
\begin{Verbatim}[fontsize=\footnotesize,frame=single,framerule=0.1mm,commandchars=&\[\]]
&fvtextcolor[mygreen][$] &fvtextcolor[myyellow][mpirun] -n 4 FreeFem++-mpi Nonlinear-solver.edp -Re 50 -v 0 
&fvtextcolor[mygreen][$] &fvtextcolor[myyellow][mpirun] -n 4 FreeFem++-mpi Nonlinear-solver.edp -Re 100 -v 0 
&fvtextcolor[mygreen][$] &fvtextcolor[myyellow][mpirun] -n 4 FreeFem++-mpi Eigensolver.edp -Re 100 -v 0 -st_ksp_rtol 1.0e-4         &fvtextcolor[myred][\] 
  -st_ksp_type fgmres -st_ksp_gmres_restart 200 -st_ksp_converged_reason
&fvtextcolor[mygreen][$] &fvtextcolor[myyellow][mpirun] -n 4 FreeFem++-mpi Eigensolver.edp -Re 100 -v 0 -st_ksp_rtol 1.0e-4         &fvtextcolor[myred][\] 
  -st_ksp_type hpddm -st_ksp_gmres_restart 200 -st_ksp_converged_reason              &fvtextcolor[myred][\] 
  -st_ksp_hpddm_variant flexible -st_ksp_hpddm_recycle 10 -st_ksp_hpddm_type gcrodr
\end{Verbatim}
The Krylov--Schur algorithm converges in six outer iterations. In each of these
iterations, multiple linear systems having the structure of~\cref{eq:ks} are
solved. In~\cref{fig:comparison-slepc}, numbers of iterates per inner solves
are reported. After the first inner solve, once recycling kicks in, one may
clearly see the benefits of switching from \pk{KSPFGMRES} to \pk{KSPHPDDM}.
Indeed, through the 46 inner solves, these numbers are roughly decreased by 25. In terms of runtime, \SI[round-mode          = places,round-precision=1]{1.5098e+03}{\sec} (resp.~\SI[round-mode          = places,round-precision=1]{1.0524e+03}{\sec}) is spent in \pk{EPSSolve} when using \pk{KSPFGMRES} (resp.~\pk{KSPHPDDM}). These timings, and all which follow, have been obtained on {Ir\`ene}, a system composed of \pgfmathprintnumber[assume math mode=true]{1656} nodes with two 24-core Intel Xeon Platinum 8168 clocked at \SI{2.7}{\giga\hertz}. 
\pgfplotstableread{
sys hpddm fgmres
1   77    77
2   54    83
3   53    83
4   52    81
5   52    81
6   51    78
7   51    79
8   51    80
9   51    80
10  51    77
11  51    79
12  51    79
13  51    79
14  53    77
15  52    79
16  49    78
17  51    77
18  51    79
19  52    79
20  52    79
21  51    77
22  51    78
23  51    79
24  53    80
25  54    79
26  52    79
27  49    75
28  51    78
29  51    78
30  51    77
31  50    78
32  55    81
33  56    79
34  54    80
35  53    80
36  53    78
37  50    78
38  52    79
39  53    77
40  52    77
41  54    80
42  53    80
43  55    80
44  53    79
45  52    77
46  53    80
}\loadedtableSLEPc
    \begin{figure}
\centering \begin{tikzpicture}
    \begin{axis}[xlabel={Krylov--Schur system index},ylabel={\# of
        iterations},xmin=1,xmax=46,ymin=0,ymajorgrids,
        legend columns=1,enlarge x limits=0.05,
        legend style={font=\small,at={(0.5,0.24)}, legend cell align=left,
        anchor=south west}]
\addplot[very thick,last]     table[x expr=\thisrowno{0},y expr=\thisrowno{2}, col sep=space] {\loadedtableSLEPc};\label{pgfplots:m0}
\addplot[very thick,myred,densely dashed]       table[x expr=\thisrowno{0},y expr=\thisrowno{1}, col sep=space] {\loadedtableSLEPc};\label{pgfplots:m1}
\legend{{\pk{KSPFGMRES}},{\pk{KSPHPDDM}}}
\end{axis}
\end{tikzpicture}
        \caption{\# of inner iterations for Krylov--Schur systems~\cref{eq:ks} with~\ref{pgfplots:m0} or without~\ref{pgfplots:m1} recycling.\label{fig:comparison-slepc}}
    \end{figure}
\pgfplotstableread{
N    baij-1-T    baij-1-F   sbaij-1-T   sbaij-1-F  aij-1-T     aij-1-F    baij-3-T    baij-3-F   sbaij-3-T   sbaij-3-F  aij-3-T     aij-3-F    baij-10-T   baij-10-F  sbaij-10-T  sbaij-10-F aij-10-T    aij-10-F
1    1.8027e-01  1.4696e+08 6.1586e-02  1.4696e+08 8.0527e-02  1.4696e+08 7.1397e-01  1.3535e+09 5.1037e-01  1.3638e+09 1.0686e+00  1.3535e+09 6.1909e+00  1.5159e+10 4.6958e+00  1.5206e+10 1.6114e+01  1.5159e+10
2    3.5192e-01  2.9392e+08 1.6650e-01  3.2482e+08 1.8615e-01  3.0422e+08 1.0437e+00  2.7071e+09 9.3575e-01  2.9234e+09 1.4753e+00  2.7380e+09 1.4970e+01  3.0319e+10 1.5530e+01  3.2482e+10 1.6267e+01  3.0422e+10
8    6.1599e-01  1.1757e+09 7.2168e-01  1.2993e+09 3.1676e-01  1.2169e+09 2.6101e+00  1.0828e+10 3.0701e+00  1.1694e+10 2.0870e+00  1.0952e+10 2.7886e+01  1.2128e+11 3.4780e+01  1.2993e+11 2.2569e+01  1.2169e+11
16   1.9594e+00  2.3513e+09 2.6186e+00  2.5986e+09 6.2770e-01  2.4337e+09 7.5274e+00  2.1656e+10 9.8057e+00  2.3387e+10 4.1388e+00  2.1904e+10 3.6085e+01  2.4255e+11 3.6290e+01  2.5986e+11 4.5153e+01  2.4337e+11
32   4.3453e+00  4.7026e+09 5.4669e+00  5.1972e+09 1.2538e+00  4.8675e+09 1.6166e+01  4.3313e+10 2.0638e+01  4.6775e+10 8.1531e+00  4.3807e+10 5.2586e+01  4.8510e+11 5.2132e+01  5.1972e+11 1.0705e+02  4.8675e+11
64   9.6799e+00  9.4053e+09 1.1616e+01  1.0394e+10 2.5713e+00  9.7350e+09 3.4513e+01  8.6626e+10 4.2465e+01  9.3549e+10 1.6444e+01  8.7615e+10 8.3267e+01  9.7020e+11 8.6683e+01  1.0394e+12 2.2792e+02  9.7350e+11
}\loadedMatProduct
    \subsubsection{Performance of primitives for block Krylov methods}
As detailed in~\cref{sec:interface-ksp}, it is currently not possible with most
PETSc preconditioners to apply them to a block of vectors. Even though there
are efforts to enable such an operation for an increasing number of \pk{PC}, in
this section, the focus is instead put on the matrix--matrix multiplication
needed in block Krylov methods or when restarting recycling Krylov methods.
In particular, the performance of three PETSc matrix types are evaluated:
\begin{itemize}
    \item \pk{MatSeqAIJ}, standard sequential sparse matrices, based on compressed sparse row format;
    \item \pk{MatSeqBAIJ}, block sparse matrices, based on block compressed sparse row format;
    \item \pk{MatSeqSBAIJ}, symmetric block sparse matrices.
\end{itemize}
Parallel matrices in PETSc heavily rely on sequential matrices. For example, \pk{MatMPIAIJ}, standard parallel sparse matrices, store on each rank two \pk{MatSeqAIJ}, one with local rows and columns, another with local rows and ``nonlocal'' columns.
To keep this study succinct, only sequential sparse matrices will be evaluated next, which can help draw conclusions for the intraprocess performance of parallel sparse matrices. The Seq substring is now dropped, since in this case, \pk{MatSeqAIJ} is equivalent to \pk{MatAIJ} and so on.

The Poisson equation on a cube is discretized with $\mathbb{P}_1$ finite elements in double precision. This yields a matrix $A$ of dimension one million. Though the numerical values of $A$ are of no interest here, the sparsity pattern is rather common and frequently encountered, e.g., when discretizing partial differential equations. Then, a random dense matrix $T$ of dimension $b$ is generated, for $b \in \{ 1,3,10 \}$, and the following matrix is evaluated $\mathbb{A} = A \otimes T$. The matrix $\mathbb{A}$ is then assembled into all three formats described above. For block matrices, a block size of $b$ is used. The performance of the \pk{MatProductNumeric} operation is evaluated for tall-and-skinny dense matrices with a varying number of columns $N \in \{1,2,8,16,32,64\}$. For $N=1$, the \pk{MatMult} operation for matrix--vector multiplication is used instead. All results are gathered in~\cref{fig:perfAX}. They can be reproduced using the following mini-app: \url{http://jolivet.perso.enseeiht.fr/MatProduct.cpp} and any input \pk{MatAIJ} stored in binary format, here using the default name binaryoutput.
\begin{Verbatim}[fontsize=\footnotesize,frame=single,framerule=0.1mm,commandchars=&\[\]]
&fvtextcolor[mygreen][$] &fvtextcolor[myyellow][mpicxx] MatProduct.cpp -O3 -std=c++11 -I&fvtextcolor[myyellow][${PETSC_DIR}]/&fvtextcolor[myyellow][${PETSC_ARCH}]/include                     &fvtextcolor[myred][\]
  -I&fvtextcolor[myyellow][${PETSC_DIR}]/include -L&fvtextcolor[myyellow][${PETSC_DIR}]/&fvtextcolor[myyellow][${PETSC_ARCH}]/lib -lpetsc -o MatProduct
&fvtextcolor[mygreen][$] &fvtextcolor[myyellow][mpirun] -n 1 ./MatProduct -f binaryoutput -log_view                                 &fvtextcolor[myred][\]
  -bs 1,3,10 -N 1,2,8,16,32,64 -type aij,baij,sbaij
\end{Verbatim}
\begin{figure}[htbp]
  \centering
    \begin{tikzpicture}
      \begin{semilogxaxis}[
        ymin=0,
        ybar=1.5pt,
        bar width=3.5pt,
        enlarge x limits=0.1, height=4cm,
        width=0.98\textwidth,legend columns=4,
        legend style={at={(0.5,1.65)}, legend cell align=left,
        anchor=north},
        xlabel={\# of right-hand sides ($N$)},
        ylabel={Efficiency},
        xtick = {1,2,4,8,16,32},
ytick = {0.2,1,2.5,5},ylabel style={yshift=6pt},xtick pos=left,
        xticklabels={1,2,8,16,32,64},
        ymajorgrids=true,yticklabel={
\pgfmathparse{int(10*\tick)}
\ifnum\pgfmathresult=0
\else
\pgfmathparse{100*\tick}
$\pgfmathprintnumber{\pgfmathresult}$\%
\fi
},
        %xmajorgrids=true,
          cycle list={blueaccent, darkgray, mediumgray},%, lightgray,last},
      ]
\pgfkeys{/pgf/fpu}
  \pgfmathparse{1.4696e+08/8.0527e-02/1000000000.0}\edef\storeresultI{\pgfmathresult}
  \pgfmathparse{1.3535e+09/7.1397e-01/1000000000.0}\edef\storeresultII{\pgfmathresult}
  \pgfmathparse{1.5159e+10/6.1909e+00/1000000000.0}\edef\storeresultIII{\pgfmathresult}
\pgfkeys{/pgf/fpu=false}
\node[font=\tiny,rotate=45,fill=white,inner sep=0pt,anchor=west] at (axis cs:0.75, 1.2) {\pgfmathprintnumber\storeresultI~GFLOP/s};
\node[font=\tiny,rotate=45,fill=white,inner sep=0pt,anchor=west] at (axis cs:0.97, 1.2) {\pgfmathprintnumber\storeresultII~GFLOP/s};
\node[font=\tiny,rotate=45,fill=white,inner sep=0pt,anchor=west] at (axis cs:1.18, 1.2) {\pgfmathprintnumber\storeresultIII~GFLOP/s};
\addlegendimage{empty legend}
          \addplot+[draw=black,fill, area legend,postaction={pattern=crosshatch dots}] table [x expr=-0.5 + \thisrow{N}>7?\thisrow{N}/2:\thisrow{N},y expr=\thisrow{aij-1-F}/\thisrow{aij-1-T}/(1.4696e+08/8.0527e-02)]{\loadedMatProduct};\label{pgfplots:aij1}
          \addplot+[draw=black,fill, area legend,postaction={pattern=crosshatch dots}] table [x expr=-0.5 + \thisrow{N}>7?\thisrow{N}/2:\thisrow{N},y expr=\thisrow{baij-1-F}/\thisrow{baij-1-T}/(1.4696e+08/8.0527e-02)]{\loadedMatProduct};\label{pgfplots:baij1}
          \addplot+[draw=black,fill, area legend,postaction={pattern=crosshatch dots}] table [x expr=-0.5 + \thisrow{N}>7?\thisrow{N}/2:\thisrow{N},y expr=\thisrow{sbaij-1-F}/\thisrow{sbaij-1-T}/(1.4696e+08/8.0527e-02)]{\loadedMatProduct};\label{pgfplots:sbaij1}
\addlegendimage{empty legend}
           \addplot+[draw=black,fill, area legend] table [x expr=\thisrow{N}>7?\thisrow{N}/2:\thisrow{N},y expr=\thisrow{aij-3-F}/\thisrow{aij-3-T}/(1.3535e+09/7.1397e-01)]{\loadedMatProduct};\label{pgfplots:aij3}
           \addplot+[draw=black,fill, area legend] table [x expr=\thisrow{N}>7?\thisrow{N}/2:\thisrow{N},y expr=\thisrow{baij-3-F}/\thisrow{baij-3-T}/(1.3535e+09/7.1397e-01)]{\loadedMatProduct};\label{pgfplots:baij3}
           \addplot+[draw=black,fill, area legend] table [x expr=\thisrow{N}>7?\thisrow{N}/2:\thisrow{N},y expr=\thisrow{sbaij-3-F}/\thisrow{sbaij-3-T}/(1.3535e+09/7.1397e-01)]{\loadedMatProduct};\label{pgfplots:sbaij3}
\addlegendimage{empty legend}
           \addplot+[draw=black,fill, area legend,postaction={pattern=bricks}] table [x expr=+0.5 + \thisrow{N}>7?\thisrow{N}/2:\thisrow{N},y expr=\thisrow{aij-10-F}/\thisrow{aij-10-T}/(1.5159e+10/6.1909e+00)]{\loadedMatProduct};\label{pgfplots:aij10}
           \addplot+[draw=black,fill, area legend,postaction={pattern=bricks}] table [x expr=+0.5 + \thisrow{N}>7?\thisrow{N}/2:\thisrow{N},y expr=\thisrow{baij-10-F}/\thisrow{baij-10-T}/(1.5159e+10/6.1909e+00)]{\loadedMatProduct};\label{pgfplots:baij10}
           \addplot+[draw=black,fill, area legend,postaction={pattern=bricks}] table [x expr=+0.5 + \thisrow{N}>7?\thisrow{N}/2:\thisrow{N},y expr=\thisrow{sbaij-10-F}/\thisrow{sbaij-10-T}/(1.5159e+10/6.1909e+00)]{\loadedMatProduct};\label{pgfplots:sbaij10}
          \addlegendentry{$b = 1$}
          \addlegendentry{              }
          \addlegendentry{               }
          \addlegendentry{                }
          \addlegendentry{$b = 3$}
          \addlegendentry{\pk{MatAIJ}}
          \addlegendentry{\pk{MatBAIJ}}
          \addlegendentry{\pk{MatSBAIJ}}
          \addlegendentry{$b = 10$~~}
          \addlegendentry{              }
          \addlegendentry{               }
          \addlegendentry{                }
    \end{semilogxaxis}
    \end{tikzpicture}
   \caption{Performance of the matrix--matrix multiplication as implemented in PETSc for three sparse matrix types. As baseline values with $N=1$, for $b=1$, performance of \pk{MatAIJ} matrix--vector multiplication is used~\ref{pgfplots:aij1}, while for $b=3$ (resp.~$b=10$), it is the performance of \pk{MatBAIJ} matrix--vector multiplication~\ref{pgfplots:baij3} (resp.~\ref{pgfplots:baij10}).\label{fig:perfAX}}
\end{figure}
Some conclusions may be drawn:
\begin{itemize}
    \item with a block size of 1, using any type but \pk{MatAIJ} is counterproductive, see~\ref{pgfplots:baij1} and~\ref{pgfplots:sbaij1}. However, with this format, as soon as the number of columns is greater or equal than 8, the efficiency reaches approximately 200\%, see~\ref{pgfplots:aij1};
    \item for block sizes lower or equal than 5, loops for matrix--vector and matrix--matrix multiplications are unrolled by hand for block formats. This yields rather disappointing performances, except for \pk{MatAIJ}, see~\ref{pgfplots:aij3}, and its 250\% efficiency, even with a moderate number of columns. For block formats \ref{pgfplots:baij3} and~\ref{pgfplots:sbaij3}, given the small value of $b$, it could be beneficial to switch to optimized libraries for small matrices, e.g., LIBXSMM~\cite{heinecke2016libxsmm};
    \item for greater block sizes, ?gemv or ?gemm are used instead for block formats. For \pk{MatAIJ}~\ref{pgfplots:aij10}, the efficiency quickly caps below 250\% for $N\in\{8,16\}$, but for block formats, this reaches almost 500\% for large number of columns, see~\ref{pgfplots:baij10} and~\ref{pgfplots:sbaij10} for $N=64$.
\end{itemize}
These results are of course specific to the sparsity pattern of the initial matrix $A$, and thus $\mathbb{A} = A\otimes T$, and matrix types. For example, if the matrices were to be offloaded to a GPU, one could instead benchmark \pk{MatAIJcuSPARSE}.
\section{PCHPDDM}\label{sec:PCHPDDM}
  \subsection{Related work}
Domain decomposition methods are, alongside multigrid methods, one of the
dominant paradigms for defining efficient and robust preconditioners in
modern large-scale applications dealing with partial differential equations.
In PETSc, preconditioners inherit from the \pk{PC} base class.
Historically, the block Jacobi and overlapping Schwarz methods were implemented
as \pk{PCBJACOBI} and \pk{PCASM} respectively. ASM stands for additive Schwarz
method, as originally introduced by Schwarz~\cite{schwarz1870alternating}, but
the PETSc implementation is more versatile in the sense that one can easily
switch between this method and other variants such as the restricted additive
Schwarz method~\cite{cai1999restricted}. More details about overlapping domain decomposition
preconditioners will be given in~\cref{sec:interface-pc}, but first the other
related preconditioners will be quickly presented.

In PETSc, there is also an implementation of two-level overlapping Schwarz
methods using a generalized Dryja, Smith, Widlund
preconditioner~\cite{dohrmann2008domain}. Its type is \pk{PCEXOTIC}. When it comes to
nonoverlapping preconditioners, \pk{PCNN} implements a basic balancing
Neumann--Neumann method~\cite{mandel1993balancing}. Much more advanced
techniques are available in \pk{PCBDDC}~\cite{zampini2016pcbddc}: adaptive selection
of primal constraints~\cite{pechstein2017unified}, change of basis for problems in
$H(\text{curl})$~\cite{dohrmann2016bddc}$\ldots$ For multigrid,
\pk{PCMG} exposes a machinery for geometric or algebraic preconditioners. In
particular, it is used for the implementation of the smoothed-aggregation
\pk{PCGAMG}~\cite{adams2004ultra}. There are also interfaces to other well-known
multigrid packages such as \pk{PCHYPRE} or \pk{PCML}~\cite{gee2006ml} from Trilinos. Of
course, there are many other available domain decomposition preconditioners,
either accessible through high-level libraries, e.g., FROSch~\cite{Heinlein:2016:PIT} in
Trilinos, or as stand-alone packages, e.g., BDDCML~\cite{vsistek2013parallel} or
FEMPAR-BDDC~\cite{badia2014highly}.

HPDDM implements the same one-level overlapping Schwarz methods as PETSc, albeit
emphasis is also put on optimized Schwarz methods~\cite{gander2006optimized}.
The main theoretical framework for defining robust two-level methods is called GenEO,
for generalized eigenvalue problem on the overlap~\cite{spillane2011robust,spillane2013abstract}.
GenEO was also implemented in other libraries~\cite{butler2020high}, but to the best of our knowledge,
the interface between HPDDM and PETSc proposes the first high-level access to
GenEO so that it can, for example, be used as a block-decomposed
solver~\cite{brown2012composable}.

  \subsection{Interfacing HPDDM overlapping Schwarz methods in PETSc\label{sec:interface-pc}}\
\pk{PCHPDDM}, the interface between HPDDM overlapping Schwarz methods and PETSc, is automatically
registered since PETSc version 3.12 when one configures PETSc with the extra
flag -{}-download-hpddm -{}-download-slepc. It is then
possible to select the corresponding \pk{PC}
using the command line option -pc\_type hpddm, or the routine \pk{PCSetType}(pc,
\pk{PCHPDDM}) and its Fortran equivalent. In this section, it is assumed that one
wants to solve 
\begin{equation}
    Ax = b,
\end{equation}
for a given coefficient matrix $A$ and right-hand side $b$ of dimension $n$
assembled on $N$ processes, using a ksp($A$, $P_A$), i.e., the matrix used to
define the preconditioner is $A$ itself. Without anything more from the user,
\pk{PCHPDDM} is strictly equivalent to \pk{PCASM}. That is, if $A$ is distributed among
$N$ processes the preconditioner $M^{-1}$ is:
\begin{equation}\label{eq:one}
    M^{-1} = \sum_{i = 1}^N \tilde{R}_i^T (R_i A R_i^T)^{-1} R_i,
\end{equation}
where $\{R_i\}_{i=1}^N$ are restriction operators that act on a global vector and
return a vector local to each process, possibly with some overlap.
$\{\tilde{R}_i\}_{i=1}^N$ are the same operators except that coefficients on the
overlap are set to 0. The action of the inverse of all $\{R_i A R_i^T\}_{i=1}^N$
can be parameterized through a local \pk{KSP}. If it is assumed that $A$ is the
discretization of a linear operator $\mathcal{L}$, e.g., $\mathcal{L}=-\Delta$, on a domain $\Omega$, it is
extremely common to exhibit parallelism by distributing $\Omega$ on $N$
processes, possibly with some overlap, and to construct the $\{R_i\}_{i=1}^N$ so
that they map global unknowns to unknowns local to each process. In addition to
the local operators $\{R_i A R_i^T\}_{i=1}^N$, GenEO needs users to supply
the local unassembled matrices $\{\mathring{A}_i\}_{i=1}^N$, also known as
Neumann matrices. These are yielded by the
discretization of $\mathcal{L}$ on each subdomain, with added natural boundary
conditions on the subdomain boundaries that are not boundaries of the initial
domain $\Omega$. In the case of the 1D Poisson
equation with homogeneous Dirichlet boundary conditions discretized using $\mathbb{P}_1$ finite elements, an explicit representation of these various operators is given~\cref{fig:poisson} for a
two-domain decomposition.
\begin{figure}[t]
    \begin{minipage}[t]{0.495\textwidth}
\begin{center}
    Domain and global unknown numbering \\[0.3cm]
\begin{tikzpicture}[darkstyle/.style={circle,draw,fill=gray!40,line
    width=0.2mm,font=\scriptsize,minimum size=12pt,inner sep=0pt}]
    \foreach \x in {0,...,\dof}
        \pgfmathsetmacro\result{int(\x + 1)}
        \node[darkstyle] (\x) at (1.25*\x,0) {\result};
    \pgfmathsetmacro\result{\dof - 1}
    \foreach \x [count=\xi] in {0,...,\result}
        \draw[double=black, double distance=0.01mm] (\x) -- (\xi);
\end{tikzpicture}
\\[0.5cm]
    Two-way element-based partitioning \\[0.1cm]
\begin{tikzpicture}[darkstyle/.style={circle,draw,fill=gray!40,line
    width=0.2mm,minimum size=12pt, inner sep=0pt}]
    \foreach \x in {0,...,\dof}
        \node[darkstyle] (\x) at (1.25*\x,0) {};
    \foreach \sub/\c in {0/red!50,1/blue!50} {
        \pgfmathsetmacro\start{int(\sub * round(\dof / 2))}
        \ifnum\sub=1
            \pgfmathsetmacro\finish{int(\dof - 1)}
        \else
            \pgfmathsetmacro\finish{int(min((\sub + 1) * round(\dof / 2), \dof) - 1)}
        \fi
        \foreach \x [count=\xi from \start+1] in {\start,...,\finish}
            \draw[double=\c, double distance=0.2mm] (\x) -- (\xi);
        \pgfmathsetmacro\finish{int(\finish+1)}
        \pgfmathsetmacro\sub{int(\sub+1)}
        \draw[{Tee Barb[left]}-{Tee Barb[right]}, shorten <= -0.1cm, shorten >= -0.1cm] ([yshift=0.5cm]\start.east) -- node[fill=white,inner sep=1mm,midway] {$\Omega_\sub$} ([yshift=0.5cm]\finish.west);
    }
\end{tikzpicture}
\end{center}
\begin{equation*}
    {\scriptsize A =
    \begin{bmatrix}1&0&0&0&0\\0&2&-1&0&0\\0&-1&2&-1&0\\0&0&-1&2&0\\0&0&0&0&1\end{bmatrix}}
\end{equation*}
    \end{minipage}
    \begin{minipage}[t]{0.495\textwidth}
\begin{center}
    Local numbering with overlap \\[0.1cm]
\begin{tikzpicture}[darkstyle/.style={circle,font=\scriptsize,draw,fill=gray!40,line
    width=0.2mm,minimum size=12pt,inner sep=0}]
    \foreach \conf/\c in {0/red!50,1/blue!50} {
        \pgfmathsetmacro\start{int(max(min(\conf * round(\dof / 2) - \overlap, \dof - 1), 0))}
        \ifnum\conf=1
            \pgfmathsetmacro\finish{int(\dof - 1)}
        \else
            \pgfmathsetmacro\finish{int(max(0, min((\conf + 1) * round(\dof / 2) + \overlap, \dof) - 1))}
        \fi
        \foreach \x in {0,...,\dof} {
            \ifnum\conf=1
                \pgfmathparse{ifthenelse(\x < \start || \x - 1 > \finish, "",
                int(\finish - \x + 2))} \let\nn\pgfmathresult
            \else
                \pgfmathparse{ifthenelse(\x < \start || \x - 1 > \finish, "",
                int(\x - \start + 1))} \let\nn\pgfmathresult
            \fi
            \pgfmathparse{ifthenelse(\x < \start || \x - 1 > \finish, "0.25", "1")} \let\oo\pgfmathresult
            \node[darkstyle, opacity = \oo] (\x\conf) at (1.25*\x,-1.25cm*\conf)
            {\nn};
        }
        \pgfmathsetmacro\loop{int(\dof - 1)}
        \foreach \x [count=\xi from 1] in {0,...,\loop} {
            \pgfmathparse{ifthenelse(\x < \start || \x > \finish, "black!25", "\c")} \let\cc\pgfmathresult
            \pgfmathparse{ifthenelse(\x < \start || \x > \finish, "0.25", "1")} \let\oo\pgfmathresult
            \ifnum\conf=1
                \pgfmathparse{ifthenelse(\x + 1 < \start || \x > \finish, "", int(\finish - \x))} \let\nn\pgfmathresult
            \else
                \pgfmathparse{ifthenelse(\x < \start || \x - 1 > \finish, "", int(\x - \start))} \let\nn\pgfmathresult
            \fi
            \pgfmathparse{ifthenelse(\conf == 1, "0.875cm", "-0.875cm"} \let\ss\pgfmathresult
            \draw[double=\cc, opacity=\oo, double distance=0.2mm] (\x\conf) -- node[xshift=\ss, yshift=0.2cm, inner sep=1mm,midway, opacity=1] {} (\xi\conf);
        }
    }
\end{tikzpicture}
\begin{equation*}
    {\scriptsize R_2 =
    \begin{bmatrix}0&0&0&0&1\\0&0&0&1&0\\0&0&1&0&0\\0&1&0&0&0\end{bmatrix} \;
        \tilde{R}_2 =
    \begin{bmatrix}0&0&0&0&1\\0&0&0&1&0\\0&0&1&0&0\\0&0&0&0&0\end{bmatrix}}
\end{equation*}
\begin{equation*}
    \hspace*{-0.5cm}  {\scriptsize \mathring{A}_i =
    \begin{bmatrix}1&0&0&0\\0&2&-1&0\\0&-1&2&-1\\0&0&-1&1\end{bmatrix} \;
      R_2 A R_2^T =
    \begin{bmatrix}1&0&0&0\\0&2&-1&0\\0&-1&2&-1\\0&0&-1&2\end{bmatrix}}
\end{equation*}
\end{center}
    \end{minipage}
\caption{Notations and basics of overlapping Schwarz methods.\label{fig:poisson}}
\end{figure}
Then, it is possible to enrich the original one-level preconditioner
from~\cref{eq:one} using a spectral coarse grid built in the following way:
\begin{enumerate}
    \item have SLEPc solve the local generalized eigenvalue problem concurrently:
        \begin{equation}\label{eq:gevp}
            \mathring{A}_i \Lambda_i = \lambda_i \tilde{R}_i R_i^T (R_i A R_i^T) \tilde{R}_i R_i^T \Lambda_i ;
        \end{equation}
    \item retrieve the $\nu_i$ smallest eigenpairs $\{\Lambda_{i_j}, \lambda_{i_j}\}_{j=1}^{\nu_i}$ and assemble a local deflation dense matrix $W_i = \begin{bmatrix} \tilde{R}_i R_i^T \Lambda_{i_1} & \cdots & \tilde{R}_i R_i^T \Lambda_{i_{\nu_j}} \end{bmatrix}$;
    \item define a global deflation matrix $P = \begin{bmatrix} R_1^T W_1 & \cdots & R_N^T W_N\end{bmatrix}$ and a new two-level preconditioner using the Galerkin product of $A$ and $P$:
        \begin{equation}\label{eq:two}
            M'^{-1}_{\text{additive}} = P \left(P^T A P\right)^{-1} P^T + \sum_{i = 1}^N \tilde{R}_i^T (R_i A R_i^T)^{-1} R_i.
        \end{equation}
\end{enumerate}
Note that $\{\tilde{R}_i R_i^T\}_{i=1}^N$ define a partition of unity, i.e.,
\begin{equation*}
    \sum_{i = 1}^N R_i^T \tilde{R}_i R_i^T R_i = I.
\end{equation*}
When used without PETSc and thus SLEPc, step \#1 is solved in HPDDM with
ARPACK~\cite{lehoucq1998arpack}. Step \#3 is common to most domain
decomposition or multigrid preconditioners. However, in HPDDM, the special
structure of the deflation matrix~$P$ which is sparse with dense blocks is
exploited. Interested readers are referred to the paper describing the technical
details of the implementation~\cite{jolivet2013scalable}. One important point to
notice is that the size of the coarse operator $P^TAP$ is much lower
than $n$, the size of $A$. It is thus common to have the coarse operator distributed on
fewer processes than for $A$, see for example the concept of telescoping later
introduced in PETSc~\cite{may2016extreme}. In HPDDM, the coarse operator is
computed and redistributed simultaneously. This could be technical to implement
in PETSc as the result of the routine \pk{MatPtAP} is distributed on the same communicator as the one of the input matrix $A$.

Of course, users do not have to implement the three previous steps themselves.
The auxiliary operators $\{\mathring{A}_i\}_{i=1}^N$ have simply to be provided
as well as the correspondence from local to global numbering,
see~\cref{fig:poisson}, using the routine \pk{PCHPDDMSetAuxiliaryMat}.  If there
is a \pk{DM} attached to the \pk{PC} of type \pk{DMPLEX}~\cite{knepley2009mesh}
or \pk{DMP4EST}~\cite{BursteddeWilcoxGhattas11}, then these auxiliary
overlapping operators and the correspondence are automatically assembled by
PETSc for users. Other \pk{DMType} may be supported by implementing the proper
\pk{DMCreateNeumannOverlap} callback. With the local matrix pencils available,
new options will be automatically registered such as
-pc\_hpddm\_levels\_1\_eps\_nev or -pc\_hpddm\_levels\_1\_eps\_threshold, which
may be used to set how many eigenpairs SLEPc should compute in step \#2.  The
default -eps\_nev value is 20, i.e., the dimension of the coarse space is
$20\times N$ under the assumption that all concurrent SLEPc solves converge.
Another important option is -pc\_hpddm\_coarse\_p, which sets the size of the
communicator on which the coarse operator should be distributed. The default
value is 1, i.e., the matrix $P^TAP$ is centralized on a single process.
The action of both inverses from~\cref{eq:two} can be parameterized through the
-pc\_hpddm\_coarse\_ksp\_type and -pc\_hpddm\_levels\_1\_ksp\_type options
respectively.
    \subsubsection{Coarse corrections and eigenvalue problems}
Some other parameters may be adjusted to define how the one-level preconditioner
and the coarse grid operator interact with each other. In fact, the additive
correction from~\cref{eq:two} is the simplest to write down, but most often not
very efficient numerically~\cite{tang2009comparison}, and thus not used as
default in \pk{PCHPDDM}. Users can choose between
$M'^{-1}_{\text{additive}}$ and:
\begin{align*}
    M'^{-1}_{\text{deflated}} &= Q + M^{-1}\left(I - AQ\right)\quad
    \text{(default)}\\
    M'^{-1}_{\text{balanced}} &= Q + \left(I - AQ\right)M^{-1}\left(I - AQ\right),
\end{align*}
with $Q = P \left(P^T A P\right)^{-1} P^T$. These preconditioners may be selected using the additional
options -pc\_hpddm\_coarse\_correction (deflated$|$additive$|$balanced) or the
routine \pk{PCHPDDMSetCoarseCorrectionType}(pc, type) and its Fortran equivalent.

The local generalized eigenvalue problems~\cref{eq:gevp} solved by SLEPc can also be adjusted.
For example, in addition to the local unassembled Neumann matrices
$\{\mathring{A}_i\}_{i=1}^N$, if one has access to local matrices with optimized
boundary conditions, e.g., Robin transmission conditions, $\{B_i\}_{i=1}^N$, it
is possible to pass these using the routine \pk{PCHPDDMSetRHSMat}(pc, $B_i$).
The following generalized eigenvalue problems, called GenEO-2 in the
literature~\cite{haferssas2017soras}, are then solved in each subdomain:
\begin{equation*}
    \mathring{A}_i \Lambda_i = \lambda_i B_i \Lambda_i.
\end{equation*}
    \subsubsection{Multilevel extension}\label{sec:ml}
In the previous sections, it was shown how to supply the required information to
\pk{PCHPDDM} in order to build robust two-level overlapping domain decomposition preconditioners.
A multilevel extension of the GenEO framework is also
possible~\cite{aldaas2019multi}. It can be used without any additional
information provided by users. Since the dimension of the
coarse operator of the second level grows with the number of subdomains of the
fine level, switching to a multilevel scheme may alleviate the increasing cost
of solving coarse systems. \pk{PCHPDDM} follows the same numbering of
\pk{PCBDDC}\footnote{which itself follows the opposite numbering of \pk{PCMG}}: the
finest level is always numbered 1, the level index increases, up until the
coarsest level whose options are always prefixed by coarse\_. In order to
register an additional level $l' = l+1$, the following conditions must be met at
level $l$:
\begin{itemize}
    \item there must be more than one subdomain;
    \item at least one local eigenvector must be computed.
\end{itemize}
Here are some examples with $N = 16$ subdomains on the finest level:
\begin{itemize}
    \item -pc\_hpddm\_levels\_1\_eps\_threshold 0.4 -pc\_hpddm\_coarse\_p 8 would
        define a two-level method with the coarse operator being distributed
        among 8 processes, assuming that there is globally enough $\lambda_i$ in~\cref{eq:gevp} lower than 0.4;
    \item -pc\_hpddm\_levels\_1\_eps\_nev 10 -pc\_hpddm\_levels\_2\_p 8
        -pc\_hpddm\_levels\_2\_eps\_nev 10 -pc\_hpddm\_coarse\_p 2 would define
        a three-level method with the second level (resp.~ coarse operator)
        being distributed among 8 (resp.~2) processes;
    \item -pc\_hpddm\_levels\_1\_eps\_nev 10 -pc\_hpddm\_levels\_2\_p 8 is not
        valid because there are only two levels, so the options levels\_2\_ are
        not registered and coarse\_ has to be used instead;
    \item -pc\_hpddm\_levels\_1\_eps\_nev 10 -pc\_hpddm\_levels\_2\_eps\_nev 10 -pc\_hpddm\_coarse\_p 2 
        is not valid because there is no option levels\_2\_p, so the default
        value of 1 is used for the second level, and it is thus not possible to
        define a third level with a single second-level subdomain.
\end{itemize}
    \subsubsection{MatIS}
Because of the intrinsic nature of balancing domain
decomposition methods, \pk{PCNN} and \pk{PCBDDC} must be supplied local unassembled
matrices. These matrices are equivalent to the $\{\mathring{A}_i\}_{i=1}^N$
defined \cref{sec:interface-pc}, assuming the domain decomposition is without
overlap, see for example $\Omega_1$ and $\Omega_2$ from~\cref{fig:poisson} in a
finite element context. Using a local to global mapping, it is possible to
reconstruct the local assembled matrices, equivalent to the $\{R_i A
R_i^T\}_{i=1}^N$ from~\cref{sec:interface-pc}. All tools needed by the GenEO
framework are thus readily available. Note however that one-level overlapping
Schwarz methods are known for converging slowly when there is in fact no
overlap. It is common in the field of BDD to see the coarse problem as an
unassembled operator as well. In fact, this is mandatory to define efficient
multilevel variants, as implemented in \pk{PCBDDC}. In~\cref{sec:bddc}, it will be
shown how a coarse problem obtained in \pk{PCBDDC} may be solved using
\pk{PCHPDDM}.
  \subsection{Applications and numerical results}
    \subsubsection{System of elasticity}\label{sec:linear-elasticity}
Considering only small displacements, this linear problem with highly heterogeneous elastic moduli is solved in 3D.
Its strong formulation is given by:
\begin{align}
	\label{eq:elasticity}
	\begin{split}
		\text{div }\sigma (u)  + f	&= 0	\quad \text{in } \Omega,\\
		u														&= 0	\quad \text{on } \Gamma_D, \\
		\sigma (u) \cdot n					&= 0	\quad \text{on } \Gamma_N.
	\end{split}
\end{align}
The physical domain $\Omega$ is a beam of dimensions $[0,6] \times [0,1] \times [0,1]$. It is decomposed in $13{,}824$ subdomains using the automatic graph partitioner ParMETIS~\cite{karypis1998fast}.
The Cauchy stress tensor $\sigma$ is given by Hooke’s law: it can be expressed in terms of Young’s modulus $E$ and Poisson’s ratio $\nu$.
\[
	\sigma_{ij}(u) = 
	\begin{cases}
		2\mu \varepsilon_{ij}(u)													\quad	& i \ne j, \\
		2\mu \varepsilon_{ii}(u) + \lambda \text{div}(u)	\quad	& i = j,
	\end{cases}
\]
where
\[
	\varepsilon_{ij}(u) = \frac{1}{2} \left(\frac{\partial u_i}{\partial x_i} + \frac{\partial u_j}{\partial x_j} \right), \mu = \frac{E}{2(1+\nu)}, \text{ and }\lambda = \frac{E\nu}{1-2\nu}.
\]
$\Gamma_D$ is the subset of the boundary of $\Omega$ corresponding to $x = 0$.
$\Gamma_N$ is defined as the complementary of $\Gamma_D$ with respect to the boundary of $\Omega$.
\Cref{eq:elasticity} is discretized using 
$(\mathbb{P}_1, \mathbb{P}_1, \mathbb{P}_1)$ finite elements.
The number of unknowns is $\pgfmathprintnumber{593} \times 10^6$, with approximately $\pgfmathprintnumber{45}$ nonzero coefficients per row.
There are heterogeneities due to jumps in $E$ and $\nu$. The following discontinuous piecewise constant values are considered: $(E_1, \nu_1) = (2\times 10^{11}, 0.25)$, $(E_2, \nu_2) = (10^{7}, 0.45)$, see~\cref{fig:elas}.
\begin{figure}[htbp]
  \centering
  \begin{tikzpicture}[remember picture]
      \node[inner sep=-0.1cm] (fig) {\includegraphics[width=0.55\textwidth]{Data/elasticity.png}};
  \end{tikzpicture}
      \begin{tikzpicture}[overlay,remember picture]
      \begin{axis}[
              at=(fig.south west),
          xshift=-2.25cm,yshift=-2cm,
    enlargelimits=false,
    colorbar,
    colormap name=paraview,
    point meta min=-4.9e-3,
    point meta max=4.9e-3,
    hide axis,
    colorbar style={
        ytick style={draw=none},
          title={$u(x, \cdot, z)$},
        scaled y ticks = false,
        height = 2.0cm,
        font = \tiny,
        width = 0.3cm,
        ytick={-4.9e-3,0.0,4.9e-3}
    }
    ]
      \end{axis}
  \end{tikzpicture}
    \caption{Deformed geometrical configuration of a 3D clamped beam subject to gravity. The striped plane in the background shows a cut of the resting configuration with the jumps in the coefficient $E_1={\color{paraviewblue}10^7}$ and $E_2={\color{paraviewred}2\times 10^{11}}$, aligned with those of $\nu$.\label{fig:elas}}
\end{figure}
The following solvers are compared:
\begin{enumerate}
    \item \pk{PCGAMG};\label{enum:gamg}
    \item \pk{PCHPDDM} with an exact coarse solver;\label{enum:exact}
    \item \pk{PCHPDDM} with an inexact coarse solver using \pk{PCHYPRE};\label{enum:inexacthypre}
    \item \pk{PCHPDDM} with inexact coarse solvers using GenEO multilevel extension, cf.~\cref{sec:ml}.\label{enum:inexact}
\end{enumerate}
The options -ksp\_gmres\_modifiedgramschmidt
-ksp\_gmres\_restart 50
-ksp\_type fgmres are used in conjunction with all \pk{PC}. The options specific to each solver described above are given next. \\[-4pt]
\begin{minipage}[t]{0.34\textwidth}
\begin{Verbatim}[fontsize=\footnotesize,frame=single,framerule=0.1mm,commandchars=\\\{\}]
\fvtextcolor{mygreen}{# cf. \cref{enum:gamg}}
-pc_type gamg
-prefix_push pc_gamg_
 -threshold 0.03
 -square_graph 4
 -sym_graph true
 -asm_use_agg true
 -repartition true
-prefix_pop
-prefix_push mg_levels_
 -pc_asm_overlap 0
 -sub_pc_type cholesky
-prefix_pop
-prefix_push mg_coarse_
 -pc_type redundant
 -redundant_pc_type cholesky
-prefix_pop
\end{Verbatim}
\end{minipage}
\hfill
\begin{minipage}[t]{0.49\textwidth}
\begin{Verbatim}[fontsize=\footnotesize,frame=single,framerule=0.1mm,commandchars=\\\{\}]
\fvtextcolor{mygreen}{# cf. \cref{enum:exact}}
-pc_type hpddm
-prefix_push pc_hpddm_
 -prefix_push levels_1_
  -pc_type     asm
  -eps_nev     40
  -sub_pc_type cholesky
  -sub_pc_factor_mat_solver_type mumps
  -st_pc_factor_mat_solver_type mumps
 -prefix_pop
 -prefix_push coarse_
  -p 24
  -pc_factor_mat_solver_type mkl_cpardiso
 -prefix_pop
 -define_subdomains
 -has_neumann
-prefix_pop
\end{Verbatim}
\end{minipage} \\
By default, the coarse problem in \pk{PCHPDDM} is solved using an exact $LU$ or Cholesky factorization, in this case using 24 processes and MKL CPARDISO. Thanks to the flexibility of PETSc, switching to an inexact solver using \pk{PCHYPRE} (resp.~GenEO multilevel extension), cf.~\cref{enum:inexacthypre} (resp.~\cref{enum:inexact}), is straightforward. \\[-4pt]
\begin{minipage}[t]{0.4\textwidth}
\begin{Verbatim}[fontsize=\footnotesize,frame=single,framerule=0.1mm,commandchars=\\\{\}, codes={\catcode`$=3\catcode`^=7}]
\fvtextcolor{mygreen}{# cf. \cref{enum:inexacthypre}}
-prefix_push pc_hpddm_coarse_
 -ksp_type    gmres
 -ksp_rtol    1.0e-2
 -ksp_max_it  100
 -ksp_pc_side right
 -pc_type     hypre
 -prefix_push pc_hypre_
  -boomeramg_interp_type  ext+i
  -boomeramg_coarsen_type hmis
 -prefix_pop
-prefix_pop
\end{Verbatim}
\end{minipage}
\hfill
\begin{minipage}[t]{0.525\textwidth}
\begin{Verbatim}[fontsize=\footnotesize,frame=single,framerule=0.1mm,commandchars=\\\{\}, codes={\catcode`$=3\catcode`^=7}]
\fvtextcolor{mygreen}{# cf. \cref{enum:inexact}}
-prefix_push pc_hpddm_levels_2_
 -p           $M$
 -ksp_type    gmres            
 -ksp_rtol    1.0e-2           
 -ksp_pc_side right            
 -pc_type     asm
 -eps_nev     80
 -sub_pc_type cholesky
 -sub_pc_factor_mat_solver_type mkl_pardiso
 -st_pc_factor_mat_solver_type mkl_pardiso
-prefix_pop
\end{Verbatim}
\vspace*{-0.5cm}
\end{minipage} \\
$M$ is the number of subdomains used to define the second-level domain decomposition, which is an aggregation of first-level subdomain.
When it comes to convergence, first note that BoomerAMG from \emph{hypre} has a really hard time reaching the stopping criterion of $10^{-2}$ when solving GenEO coarse problems. This is emphasized in~\cref{fig:hypre-conv}, where the lack of convergence at the coarse level with \pk{PCHYPRE} greatly deteriorates the outer solver convergence, this configuration is thus discarded next and motivates the need for GenEO multilevel extension.
\pgfplotstableread{
it res inner
  0  3.059298477541e+01 nan
  1  3.059296797779e+01 100
  2  3.059296195928e+01 15
  3  3.059296184099e+01 15
  4  3.059295742636e+01 13
  5  3.059295742508e+01 15
  6  3.059294966626e+01 14
  7  3.059293719161e+01 95
  8  3.059292563008e+01 100
  9  3.059286757287e+01 100
 10  3.059276348161e+01 100
 11  3.059259253443e+01 100
 12  3.059200963623e+01 100
 13  3.059116858285e+01 100
 14  3.059008706025e+01 100
 15  3.058899295395e+01 100
 16  3.058765538548e+01 100
 17  3.058718339920e+01 100
 18  3.058690573439e+01 100
 19  3.058631705473e+01 100
 20  3.058529856313e+01 100
 21  3.058458303471e+01 100
 22  3.058441527393e+01 100
 23  3.058404813350e+01 100
 24  3.058350395496e+01 100
 25  3.058297671639e+01 100
 26  3.058267730942e+01 100
 27  3.058267032565e+01 100
 28  3.058260186627e+01 19
 29  3.058233416052e+01 100
 30  3.058177883143e+01 100
 31  3.058107492291e+01 100
 32  3.058053304435e+01 100
 33  3.058021917750e+01 100
 34  3.057991067536e+01 100
 35  3.057984123224e+01 100
 36  3.057981166461e+01 100
 37  3.057969981702e+01 100
 38  3.057951130340e+01 100
 39  3.057909738295e+01 100
 40  3.057879065735e+01 100
 41  3.057844662403e+01 100
 42  3.057813110973e+01 100
 43  3.057805844313e+01 100
 44  3.057748148817e+01 100
 45  3.057716298347e+01 100
 46  3.057695365582e+01 100
 47  3.057687162665e+01 100
 48  3.057685976587e+01 100
 49  3.057676977855e+01 100
 50  3.057666112196e+01 100
 51  3.057666112157e+01 100
 52  3.057666066500e+01 15
 53  3.057666020080e+01 15
 54  3.057666000522e+01 12
 55  3.057666000081e+01 15
 56  3.057665991326e+01 15
 57  3.057665913485e+01 95
 58  3.057665848806e+01 100
 59  3.057665837181e+01 100
 60  3.057665830553e+01 100
 61  3.057665733071e+01 100
 62  3.057665722344e+01 100
 63  3.057665608783e+01 100
 64  3.057664613215e+01 100
 65  3.057661541445e+01 100
 66  3.057659606149e+01 100
 67  3.057657581806e+01 100
 68  3.057648665398e+01 100
 69  3.057648651148e+01 100
 70  3.057647274070e+01 100
 71  3.057646787321e+01 100
 72  3.057646700351e+01 100
 73  3.057646583987e+01 100
 74  3.057640958561e+01 100
 75  3.057640655402e+01 100
 76  3.057640464884e+01 100
 77  3.057634384959e+01 100
 78  3.057630002417e+01 100
 79  3.057628670010e+01 100
 80  3.057628640121e+01 100
 81  3.057628480757e+01 100
 82  3.057625930943e+01 100
 83  3.057619950478e+01 100
 84  3.057615004023e+01 100
 85  3.057605590648e+01 100
 86  3.057603257752e+01 100
 87  3.057575042666e+01 100
 88  3.057537924704e+01 100
 89  3.057516167721e+01 100
 90  3.057445320162e+01 100
 91  3.057434150644e+01 100
 92  3.057433943630e+01 100
 93  3.057431618884e+01 100
 94  3.057419597979e+01 100
 95  3.057412470447e+01 100
 96  3.057409280981e+01 100
 97  3.057407988562e+01 100
 98  3.057397192115e+01 100
 99  3.057365742468e+01 100
100  3.057298339599e+01 100
}\loadedtableHYPRE
\begin{figure}[htbp]
    \centering
        \begin{tikzpicture}[font=\scriptsize]
\begin{axis}[%title=\framebox{{\upcase diffusion 2D, $\mathbb{P}_4$ FE, $\varepsilon = \num[exponent-product=\cdot]{1e-06}$}},
        axis y line*=left,ymajorgrids, unbounded coords=jump, xmin=0,xmax=100,extra y ticks={100}, width=0.5\textwidth,extra y tick labels=\empty, extra y tick style={grid=major}, ylabel = {\# of inner iterations~\ref{pgfplots:inner-hypre}}, xlabel = {Outer iteration index}]
\addplot[last, very thick] table[x=it, y=inner] {\loadedtableHYPRE};\label{pgfplots:inner-hypre}
% \node[rectangle, draw, font=\scriptsize, text width = 4.6cm] at (axis cs:12.9, 18) {\texttt{-hpddm\_master\_p 32 \\ -hpddm\_master\_aggregate\_size 16}};
\end{axis}
\begin{axis}[axis x line=none,ymax=31,ymin=30, axis y line*=right, xmin=0,xmax=100,width=0.5\textwidth, ylabel near ticks=right, ylabel = {Unpreconditioned residual~\ref{pgfplots:res-hypre}}]
\addplot[myred, densely dashed, very thick] table[x=it, y=res] {\loadedtableHYPRE};\label{pgfplots:res-hypre}
\end{axis}
\end{tikzpicture}
    \caption{Convergence history of \pk{PCHPDDM} using \pk{PCHYPRE} on 24 processes
    as a coarse solver. The maximum number of iterations of both the inner coarse solver and the outer solver is set to 100.\label{fig:hypre-conv}}
\end{figure}

Second, when properly parametrized, \cref{enum:exact} and \cref{enum:inexact} yield the same number of outer iterates. In~\cref{fig:gamg-geneo}, the convergence history of \pk{PCGAMG} and \pk{PCHPDDM} are reported. In~\cref{fig:inexact}, the number of inner coarse iterations are reported for various values of $M$. In the case of a two-level method,~\cref{enum:exact}, this number is equal to one and is thus not reported in the figure.
\pgfplotstableread{
    it gamg hpddm
  0 3.859411178402e+00 3.859411178402e+00
  1 3.859411159689e+00 3.859411176864e+00
  2 3.859410249427e+00 3.859411160502e+00
  3 3.859409408276e+00 3.859411009079e+00
  4 3.859405618616e+00 3.859410641267e+00
  5 3.859384473758e+00 3.859409504108e+00
  6 3.859336132196e+00 3.859406284759e+00
  7 3.859180574117e+00 3.859396774066e+00
  8 3.858527456325e+00 3.859371139173e+00
  9 3.855946156750e+00 3.859291203708e+00
 10 3.847882304613e+00 3.859050327998e+00
 11 3.823783202371e+00 3.858349404873e+00
 12 3.761793608189e+00 3.856196410336e+00
 13 3.604276380304e+00 3.850223368267e+00
 14 3.219169630144e+00 3.833574948206e+00
 15 2.524138668209e+00 3.789137446493e+00
 16 1.612332237539e+00 3.677085304126e+00
 17 9.250281017522e-01 3.416685094366e+00
 18 5.244208314872e-01 2.914257430799e+00
 19 2.900205723223e-01 2.201797784976e+00
 20 1.639686082701e-01 1.504764597680e+00
 21 8.999329395027e-02 9.651409800649e-01
 22 4.806932060093e-02 5.991352736022e-01
 23 2.623490189992e-02 3.709375100943e-01
 24 1.420399852677e-02 2.262545029178e-01
 25 7.785203261213e-03 1.386897546734e-01
 26 4.279336819903e-03 8.468744514883e-02
 27 2.392528915920e-03 5.199440557582e-02
 28 1.398509701362e-03 3.167797843106e-02
 29 7.951946917375e-04 1.958093540992e-02
 30 4.302913355532e-04 1.195194560026e-02
 31 2.336327784594e-04 7.371078209494e-03
 32 1.242918661343e-04 4.565136636812e-03
 33 6.466088102124e-05 2.757997749188e-03
 34 3.347704352999e-05 1.722488935062e-03
 35               nan  1.040719253466e-03 
 36               nan  6.471646548950e-04 
 37               nan  3.944692230299e-04 
 38               nan  2.413209464780e-04 
 39               nan  1.481390561258e-04 
 40               nan  9.041135719421e-05 
 41               nan  5.606739576810e-05 
 42               nan  3.445256788047e-05 
}\comparisontable
\pgfplotstableread{
it p-16  p-256  p-1024
 1 26    16     13
 2 9     6      6
 3 10    7      6
 4 8     6      6
 5 9     6      5
 6 5     5      5
 7 8     6      5
 8 9     6      5
 9 7     5      5
10 7     6      5
11 7     6      5
12 8     6      5
13 5     5      5
14 6     6      4
15 6     5      5
16 8     6      5
17 6     5      5
18 6     5      5
19 7     5      4
20 8     6      5
21 8     6      5
22 7     5      4
23 7     6      5
24 6     6      5
25 6     6      5
26 6     5      5
27 6     5      4
28 5     5      5
29 6     5      5
30 6     5      5
31 7     5      5
32 8     6      5
33 7     6      4
34 5     5      5
35 5     5      5
36 6     5      4
37 6     5      4
38 6     5      5
39 7     5      4
40 5     5      5
41 5     5      5
42 6     5      nan
43 5     nan    nan
}\comparisonP
\begin{figure}[htbp]
  \centering
    \subfloat[\parbox{5cm}{Numerical comparison of \pk{PCGAMG} and \pk{PCHPDDM}.\label{fig:gamg-geneo}}]{
\begin{tikzpicture}[font=\scriptsize]
\begin{semilogyaxis}[xmin=0,xmax=50,width=0.46\textwidth,
    legend style={at={(1.05,0.95)},cells={anchor=west}},ymajorgrids,
    xlabel={Iteration number},
    ylabel={Unpreconditioned residual},
    enlarge x limits=0.05,
% yticklabel={
% \pgfmathparse{int(\tick)}
% \ifnum\pgfmathresult=0
% $1$
% \else
% \pgfmathparse{int(\tick/ln(10))}{$10^{\pgfmathresult}$}
% \fi
% }
]
\addplot[very thick,last] table [x=it, y index=1] {\comparisontable};
\addplot[very thick,myred, densely dashed] table [x=it, y index=2]       {\comparisontable};
    \legend{{\pk{PCGAMG}},{\pk{PCHPDDM}}}
% lightgray
\end{semilogyaxis}
\end{tikzpicture}
}
    \subfloat[\parbox{5cm}{\# of inner coarse iterations for various values of $M$, the number of second-level subdomains.}\label{fig:inexact}]{
        \begin{tikzpicture}[font=\scriptsize]
\begin{axis}[%title=\framebox{{\upcase diffusion 2D, $\mathbb{P}_4$ FE, $\varepsilon = \num[exponent-product=\cdot]{1e-06}$}},
        ymajorgrids, unbounded coords=jump, xmin=-1,extra y ticks={100}, width=0.46\textwidth,extra y tick labels=\empty, extra y tick style={grid=major},legend columns=3,legend cell align=left, ylabel = {\# of inner iterations}, xlabel = {Outer iteration index}]
\addlegendimage{empty legend}
\addlegendimage{empty legend}
\addlegendimage{empty legend}
\addplot[last, very thick] table[x=it, y=p-16] {\comparisonP};\label{pgfplots:p-16}
\addplot[myred, densely dashed, very thick] table[x=it, y=p-256] {\comparisonP};\label{pgfplots:p-256}
\addplot[myyellow, densely dotted, very thick] table[x=it, y=p-1024] {\comparisonP};\label{pgfplots:p-1024}

   \addlegendentry{\hspace{-.6cm}-pc\_hpddm\_levels\_2\_p $M$ \hspace*{-2.5cm}}
   \addlegendentry{}
   \addlegendentry{}
   \addlegendentry{16}
   \addlegendentry{256}
    \addlegendentry{\pgfmathprintnumber{1024}}
\end{axis}
        \end{tikzpicture}
    }
    \caption{Convergence histories of \pk{PCGAMG} and \pk{PCHPDDM} for solving the 3D system of linear elasticity. For \pk{PCHPDDM}, a comparison of the number of inner coarse iterations is displayed when using an inexact method, cf.~\cref{enum:inexact}.}
\end{figure}
Note that in this test case, \pk{PCGAMG} is faster than all \pk{PCHPDDM}
alternatives. Its setup (resp.~solution) phase takes 70 (resp.~19) seconds while
for the best parameters of \pk{PCHPDDM}, which is
configuration~\ref{pgfplots:p-256} with 256 second-level subdomains, its
respectively 64 and 36 seconds. Out of these 64 seconds, 31 are spent in SLEPc
\pk{EPSSolve} for solving GenEO at each level but the coarsest. The code used to
produce these results was borrowed from the original paper about the multilevel
extension of GenEO~\cite{aldaas2019multi} and is available
at~\url{https://github.com/prj-/aldaas2019multi}.
    \subsubsection{Liouville--Bratu--Gelfand equation} The strong formulation of this nonlinear problem is given by:
\begin{equation}
    -\nabla \cdot (\kappa \nabla u) - 6.2 \exp^u = 0,\label{eq:bratu}
\end{equation}
where $\kappa$ is a heterogeneous coefficient distribution,
see~\cref{fig:kappa}. This equation may model the temperature distribution in
combustion models. Here, it is merely used to test \pk{PCHPDDM} in the context
of solving successive linearized equations, but there are many other techniques
for solving such a nonlinear equation, such as first decomposing the domain and
then linearizing the equation~\cite{dryja1997nonlinear}.
The physical domain $\Omega$ is the unit cube. It is decomposed in $10{,}272$ subdomains.
\Cref{eq:bratu} is discretized using $\mathbb{P}_2$ finite elements.
The number of unknowns is $\pgfmathprintnumber{217} \times 10^6$, with
approximately $\pgfmathprintnumber{29}$ nonzero coefficients per row.
It is solved using a \pk{SNES}, with \pk{PCHPDDM} being the preconditioner for
solving each linearized system, see~\cref{fig:u}.
\begin{figure}[htbp]
  \centering
  \subfloat[Coefficient distribution.\label{fig:kappa}]{
  \begin{tikzpicture}[remember picture]
      \node[inner sep=-0.1cm] (fig) {\includegraphics[width=0.45\textwidth]{Data/kappa.png}};
  \end{tikzpicture}
  }
    \subfloat[Isosurfaces of the solution.\label{fig:u}]{
  \begin{tikzpicture}[remember picture]
      \node[inner sep=-0.1cm] (fq) {\includegraphics[width=0.45\textwidth]{Data/u.png}};
  \end{tikzpicture}
      \begin{tikzpicture}[overlay,remember picture]
      \begin{axis}[
              at=(fig.south west),
          xshift=-1.5cm,yshift=-2cm,
    enlargelimits=false,
    colorbar,
    colormap name=paraview,
    point meta min=6e-2,
    point meta max=8.5,
    hide axis,
    colorbar style={
        ytick style={draw=none},
        title={$\kappa$},
        scaled y ticks = false,
        height = 2.0cm,
        font = \tiny,
        width = 0.3cm,
        ytick={6e-2,4.0,8.5}
    }
    ]
      \end{axis}
      \begin{axis}[
              at=(fq.south east),
          xshift=0.0cm,yshift=-2cm,
    enlargelimits=false,
    colorbar,
    colormap name=paraview,
    point meta min=-1.9,
    point meta max=1.9,
    hide axis,
    colorbar style={
        ytick style={draw=none},
        title={$u$},
        scaled y ticks = false,
        height = 2.0cm,
        font = \tiny,
        width = 0.3cm,
        ytick={-1.9,0,1.9}
    }
    ]
      \end{axis}
  \end{tikzpicture}
  }
    \caption{Solving the Liouville--Bratu--Gelfand equation using \pk{SNESNEWTONLS} and \pk{PCHPDDM}.}
\end{figure}
The GenEO framework has seldom been used in the context of nonlinear problems, but
here, it is shown that it can solve the linearized equations
from~\cref{eq:bratu} in few iterates.  There are two strategies to define the
\pk{PC} here:
\begin{itemize}
\item supply whenever the Jacobian is being updated in the \pk{SNESSetJacobian}
function new unassembled Neumann matrices at the current linearization point with the previously
explained \pk{PCHPDDMSetAuxiliaryMat} function, this is done automatically with a
\pk{DMPLEX} or a \pk{DMP4EST};
\item by reusing the \pk{PCHPDDM} hierarchy assembled when solving the first
    linearized system, using the \pk{SNESSetLagPreconditioner} option.
\end{itemize}
Since the system is not too stiff here, both strategies lead to the same
convergence history, which is displayed in~\cref{fig:bratu-conv}. The complete set of
options is given in~\cref{fig:bratu-options}.
\pgfplotstableread{
it first                second               third                  fourth
 0 1.199398376742e+01   3.397386070782e-04   1.863266214507e-05     8.796157053238e-07
 1 3.105058690048e+00   3.396971540287e-04   1.863121679230e-05     8.795510768269e-07
 2 7.493003609951e-01   3.394715205399e-04   1.862509132651e-05     8.792972699303e-07
 3 2.066956200353e-01   3.372614259474e-04   1.860041089031e-05     8.783878285049e-07
 4 6.309969159499e-02   3.127230257133e-04   1.838790479520e-05     8.701902993653e-07
 5 1.933161037163e-02   2.148249088439e-04   1.707968608004e-05     8.181649294279e-07
 6 6.550219066971e-03   9.681223676375e-05   1.238892191333e-05     6.183954841666e-07
 7 2.324009861121e-03   3.933958932173e-05   6.179174147298e-06     3.194732043996e-07
 8 8.476764474277e-04   1.558086376310e-05   2.605805785728e-06     1.363806062938e-07
 9 3.138242119678e-04   6.269961516841e-06   1.052999054933e-06     5.522010168719e-08
10 1.183695049547e-04   2.522978646378e-06   4.300851965899e-07     2.257459993685e-08
11                nan   1.024961124607e-06   1.753694476531e-07     9.219016303441e-09
12                nan   4.249085515277e-07   7.338448965970e-08     3.852685693794e-09
13                nan   1.763394120011e-07   3.059024771687e-08     1.609255024084e-09
14                nan   7.432342714299e-08   1.297122798875e-08     6.817901759942e-10
15                nan   3.135242203325e-08   5.496143820621e-09     2.896018108835e-10
16                nan   1.326698994886e-08   2.331975304171e-09     1.226613149417e-10
17                nan   5.639489438404e-09   9.871150757910e-10     5.194784893432e-11
18                nan   2.342602656325e-09   4.069402853378e-10     2.137940700304e-11
19                nan                  nan   1.679412238393e-10     8.828401758238e-12
20                nan                  nan                  nan     3.628457807677e-12
}\comparisonBratu
% 4 SNES Function norm 2.502417058684e-09
\begin{SaveVerbatim}[commandchars=\\\{\}]{VerbEnv}
-snes_type            newtonls
-snes_linesearch_type basic
-ksp_pc_side          right
-pc_type              hpddm
-prefix_push pc_hpddm_
 -prefix_push levels_1_
  -pc_type     asm
  -eps_nev     10
  -sub_pc_type cholesky
  -sub_pc_factor_mat_solver_type mumps
 -prefix_pop
 -coarse_p 128
 -define_subdomains
 -has_neumann
-prefix_pop
\end{SaveVerbatim}
\begin{figure}[htbp]
  \centering
    \subfloat[Convergence histories of the four linearized systems.
    At the top of each curve, for the first iterate, the value of the function
    norm is displayed.\label{fig:bratu-conv}]{
\begin{tikzpicture}[font=\scriptsize]
\begin{semilogyaxis}[clip=false,xmin=0,xmax=68,width=0.49\textwidth,
    legend style={at={(1.05,0.95)},cells={anchor=west}},ymajorgrids,
    xlabel={Iteration number},
    extra x ticks={0, 10, 28, 47, 67},extra x tick labels=\empty,
    extra x tick style={grid=major},
    ylabel={Relative unpreconditioned residual},
    enlarge x limits=0.05,
yticklabel={
\pgfmathparse{int(\tick)}
\ifnum\pgfmathresult=0
$1$
\else
\pgfmathparse{int(\tick/ln(10))}{$10^{\pgfmathresult}$}
\fi
}
]
\addplot[very thick,last] table [x=it, y expr=\thisrowno{1}/1.199398376742e+01] {\comparisonBratu};
\addplot[very thick,last] table [x expr=\thisrowno{0}+10, y expr=\thisrowno{2}/3.397386070782e-04] {\comparisonBratu};
\addplot[very thick,last] table [x expr=\thisrowno{0}+28, y expr=\thisrowno{3}/1.863266214507e-05] {\comparisonBratu};
\addplot[very thick,last] table [x expr=\thisrowno{0}+47, y expr=\thisrowno{4}/8.796157053238e-07] {\comparisonBratu};
\node[font=\tiny,fill=white,inner sep=1pt] at (axis cs:0, 2) {12.0};
\node[font=\tiny,fill=white,inner sep=1pt] at (axis cs:12, 2) {3.4E-4};
\node[font=\tiny,fill=white,inner sep=1pt] at (axis cs:28, 2) {1.9E-5};
\node[font=\tiny,fill=white,inner sep=1pt] at (axis cs:47, 2) {8.8E-7};
\node[font=\tiny,fill=white,inner sep=1pt] at (axis cs:67, 2) {2.5E-9};
\end{semilogyaxis}
\end{tikzpicture}
}
    \subfloat[\parbox{5cm}{\pk{SNES}, \pk{KSP}, and \pk{PC} options for the
    solving~\cref{eq:bratu}}\label{fig:bratu-options}]{
        {\setlength{\fboxrule}{0.1mm}
        \fbox{\BUseVerbatim[fontsize=\footnotesize]{VerbEnv}}
        }
    }
    \caption{Numerical performance of \pk{PCHPDDM} in a nonlinear context for
    solving the Liouville--Bratu--Gelfand equation. Default PETSc tolerances are used:
    $10^{-5}$ decrease of relative residuals for linear solves, $10^{-8}$ decrease
    of function norms for nonlinear solves.}
\end{figure}
    \subsubsection{Using PCHPDDM as a coarse grid solver for PCBDDC\label{sec:bddc}}
A last toy problem is explained to show how one may switch between a multilevel
BDDC solver to a two-level BDDC solver using \pk{PCHPDDM} for solving the
coarse problem. The problem is generated using MFEM~\cite{anderson2019mfem}
definite Maxwell example available at \url{https://github.com/mfem/mfem/blob/master/examples/petsc/ex3p.cpp}.
The strong formulation of the continuous problem is given by:
\begin{align}
	\label{eq:maxwell}
	\begin{split}
        \nabla \times (\nabla \times E) + E	&= 0	\quad \text{in } \Omega,\\
        E \times n					&= (1+16\pi^2)\begin{bmatrix} \sin{4\pi y}&
        \sin{4\pi z}&\sin{4\pi x} \end{bmatrix}^T \text{on } \partial \Omega.
	\end{split}
\end{align}
It is discretized using order 1 N\'ed\'elec finite elements. There is a command
line option -{}-nonoverlapping to activate \pk{PCBDDC}. The following command line
is thus used. \\[-4pt]
\begin{Verbatim}[fontsize=\footnotesize,frame=single,framerule=0.1mm,commandchars=&\[\]]
&fvtextcolor[mygreen][$] &fvtextcolor[myyellow][mpirun] -n 512 ./ex3p -m ../../data/fichera.mesh -f 4 --petscopts rc_ex3p_bddc      &fvtextcolor[myred][\]
  --nonoverlapping
\end{Verbatim}
The option file shows how \pk{PCHPDDM} may be composed into \pk{PCBDDC}. \\[-4pt]
\begin{minipage}[t]{0.565\textwidth}
\begin{Verbatim}[fontsize=\footnotesize,frame=single,framerule=0.1mm,commandchars=&\[\]]
&fvtextcolor[mygreen][$] &fvtextcolor[myyellow][cat] rc_ex3p_bddc
-ksp_type      fgmres
-ksp_norm_type unpreconditioned
-ksp_rtol      1.0e-8

-prefix_push pc_bddc_
 -use_deluxe_scaling
 -levels             1
 -adaptive_threshold 2.0
 -coarsening_ratio   8

 -neumann_pc_factor_mat_solver_type   mumps
 -dirichlet_pc_factor_mat_solver_type mumps
&fvtextcolor[mygreen][# continue on the left column]
\end{Verbatim}
\end{minipage}
\begin{minipage}[t]{0.425\textwidth}
\begin{Verbatim}[fontsize=\footnotesize,frame=single,framerule=0.1mm,commandchars=&\[\]]
&fvtextcolor[mygreen][# continued from the right column]
 -prefix_push coarse_
  -ksp_converged_reason
  -ksp_type      gmres
  -ksp_max_it    100
  -ksp_rtol      1.0e-1
  -pc_type       hpddm
  -ksp_norm_type preconditioned
  -prefix_push pc_hpddm_
   -levels_1_pc_type     asm
   -levels_1_eps_nev     10
   -levels_1_sub_pc_type cholesky
   -coarse_pc_type       cholesky
  -prefix_pop
 -prefix_pop
-prefix_pop
\end{Verbatim}
\end{minipage}
Deluxe scaling is used to build an adaptive second level with \pk{PCBDDC}. On this
second level, new subdomains are built by aggregating the constraints
of 8 fine-level subdomains. The BDDC coarse problem composed of $\frac{512}{8} =
64$ subdomains is then solved using \pk{PCHPDDM}, which itself builds another adaptive
level using $10$ local eigenvectors per subdomains. The coarsest level is
aggregated on a single process and solved using an exact Cholesky factorization.
The magnitude of the electric field is plotted~\cref{fig:E}. For the
convergence history of the solver, it is~\cref{fig:BDDC}. The outer solver
reaches the prescribed convergence criterion in 17 iterations~\ref{pgfplots:res}. The number of
inner iterations for solving the BDDC coarse problem is reported as well~\ref{pgfplots:inner}.
\pgfplotstableread{
it res inner
  0 2.362991511979e+02 nan
  1 3.301059014393e+01 3
  2 9.008361295505e+00 4
  3 4.198372502615e+00 7
  4 4.193238438798e+00 8
  5 3.757039752765e+00 7
  6 1.610223090345e+00 23
  7 4.119318170796e-01 5
  8 1.304562389595e-01 7
  9 3.996802619765e-02 6
 10 1.230923272811e-02 6
 11 4.922752304029e-03 6
 12 1.532991950467e-03 6
 13 3.772099626068e-04 6
 14 1.206814967573e-04 6
 15 3.229192953782e-05 6
 16 7.563953270757e-06 6
 17 1.638139681380e-06 6
}\loadedtableBDDC
\begin{figure}[htbp]
  \centering
  \subfloat[Magnitude of the electric field.\label{fig:E}]{
  \begin{tikzpicture}[remember picture]
      \node[inner sep=-0.1cm] (fig) {\includegraphics[width=0.45\textwidth]{Data/fichera.png}};
  \end{tikzpicture}
      \begin{tikzpicture}[overlay,remember picture]
      \begin{axis}[
              at=(fig.south west),
          xshift=-1.25cm,yshift=-2cm,
    enlargelimits=false,
    colorbar,
    colormap name=paraview,
    point meta min=0,
    point meta max=1.4,
    hide axis,
    colorbar style={
        ytick style={draw=none},
        title={$||E||$},
        scaled y ticks = false,
        height = 2.0cm,
        font = \tiny,
        width = 0.3cm,
        ytick={0,0.7,1.4}
    }
    ]
      \end{axis}
  \end{tikzpicture}
  }
    \subfloat[\parbox{5cm}{Convergence history of \pk{PCBDDC} using \pk{PCHPDDM}
    as a coarse solver.\label{fig:BDDC}}]{
        \begin{tikzpicture}[font=\scriptsize]
\begin{axis}[%title=\framebox{{\upcase diffusion 2D, $\mathbb{P}_4$ FE, $\varepsilon = \num[exponent-product=\cdot]{1e-06}$}},
        axis y line*=left,ymajorgrids, unbounded coords=jump, xmin=-1,extra y ticks={100}, width=0.5\textwidth,extra y tick labels=\empty, extra y tick style={grid=major}, ylabel = {\# of inner iterations~\ref{pgfplots:inner}}, xlabel = {Outer iteration index}]
\addplot[last, very thick] table[x=it, y=inner] {\loadedtableBDDC};\label{pgfplots:inner}
% \node[rectangle, draw, font=\scriptsize, text width = 4.6cm] at (axis cs:12.9, 18) {\texttt{-hpddm\_master\_p 32 \\ -hpddm\_master\_aggregate\_size 16}};
\end{axis}
\begin{semilogyaxis}[axis x line=none, axis y line*=right, xmin=-1,width=0.5\textwidth, ylabel near ticks=right, ylabel = {Unpreconditioned residual~\ref{pgfplots:res}}]
\addplot[myred, densely dashed, very thick] table[x=it, y=res] {\loadedtableBDDC};\label{pgfplots:res}
\end{semilogyaxis}
\end{tikzpicture}
  }
    \caption{MFEM example ex3p for solving the Fichera corner problem.}
\end{figure}
\section{Conclusion and perspectives}\label{sec:conclusion}
In this paper, the interface between PETSc and HPDDM is presented. The new
registered Krylov and overlapping Schwarz methods are explained and numerical
examples are provided to show the applicability of these. Solving well-known
problems, various options have been described. Some more unconventional test
cases were commented as well, for example using \pk{MatKAIJ} or \pk{MatIS} matrix types.
Overall, the interface paves the way for having recycling and block Krylov
methods plus robust overlapping Schwarz methods using the GenEO framework in
PETSc.

Concerning PETSc, one may notice that it is now possible to do some code
refactorization for applications that had to deal with multiple right-hand sides.
As a striking example, LOBPCG~\cite{knyazev2001toward} or
CISS~\cite{sakurai2003projection} implementations in SLEPc are
currently\footnote{in version 3.13.2} not using blocking when applying 
spectral transformations, but this is being rewritten to use the new routine
\pk{KSPMatSolve}\footnote{\url{https://gitlab.com/slepc/slepc/-/merge_requests/63}} and should be available in version 3.14.0.

Concerning HPDDM, its integration inside PETSc offers a much-needed flexibility
compared to the standard implementation. Indeed, it is now possible to switch at
runtime many parameters that are determined at compile-time in the standard implementation, e.g.,
subdomain solvers, matrix types, and such.

\section*{Acknowledgments}
The authors would like to thank S. Balay, J. Brown, V. Hapla, M. Knepley, J.
Roman, and B. Smith for reviewing the successive merge requests in PETSc and
SLEPc repositories. This work was granted access to the HPC resources of
TGCC@CEA under the allocations A0070607519 made by GENCI.
% We would like to acknowledge the assistance of volunteers in putting
% together this example manuscript and supplement.

\bibliographystyle{siamplain}
\begin{thebibliography}{10}

\bibitem{adams2004ultra}
{\sc M.~Adams, H.~H. Bayraktar, T.~M. Keaveny, and P.~Papadopoulos}, {\em
  {Ultrascalable Implicit Finite Element Analyses in Solid Mechanics with over
  a Half a Billion Degrees of Freedom}}, in Proceedings of the 2004 ACM/IEEE
  Conference on Supercomputing, SC04, IEEE Computer Society, 2004,
  pp.~\mbox{34:1--34:15}.

\bibitem{aldaas2019multi}
{\sc H.~Al~Daas, L.~Grigori, P.~Jolivet, and P.-H. Tournier}, {\em {A
  Multilevel Schwarz Preconditioner Based on a Hierarchy of Robust Coarse
  Spaces}}, Journal of Scientific Computing,  (2019), p.~submitted for
  publication, \url{https://github.com/prj-/aldaas2019multi}.

\bibitem{amestoy2001fully}
{\sc P.~Amestoy, I.~Duff, J.-Y. L'Excellent, and J.~Koster}, {\em A fully
  asynchronous multifrontal solver using distributed dynamic scheduling}, SIAM
  Journal on Matrix Analysis and Applications, 23 (2001), pp.~15--41,
  \url{http://mumps.enseeiht.fr}.

\bibitem{anderson2019mfem}
{\sc R.~Anderson, J.~Andrej, A.~Barker, J.~Bramwell, J.-S. Camier, J.~Cerveny,
  V.~Dobrev, Y.~Dudouit, A.~Fisher, T.~Kolev, et~al.}, {\em {MFEM: a modular
  finite element methods library}}, arXiv:1911.09220,  (2019),
  \url{http://mfem.org}.

\bibitem{badia2014highly}
{\sc S.~Badia, A.~F. Mart{\'\i}n, and J.~Principe}, {\em A highly scalable
  parallel implementation of balancing domain decomposition by constraints},
  SIAM Journal on Scientific Computing, 36 (2014), pp.~C190--C218.

\bibitem{baker2005technique}
{\sc A.~H. Baker, E.~R. Jessup, and T.~Manteuffel}, {\em {A Technique for
  Accelerating the Convergence of Restarted GMRES}}, SIAM Journal on Matrix
  Analysis and Applications, 26 (2005), pp.~962--984.

\bibitem{balay1997cient}
{\sc S.~Balay, W.~Gropp, L.~Curfman~McInnes, and B.~Smith}, {\em Efficient
  management of parallelism in object-oriented numerical software libraries},
  Modern Software Tools in Scientific Computing,  (1997), pp.~163--202.

\bibitem{bastian2010generic}
{\sc P.~Bastian, F.~Heimann, and S.~Marnach}, {\em {Generic implementation of
  finite element methods in the distributed and unified numerics environment
  (DUNE)}}, Kybernetika, 46 (2010), pp.~294--315,
  \url{https://www.dune-project.org}.

\bibitem{bavier2012amesos2}
{\sc E.~Bavier, M.~Hoemmen, S.~Rajamanickam, and H.~Thornquist}, {\em {Amesos2
  and Belos: Direct and Iterative Solvers for Large Sparse Linear Systems}},
  Scientific Programming, 20 (2012), pp.~241--255.

\bibitem{Benzi2011a}
{\sc M.~Benzi, M.~A. Olshanskii, and Z.~Wang}, {\em {Modified augmented
  Lagrangian preconditioners for the incompressible Navier--Stokes equations}},
  International Journal for Numerical Methods in Fluids, 66 (2011),
  pp.~486--508.

\bibitem{brown2012composable}
{\sc J.~Brown, M.~G. Knepley, D.~A. May, L.~C. McInnes, and B.~F. Smith}, {\em
  Composable linear solvers for multiphysics}, in 2012 11th International
  Symposium on Parallel and Distributed Computing, IEEE, 2012, pp.~55--62.

\bibitem{BursteddeWilcoxGhattas11}
{\sc C.~Burstedde, L.~C. Wilcox, and O.~Ghattas}, {\em {\texttt{p4est}}:
  Scalable algorithms for parallel adaptive mesh refinement on forests of
  octrees}, SIAM Journal on Scientific Computing, 33 (2011), pp.~1103--1133,
  \url{http://www.p4est.org}.

\bibitem{butler2020high}
{\sc R.~Butler, T.~Dodwell, A.~Reinarz, A.~Sandhu, R.~Scheichl, and
  L.~Seelinger}, {\em High-performance dune modules for solving large-scale,
  strongly anisotropic elliptic problems with applications to aerospace
  composites}, Computer Physics Communications, 249 (2020).

\bibitem{cai1999restricted}
{\sc X.-C. Cai and M.~Sarkis}, {\em {A restricted additive Schwarz
  preconditioner for general sparse linear systems}}, SIAM Journal on
  Scientific Computing, 21 (1999), pp.~792--797.

\bibitem{calandra2012flexible}
{\sc H.~Calandra, S.~Gratton, J.~Langou, X.~Pinel, and X.~Vasseur}, {\em
  {Flexible Variants of Block Restarted GMRES Methods with Application to
  Geophysics}}, SIAM Journal on Scientific Computing, 34 (2012),
  pp.~A714--A736.

\bibitem{carvalho2011flexible}
{\sc L.~M. Carvalho, S.~Gratton, R.~Lago, and X.~Vasseur}, {\em {A Flexible
  Generalized Conjugate Residual Method With Inner Orthogonalization and
  Deflated Restarting}}, SIAM Journal on Matrix Analysis and Applications, 32
  (2011), pp.~1212--1235.

\bibitem{dohrmann2008domain}
{\sc C.~R. Dohrmann, A.~Klawonn, and O.~B. Widlund}, {\em {Domain decomposition
  for less regular subdomains: Overlapping Schwarz in two dimensions}}, SIAM
  Journal on Numerical Analysis, 46 (2008), pp.~2153--2168.

\bibitem{dohrmann2016bddc}
{\sc C.~R. Dohrmann and O.~B. Widlund}, {\em {A BDDC Algorithm with Deluxe
  Scaling for Three-Dimensional $H(\textnormal{curl})$ Problems}},
  Communications on Pure and Applied Mathematics, 69 (2016), pp.~745--770.

\bibitem{dolean2014ddm}
{\sc V.~Dolean, P.~Jolivet, and F.~Nataf}, {\em An Introduction to Domain
  Decomposition Methods\col\ Algorithms, Theory and Parallel Implementation},
  SIAM, 2015.

\bibitem{dryja1997nonlinear}
{\sc M.~Dryja and W.~Hackbusch}, {\em On the nonlinear domain decomposition
  method}, BIT Numerical Mathematics, 37 (1997), pp.~296--311.

\bibitem{erhel1996restarted}
{\sc J.~Erhel, K.~Burrage, and B.~Pohl}, {\em {Restarted GMRES preconditioned
  by deflation}}, Journal of Computational and Applied Mathematics, 69 (1996),
  pp.~303--318.

\bibitem{falgout2002hypre}
{\sc R.~Falgout and U.~Yang}, {\em \emph{hypre}: a library of high performance
  preconditioners}, Computational Science---ICCS 2002,  (2002), pp.~632--641,
  \url{https://www.llnl.gov/casc/hypre}.

\bibitem{gander2006optimized}
{\sc M.~J. Gander}, {\em Optimized {S}chwarz {M}ethods}, SIAM Journal on
  Numerical Analysis, 44 (2006), pp.~699--731.

\bibitem{gee2006ml}
{\sc M.~Gee, C.~Siefert, J.~Hu, R.~Tuminaro, and M.~Sala}, {\em {ML} 5.0
  smoothed aggregation user's guide}, Tech. Report SAND2006-2649, Sandia
  National Laboratories, 2006, \url{http://trilinos.sandia.gov/packages/ml}.

\bibitem{gutknecht2006block}
{\sc M.~H. Gutknecht}, {\em Block {K}rylov space methods for linear systems
  with multiple right-hand sides: an introduction}, in Modern Mathematical
  Models, Methods and Algorithms for Real World Systems, A.~Siddiqui, I.~Duff,
  and O.~Christensen, eds., 2006, pp.~420--447.

\bibitem{gutknecht2008updating}
{\sc M.~H. Gutknecht and T.~Schmelzer}, {\em {Updating the $QR$ decomposition
  of block tridiagonal and block Hessenberg matrices}}, {Applied Numerical
  Mathematics}, 58 (2008), pp.~871--883.

\bibitem{haferssas2017soras}
{\sc R.~Haferssas, P.~Jolivet, and F.~Nataf}, {\em {An additive Schwarz method
  type theory for Lions's algorithm and a Symmetrized Optimized Restricted
  Additive Schwarz Method}}, Journal on Scientific Computing, 39 (2017),
  pp.~A1345--A1365.

\bibitem{hecht2012new}
{\sc F.~Hecht}, {\em New development in {FreeFem++}}, Journal of Numerical
  Mathematics, 20 (2012), pp.~251--266, \url{http://freefem.org}.

\bibitem{heinecke2016libxsmm}
{\sc A.~Heinecke, G.~Henry, M.~Hutchinson, and H.~Pabst}, {\em {LIBXSMM:
  accelerating small matrix multiplications by runtime code generation}}, in
  Proceedings of the 2016 International Conference for High Performance
  Computing, Networking, Storage and Analysis, SC16, IEEE, 2016,
  \url{https://github.com/hfp/libxsmm}.

\bibitem{Heinlein:2016:PIT}
{\sc A.~Heinlein, A.~Klawonn, and O.~Rheinbach}, {\em A parallel implementation
  of a two-level overlapping {S}chwarz method with energy-minimizing coarse
  space based on {T}rilinos}, SIAM Journal on Scientific Computing, 38 (2016),
  pp.~C713--C747.

\bibitem{hernandez2005ssf}
{\sc V.~Hernandez, J.~E. Roman, and V.~Vidal}, {\em {SLEP}c: A scalable and
  flexible toolkit for the solution of eigenvalue problems}, ACM Transactions
  on Mathematical Software, 31 (2005), pp.~351--362,
  \url{https://slepc.upv.es}.

\bibitem{heroux2005overview}
{\sc M.~A. Heroux, R.~A. Bartlett, V.~E. Howle, R.~J. Hoekstra, J.~J. Hu, T.~G.
  Kolda, R.~B. Lehoucq, K.~R. Long, R.~P. Pawlowski, E.~T. Phipps, et~al.},
  {\em {An overview of the Trilinos project}}, ACM Transactions on Mathematical
  Software (TOMS), 31 (2005), pp.~397--423, \url{https://trilinos.github.io}.

\bibitem{hestenes1952methods}
{\sc M.~R. Hestenes and E.~Stiefel}, {\em {Methods of Conjugate Gradients for
  Solving Linear Systems}}, Journal of Research of the National Bureau of
  Standards, 49 (1952), pp.~409--436.

\bibitem{ji2017breakdown}
{\sc H.~Ji and Y.~Li}, {\em A breakdown-free block conjugate gradient method},
  BIT Numerical Mathematics, 57 (2017), pp.~379--403.

\bibitem{jolivet2013scalable}
{\sc P.~Jolivet, F.~Hecht, F.~Nataf, and C.~Prud'homme}, {\em {Scalable Domain
  Decomposition Preconditioners for Heterogeneous Elliptic Problems}}, in
  Proceedings of the International Conference on High Performance Computing,
  Networking, Storage and Analysis, SC13, ACM, 2013.

\bibitem{jolivet2016block}
{\sc P.~Jolivet and P.-H. Tournier}, {\em {Block Iterative Methods and
  Recycling for Improved Scalability of Linear Solvers}}, in Proceedings of the
  2016 International Conference for High Performance Computing, Networking,
  Storage and Analysis, SC16, IEEE, 2016.

\bibitem{KALANTZIS2018136}
{\sc V.~Kalantzis, A.~C.~I. Malossi, C.~Bekas, A.~Curioni, E.~Gallopoulos, and
  Y.~Saad}, {\em A scalable iterative dense linear system solver for multiple
  right-hand sides in data analytics}, Parallel Computing, 74 (2018),
  pp.~136--153.

\bibitem{karypis1998fast}
{\sc G.~Karypis and V.~Kumar}, {\em A fast and high quality multilevel scheme
  for partitioning irregular graphs}, SIAM Journal on Scientific Computing, 20
  (1998), pp.~359--392,
  \url{http://glaros.dtc.umn.edu/gkhome/metis/parmetis/overview}.

\bibitem{knepley2009mesh}
{\sc M.~G. Knepley and D.~A. Karpeev}, {\em {Mesh algorithms for PDE with Sieve
  I: Mesh distribution}}, Scientific Programming, 17 (2009), pp.~215--230.

\bibitem{knyazev2001toward}
{\sc A.~V. Knyazev}, {\em {Toward the optimal preconditioned eigensolver:
  Locally optimal block preconditioned conjugate gradient method}}, SIAM
  journal on scientific computing, 23 (2001), pp.~517--541.

\bibitem{lehoucq1998arpack}
{\sc R.~Lehoucq, D.~Sorensen, and C.~Yang}, {\em {ARPACK} users’ guide\col\
  solution of large-scale eigenvalue problems with implicitly restarted Arnoldi
  methods}, vol.~6, Society for Industrial and Applied Mathematics, 1998,
  \url{http://www.caam.rice.edu/software/ARPACK}.

\bibitem{mandel1993balancing}
{\sc J.~Mandel}, {\em Balancing domain decomposition}, Communications in
  Numerical Methods in Engineering, 9 (1993), pp.~233--241.

\bibitem{may2016extreme}
{\sc D.~A. May, P.~Sanan, K.~Rupp, M.~G. Knepley, and B.~F. Smith}, {\em
  {Extreme-scale multigrid components within PETSc}}, in Proceedings of the
  Platform for Advanced Scientific Computing Conference, 2016.

\bibitem{moulin2018al}
{\sc J.~Moulin, P.~Jolivet, and O.~Marquet}, {\em {Augmented Lagrangian
  Preconditioner for Large-Scale Hydrodynamic Stability Analysis}}, Computer
  Methods in Applied Mechanics and Engineering, 351 (2019), pp.~718--743,
  \url{https://github.com/prj-/moulin2019al}.

\bibitem{notay2000flexible}
{\sc Y.~Notay}, {\em Flexible conjugate gradients}, SIAM Journal on Scientific
  Computing, 22 (2000), pp.~1444--1460.

\bibitem{o1980block}
{\sc D.~P. O'Leary}, {\em {The Block Conjugate Gradient Algorithm and Related
  Methods}}, Linear Algebra and its Applications, 29 (1980), pp.~293--322.

\bibitem{parks2006recycling}
{\sc M.~L. Parks, E.~de~Sturler, G.~Mackey, D.~D. Johnson, and S.~Maiti}, {\em
  {Recycling Krylov Subspaces for Sequences of Linear Systems}}, SIAM Journal
  on Scientific Computing, 28 (2006), pp.~1651--1674.

\bibitem{pechstein2017unified}
{\sc C.~Pechstein and C.~R. Dohrmann}, {\em {A unified framework for adaptive
  BDDC}}, Electronic Transactions on Numerical Analysis, 46 (2017),
  pp.~273--336.

\bibitem{saad1993flexible}
{\sc Y.~Saad}, {\em {A Flexible Inner--Outer Preconditioned GMRES Algorithm}},
  SIAM Journal on Scientific Computing, 14 (1993), pp.~461--469.

\bibitem{saad2003iterative}
{\sc Y.~Saad}, {\em {Iterative Methods for Sparse Linear Systems}}, SIAM, 2003.

\bibitem{saad1986gmres}
{\sc Y.~Saad and M.~H. Schultz}, {\em {GMRES: a generalized minimal residual
  algorithm for solving nonsymmetric linear systems}}, SIAM Journal on
  Scientific and Statistical Computing, 7 (1986), pp.~856--869.

\bibitem{sakurai2003projection}
{\sc T.~Sakurai and H.~Sugiura}, {\em A projection method for generalized
  eigenvalue problems using numerical integration}, Journal of Computational
  and Applied Mathematics, 159 (2003), pp.~119--128.

\bibitem{schwarz1870alternating}
{\sc H.~Schwarz}, {\em {\"Uber einen Grenz\"ubergang durch alternierendes
  Verfahren}}, Vierteljahrsschrift der Naturforschenden Gesellschaft in
  Z\"urich, 15 (1870), pp.~272--286.

\bibitem{si2013tet}
{\sc H.~Si}, {\em {TetGen: A Quality Tetrahedral Mesh Generator and 3D Delaunay
  Triangulator}}, Tech. Report~13, 2013,
  \url{http://wias-berlin.de/software/tetgen}.

\bibitem{vsistek2013parallel}
{\sc J.~{\v{S}}{\'\i}stek, J.~Mandel, B.~Soused{\'\i}k, and P.~Burda}, {\em
  Parallel implementation of multilevel {BDDC}}, in Numerical Mathematics and
  Advanced Applications 2011, Springer, 2013, pp.~681--689,
  \url{http://users.math.cas.cz/~sistek/software/bddcml.html}.

\bibitem{smith2004domain}
{\sc B.~F. Smith, P.~Bj{\o}rstad, and B.~Gropp}, {\em Domain decomposition:
  parallel multilevel methods for elliptic partial differential equations},
  Cambridge University Press, 2004.

\bibitem{spillane2011robust}
{\sc N.~Spillane, V.~Dolean, P.~Hauret, F.~Nataf, C.~Pechstein, and
  R.~Scheichl}, {\em A robust two-level domain decomposition preconditioner for
  systems of {PDE}s}, Comptes Rendus Mathématique, 349 (2011), pp.~1255--1259.

\bibitem{spillane2013abstract}
{\sc N.~Spillane, V.~Dolean, P.~Hauret, F.~Nataf, C.~Pechstein, and
  R.~Scheichl}, {\em Abstract robust coarse spaces for systems of {PDEs} via
  generalized eigenproblems in the overlaps}, Numerische Mathematik, 126
  (2013), pp.~741--700.

\bibitem{stathopoulos2002block}
{\sc A.~Stathopoulos and K.~Wu}, {\em {A Block Orthogonalization Procedure with
  Constant Synchronization Requirements}}, SIAM Journal on Scientific
  Computing, 23 (2002), pp.~2165--2182.

\bibitem{Stewart2002}
{\sc G.~W. Stewart}, {\em {A Krylov--Schur Algorithm for Large Eigenproblems}},
  SIAM Journal on Matrix Analysis and Applications, 23 (2002), pp.~601--614.

\bibitem{tang2009comparison}
{\sc J.~Tang, R.~Nabben, C.~Vuik, and Y.~Erlangga}, {\em Comparison of
  two-level preconditioners derived from deflation, domain decomposition and
  multigrid methods}, Journal of Scientific Computing, 39 (2009), pp.~340--370.

\bibitem{tournier2016micro}
{\sc P.-H. Tournier, I.~Aliferis, M.~Bonazzoli, M.~de~Buhan, M.~Darbas,
  V.~Dolean, F.~Hecht, P.~Jolivet, I.~El~Kanfoud, C.~Migliaccio, F.~Nataf,
  C.~Pichot, and S.~Semenov}, {\em {Microwave Tomographic Imaging of
  Cerebrovascular Accidents by Using High-Performance Computing}}, Parallel
  Computing, 85 (2019), pp.~88--97.

\bibitem{van1992bi}
{\sc H.~A. Van~der Vorst}, {\em {Bi-CGSTAB: a fast and smoothly converging
  variant of Bi-CG for the solution of nonsymmetric linear systems}}, SIAM
  Journal on Scientific and Statistical Computing, 13 (1992), pp.~631--644.

\bibitem{wakam2013158}
{\sc D.~N. Wakam and F.~Pacull}, {\em Memory efficient hybrid algebraic solvers
  for linear systems arising from compressible flows}, Computers \& Fluids, 80
  (2013), pp.~158--167.

\bibitem{zampini2016pcbddc}
{\sc S.~Zampini}, {\em {PCBDDC: a class of robust dual--primal methods in
  PETSc}}, SIAM Journal on Scientific Computing, 38 (2016), pp.~S282--S306,
  \url{https://www.mcs.anl.gov/petsc/petsc-master/docs/manualpages/PC/PCBDDC.html}.

\end{thebibliography}
\end{document}

### 2020-06-11 MatProduct.cpp

@@ -1,0 +1,134 @@
+#include <petsc.h>
+#include <iostream>
+#include <vector>
+
+int main(int argc, char** argv) {
+  Mat            A, B, C, D;
+  PetscInt       nbs = 10, ntype = 10, nN = 8, m, M, trial = 5;
+  PetscViewer    viewer;
+  PetscInt       *bs,*N;
+  char           **type;
+  PetscMPIInt    size;
+  PetscBool      flg;
+  char           file[PETSC_MAX_PATH_LEN];
+  PetscErrorCode ierr;
+  ierr = PetscInitialize(&argc, &argv, nullptr, nullptr);CHKERRQ(ierr);
+  if (ierr) return ierr;
+  ierr = MPI_Comm_size(PETSC_COMM_WORLD, &size);CHKERRQ(ierr);
+  if (size != 1) SETERRQ(PETSC_COMM_WORLD, PETSC_ERR_WRONG_MPI_SIZE, "This is a uniprocessor example only");
+  ierr = PetscOptionsGetString(nullptr, nullptr, "-f", file, PETSC_MAX_PATH_LEN, &flg);CHKERRQ(ierr);
+  if (!flg) SETERRQ(PETSC_COMM_WORLD, PETSC_ERR_USER_INPUT, "Must indicate binary file with the -f option");
+  ierr = PetscOptionsGetInt(nullptr, nullptr, "-trial", &trial, &flg);CHKERRQ(ierr);
+  bs = new PetscInt[nbs]();
+  ierr = PetscOptionsGetIntArray(nullptr, nullptr, "-bs", bs, &nbs, &flg);CHKERRQ(ierr);
+  if (!flg) {
+    nbs = 1;
+    *bs = 1;
+  }
+  N = new PetscInt[nN]();
+  ierr = PetscOptionsGetIntArray(nullptr, nullptr, "-N", N, &nN, &flg);CHKERRQ(ierr);
+  if (!flg) {
+    nN = 8;
+    N[0] = 1;  N[1] = 2;  N[2] = 4;  N[3] = 8;
+    N[4] = 16; N[5] = 32; N[6] = 64; N[7] = 128;
+  }
+  type = new char*[ntype]();
+  ierr = PetscOptionsGetStringArray(nullptr, nullptr, "-type", type, &ntype, &flg);CHKERRQ(ierr);
+  if (!flg) {
+    ntype = 1;
+    *type = new char[6];
[…]

### 2020-06-11 MatProduct.c
Initial commit

@@ -1,0 +1,413 @@
+/*
+   Author(s): Pierre Jolivet <pierre.jolivet@enseeiht.fr>
+              Jose E. Roman <jroman@dsic.upv.es>
+              Stefano Zampini <stefano.zampini@kaust.edu.sa>
+        Date: 2020-07-10
+
+   Copyright (C) 2020-     Centre National de la Recherche Scientifique
+                 2020-     Universitat Politècnica de València
+                 2020-     King Abdullah University of Science and Technology
+
+   This script is distributed in the hope that it will be useful,
+   but WITHOUT ANY WARRANTY; without even the implied warranty of
+   MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.
+
+   If you use this script, you are kindly asked to cite the following article:
+
+   "KSPHPDDM and PCHPDDM: Extending PETSc with Robust Overlapping Schwarz Preconditioners and Advanced Krylov Methods",
+   P. Jolivet, J. E. Roman, and S. Zampini (2020).
+ */
+
+#include <petsc.h>
+
+#if defined(PETSC_HAVE_MKL)
+#include <mkl.h>
+#define PetscStackCallMKLSparse(func, args) do {               \
+    sparse_status_t __ierr;                                    \
+    PetscStackPush(#func);                                     \
+    __ierr = func args;                                        \
+    PetscStackPop;                                             \
+    if (__ierr != SPARSE_STATUS_SUCCESS) SETERRQ2(PETSC_COMM_SELF, PETSC_ERR_LIB, "Error in %s(): error code %d", #func, (int)__ierr); \
+  } while (0)
+#else
+#define PetscStackCallMKLSparse(func, args) do {               \
+    SETERRQ(PETSC_COMM_SELF, PETSC_ERR_SUP, "No MKL support"); \
+  } while (0)
+#endif
+
[…]

### 2020-06-19 article.tex
From .cpp to .c

@@
 \hypersetup{
-  pdftitle={KSPHPDDM and PCHPDDM: A Class of Adaptive Krylov and Overlapping Schwarz Methods in PETSc},
+  pdftitle={KSPHPDDM and PCHPDDM: Two Classes of Adaptive Krylov and Overlapping Schwarz Methods in PETSc},
   pdfauthor={P. Jolivet and S. Zampini}
@@
 \definecolor{mediumgray}{RGB}{255,255,191}
+\usepackage[outline]{contour}
 \usepackage{tikz}
 \usepackage[binary-units=true]{siunitx}
-\usetikzlibrary{decorations.text,arrows.meta,calc,plotmarks,backgrounds,shadows,spy,fixedpointarithmetic,patterns,intersections,arrows.meta,shapes,decorations.text,decorations.pathmorphing,backgrounds,fit,positioning,shapes.symbols,chains,3d,calc,external}
+\usetikzlibrary{decorations.text,arrows.meta,calc,plotmarks,shadows,spy,fixedpointarithmetic,patterns,intersections,arrows.meta,shapes,decorations.text,decorations.pathmorphing,backgrounds,fit,positioning,shapes.symbols,chains,3d,calc,external}
+\pgfdeclarelayer{background}
+\pgfdeclarelayer{foreground}
+\pgfsetlayers{background,main,foreground}
 \long\def\ifnodedefined#1#2#3{%
@@
     to end-users adaptive Krylov and overlapping Schwarz methods.  The
-    flexibility of the implementation, which for example may deal with various
-    matrix formats, are showcased and explained.  Diverse examples are
+    flexibility of the implementation %, which for example may deal with various matrix formats, 
+    are showcased and explained.  Diverse examples are
     displayed, ranging from fracture mechanics to linear stability analysis, and
     the sets of options are thoroughly documented, so that they may be readily
-    tried on other applications.
+    tried on other applications and reproduce results.
 \end{abstract}
@@
 nonlinear systems of equations. It comes with many built-in features such as
-advanced preconditioning or tailored matrix formats. Also, it can be used to
-compile and interact with third-party libraries, such as
+advanced preconditioning or tailored matrix formats. Also, it can % be used to compile and 
+interact with third-party libraries, such as
 \emph{hypre}~\cite{falgout2002hypre} for linear solvers, TetGen~\cite{si2013tet}
@@
 packages. On the other hand, HPDDM is a much smaller project purely focused on
-domain decomposition preconditioners~\cite{jolivet2013scalable}, a small
-subclass of preconditioners available in PETSc as well, and advanced iterative
+domain decomposition preconditioners~\cite{jolivet2013scalable} % , a small subclass of preconditioners available in PETSc as well, 
+and advanced iterative
 methods~\cite{jolivet2016block}. Such methods deal efficiently with linear
@@
 methods and its domain decomposition preconditioners, into PETSc is explained.
-The design principles of the interface are showcased, alongside minimalist and
-easy-to-run examples. The paper is divided in two main parts:
+The design principles of the interface are showcased, with % alongside
+minimalist and easy-to-run examples. The paper is divided in two main parts:
 \cref{sec:KSPHPDDM} which introduces \pk{KSPHPDDM}, and \cref{sec:PCHPDDM} which
@@
 challenging high-dimensional systems, finding appropriate preconditioners is
-mandatory to make these methods robust, but this goes beyond the scope of this
-section. In PETSc, Krylov methods inherit from the \pk{KSP} base class. Some of the
+mandatory to make these methods robust. %, but this goes beyond the scope of this section.
+In PETSc, Krylov methods belong to the \pk{KSP} base class. Some of the
 most used types are \pk{KSPGMRES}, \pk{KSPCG}, or \pk{KSPBCGS}, which respectively implement
 the generalized minimal residual method~\cite{saad1986gmres}~(GMRES), conjugate
-gradient~\cite{hestenes1952methods}~(CG), or biconjugate gradient stabilized method~\cite{van1992bi}~(BCGS). Some of these methods are also
-available in other linear algebra backends such as~\emph{hypre} or DUNE~\cite{bastian2010generic}.
+gradient~\cite{hestenes1952methods}~(CG), or biconjugate gradient stabilized method~\cite{van1992bi}~(BCGS). % Some of these methods are also
+% available in other linear algebra backends such as~\emph{hypre} or DUNE~\cite{bastian2010generic}.
 
@@
 
-Recycling put aside, another important aspect of Krylov methods is their ability
+% Recycling put aside, 
+Another important aspect of Krylov methods is their ability
 to deal with multiple right-hand sides simultaneously. Block Krylov methods~\cite{gutknecht2006block} are
@@
 of the advanced iterative methods aforementioned. Although having PETSc and
-Trilinos interoperate is possible, there is currently no \pk{KSP} interface to Belos.
-This limits its composability inside complete PETSc applications. Furthermore,
+Trilinos interoperate is possible, there is currently no \pk{KSP} interface to Belos. % This limits its composability inside complete PETSc applications.
+Furthermore,
 as displayed in~\cref{sec:interface-ksp}, some Krylov methods implemented in
@@
 \pk{KSPHPDDM}, the interface between HPDDM Krylov methods and PETSc, is automatically
-registered since PETSc version 3.12 when one configures PETSc with the extra
+registered since PETSc version 3.12 when configuring PETSc with the extra
 flag -{}-download-hpddm. It is then possible to select the corresponding \pk{KSP}
 using the command line option -ksp\_type hpddm, or the routine \pk{KSPSetType}(ksp,
-\pk{KSPHPDDM}) and its Fortran equivalent.
+\pk{KSPHPDDM}).% and its Fortran equivalent.
 When \pk{KSPHPDDM} is being used, it is then possible to select between the following
@@
 \end{itemize}
-Unfortunately, PETSc does not have a general routine for applying $P_B$ to a
+Currently, PETSc does not have a general routine for applying $P_B$ to a
 tall-and-skinny dense matrix $Y$. In HPDDM, a specialized routine has been implemented
@@
 \pgfplotstableread{
-N    baij-1-T    baij-1-F   sbaij-1-T   sbaij-1-F  aij-1-T     aij-1-F    baij-3-T    baij-3-F   sbaij-3-T   sbaij-3-F  aij-3-T     aij-3-F    baij-10-T   baij-10-F  sbaij-10-T  sbaij-10-F aij-10-T    aij-10-F
-1    1.8027e-01  1.4696e+08 6.1586e-02  1.4696e+08 8.0527e-02  1.4696e+08 7.1397e-01  1.3535e+09 5.1037e-01  1.3638e+09 1.0686e+00  1.3535e+09 6.1909e+00  1.5159e+10 4.6958e+00  1.5206e+10 1.6114e+01  1.5159e+10
-2    3.5192e-01  2.9392e+08 1.6650e-01  3.2482e+08 1.8615e-01  3.0422e+08 1.0437e+00  2.7071e+09 9.3575e-01  2.9234e+09 1.4753e+00  2.7380e+09 1.4970e+01  3.0319e+10 1.5530e+01  3.2482e+10 1.6267e+01  3.0422e+10
-8    6.1599e-01  1.1757e+09 7.2168e-01  1.2993e+09 3.1676e-01  1.2169e+09 2.6101e+00  1.0828e+10 3.0701e+00  1.1694e+10 2.0870e+00  1.0952e+10 2.7886e+01  1.2128e+11 3.4780e+01  1.2993e+11 2.2569e+01  1.2169e+11
-16   1.9594e+00  2.3513e+09 2.6186e+00  2.5986e+09 6.2770e-01  2.4337e+09 7.5274e+00  2.1656e+10 9.8057e+00  2.3387e+10 4.1388e+00  2.1904e+10 3.6085e+01  2.4255e+11 3.6290e+01  2.5986e+11 4.5153e+01  2.4337e+11
-32   4.3453e+00  4.7026e+09 5.4669e+00  5.1972e+09 1.2538e+00  4.8675e+09 1.6166e+01  4.3313e+10 2.0638e+01  4.6775e+10 8.1531e+00  4.3807e+10 5.2586e+01  4.8510e+11 5.2132e+01  5.1972e+11 1.0705e+02  4.8675e+11
-64   9.6799e+00  9.4053e+09 1.1616e+01  1.0394e+10 2.5713e+00  9.7350e+09 3.4513e+01  8.6626e+10 4.2465e+01  9.3549e+10 1.6444e+01  8.7615e+10 8.3267e+01  9.7020e+11 8.6683e+01  1.0394e+12 2.2792e+02  9.7350e+11
+N    baij-1-T    baij-1-F   sbaij-1-T   sbaij-1-F  aij-1-T     aij-1-F    baij-3-T    baij-3-F   sbaij-3-T   sbaij-3-F  aij-3-T     aij-3-F    baij-10-T   baij-10-F  sbaij-10-T  sbaij-10-F aij-10-T    aij-10-F     cusparse-1-T    cusparse-1-F    cusparse-3-T    cusparse-3-F    cusparse-10-T   cusparse-10-F    aijmkl-seq-1-T      aijmkl-seq-3-T      aijmkl-seq-10-T     baijmkl-seq-1-T      baijmkl-seq-3-T      baijmkl-seq-10-T   sbaijmkl-seq-1-T      sbaijmkl-seq-3-T      sbaijmkl-seq-10-T         aijmkl-1-T      aijmkl-3-T      aijmkl-10-T     baijmkl-1-T      baijmkl-3-T      baijmkl-10-T   sbaijmkl-1-T      sbaijmkl-3-T      sbaijmkl-10-T
+1      2.0126e-01   1.4696e+08 8.0098e-02   1.4696e+08   9.3388e-02   1.4696e+08     8.9584e-01   1.3535e+09    6.2428e-01   1.3638e+09     1.2072e+00   1.3535e+09   4.0734e+00   5.4450e+09    2.2476e+00   5.4708e+09    5.0428e+00   5.4450e+09   4.4800e-02   1.5211e+08   3.7933e-01   1.3690e+09   1.5043e+00   5.4759e+09   6.6857e-02  7.4649e-01   3.1128e+00   4.7672e-01   6.2903e-01  2.2271e+00   4.7675e-01   6.2990e-01    2.2273e+00    7.8399e-03    8.6487e-02    3.5536e-01   5.2814e-02 6.6225e-02   2.4584e-01   4.5115e-02   6.6404e-02   2.4489e-01
+2      3.2635e-01   2.9392e+08 2.4204e-01   3.2482e+08   1.1522e-01   3.0422e+08     1.2835e+00   2.7071e+09    1.0697e+00   2.9234e+09     9.9480e-01   2.7380e+09   1.4504e+01   1.0890e+10    1.5265e+01   1.1694e+10    3.7799e+00   1.0952e+10   5.2164e-03   3.0422e+08   3.4056e-02   2.7380e+09   7.7659e-02   1.0952e+10   1.1703e-01  9.3872e-01   3.5976e+00   4.7905e-01   7.3497e-01  3.3331e+00   4.7886e-01   7.3415e-01    3.3297e+00    1.2450e-02    1.0124e-01    3.9873e-01   4.7633e-02 7.4107e-02   2.8513e-01   4.7733e-02   7.4693e-02   2.8507e-01
+8      7.9573e-01   1.1757e+09 1.0070e+00   1.2993e+09   3.3729e-01   1.2169e+09     3.3663e+00   1.0828e+10    3.9830e+00   1.1694e+10     2.4656e+00   1.0952e+10   2.2209e+01   4.3560e+10    2.2545e+01   4.6775e+10    9.2040e+00   4.3807e+10   1.3652e-02   1.2169e+09   8.2139e-02   1.0952e+10   1.8984e-01   4.3807e+10   4.2722e-01  2.8760e+00   1.0913e+01   1.3256e+00   1.7457e+00  7.9644e+00   1.3210e+00   1.7410e+00    7.9679e+00    4.2997e-02    2.6899e-01    1.6814e+00   1.3183e-01 1.7832e-01   8.8593e-01   1.3082e-01   1.7947e-01   8.8420e-01
+16     2.1378e+00   2.3513e+09 3.0837e+00   2.5986e+09   6.7295e-01   2.4337e+09     9.1522e+00   2.1656e+10    1.1762e+01   2.3387e+10     4.9326e+00   2.1904e+10   3.1530e+01   8.7120e+10    2.9334e+01   9.3549e+10    1.8462e+01   8.7615e+10   2.4907e-02   2.4337e+09   1.3832e-01   2.1904e+10   3.3609e-01   8.7615e+10   8.5514e-01  5.7493e+00   2.1980e+01   2.6391e+00   3.4887e+00  1.5947e+01   2.6455e+00   3.4893e+00    1.5928e+01    8.4255e-02    5.4284e-01    3.3861e+00   2.6162e-01 3.5777e-01   1.7955e+00   2.6238e-01   3.5880e-01   1.7923e+00
+32     4.2781e+00   4.7026e+09 6.5426e+00   5.1972e+09   1.3447e+00   4.8675e+09     1.8545e+01   4.3313e+10    2.4015e+01   4.6775e+10     9.8656e+00   4.3807e+10   4.4678e+01   1.7424e+11    4.1021e+01   1.8710e+11    3.6822e+01   1.7523e+11   4.7533e-02   4.8675e+09   2.6294e-01   4.3807e+10   6.3836e-01   1.7523e+11   1.7075e+00  1.1487e+01   4.3644e+01   5.2831e+00   6.9603e+00  3.1855e+01   5.2907e+00   6.9739e+00    3.1852e+01    1.6480e-01    1.0807e+00    6.7653e+00   5.2012e-01 7.1464e-01   3.5401e+00   5.2109e-01   7.1542e-01   3.5382e+00
+64     9.9341e+00   9.4053e+09 1.3822e+01   1.0394e+10   2.6927e+00   9.7350e+09     3.5490e+01   8.6626e+10    4.6966e+01   9.3549e+10     1.9724e+01   8.7615e+10   6.7668e+01   3.4848e+11  6.3229e+01   3.7420e+11  7.4386e+01   3.5046e+11   8.3249e-02   9.7350e+09   5.0956e-01   8.7615e+10   1.2334e+00   3.5046e+11   3.4264e+00  2.3052e+01   8.7358e+01   1.0573e+01   1.3960e+01  6.3695e+01   1.0581e+01   1.3961e+01    6.3949e+01    3.2990e-01    2.1734e+00    1.3526e+01   1.0453e+00 1.4281e+00   7.0763e+00   1.0497e+00   1.4326e+00   7.1714e+00
 }\loadedMatProduct
@@
 Though the numerical values of $A$ are of no interest here, the sparsity pattern is rather common and frequently encountered, e.g., when discretizing partial differential equations.
-Then, a random dense matrix $T$ of dimension $b$ is generated, for $b \in \{ 1,3,10 \}$, and the following matrix is evaluated $\mathbb{A} = A \otimes T$.
+Then, a random dense matrix $T$ of dimension $b$ is generated, for $b \in \{ 1,3,6 \}$, and the following matrix is evaluated $\mathbb{A} = A \otimes T$.
 The matrix $\mathbb{A}$ is then assembled into all three formats described above.
@@
 \begin{Verbatim}[fontsize=\footnotesize,frame=single,framerule=0.1mm,commandchars=&\[\]]
-&fvtextcolor[mygreen][$] &fvtextcolor[myyellow][mpicxx] MatProduct.cpp -O3 -std=c++11 -I&fvtextcolor[myyellow][${PETSC_DIR}]/&fvtextcolor[myyellow][${PETSC_ARCH}]/include                     &fvtextcolor[myred][\]
+&fvtextcolor[mygreen][$] &fvtextcolor[myyellow][mpicc] MatProduct.c -O3 -I&fvtextcolor[myyellow][${PETSC_DIR}]/&fvtextcolor[myyellow][${PETSC_ARCH}]/include                        &fvtextcolor[myred][\]
   -I&fvtextcolor[myyellow][${PETSC_DIR}]/include -L&fvtextcolor[myyellow][${PETSC_DIR}]/&fvtextcolor[myyellow][${PETSC_ARCH}]/lib -lpetsc -o MatProduct
 &fvtextcolor[mygreen][$] &fvtextcolor[myyellow][mpirun] -n 1 ./MatProduct -f binaryoutput -log_view                                 &fvtextcolor[myred][\]
-  -bs 1,3,10 -N 1,2,8,16,32,64 -type aij,baij,sbaij
+  -bs 1,3,6 -N 1,2,8,16,32,64 -type aij,aijmkl,baij,baijmkl,sbaij,sbaijmkl
 \end{Verbatim}
 \begin{figure}[htbp]
+\pgfkeys{/pgf/fpu}
+  \pgfmathparse{1.4696e+08/9.3388e-02/1000000000.0}\edef\storeresultI{\pgfmathresult}
+  \pgfmathparse{1.3535e+09/8.9584e-01/1000000000.0}\edef\storeresultII{\pgfmathresult}
+  \pgfmathparse{5.4450e+09/4.0734e+00/1000000000.0}\edef\storeresultIII{\pgfmathresult}
+\pgfkeys{/pgf/fpu=false}
+  % \hspace*{-1.5cm}
   \centering
@@
         bar width=3.5pt,
-        enlarge x limits=0.1, height=4cm,
-        width=0.98\textwidth,legend columns=4,
-        legend style={at={(0.5,1.65)}, legend cell align=left,
-        anchor=north},
+        enlarge x limits=0.101,height=4cm,
+        width=0.8\textwidth,
+        ylabel={},
+        legend columns=3, legend style={legend cell align=left,at={(0.5,1.03)},anchor=south},
+        xtick = {\empty},
+        xmajorgrids=true,
+        extra x ticks = {1,2,4,8,16,32},
+        extra x tick labels = {\empty},
+        extra tick style={
+            grid style={dotted,black}
+        },
+        xtick pos=left,name=plot1,
+        xticklabels={\empty},xmajorticks=false,
+        ytick = {0.5,1,1.5,2},
+        ymajorgrids=true,yticklabel={
+\pgfmathparse{int(10*\tick)}
+\ifnum\pgfmathresult=0
+\else
+\pgfmathparse{100*\tick}
+$\pgfmathprintnumber{\pgfmathresult}$\%
+\fi
+},
+          cycle list={blueaccent, darkgray, mediumgray},%,last},
+          clip=false,point meta=exp(x),
+      ]
+          \addplot+[nodes near coords style={xshift=0.5cm,yshift=-0.1cm,above}, nodes near coords*={\rotatebox{45}{\tiny\color{black}{\pgfmathfloattofixed{\pgfplotspointmeta}\pgfmathparse{int(round(\pgfmathresult))}\ifnum\pgfmathresult=1\pgfmathprintnumber\storeresultI~GFLOP/s\fi}}},
+draw=black,fill, area legend,postaction={pattern=crosshatch dots}] table [x expr=\thisrow{N}>7?\thisrow{N}/2:\thisrow{N},y expr=\thisrow{aij-1-F}/\thisrow{aij-1-T}/(\storeresultI*1000000000.0)]{\loadedMatProduct};\label{pgfplots:aij1}
+          \addplot+[draw=black,fill, area legend,postaction={pattern=crosshatch dots}] table [x expr=\thisrow{N}>7?\thisrow{N}/2:\thisrow{N},y expr=\thisrow{baij-1-F}/\thisrow{baij-1-T}/(\storeresultI*1000000000.0)]{\loadedMatProduct};\label{pgfplots:baij1}
+          \addplot+[draw=black,fill, area legend,postaction={pattern=crosshatch dots}] table [x expr=\thisrow{N}>7?\thisrow{N}/2:\thisrow{N},y expr=\thisrow{sbaij-1-F}/\thisrow{sbaij-1-T}/(\storeresultI*1000000000.0)]{\loadedMatProduct};\label{pgfplots:sbaij1}
+           \addplot+[draw=black,fill, area legend,postaction={pattern=bricks}] table [x expr=\thisrow{N}>7?\thisrow{N}/2:\thisrow{N},y expr=\thisrow{aij-1-F}/\thisrow{aijmkl-seq-1-T}/(\storeresultI*1000000000.0)]{\loadedMatProduct};\label{pgfplots:mklaijseq1}
+           \addplot+[draw=black,fill, area legend,postaction={pattern=bricks}] table [x expr=\thisrow{N}>7?\thisrow{N}/2:\thisrow{N},y expr=\thisrow{baij-1-F}/\thisrow{baijmkl-seq-1-T}/(\storeresultI*1000000000.0)]{\loadedMatProduct};\label{pgfplots:mklbaijseq1}
+           \addplot+[draw=black,fill, area legend,postaction={pattern=bricks}] table [x expr=\thisrow{N}>7?\thisrow{N}/2:\thisrow{N},y expr=\thisrow{sbaij-1-F}/\thisrow{sbaijmkl-seq-1-T}/(\storeresultI*1000000000.0)]{\loadedMatProduct};\label{pgfplots:mklsbaijseq1}
+           \addlegendentry{\pk{MatAIJ}}
+           \addlegendentry{\pk{MatBAIJ}}
+           \addlegendentry{\pk{MatSBAIJ}}
+           \addlegendentry{\pk{MatAIJMKL}}
+           \addlegendentry{\pk{MatBAIJMKL}}
+           \addlegendentry{MatSBAIJMKL}
+\node[rotate=0,fill=white,inner sep=0pt,anchor=west] at (rel axis cs:1.02, 0.5) {$b=1$};
+    \end{semilogxaxis}
+      \begin{semilogxaxis}[
+        ymin=0,
+        ybar=1.5pt,
+        bar width=3.5pt,
+        enlarge x limits=0.101,height=4cm,
+        width=0.8\textwidth,
+        % legend columns=1, legend style={legend cell align=left,at={(1.02,0.5)},anchor=west},
+        ylabel={Efficiency},
+        xtick = {\empty},
+        xmajorgrids=true,
+        extra x ticks = {1,2,4,8,16,32},
+        extra x tick labels = {\empty},
+        extra tick style={
+            grid style={dotted,black}
+        },
+        xtick pos=left,name=plot2,at={($(plot1.south)-(0,0.2cm)$)},anchor=north,
+        xticklabels={\empty},xmajorticks=false,
+        ytick = {1,2,3,4},
+        ymajorgrids=true,yticklabel={
+\pgfmathparse{int(10*\tick)}
+\ifnum\pgfmathresult=0
+\else
+\pgfmathparse{100*\tick}
+$\pgfmathprintnumber{\pgfmathresult}$\%
+\fi
+},
+          cycle list={blueaccent, darkgray, mediumgray},%,last},
+          clip=false,point meta=exp(x)
+      ]
+          \addplot+[draw=black,fill, area legend,postaction={pattern=crosshatch dots}] table [x expr=\thisrow{N}>7?\thisrow{N}/2:\thisrow{N},y expr=\thisrow{aij-3-F}/\thisrow{aij-3-T}/(\storeresultII*1000000000.0)]{\loadedMatProduct};\label{pgfplots:aij3}
+          \addplot+[nodes near coords style={xshift=0.5cm,yshift=-0.1cm,above}, nodes near coords*={\rotatebox{45}{\tiny\color{black}{\pgfmathfloattofixed{\pgfplotspointmeta}\pgfmathparse{int(round(\pgfmathresult))}\ifnum\pgfmathresult=1\pgfmathprintnumber\storeresultII~GFLOP/s\fi}}},
+              draw=black,fill, area legend,postaction={pattern=crosshatch dots}] table [x expr=\thisrow{N}>7?\thisrow{N}/2:\thisrow{N},y expr=\thisrow{baij-3-F}/\thisrow{baij-3-T}/(\storeresultII*1000000000.0)]{\loadedMatProduct};\label{pgfplots:baij3}
+          \addplot+[draw=black,fill, area legend,postaction={pattern=crosshatch dots}] table [x expr=\thisrow{N}>7?\thisrow{N}/2:\thisrow{N},y expr=\thisrow{sbaij-3-F}/\thisrow{sbaij-3-T}/(\storeresultII*1000000000.0)]{\loadedMatProduct};\label{pgfplots:sbaij3}
+           \addplot+[draw=black,fill, area legend,postaction={pattern=bricks}] table [x expr=\thisrow{N}>7?\thisrow{N}/2:\thisrow{N},y expr=\thisrow{aij-3-F}/\thisrow{aijmkl-seq-3-T}/(\storeresultII*1000000000.0)]{\loadedMatProduct};\label{pgfplots:mklaijseq3}
+           \addplot+[draw=black,fill, area legend,postaction={pattern=bricks}] table [x expr=\thisrow{N}>7?\thisrow{N}/2:\thisrow{N},y expr=\thisrow{baij-3-F}/\thisrow{baijmkl-seq-3-T}/(\storeresultII*1000000000.0)]{\loadedMatProduct};\label{pgfplots:mklbaijseq3}
+           \addplot+[draw=black,fill, area legend,postaction={pattern=bricks}] table [x expr=\thisrow{N}>7?\thisrow{N}/2:\thisrow{N},y expr=\thisrow{sbaij-3-F}/\thisrow{sbaijmkl-seq-3-T}/(\storeresultII*1000000000.0)]{\loadedMatProduct};\label{pgfplots:mklsbaijseq3}
+           % \addlegendentry{\pk{MatAIJ}}
+           % \addlegendentry{\pk{MatBAIJ}}
+           % \addlegendentry{\pk{MatSBAIJ}}
+           % \addlegendentry{\pk{MatAIJMKL}}
+           % \addlegendentry{\pk{MatBAIJMKL}}
+           % \addlegendentry{MatSBAIJMKL}
+\node[rotate=0,fill=white,inner sep=0pt,anchor=west] at (rel axis cs:1.02, 0.5) {$b=3$};
+    \end{semilogxaxis}
+      \begin{semilogxaxis}[
+        ymin=0,
+        ybar=1.5pt,
+        bar width=3.5pt,
+        enlarge x limits=0.101,height=4cm,
+        width=0.8\textwidth,
         xlabel={\# of right-hand sides ($N$)},
-        ylabel={Efficiency},
-        xtick = {1,2,4,8,16,32},
-ytick = {0.2,1,2.5,5},ylabel style={yshift=6pt},xtick pos=left,
+        ylabel={},
+        xtick = {\empty},
+        xmajorgrids=true,
+        extra x ticks = {1,2,4,8,16,32},
+        extra x tick labels = {1,2,8,16,32,64},
+        extra tick style={
+            grid style={dotted,black}
+        },
+        xtick pos=left,name=plot3,at={($(plot2.south)-(0,0.2cm)$)},anchor=north,
         xticklabels={1,2,8,16,32,64},
+        ytick = {1,2,3,4},
         ymajorgrids=true,yticklabel={
@@
         %xmajorgrids=true,
-          cycle list={blueaccent, darkgray, mediumgray},%, lightgray,last},
+          cycle list={blueaccent, darkgray, mediumgray},%,last},
+          clip=false, point meta = exp(x)
       ]
-\pgfkeys{/pgf/fpu}
-  \pgfmathparse{1.4696e+08/8.0527e-02/1000000000.0}\edef\storeresultI{\pgfmathresult}
-  \pgfmathparse{1.3535e+09/7.1397e-01/1000000000.0}\edef\storeresultII{\pgfmathresult}
-  \pgfmathparse{1.5159e+10/6.1909e+00/1000000000.0}\edef\storeresultIII{\pgfmathresult}
-\pgfkeys{/pgf/fpu=false}
-\node[font=\tiny,rotate=45,fill=white,inner sep=0pt,anchor=west] at (axis cs:0.75, 1.2) {\pgfmathprintnumber\storeresultI~GFLOP/s};
-\node[font=\tiny,rotate=45,fill=white,inner sep=0pt,anchor=west] at (axis cs:0.97, 1.2) {\pgfmathprintnumber\storeresultII~GFLOP/s};
-\node[font=\tiny,rotate=45,fill=white,inner sep=0pt,anchor=west] at (axis cs:1.18, 1.2) {\pgfmathprintnumber\storeresultIII~GFLOP/s};
-\addlegendimage{empty legend}
-          \addplot+[draw=black,fill, area legend,postaction={pattern=crosshatch dots}] table [x expr=-0.5 + \thisrow{N}>7?\thisrow{N}/2:\thisrow{N},y expr=\thisrow{aij-1-F}/\thisrow{aij-1-T}/(1.4696e+08/8.0527e-02)]{\loadedMatProduct};\label{pgfplots:aij1}
-          \addplot+[draw=black,fill, area legend,postaction={pattern=crosshatch dots}] table [x expr=-0.5 + \thisrow{N}>7?\thisrow{N}/2:\thisrow{N},y expr=\thisrow{baij-1-F}/\thisrow{baij-1-T}/(1.4696e+08/8.0527e-02)]{\loadedMatProduct};\label{pgfplots:baij1}
-          \addplot+[draw=black,fill, area legend,postaction={pattern=crosshatch dots}] table [x expr=-0.5 + \thisrow{N}>7?\thisrow{N}/2:\thisrow{N},y expr=\thisrow{sbaij-1-F}/\thisrow{sbaij-1-T}/(1.4696e+08/8.0527e-02)]{\loadedMatProduct};\label{pgfplots:sbaij1}
-\addlegendimage{empty legend}
-           \addplot+[draw=black,fill, area legend] table [x expr=\thisrow{N}>7?\thisrow{N}/2:\thisrow{N},y expr=\thisrow{aij-3-F}/\thisrow{aij-3-T}/(1.3535e+09/7.1397e-01)]{\loadedMatProduct};\label{pgfplots:aij3}
-           \addplot+[draw=black,fill, area legend] table [x expr=\thisrow{N}>7?\thisrow{N}/2:\thisrow{N},y expr=\thisrow{baij-3-F}/\thisrow{baij-3-T}/(1.3535e+09/7.1397e-01)]{\loadedMatProduct};\label{pgfplots:baij3}
-           \addplot+[draw=black,fill, area legend] table [x expr=\thisrow{N}>7?\thisrow{N}/2:\thisrow{N},y expr=\thisrow{sbaij-3-F}/\thisrow{sbaij-3-T}/(1.3535e+09/7.1397e-01)]{\loadedMatProduct};\label{pgfplots:sbaij3}
-\addlegendimage{empty legend}
-           \addplot+[draw=black,fill, area legend,postaction={pattern=bricks}] table [x expr=+0.5 + \thisrow{N}>7?\thisrow{N}/2:\thisrow{N},y expr=\thisrow{aij-10-F}/\thisrow{aij-10-T}/(1.5159e+10/6.1909e+00)]{\loadedMatProduct};\label{pgfplots:aij10}
-           \addplot+[draw=black,fill, area legend,postaction={pattern=bricks}] table [x expr=+0.5 + \thisrow{N}>7?\thisrow{N}/2:\thisrow{N},y expr=\thisrow{baij-10-F}/\thisrow{baij-10-T}/(1.5159e+10/6.1909e+00)]{\loadedMatProduct};\label{pgfplots:baij10}
-           \addplot+[draw=black,fill, area legend,postaction={pattern=bricks}] table [x expr=+0.5 + \thisrow{N}>7?\thisrow{N}/2:\thisrow{N},y expr=\thisrow{sbaij-10-F}/\thisrow{sbaij-10-T}/(1.5159e+10/6.1909e+00)]{\loadedMatProduct};\label{pgfplots:sbaij10}
-          \addlegendentry{$b = 1$}
-          \addlegendentry{              }
-          \addlegendentry{               }
-          \addlegendentry{                }
-          \addlegendentry{$b = 3$}
-          \addlegendentry{\pk{MatAIJ}}
-          \addlegendentry{\pk{MatBAIJ}}
-          \addlegendentry{\pk{MatSBAIJ}}
-          \addlegendentry{$b = 10$~~}
-          \addlegendentry{              }
-          \addlegendentry{               }
-          \addlegendentry{                }
+      \contourlength{.6pt}
+\node[rotate=0,fill=white,inner sep=0pt,anchor=west] at (rel axis cs:1.02, 0.5) {$b=6$};
+          \addplot+[draw=black,fill, area legend,postaction={pattern=crosshatch dots}] table [x expr=\thisrow{N}>7?\thisrow{N}/2:\thisrow{N},y expr=\thisrow{aij-10-F}/\thisrow{aij-10-T}/(\storeresultIII*1000000000.0)]{\loadedMatProduct};\label{pgfplots:aij10}
+          \addplot+[nodes near coords style={xshift=0.5cm,yshift=-0.1cm,above}, nodes near coords*={\rotatebox{45}{\tiny\color{black}{\pgfmathfloattofixed{\pgfplotspointmeta}\pgfmathparse{int(round(\pgfmathresult))}\ifnum\pgfmathresult=1\contour{white}{\pgfmathprintnumber\storeresultIII}~GFLOP/s\fi}}},draw=black,fill, area legend,postaction={pattern=crosshatch dots}] table [x expr=\thisrow{N}>7?\thisrow{N}/2:\thisrow{N},y expr=\thisrow{baij-10-F}/\thisrow{baij-10-T}/(\storeresultIII*1000000000.0)]{\loadedMatProduct};\label{pgfplots:baij10}
+          \addplot+[draw=black,fill, area legend,postaction={pattern=crosshatch dots}] table [x expr=\thisrow{N}>7?\thisrow{N}/2:\thisrow{N},y expr=\thisrow{sbaij-10-F}/\thisrow{sbaij-10-T}/(\storeresultIII*1000000000.0)]{\loadedMatProduct};\label{pgfplots:sbaij10}
+           \addplot+[draw=black,fill, area legend,postaction={pattern=bricks}] table [x expr=\thisrow{N}>7?\thisrow{N}/2:\thisrow{N},y expr=\thisrow{aij-10-F}/\thisrow{aijmkl-seq-10-T}/(\storeresultIII*1000000000.0)]{\loadedMatProduct};\label{pgfplots:mklaijseq10}
+           \addplot+[draw=black,fill, area legend,postaction={pattern=bricks}] table [x expr=\thisrow{N}>7?\thisrow{N}/2:\thisrow{N},y expr=\thisrow{baij-10-F}/\thisrow{baijmkl-seq-10-T}/(\storeresultIII*1000000000.0)]{\loadedMatProduct};\label{pgfplots:mklbaijseq10}
+           \addplot+[draw=black,fill, area legend,postaction={pattern=bricks}] table [x expr=\thisrow{N}>7?\thisrow{N}/2:\thisrow{N},y expr=\thisrow{sbaij-10-F}/\thisrow{sbaijmkl-seq-10-T}/(\storeresultIII*1000000000.0)]{\loadedMatProduct};\label{pgfplots:mklsbaijseq10}
     \end{semilogxaxis}
@@
    \caption{Performance of the matrix--matrix multiplication as implemented in PETSc for three sparse matrix types.
-As baseline values with $N=1$, for $b=1$, performance of \pk{MatAIJ} matrix--vector multiplication is used~\ref{pgfplots:aij1}, while for $b=3$ (resp.~$b=10$), it is the performance of \pk{MatBAIJ} matrix--vector multiplication~\ref{pgfplots:baij3} (resp.~\ref{pgfplots:baij10}).\label{fig:perfAX}}
+As baseline values with $N=1$, for $b=1$, performance of \pk{MatAIJ} matrix--vector multiplication is used~\ref{pgfplots:aij1}, while for $b=\{3,6\}$, it is the performance of \pk{MatBAIJ} matrix--vector multiplication~\ref{pgfplots:baij3}.\label{fig:perfAX}}
 \end{figure}
-Some conclusions may be drawn:
-\begin{itemize}
-    \item with a block size of 1, using any type but \pk{MatAIJ} is counterproductive, see~\ref{pgfplots:baij1} and~\ref{pgfplots:sbaij1}.
-However, with this format, as soon as the number of columns is greater or equal than 8, the efficiency reaches approximately 200\%, see~\ref{pgfplots:aij1};
-    \item for block sizes lower or equal than 5, loops for matrix--vector and matrix--matrix multiplications are unrolled by hand for block formats.
-This yields rather disappointing performances, except for \pk{MatAIJ}, see~\ref{pgfplots:aij3}, and its 250\% efficiency, even with a moderate number of columns.
-For block formats \ref{pgfplots:baij3} and~\ref{pgfplots:sbaij3}, given the small value of $b$, it could be beneficial to switch to optimized libraries for small matrices, e.g., LIBXSMM~\cite{heinecke2016libxsmm};
-    \item for greater block sizes, ?gemv or ?gemm are used instead for block formats.
-For \pk{MatAIJ}~\ref{pgfplots:aij10}, the efficiency quickly caps below 250\% for $N\in\{8,16\}$, but for block formats, this reaches almost 500\% for large number of columns, see~\ref{pgfplots:baij10} and~\ref{pgfplots:sbaij10} for $N=64$.
-\end{itemize}
-These results are of course specific to the sparsity pattern of the initial matrix $A$, and thus $\mathbb{A} = A\otimes T$, and matrix types.
-For example, if the matrices were to be offloaded to a GPU, one could instead benchmark \pk{MatAIJcuSPARSE}.
+
+\begin{figure}[htbp]
+\pgfkeys{/pgf/fpu}
+  \pgfmathparse{1.4696e+08/7.8399e-03/1000000000.0}\edef\storeresultI{\pgfmathresult}
+  \pgfmathparse{1.3535e+09/6.6225e-02/1000000000.0}\edef\storeresultII{\pgfmathresult}
+  \pgfmathparse{5.4450e+09/2.4584e-01/1000000000.0}\edef\storeresultIII{\pgfmathresult}
+\pgfkeys{/pgf/fpu=false}
+  \centering
+    \begin{tikzpicture}
+      \begin{semilogxaxis}[
+        ymin=0,
+        ybar=1.5pt,
+        bar width=3.5pt,
+        enlarge x limits=0.101,height=4cm,
+        width=0.8\textwidth,
+        ylabel={},
+        legend columns=4, legend style={legend cell align=left,at={(0.5,1.03)},anchor=south},
+        xtick = {\empty},
+        xmajorgrids=true,
+        extra x ticks = {1,2,4,8,16,32},
+        extra x tick labels = {\empty},
+        extra tick style={
+            grid style={dotted,black}
+        },
+        xtick pos=left,name=plot4,
+        xticklabels={\empty},xmajorticks=false,
+        ytick = {1,3.5,6},
+        ymajorgrids=true,yticklabel={
+\pgfmathparse{int(10*\tick)}
+\ifnum\pgfmathresult=0
+\else
+\pgfmathparse{100*\tick}
+$\pgfmathprintnumber{\pgfmathresult}$\%
+\fi
+},
+        %xmajorgrids=true,
+          cycle list={blueaccent, darkgray, mediumgray,last},
+          point meta=exp(x),clip=false
+      ]
+% \node[font=\tiny,rotate=45,fill=white,inner sep=0pt,anchor=west] at (axis cs:0.725, 1.2) {\pgfmathprintnumber\storeresultI~GFLOP/s};
+% \node[font=\tiny,rotate=45,fill=white,inner sep=0pt,anchor=west] at (axis cs:0.95, 1.2) {\pgfmathprintnumber\storeresultII~GFLOP/s};
+% \node[font=\tiny,rotate=45,fill=white,inner sep=0pt,anchor=west] at (axis cs:1.2, 1.2) {\pgfmathprintnumber\storeresultIII~GFLOP/s};
+          \addplot+[nodes near coords style={xshift=0.55cm,yshift=-0.1cm,above}, nodes near coords*={\rotatebox{45}{\tiny\color{black}{\pgfmathfloattofixed{\pgfplotspointmeta}\pgfmathparse{int(round(\pgfmathresult))}\ifnum\pgfmathresult=1\pgfmathprintnumber\storeresultI~GFLOP/s\fi}}},draw=black,fill, area legend] table [x expr=\thisrow{N}>7?\thisrow{N}/2:\thisrow{N},y expr=\thisrow{aij-1-F}/\thisrow{aijmkl-1-T}/(\storeresultI*1000000000.0)]{\loadedMatProduct};\label{pgfplots:aijmkl1}
+          \addplot+[draw=black,fill, area legend] table [x expr=\thisrow{N}>7?\thisrow{N}/2:\thisrow{N},y expr=\thisrow{baij-1-F}/\thisrow{baijmkl-1-T}/(\storeresultI*1000000000.0)]{\loadedMatProduct};\label{pgfplots:baijmkl1}
+          \addplot+[draw=black,fill, area legend] table [x expr=\thisrow{N}>7?\thisrow{N}/2:\thisrow{N},y expr=\thisrow{sbaij-1-F}/\thisrow{sbaijmkl-1-T}/(\storeresultI*1000000000.0)]{\loadedMatProduct};\label{pgfplots:sbaijmkl1}
+           \addplot+[draw=black,fill, area legend] table [x expr=\thisrow{N}>7?\thisrow{N}/2:\thisrow{N},y expr=\thisrow{aij-1-F}/\thisrow{cusparse-1-T}/(\storeresultI*1000000000.0)]{\loadedMatProduct};\label{pgfplots:cusparse1}
+           \addlegendentry{\pk{MatAIJMKL}}
+           \addlegendentry{\pk{MatBAIJMKL}}
+           \addlegendentry{MatSBAIJMKL}
+           \addlegendentry{\pk{MatAIJcuSPARSE}}
+\node[rotate=0,fill=white,inner sep=0pt,anchor=west] at (rel axis cs:1.02, 0.5) {$b=1$};
+    \end{semilogxaxis}
+      \begin{semilogxaxis}[
+        ymin=0,
+        ybar=1.5pt,
+        bar width=3.5pt,
+        enlarge x limits=0.101,height=4cm,
+        width=0.8\textwidth,
+        % legend columns=1, legend style={legend cell align=left,at={(1.02,0.5)},anchor=west},
+        ylabel={Efficiency},
+        xtick = {\empty},
+        xmajorgrids=true,
+        extra x ticks = {1,2,4,8,16,32},
+        extra x tick labels = {\empty},
+        extra tick style={
+            grid style={dotted,black}
+        },
+        xtick pos=left,name=plot5,at={($(plot4.south)-(0,0.2cm)$)},anchor=north,
+        xticklabels={\empty},xmajorticks=false,
+        ytick = {1,4.5,8},
+        ymajorgrids=true,yticklabel={
+\pgfmathparse{int(10*\tick)}
+\ifnum\pgfmathresult=0
+\else
+\pgfmathparse{100*\tick}
+$\pgfmathprintnumber{\pgfmathresult}$\%
+\fi
+},
+        %xmajorgrids=true,
+          cycle list={blueaccent, darkgray, mediumgray,last},
+          point meta=exp(x),clip=false
+      ]
+% \node[font=\tiny,rotate=45,fill=white,inner sep=0pt,anchor=west] at (axis cs:0.725, 1.2) {\pgfmathprintnumber\storeresultI~GFLOP/s};
+% \node[font=\tiny,rotate=45,fill=white,inner sep=0pt,anchor=west] at (axis cs:0.95, 1.2) {\pgfmathprintnumber\storeresultII~GFLOP/s};
+% \node[font=\tiny,rotate=45,fill=white,inner sep=0pt,anchor=west] at (axis cs:1.2, 1.2) {\pgfmathprintnumber\storeresultIII~GFLOP/s};
+          \addplot+[draw=black,fill, area legend] table [x expr=\thisrow{N}>7?\thisrow{N}/2:\thisrow{N},y expr=\thisrow{aij-3-F}/\thisrow{aijmkl-3-T}/(\storeresultII*1000000000.0)]{\loadedMatProduct};\label{pgfplots:aijmkl3}
+          \addplot+[nodes near coords style={xshift=0.55cm,yshift=-0.1cm,above}, nodes near coords*={\rotatebox{45}{\tiny\color{black}{\pgfmathfloattofixed{\pgfplotspointmeta}\pgfmathparse{int(round(\pgfmathresult))}\ifnum\pgfmathresult=1\pgfmathprintnumber\storeresultII~GFLOP/s\fi}}},draw=black,fill, area legend] table [x expr=\thisrow{N}>7?\thisrow{N}/2:\thisrow{N},y expr=\thisrow{baij-3-F}/\thisrow{baijmkl-3-T}/(\storeresultII*1000000000.0)]{\loadedMatProduct};\label{pgfplots:baijmkl3}
+          \addplot+[draw=black,fill, area legend] table [x expr=\thisrow{N}>7?\thisrow{N}/2:\thisrow{N},y expr=\thisrow{sbaij-3-F}/\thisrow{sbaijmkl-3-T}/(\storeresultII*1000000000.0)]{\loadedMatProduct};\label{pgfplots:sbaijmkl3}
+           \addplot+[draw=black,fill, area legend] table [x expr=\thisrow{N}>7?\thisrow{N}/2:\thisrow{N},y expr=\thisrow{aij-3-F}/\thisrow{cusparse-3-T}/(\storeresultII*1000000000.0)]{\loadedMatProduct};\label{pgfplots:cusparse3}
+           % \addlegendentry{\pk{MatAIJ}}
+           % \addlegendentry{\pk{MatBAIJ}}
+           % \addlegendentry{\pk{MatSBAIJ}}
+           % \addlegendentry{\pk{MatAIJMKL}}
+           % \addlegendentry{\pk{MatBAIJMKL}}
+           % \addlegendentry{MatSBAIJMKL}
+\node[rotate=0,fill=white,inner sep=0pt,anchor=west] at (rel axis cs:1.02, 0.5) {$b=3$};
+    \end{semilogxaxis}
+      \begin{semilogxaxis}[
+        ymin=0,ymax=5.5,
+        ybar=1.5pt,
+        bar width=3.5pt,
+        enlarge x limits=0.101,height=4cm,
+        width=0.8\textwidth,
+        xlabel={\# of right-hand sides ($N$)},
+        ylabel={},
+        xtick = {\empty},
+        xmajorgrids=true,
+        extra x ticks = {1,2,4,8,16,32},
+        extra x tick labels = {1,2,8,16,32,64},
+        extra tick style={
+            grid style={dotted,black}
+        },
+        xtick pos=left,name=plot3,at={($(plot2.south)-(0,0.2cm)$)},anchor=north,
+        xticklabels={1,2,8,16,32,64},
+        ytick = {1,2,ln(6.5-5)+3,ln(10-5)+3},%,ln(30-5)+3},
+        ylabel style={yshift=6pt},
+        ymajorgrids=true,yticklabel={
+\pgfmathparse{int(10*\tick)}
+\ifnum\pgfmathresult=0
+\else
+\ifnum\pgfmathresult>20
+\pgfmathparse{ceil((exp(\tick-3)+5))*100}
+$\pgfmathprintnumber{\pgfmathresult}$\%
+\else
+\pgfmathparse{100*\tick}
+$\pgfmathprintnumber{\pgfmathresult}$\%
+\fi
+\fi
+},
+        %xmajorgrids=true,
+          cycle list={blueaccent, darkgray, mediumgray,last},
+          point meta=exp(x),clip=false,
+      ]
+\node[rotate=0,fill=white,inner sep=0pt,anchor=west] at (rel axis cs:1.02, 0.5) {$b=6$};
+% \node[font=\tiny,rotate=45,fill=white,inner sep=0pt,anchor=west] at (axis cs:0.725, 1.2) {\pgfmathprintnumber\storeresultI~GFLOP/s};
+% \node[font=\tiny,rotate=45,fill=white,inner sep=0pt,anchor=west] at (axis cs:0.95, 1.2) {\pgfmathprintnumber\storeresultII~GFLOP/s};
+% \node[font=\tiny,rotate=45,fill=white,inner sep=0pt,anchor=west] at (axis cs:1.2, 1.2) {\pgfmathprintnumber\storeresultIII~GFLOP/s};
+          \addplot+[draw=black,fill, area legend] table [x expr=\thisrow{N}>7?\thisrow{N}/2:\thisrow{N},y expr=\thisrow{aij-10-F}/\thisrow{aijmkl-10-T}/(\storeresultIII*1000000000.0)]{\loadedMatProduct};\label{pgfplots:aijmkl10}
+          \addplot+[nodes near coords style={xshift=0.55cm,yshift=-0.1cm,above}, nodes near coords*={\rotatebox{45}{\tiny\color{black}{\pgfmathfloattofixed{\pgfplotspointmeta}\pgfmathparse{int(round(\pgfmathresult))}\ifnum\pgfmathresult=1\pgfmathprintnumber\storeresultIII~GFLOP/s\fi}}},draw=black,fill, area legend] table [x expr=\thisrow{N}>7?\thisrow{N}/2:\thisrow{N},y expr=\thisrow{baij-10-F}/\thisrow{baijmkl-10-T}/(\storeresultIII*1000000000.0)]{\loadedMatProduct};\label{pgfplots:baijmkl10}
+          \addplot+[draw=black,fill, area legend] table [x expr=\thisrow{N}>7?\thisrow{N}/2:\thisrow{N},y expr=\thisrow{sbaij-10-F}/\thisrow{sbaijmkl-10-T}/(\storeresultIII*1000000000.0)]{\loadedMatProduct};\label{pgfplots:sbaijmkl10}
+          %\addplot+[draw=black,fill, area legend] table [x expr=\thisrow{N}>7?\thisrow{N}/2:\thisrow{N},y expr=\thisrow{aij-10-F}/\thisrow{cusparse-10-T}/(\storeresultIII*1000000000.0)]{\loadedMatProduct};\label{pgfplots:cusparse10}
+          \addplot+[draw=black,fill, area legend] table [x expr=\thisrow{N}>7?\thisrow{N}/2:\thisrow{N},y expr=\thisrow{N}>1.5?ln(\thisrow{aij-10-F}/\thisrow{cusparse-10-T}/(\storeresultIII*1000000000.0)-5.0)+3:\thisrow{aij-10-F}/\thisrow{cusparse-10-T}/(\storeresultIII*1000000000.0)]{\loadedMatProduct};\label{pgfplots:cusparse10}
+          \begin{pgfonlayer}{foreground}
+\begin{scope}[decoration={snake,amplitude=1pt,pre length=0,post length=0}]
+          \path[shape=coordinate]
+(rel axis cs:-0.001,0.49) coordinate(b1) (rel axis cs: 1.001,0.49) coordinate(b2)
+(rel axis cs: 1.001,0.51) coordinate(b3) (rel axis cs:-0.001,0.51) coordinate(b4);
+  \path[fill=white] (b2) -- (b3) 
+                decorate {-- (b4)}  
+                -- (b1) 
+                decorate {-- (b2)};
+  \draw         (b3) 
+                decorate {-- (b4)};  
+  \draw         (b1) 
+                decorate {-- (b2)};
+\end{scope}
+          \end{pgfonlayer}{foreground}
+    \end{semilogxaxis}
+    \end{tikzpicture}
+   \caption{Performance of the matrix--matrix multiplication as implemented in PETSc for three sparse matrix types.
+As baseline values with $N=1$, for $b=1$, performance of \pk{MatAIJMKL} matrix--vector multiplication is used~\ref{pgfplots:aijmkl1}, while for $b=\{3,6\}$, it is the performance of \pk{MatBAIJMKL} matrix--vector multiplication~\ref{pgfplots:baijmkl3}.\label{fig:perfAXpar}}
+\end{figure}
 \section{PCHPDDM}\label{sec:PCHPDDM}
@@
 
-\bibitem{bastian2010generic}
-{\sc P.~Bastian, F.~Heimann, and S.~Marnach}, {\em {Generic implementation of
-  finite element methods in the distributed and unified numerics environment
-  (DUNE)}}, Kybernetika, 46 (2010), pp.~294--315,
-  \url{https://www.dune-project.org}.
-
 \bibitem{bavier2012amesos2}
@@
   Mathematics, 20 (2012), pp.~251--266, \url{http://freefem.org}.
-
-\bibitem{heinecke2016libxsmm}
-{\sc A.~Heinecke, G.~Henry, M.~Hutchinson, and H.~Pabst}, {\em {LIBXSMM:
-  accelerating small matrix multiplications by runtime code generation}}, in
-  Proceedings of the 2016 International Conference for High Performance
-  Computing, Networking, Storage and Analysis, SC16, IEEE, 2016,
-  \url{https://github.com/hfp/libxsmm}.
 

### 2020-06-19 MatProduct.c

@@ -1,0 +1,349 @@
+#include <petsc.h>
+
+#if defined(PETSC_HAVE_MKL)
+#include <mkl.h>
+#define PetscStackCallMKLSparse(func,args) do {                                                           \
+    sparse_status_t __ierr;                                                                               \
+    PetscStackPush(#func);                                                                                \
+    __ierr = func args;                                                                                   \
+    PetscStackPop;                                                                                        \
+    if (__ierr != SPARSE_STATUS_SUCCESS) SETERRQ2(PETSC_COMM_SELF,PETSC_ERR_LIB,"Error in %s(): error code %d",#func,(int)__ierr); \
+  } while (0)
+#else
+#define PetscStackCallMKLSparse(func,args) do {                                                           \
+    SETERRQ(PETSC_COMM_SELF,PETSC_ERR_SUP,"No MKL support"); \
+  } while (0)
+#endif
+
+int main(int argc, char** argv) {
+  Mat            A, C, D, Ct = NULL;
+  PetscInt       nbs = 10, ntype = 10, nN = 8, m, M, trial = 5;
+  PetscViewer    viewer;
+  PetscInt       bs[10], N[8];
+  char           *type[10];
+  PetscMPIInt    size;
+  PetscBool      flg, iscuda, check = PETSC_FALSE, trans = PETSC_FALSE, mkl;
+  char           file[PETSC_MAX_PATH_LEN];
+  PetscErrorCode ierr;
+
+  ierr = PetscInitialize(&argc, &argv, NULL, NULL);CHKERRQ(ierr);
+  if (ierr) return ierr;
+  ierr = MPI_Comm_size(PETSC_COMM_WORLD, &size);CHKERRQ(ierr);
[…]

### 2020-06-19 MatProduct.cpp

@@ -1,134 +1,0 @@
-#include <petsc.h>
-#include <iostream>
-#include <vector>
-
-int main(int argc, char** argv) {
-  Mat            A, B, C, D;
-  PetscInt       nbs = 10, ntype = 10, nN = 8, m, M, trial = 5;
-  PetscViewer    viewer;
-  PetscInt       *bs,*N;
-  char           **type;
-  PetscMPIInt    size;
-  PetscBool      flg;
-  char           file[PETSC_MAX_PATH_LEN];
-  PetscErrorCode ierr;
-  ierr = PetscInitialize(&argc, &argv, nullptr, nullptr);CHKERRQ(ierr);
-  if (ierr) return ierr;
-  ierr = MPI_Comm_size(PETSC_COMM_WORLD, &size);CHKERRQ(ierr);
-  if (size != 1) SETERRQ(PETSC_COMM_WORLD, PETSC_ERR_WRONG_MPI_SIZE, "This is a uniprocessor example only");
-  ierr = PetscOptionsGetString(nullptr, nullptr, "-f", file, PETSC_MAX_PATH_LEN, &flg);CHKERRQ(ierr);
-  if (!flg) SETERRQ(PETSC_COMM_WORLD, PETSC_ERR_USER_INPUT, "Must indicate binary file with the -f option");
-  ierr = PetscOptionsGetInt(nullptr, nullptr, "-trial", &trial, &flg);CHKERRQ(ierr);
-  bs = new PetscInt[nbs]();
-  ierr = PetscOptionsGetIntArray(nullptr, nullptr, "-bs", bs, &nbs, &flg);CHKERRQ(ierr);
-  if (!flg) {
-    nbs = 1;
-    *bs = 1;
-  }
-  N = new PetscInt[nN]();
-  ierr = PetscOptionsGetIntArray(nullptr, nullptr, "-N", N, &nN, &flg);CHKERRQ(ierr);
-  if (!flg) {
-    nN = 8;
-    N[0] = 1;  N[1] = 2;  N[2] = 4;  N[3] = 8;
-    N[4] = 16; N[5] = 32; N[6] = 64; N[7] = 128;
-  }
-  type = new char*[ntype]();
-  ierr = PetscOptionsGetStringArray(nullptr, nullptr, "-type", type, &ntype, &flg);CHKERRQ(ierr);
-  if (!flg) {
-    ntype = 1;
-    *type = new char[6];
[…]

### 2020-06-19 article.tex
Some more recent changes

@@
 e.g., in tomography~\cite{tournier2016micro} or data
-analytics~\cite{KALANTZIS2018136}. Of course, it is possible to cast a system
-with multiple right-hand sides into a sequence of systems with the same
-coefficient matrix but changing right-hand sides $A x_i = b_i, i =
-1,2,\ldots,k$. However, block Krylov methods, at the cost of higher arithmetic
+analytics~\cite{KALANTZIS2018136}.
+% Of course, it is possible to cast a system with multiple right-hand sides into a sequence of systems with the same coefficient matrix but changing right-hand sides $A x_i = b_i, i = 1,2,\ldots,k$.
+However,
+Block Krylov methods, while having higher arithmetic
 intensities, generate larger subspaces and typically converge in fewer iterates.
@@
 this paragraph is to explain how the results have been reproduced in PETSc.
-Thanks to \pk{KSPHPDDM}, it is quite straightforward. First, the sequence of
-MATLAB systems had to be converted to PETSc \pk{Mat} and \pk{Vec}. Then, it is
-just a matter of calling \pk{KSPSolve} in sequence. The dataset and the source
-code are available at
+% Thanks to \pk{KSPHPDDM}, it is quite straightforward.
+First, the sequence of MATLAB systems had to be converted to PETSc \pk{Mat} and \pk{Vec}.
+Then, it is just a matter of calling \pk{KSPSolve} in sequence.
+The dataset and the source code are available at
 \url{https://gitlab.com/petsc/datafiles/-/tree/master/matrices/hpddm/GCRODR}
@@
 \end{itemize}
-Except for the right-preconditioned GCRODR, these tests are the same as the
-ones from the original GCRODR paper~\cite{parks2006recycling}. Furthermore,
-\pk{PCICC} and ichol from MATLAB are not consistent. All PETSc tests are run on four MPI
+All PETSc tests are run on four MPI
 processes. They can be launched using the following command line. \\[-4pt]
@@
 The various number of iterations needed to reach the convergence criterion as
-originally set to $10^{-10}$ are gathered in~\cref{fig:comparison}. Apart from
+originally set to $10^{-10}$ are gathered in~\cref{fig:comparison}. 
+Except for the right-preconditioned GCRODR, these tests are the same as the
+ones from the original GCRODR paper~\cite{parks2006recycling}. Furthermore,
+\pk{PCICC} and ichol from MATLAB are not consistent.  Apart from
 the tests \ref{pgfplots:r2} and \ref{pgfplots:r3} using $ICC(0)$ and \ref{pgfplots:r4} and \ref{pgfplots:r5}
@@
 Though the numerical values of $A$ are of no interest here, the sparsity pattern is rather common and frequently encountered, e.g., when discretizing partial differential equations.
-Then, a random dense matrix $T$ of dimension $b$ is generated, for $b \in \{ 1,3,6 \}$, and the following matrix is evaluated $\mathbb{A} = A \otimes T$.
+Then, a random symmetric dense matrix $T$ of dimension $b$ is generated, for $b \in \{ 1,3,6 \}$, and the following matrix is evaluated $\mathbb{A} = A \otimes T$.
 The matrix $\mathbb{A}$ is then assembled into all three formats described above.
@@
 \end{align}
-The physical domain $\Omega$ is a beam of dimensions $[0,6] \times [0,1] \times [0,1]$. It is decomposed in $13{,}824$ subdomains using the automatic graph partitioner ParMETIS~\cite{karypis1998fast}.
+The physical domain $\Omega$ is a beam of dimensions $[0,6] \times [0,1] \times [0,1]$.
 The Cauchy stress tensor $\sigma$ is given by Hooke’s law: it can be expressed in terms of Young’s modulus $E$ and Poisson’s ratio $\nu$.
@@
 $(\mathbb{P}_1, \mathbb{P}_1, \mathbb{P}_1)$ finite elements.
+The physical domain $\Omega$ is decomposed in $13{,}824$ subdomains using the automatic graph partitioner ParMETIS~\cite{karypis1998fast}.
 The number of unknowns is $\pgfmathprintnumber{593} \times 10^6$, with approximately $\pgfmathprintnumber{45}$ nonzero coefficients per row.
@@
 By default, the coarse problem in \pk{PCHPDDM} is solved using an exact $LU$ or Cholesky factorization, in this case using 24 processes and MKL CPARDISO.
-Thanks to the flexibility of PETSc, switching to an inexact solver using \pk{PCHYPRE} (resp.~GenEO multilevel extension), cf.~\cref{enum:inexacthypre} (resp.~\cref{enum:inexact}), is straightforward.
+Switching to an inexact solver using \pk{PCHYPRE} (resp.~GenEO multilevel extension), cf.~\cref{enum:inexacthypre} (resp.~\cref{enum:inexact}), is straightforward.
 \\[-4pt]

### 2020-06-20 article.tex
Avoid copy to GPU in PetscLogStage

@@
 \newcommand\col{:}
-\newcommand{\pk}[1]{{\color{gray!90!black} #1}}
+\newcommand{\pk}[1]{\texttt{#1}}
 \usepackage{stmaryrd}
@@
     flexibility of the implementation %, which for example may deal with various matrix formats, 
-    are showcased and explained.  Diverse examples are
+    is showcased and explained.  Diverse examples are
     displayed, ranging from fracture mechanics to linear stability analysis, and
@@
 introduces \pk{PCHPDDM}. Eventually, concluding remarks may be found
-\cref{sec:conclusion}. All PETSc keywords are highlighted in \pk{gray}, these
+\cref{sec:conclusion}. All PETSc keywords and options are typeset in a \texttt{typewriter} front, these
 are public and documented at \url{https://www.mcs.anl.gov/petsc/petsc-master/docs}.
@@
 matrix--vector. Still, systems with multiple right-hand sides are ubiquitous,
-e.g., in tomography~\cite{tournier2016micro} or data
-analytics~\cite{KALANTZIS2018136}.
+e.g., in tomography~\cite{tournier2016micro}, data
+analytics~\cite{KALANTZIS2018136}, eigensolvers~\cite{knyazev2001toward}, geophysics~\cite{calandra2012flexible}, quantum chromodynamics~\cite{sakurai2010application}, etc.
 % Of course, it is possible to cast a system with multiple right-hand sides into a sequence of systems with the same coefficient matrix but changing right-hand sides $A x_i = b_i, i = 1,2,\ldots,k$.
@@
 registered since PETSc version 3.12 when configuring PETSc with the extra
-flag -{}-download-hpddm. It is then possible to select the corresponding \pk{KSP}
-using the command line option -ksp\_type hpddm, or the routine \pk{KSPSetType}(ksp,
+flag \pk{-{}-download-hpddm}. It is then possible to select the corresponding \pk{KSP}
+using the command line option \pk{-ksp\_type hpddm}, or the routine \pk{KSPSetType}(ksp,
 \pk{KSPHPDDM}).% and its Fortran equivalent.
@@
 number of global synchronizations, e.g., reductions for scalar products. But they are mathematically equivalent to
-their ``standard'' counterpart. These methods may be selected using the additional
-options -ksp\_hpddm\_type (gmres$|$cg$|$gcrodr$|$bgmres$|$bcg$|$bfbcg$|$bgcrodr$|$preonly) and
--ksp\_hpddm\_variant (left$|$right$|$flexible) to select whether preconditioning
+their ``standard'' counterpart. These methods may be selected using the 
+options \pk{-ksp\_hpddm\_type} \pk{(gmres$|$cg$|$gcrodr$|$bgmres$|$bcg$|$bfbcg$|$bgcrodr$|$preonly)} and
+\pk{-ksp\_hpddm\_variant} \pk{(left$|$right$|$flexible)} to select whether preconditioning
 is applied on the left, on the right, or variable. It is also possible to set
-this option through the more common PETSc option -ksp\_pc\_side (left$|$right). 
+this option through the more common PETSc option \pk{-ksp\_pc\_side (left$|$right)}. 
 Note that the value of the latter option is useless for BCG or BFBCG, and HPDDM CG only handles left and
@@
 tall-and-skinny dense matrix $Y$. In HPDDM, a specialized routine has been implemented
-to compute $X = P_BY$ if the preconditioner is of one of the following type:
+to compute $X = P_BY$ if the preconditioner is of one of the following types:
 \begin{itemize}
@@
     \item incomplete $LU$ or Cholesky factorization, \pk{PCILU} or \pk{PCICC};
-    \item block Jacobi with a single subdomain per process and exact or
-        incomplete subdomain solvers, \pk{PCBJACOBI};
-    \item overlapping Schwarz methods with a single subdomain per process and
+    \item block Jacobi with % a single subdomain per process and
+        exact or incomplete subdomain solvers, \pk{PCBJACOBI};
+    \item overlapping Schwarz methods with % a single subdomain per process and
         exact or incomplete subdomain solvers, \pk{PCASM}.
@@
 Indeed, in these cases, it is possible to get access to a factored matrix $F$
-and call \pk{MatMatSolve}($F$, $\tilde{Y}$, $\tilde{X}$). With the two first sets of
-preconditioners, $\tilde{Y} = Y$ and $\tilde{X} = X$. With block Jacobi,
-$\tilde{Y}$, resp. $\tilde{X}$, represents only the part of the matrix $Y$,
-resp. $X$, local to a given process. With overlapping Schwarz methods, there are
-furthermore overlapping ``ghost'' unknowns, interested readers are referred to
-monographs on practical aspects of domain decomposition
-methods~\cite{smith2004domain,dolean2014ddm}. For all other types of
-preconditioners, HPDDM will loop over the columns of $Y$ and apply $P_B$
-successively to get all columns of $X$.
+and call \pk{MatMatSolve}($F$, $\tilde{Y}$, $\tilde{X}$).
+% With the two first sets of preconditioners, $\tilde{Y} = Y$ and $\tilde{X} = X$.
+With block Jacobi, $\tilde{Y}$, resp.
+$\tilde{X}$, represents only the part of the matrix $Y$, resp.
+$X$, local to a given process.
+With overlapping Schwarz methods, there are furthermore overlapping ``ghost'' unknowns, interested readers are referred to monographs on practical aspects of domain decomposition methods~\cite{smith2004domain,dolean2014ddm}.
+For all other types of preconditioners, HPDDM will loop over the columns of $Y$
+and apply $P_B$ successively to get all columns of $X$.
 
@@
 example, in all Krylov methods using the Arnoldi process, it is mandatory to orthonormalize
-candidate basis vectors. It would be impractical to try to reuse PETSc routines
-like \pk{KSPGMRESModifiedGramSchmidtOrthogonalization} for two reasons: 1) it does
-not handle the orthonormalization of multiple vectors at once which is needed in
-pseudo-block methods 2) HPDDM may be used without PETSc as a stand-alone
-library. To give some more technical details to curious readers, in HPDDM, $QR$
-factorizations of tall-and-skinny dense matrices, which are obviously needed for
-block Krylov methods but also for recycling Krylov methods are computed using
+candidate basis vectors.
+% It would be impractical to try to reuse PETSc routines like \pk{KSPGMRESModifiedGramSchmidtOrthogonalization} for two reasons: 1) it does not handle the orthonormalization of multiple vectors at once which is needed in pseudo-block methods 2) HPDDM may be used without PETSc as a stand-alone library.
+To give some more technical details to curious readers, in HPDDM,
+$QR$ factorizations of tall-and-skinny dense matrices, which are needed by
+block Krylov methods and for recycling Krylov methods are computed using
 the CholQR algorithm~\cite{stathopoulos2002block}, or the block modified or classical
-Gram--Schmidt method. This may be adjusted with the option -ksp\_hpddm\_qr (cholqr$|$mgs$|$cgs).
+Gram--Schmidt method. This may be adjusted with the option \pk{-ksp\_hpddm\_qr (cholqr$|$mgs$|$cgs)}.
 For Hessenberg matrices generated by the Arnoldi
@@
 
-To monitor convergence of a \pk{KSP} object, calls to the PETSc \pk{KSPMonitor} routine
-were added inside HPDDM Krylov methods. 
+Monitoring convergence with the \pk{KSPMonitor} and \pk{KSPSetConvergenceTest} callbacks is fully supported.
+
     \subsubsection{User-defined deflation space} Recycling Krylov methods try to
@@
 solving small standard or generalized dense eigenproblems. Then, during a
-subsequent solve, this information is reused appropriately. Two new routines were
+subsequent solve, this information is reused appropriately~\cite{saad2000deflated,stathopoulos2002block,soodhalter2014krylov}. Two new routines were
 added in PETSc to either get the information generated by such recycling
@@
 the single right-hand side system is automatically cast into a system with
-multiple right-hand sides instead. The standard function \pk{KSPSolve} takes as
-input argument a \pk{KSP} and two \pk{Vec}. A new function
-\pk{KSPMatSolve} was added in PETSc 3.14.0, which takes as input argument
+multiple right-hand sides instead. % The standard function \pk{KSPSolve} takes as input argument a \pk{KSP} and two \pk{Vec}.
+A new function \pk{KSPMatSolve} was added in PETSc 3.14.0, which takes as input argument
 a \pk{KSP} and two dense \pk{Mat}, for solving true
@@
 \pk{KSPHPDDM}, the solution phase is done in a column by column fashion.
-Otherwise, it is possible to use any \pk{KSPHPDDM} (pseudo-)block Krylov methods.
+Otherwise, it is possible to use any \pk{KSPHPDDM} \mbox{(pseudo-)block} Krylov methods.
 Furthermore, a function \pk{KSPSetMatSolveBlockSize} is also provided to decompose
@@
 \begin{itemize}
-    \item in MATLAB, unpreconditioned GCRODR(40, 20), $ICC(0)$ left-preconditioned
+    \item in MATLAB, unpreconditioned GCRODR(40, 20), ICC(0) left-preconditioned
         GCRODR(40, 20), and Jacobi right-preconditioned GCRODR(40, 20);
@@
         legend columns=1,enlarge x limits=0.05,
-        legend style={font=\small,at={(1.02,0.25)}, legend cell align=left,
-        anchor=south west}]
+        legend style={font=\small,at={(1.02,0.5)},anchor=west, legend cell align=left}]
 \addplot[very thick,myred]                       table[x expr=\thisrowno{0},y expr=\thisrowno{1}, col sep=space] {\loadedtable};\label{pgfplots:r0}
@@
 \end{equation}
-In SLEPc, these systems may be parameterized using the -st\_ prefix, e.g.,
--st\_pc\_type~lu. If they are solved iteratively, it is possible to
+In SLEPc, these systems are parameterized using the \pk{-st\_} prefix, e.g.,
+\pk{-st\_pc\_type~lu}. If they are solved iteratively, it is possible to
 use a \pk{KSP} of type \pk{KSPHPDDM} to recycle information between
@@
 In terms of runtime, \SI[round-mode          = places,round-precision=1]{1.5098e+03}{\sec} (resp.~\SI[round-mode          = places,round-precision=1]{1.0524e+03}{\sec}) is spent in \pk{EPSSolve} when using \pk{KSPFGMRES} (resp.~\pk{KSPHPDDM}).
-These timings, and all which follow, have been obtained on {Ir\`ene}, a system composed of \pgfmathprintnumber[assume math mode=true]{1656} nodes with two 24-core Intel Xeon Platinum 8168 clocked at \SI{2.7}{\giga\hertz}.
+These timings, and all which follow except in~\cref{sec:primitives}, have been obtained on {Ir\`ene}, a system composed of \pgfmathprintnumber[assume math mode=true]{1656} nodes with two 24-core Intel Xeon Platinum 8168 clocked at \SI{2.7}{\giga\hertz}.
 \pgfplotstableread{
@@
 }\loadedtableSLEPc
-    \begin{figure}
+    \begin{figure}[!b]
 \centering \begin{tikzpicture}
@@
         iterations},xmin=1,xmax=46,ymin=0,ymajorgrids,
-        legend columns=1,enlarge x limits=0.05,
-        legend style={font=\small,at={(0.5,0.24)}, legend cell align=left,
-        anchor=south west}]
+        legend columns=1,enlarge x limits=0.05, height=5.5cm,
+        legend style={font=\small,at={(0.98,0.02)}, legend cell align=left,
+        anchor=south east}]
 \addplot[very thick,last]     table[x expr=\thisrowno{0},y expr=\thisrowno{2}, col sep=space] {\loadedtableSLEPc};\label{pgfplots:m0}
@@
 \end{tikzpicture}
-        \caption{\# of inner iterations for Krylov--Schur systems~\cref{eq:ks} with~\ref{pgfplots:m0} or without~\ref{pgfplots:m1} recycling.\label{fig:comparison-slepc}}
+        \caption{\# of inner iterations for systems~\cref{eq:ks} with~\ref{pgfplots:m0} or without~\ref{pgfplots:m1} recycling.\label{fig:comparison-slepc}}
     \end{figure}
@@
 N    baij-1-T    baij-1-F   sbaij-1-T   sbaij-1-F  aij-1-T     aij-1-F    baij-3-T    baij-3-F   sbaij-3-T   sbaij-3-F  aij-3-T     aij-3-F    baij-10-T   baij-10-F  sbaij-10-T  sbaij-10-F aij-10-T    aij-10-F     cusparse-1-T    cusparse-1-F    cusparse-3-T    cusparse-3-F    cusparse-10-T   cusparse-10-F    aijmkl-seq-1-T      aijmkl-seq-3-T      aijmkl-seq-10-T     baijmkl-seq-1-T      baijmkl-seq-3-T      baijmkl-seq-10-T   sbaijmkl-seq-1-T      sbaijmkl-seq-3-T      sbaijmkl-seq-10-T         aijmkl-1-T      aijmkl-3-T      aijmkl-10-T     baijmkl-1-T      baijmkl-3-T      baijmkl-10-T   sbaijmkl-1-T      sbaijmkl-3-T      sbaijmkl-10-T
-1      2.0126e-01   1.4696e+08 8.0098e-02   1.4696e+08   9.3388e-02   1.4696e+08     8.9584e-01   1.3535e+09    6.2428e-01   1.3638e+09     1.2072e+00   1.3535e+09   4.0734e+00   5.4450e+09    2.2476e+00   5.4708e+09    5.0428e+00   5.4450e+09   4.4800e-02   1.5211e+08   3.7933e-01   1.3690e+09   1.5043e+00   5.4759e+09   6.6857e-02  7.4649e-01   3.1128e+00   4.7672e-01   6.2903e-01  2.2271e+00   4.7675e-01   6.2990e-01    2.2273e+00    7.8399e-03    8.6487e-02    3.5536e-01   5.2814e-02 6.6225e-02   2.4584e-01   4.5115e-02   6.6404e-02   2.4489e-01
+1      2.0126e-01   1.4696e+08 8.0098e-02   1.4696e+08   9.3388e-02   1.4696e+08     8.9584e-01   1.3535e+09    6.2428e-01   1.3638e+09     1.2072e+00   1.3535e+09   4.0734e+00   5.4450e+09    2.2476e+00   5.4708e+09    5.0428e+00   5.4450e+09   2.1248e-03   1.5211e+08   1.3669e-02   1.3690e+09   4.7833e-02   5.4759e+09   6.6857e-02  7.4649e-01   3.1128e+00   4.7672e-01   6.2903e-01  2.2271e+00   4.7675e-01   6.2990e-01    2.2273e+00    7.8399e-03    8.6487e-02    3.5536e-01   5.2814e-02 6.6225e-02   2.4584e-01   4.5115e-02   6.6404e-02   2.4489e-01
 2      3.2635e-01   2.9392e+08 2.4204e-01   3.2482e+08   1.1522e-01   3.0422e+08     1.2835e+00   2.7071e+09    1.0697e+00   2.9234e+09     9.9480e-01   2.7380e+09   1.4504e+01   1.0890e+10    1.5265e+01   1.1694e+10    3.7799e+00   1.0952e+10   5.2164e-03   3.0422e+08   3.4056e-02   2.7380e+09   7.7659e-02   1.0952e+10   1.1703e-01  9.3872e-01   3.5976e+00   4.7905e-01   7.3497e-01  3.3331e+00   4.7886e-01   7.3415e-01    3.3297e+00    1.2450e-02    1.0124e-01    3.9873e-01   4.7633e-02 7.4107e-02   2.8513e-01   4.7733e-02   7.4693e-02   2.8507e-01
@@
 }\loadedMatProduct
-    \subsubsection{Performance of primitives for block Krylov methods}
+    \subsubsection{Performance of primitives for block Krylov methods}\label{sec:primitives}
 As detailed in~\cref{sec:interface-ksp}, it is currently not possible with most
@@
 For example, \pk{MatMPIAIJ}, standard parallel sparse matrices, store on each rank two \pk{MatSeqAIJ}, one with local rows and columns, another with local rows and ``nonlocal'' columns.
-To keep this study succinct, only sequential sparse matrices will be evaluated next, which can help draw conclusions for the intraprocess performance of parallel sparse matrices.
+To keep this study succinct, only nondistributed sparse matrices will be evaluated next, which can help draw conclusions for the intraprocess performance of parallel sparse matrices.
 The Seq substring is now dropped, since in this case, \pk{MatSeqAIJ} is equivalent to \pk{MatAIJ} and so on.
-
-The Poisson equation on a cube is discretized with $\mathbb{P}_1$ finite elements in double precision.
+The performance of the multiplication primitives of the three aforementioned types will also be compared to MKL inspector--executor sparse BLAS routines~\cite{mkl-web-page}, using a single OpenMP thread.
+There is an ongoing effort to better integrate these routines in PETSc respective types \pk{MatAIJMKL} and \pk{MatBAIJMKL}, but at the time of writing, direct calls to the MKL were used.
+
+The Poisson equation on a cube is discretized with $\mathbb{P}_1$ finite elements in double precision with 32-bit integers.
 This yields a matrix $A$ of dimension one million.
@@
 All results are gathered in~\cref{fig:perfAX}.
-They can be reproduced using the following mini-app: \url{http://jolivet.perso.enseeiht.fr/MatProduct.cpp} and any input \pk{MatAIJ} stored in binary format, here using the default name binaryoutput.
+They can be reproduced using the following mini-app: \url{http://jolivet.perso.enseeiht.fr/MatProduct.c} and any input \pk{MatAIJ} stored in binary format, here using the default name \pk{binaryoutput}.
 \begin{Verbatim}[fontsize=\footnotesize,frame=single,framerule=0.1mm,commandchars=&\[\]]
@@
   \pgfmathparse{1.4696e+08/9.3388e-02/1000000000.0}\edef\storeresultI{\pgfmathresult}
-  \pgfmathparse{1.3535e+09/8.9584e-01/1000000000.0}\edef\storeresultII{\pgfmathresult}
-  \pgfmathparse{5.4450e+09/4.0734e+00/1000000000.0}\edef\storeresultIII{\pgfmathresult}
+  % \pgfmathparse{1.3535e+09/8.9584e-01/1000000000.0}\edef\storeresultII{\pgfmathresult}
+  % \pgfmathparse{5.4450e+09/4.0734e+00/1000000000.0}\edef\storeresultIII{\pgfmathresult}
+  \pgfmathparse{1.3535e+09/1.2072e+00/1000000000.0}\edef\storeresultII{\pgfmathresult}
+  \pgfmathparse{5.4450e+09/5.0428e+00/1000000000.0}\edef\storeresultIII{\pgfmathresult}
 \pgfkeys{/pgf/fpu=false}
-  % \hspace*{-1.5cm}
   \centering
@@
         xticklabels={\empty},xmajorticks=false,
-        ytick = {1,2,3,4},
+        ytick = {1,3.5,6},
         ymajorgrids=true,yticklabel={
@@
       ]
-          \addplot+[draw=black,fill, area legend,postaction={pattern=crosshatch dots}] table [x expr=\thisrow{N}>7?\thisrow{N}/2:\thisrow{N},y expr=\thisrow{aij-3-F}/\thisrow{aij-3-T}/(\storeresultII*1000000000.0)]{\loadedMatProduct};\label{pgfplots:aij3}
-          \addplot+[nodes near coords style={xshift=0.5cm,yshift=-0.1cm,above}, nodes near coords*={\rotatebox{45}{\tiny\color{black}{\pgfmathfloattofixed{\pgfplotspointmeta}\pgfmathparse{int(round(\pgfmathresult))}\ifnum\pgfmathresult=1\pgfmathprintnumber\storeresultII~GFLOP/s\fi}}},
+          \addplot+[nodes near coords style={xshift=0.5cm,yshift=-0.1cm,above}, nodes near coords*={\rotatebox{45}{\tiny\color{black}{\pgfmathfloattofixed{\pgfplotspointmeta}\pgfmathparse{int(round(\pgfmathresult))}\ifnum\pgfmathresult=1\pgfmathprintnumber\storeresultII~GFLOP/s\fi}}},draw=black,fill, area legend,postaction={pattern=crosshatch dots}] table [x expr=\thisrow{N}>7?\thisrow{N}/2:\thisrow{N},y expr=\thisrow{aij-3-F}/\thisrow{aij-3-T}/(\storeresultII*1000000000.0)]{\loadedMatProduct};\label{pgfplots:aij3}
+          \addplot+[
               draw=black,fill, area legend,postaction={pattern=crosshatch dots}] table [x expr=\thisrow{N}>7?\thisrow{N}/2:\thisrow{N},y expr=\thisrow{baij-3-F}/\thisrow{baij-3-T}/(\storeresultII*1000000000.0)]{\loadedMatProduct};\label{pgfplots:baij3}
@@
            \addplot+[draw=black,fill, area legend,postaction={pattern=bricks}] table [x expr=\thisrow{N}>7?\thisrow{N}/2:\thisrow{N},y expr=\thisrow{sbaij-3-F}/\thisrow{sbaijmkl-seq-3-T}/(\storeresultII*1000000000.0)]{\loadedMatProduct};\label{pgfplots:mklsbaijseq3}
-           % \addlegendentry{\pk{MatAIJ}}
-           % \addlegendentry{\pk{MatBAIJ}}
-           % \addlegendentry{\pk{MatSBAIJ}}
-           % \addlegendentry{\pk{MatAIJMKL}}
-           % \addlegendentry{\pk{MatBAIJMKL}}
-           % \addlegendentry{MatSBAIJMKL}
 \node[rotate=0,fill=white,inner sep=0pt,anchor=west] at (rel axis cs:1.02, 0.5) {$b=3$};
@@
         xticklabels={1,2,8,16,32,64},
-        ytick = {1,2,3,4},
+        ytick = {1,3,5},
         ymajorgrids=true,yticklabel={
@@
 },
-        %xmajorgrids=true,
           cycle list={blueaccent, darkgray, mediumgray},%,last},
@@
 \node[rotate=0,fill=white,inner sep=0pt,anchor=west] at (rel axis cs:1.02, 0.5) {$b=6$};
-          \addplot+[draw=black,fill, area legend,postaction={pattern=crosshatch dots}] table [x expr=\thisrow{N}>7?\thisrow{N}/2:\thisrow{N},y expr=\thisrow{aij-10-F}/\thisrow{aij-10-T}/(\storeresultIII*1000000000.0)]{\loadedMatProduct};\label{pgfplots:aij10}
-          \addplot+[nodes near coords style={xshift=0.5cm,yshift=-0.1cm,above}, nodes near coords*={\rotatebox{45}{\tiny\color{black}{\pgfmathfloattofixed{\pgfplotspointmeta}\pgfmathparse{int(round(\pgfmathresult))}\ifnum\pgfmathresult=1\contour{white}{\pgfmathprintnumber\storeresultIII}~GFLOP/s\fi}}},draw=black,fill, area legend,postaction={pattern=crosshatch dots}] table [x expr=\thisrow{N}>7?\thisrow{N}/2:\thisrow{N},y expr=\thisrow{baij-10-F}/\thisrow{baij-10-T}/(\storeresultIII*1000000000.0)]{\loadedMatProduct};\label{pgfplots:baij10}
+          \addplot+[nodes near coords style={xshift=0.5cm,yshift=-0.1cm,above}, nodes near coords*={\begin{pgfonlayer}{main}\rotatebox{45}{\tiny\color{black}{\pgfmathfloattofixed{\pgfplotspointmeta}\pgfmathparse{int(round(\pgfmathresult))}\ifnum\pgfmathresult=1\contour{white}{\pgfmathprintnumber\storeresultIII}~GFLOP/s\fi}}\end{pgfonlayer}},draw=black,fill, area legend,postaction={pattern=crosshatch dots}] table [x expr=\thisrow{N}>7?\thisrow{N}/2:\thisrow{N},y expr=\thisrow{aij-10-F}/\thisrow{aij-10-T}/(\storeresultIII*1000000000.0)]{\loadedMatProduct};\label{pgfplots:aij10}
+          \addplot+[draw=black,fill, area legend,postaction={pattern=crosshatch dots}] table [x expr=\thisrow{N}>7?\thisrow{N}/2:\thisrow{N},y expr=\thisrow{baij-10-F}/\thisrow{baij-10-T}/(\storeresultIII*1000000000.0)]{\loadedMatProduct};\label{pgfplots:baij10}
+\begin{pgfonlayer}{foreground}
           \addplot+[draw=black,fill, area legend,postaction={pattern=crosshatch dots}] table [x expr=\thisrow{N}>7?\thisrow{N}/2:\thisrow{N},y expr=\thisrow{sbaij-10-F}/\thisrow{sbaij-10-T}/(\storeresultIII*1000000000.0)]{\loadedMatProduct};\label{pgfplots:sbaij10}
+\end{pgfonlayer}{foreground}
            \addplot+[draw=black,fill, area legend,postaction={pattern=bricks}] table [x expr=\thisrow{N}>7?\thisrow{N}/2:\thisrow{N},y expr=\thisrow{aij-10-F}/\thisrow{aijmkl-seq-10-T}/(\storeresultIII*1000000000.0)]{\loadedMatProduct};\label{pgfplots:mklaijseq10}
@@
    \caption{Performance of the matrix--matrix multiplication as implemented in PETSc for three sparse matrix types.
-As baseline values with $N=1$, for $b=1$, performance of \pk{MatAIJ} matrix--vector multiplication is used~\ref{pgfplots:aij1}, while for $b=\{3,6\}$, it is the performance of \pk{MatBAIJ} matrix--vector multiplication~\ref{pgfplots:baij3}.\label{fig:perfAX}}
+As baseline values with $N=1$, performance of \pk{MatAIJ} matrix--vector multiplication is used~\ref{pgfplots:aij1}.
+Results are averages of 5 consecutives multiplications.\label{fig:perfAX}}
 \end{figure}
-
-\begin{figure}[htbp]
+Some conclusions may be drawn:
+\begin{itemize}
+    \item with a block size of 1, top plot $b=1$ in~\cref{fig:perfAX}, using any type but \pk{MatAIJ} is counterproductive, see~\ref{pgfplots:baij1} and~\ref{pgfplots:sbaij1}, even with the MKL, see~\ref{pgfplots:mklbaijseq1} and~\ref{pgfplots:mklsbaijseq1}.
+However, with this format, as soon as the number of columns is greater or equal than 8, the efficiency reaches approximately 200\%, see~\ref{pgfplots:aij1}; PETSc implementation seems slightly better than MKL except for the single-column case;
+    \item for block sizes lower or equal than 5, loops for matrix--vector and matrix--matrix multiplications are unrolled by hand for block formats.
+This yields rather disappointing performance, except for \pk{MatAIJ}, see~\ref{pgfplots:aij3} in middle plot $b=3$, and its 350\% efficiency, even with a moderate number of columns.
+For block formats \ref{pgfplots:baij3} and~\ref{pgfplots:sbaij3}, given the small value of $b$, it could be beneficial to switch to optimized libraries for small matrices, e.g., LIBXSMM~\cite{heinecke2016libxsmm}; in fact, MKL implementation for block formats is here clearly outperforming PETSc, see~\ref{pgfplots:mklbaijseq3} and~\ref{pgfplots:mklsbaijseq3};
+    \item for greater block sizes, \texttt{?gemv} or \texttt{?gemm} are used instead for block formats in PETSc.
+For \pk{MatAIJ}~\ref{pgfplots:aij10} in bottom plot $b=10$, the efficiency quickly caps near 400\% for $N\in\{8,16,32,64\}$, but for block formats, this reaches 500\% for large number of columns, see~\ref{pgfplots:baij10} and~\ref{pgfplots:sbaij10} for $N=64$.
+With these numbers of columns, PETSc and MKL perform similarly.
+\end{itemize}
+% These results are of course specific to the sparsity pattern of the initial matrix $A$, and thus $\mathbb{A} = A\otimes T$, and matrix types.
+For example, if the matrices were to be offloaded to a GPU, one could instead benchmark \pk{MatAIJcuSPARSE}.
+The mini-app may also be used to benchmark intranode performance of matrix--matrix multiplications, either with multiple OpenMP threads, or using CUDA.
+\begin{Verbatim}[fontsize=\footnotesize,frame=single,framerule=0.1mm,commandchars=&\[\]]
+&fvtextcolor[mygreen][$] &fvtextcolor[myyellow][export] OMP_NUM_THREADS=20 &&&& &fvtextcolor[myyellow][export] MKL_NUM_THREADS=20 
+&fvtextcolor[mygreen][$] &fvtextcolor[myyellow][mpirun] -n 1 ./MatProduct -f binaryoutput -log_view                                 &fvtextcolor[myred][\]
+  -bs 1,3,6 -N 1,2,8,16,32,64 -type aijmkl,baijmkl,sbaijmkl,aijcusparse
+\end{Verbatim}
+Results are reported in~\cref{fig:perfAXpar}.
+They have been obtained on a single node of {Jean Zay}, a system composed of \pgfmathprintnumber[assume math mode=true]{261} nodes with two 20-core Intel Xeon Gold 6248 clocked at \SI{2.5}{\giga\hertz} and four NVIDIA Tesla V100 SXM2.
+Neither transfers between host and device nor time spent in \texttt{mkl\_sparse\_optimize} are taken into account.
+Once again, it is clear that the larger the number of right-hand sides $N$, the better the efficiency, no matter the block size.
+\begin{figure}[tbp]
 \pgfkeys{/pgf/fpu}
   \pgfmathparse{1.4696e+08/7.8399e-03/1000000000.0}\edef\storeresultI{\pgfmathresult}
-  \pgfmathparse{1.3535e+09/6.6225e-02/1000000000.0}\edef\storeresultII{\pgfmathresult}
-  \pgfmathparse{5.4450e+09/2.4584e-01/1000000000.0}\edef\storeresultIII{\pgfmathresult}
+  % \pgfmathparse{1.3535e+09/6.6225e-02/1000000000.0}\edef\storeresultII{\pgfmathresult}
+  % \pgfmathparse{5.4450e+09/2.4584e-01/1000000000.0}\edef\storeresultIII{\pgfmathresult}
+  \pgfmathparse{1.3535e+09/8.6487e-02/1000000000.0}\edef\storeresultII{\pgfmathresult}
+  \pgfmathparse{5.4450e+09/3.5536e-01/1000000000.0}\edef\storeresultIII{\pgfmathresult}
 \pgfkeys{/pgf/fpu=false}
@@
 },
-        %xmajorgrids=true,
           cycle list={blueaccent, darkgray, mediumgray,last},
@@
       ]
-% \node[font=\tiny,rotate=45,fill=white,inner sep=0pt,anchor=west] at (axis cs:0.725, 1.2) {\pgfmathprintnumber\storeresultI~GFLOP/s};
-% \node[font=\tiny,rotate=45,fill=white,inner sep=0pt,anchor=west] at (axis cs:0.95, 1.2) {\pgfmathprintnumber\storeresultII~GFLOP/s};
-% \node[font=\tiny,rotate=45,fill=white,inner sep=0pt,anchor=west] at (axis cs:1.2, 1.2) {\pgfmathprintnumber\storeresultIII~GFLOP/s};
           \addplot+[nodes near coords style={xshift=0.55cm,yshift=-0.1cm,above}, nodes near coords*={\rotatebox{45}{\tiny\color{black}{\pgfmathfloattofixed{\pgfplotspointmeta}\pgfmathparse{int(round(\pgfmathresult))}\ifnum\pgfmathresult=1\pgfmathprintnumber\storeresultI~GFLOP/s\fi}}},draw=black,fill, area legend] table [x expr=\thisrow{N}>7?\thisrow{N}/2:\thisrow{N},y expr=\thisrow{aij-1-F}/\thisrow{aijmkl-1-T}/(\storeresultI*1000000000.0)]{\loadedMatProduct};\label{pgfplots:aijmkl1}
@@
           \addplot+[draw=black,fill, area legend] table [x expr=\thisrow{N}>7?\thisrow{N}/2:\thisrow{N},y expr=\thisrow{sbaij-1-F}/\thisrow{sbaijmkl-1-T}/(\storeresultI*1000000000.0)]{\loadedMatProduct};\label{pgfplots:sbaijmkl1}
+\begin{pgfonlayer}{foreground}
            \addplot+[draw=black,fill, area legend] table [x expr=\thisrow{N}>7?\thisrow{N}/2:\thisrow{N},y expr=\thisrow{aij-1-F}/\thisrow{cusparse-1-T}/(\storeresultI*1000000000.0)]{\loadedMatProduct};\label{pgfplots:cusparse1}
+\end{pgfonlayer}{foreground}
            \addlegendentry{\pk{MatAIJMKL}}
@@
         xticklabels={\empty},xmajorticks=false,
-        ytick = {1,4.5,8},
+        ytick = {1,5,9},
         ymajorgrids=true,yticklabel={
@@
       ]
-% \node[font=\tiny,rotate=45,fill=white,inner sep=0pt,anchor=west] at (axis cs:0.725, 1.2) {\pgfmathprintnumber\storeresultI~GFLOP/s};
-% \node[font=\tiny,rotate=45,fill=white,inner sep=0pt,anchor=west] at (axis cs:0.95, 1.2) {\pgfmathprintnumber\storeresultII~GFLOP/s};
-% \node[font=\tiny,rotate=45,fill=white,inner sep=0pt,anchor=west] at (axis cs:1.2, 1.2) {\pgfmathprintnumber\storeresultIII~GFLOP/s};
-          \addplot+[draw=black,fill, area legend] table [x expr=\thisrow{N}>7?\thisrow{N}/2:\thisrow{N},y expr=\thisrow{aij-3-F}/\thisrow{aijmkl-3-T}/(\storeresultII*1000000000.0)]{\loadedMatProduct};\label{pgfplots:aijmkl3}
-          \addplot+[nodes near coords style={xshift=0.55cm,yshift=-0.1cm,above}, nodes near coords*={\rotatebox{45}{\tiny\color{black}{\pgfmathfloattofixed{\pgfplotspointmeta}\pgfmathparse{int(round(\pgfmathresult))}\ifnum\pgfmathresult=1\pgfmathprintnumber\storeresultII~GFLOP/s\fi}}},draw=black,fill, area legend] table [x expr=\thisrow{N}>7?\thisrow{N}/2:\thisrow{N},y expr=\thisrow{baij-3-F}/\thisrow{baijmkl-3-T}/(\storeresultII*1000000000.0)]{\loadedMatProduct};\label{pgfplots:baijmkl3}
+          \addplot+[nodes near coords style={xshift=0.55cm,yshift=-0.1cm,above}, nodes near coords*={\rotatebox{45}{\tiny\color{black}{\pgfmathfloattofixed{\pgfplotspointmeta}\pgfmathparse{int(round(\pgfmathresult))}\ifnum\pgfmathresult=1\pgfmathprintnumber\storeresultII~GFLOP/s\fi}}},draw=black,fill, area legend] table [x expr=\thisrow{N}>7?\thisrow{N}/2:\thisrow{N},y expr=\thisrow{aij-3-F}/\thisrow{aijmkl-3-T}/(\storeresultII*1000000000.0)]{\loadedMatProduct};\label{pgfplots:aijmkl3}
+          \addplot+[draw=black,fill, area legend] table [x expr=\thisrow{N}>7?\thisrow{N}/2:\thisrow{N},y expr=\thisrow{baij-3-F}/\thisrow{baijmkl-3-T}/(\storeresultII*1000000000.0)]{\loadedMatProduct};\label{pgfplots:baijmkl3}
           \addplot+[draw=black,fill, area legend] table [x expr=\thisrow{N}>7?\thisrow{N}/2:\thisrow{N},y expr=\thisrow{sbaij-3-F}/\thisrow{sbaijmkl-3-T}/(\storeresultII*1000000000.0)]{\loadedMatProduct};\label{pgfplots:sbaijmkl3}
+\begin{pgfonlayer}{foreground}
            \addplot+[draw=black,fill, area legend] table [x expr=\thisrow{N}>7?\thisrow{N}/2:\thisrow{N},y expr=\thisrow{aij-3-F}/\thisrow{cusparse-3-T}/(\storeresultII*1000000000.0)]{\loadedMatProduct};\label{pgfplots:cusparse3}
-           % \addlegendentry{\pk{MatAIJ}}
-           % \addlegendentry{\pk{MatBAIJ}}
-           % \addlegendentry{\pk{MatSBAIJ}}
-           % \addlegendentry{\pk{MatAIJMKL}}
-           % \addlegendentry{\pk{MatBAIJMKL}}
-           % \addlegendentry{MatSBAIJMKL}
+\end{pgfonlayer}{foreground}
 \node[rotate=0,fill=white,inner sep=0pt,anchor=west] at (rel axis cs:1.02, 0.5) {$b=3$};
@@
       \begin{semilogxaxis}[
-        ymin=0,ymax=5.5,
+        ymin=0,
         ybar=1.5pt,
@@
         xtick pos=left,name=plot3,at={($(plot2.south)-(0,0.2cm)$)},anchor=north,
-        xticklabels={1,2,8,16,32,64},
-        ytick = {1,2,ln(6.5-5)+3,ln(10-5)+3},%,ln(30-5)+3},
+        xticklabels={1,2,8,16,32,64},ymax=7,
+        ytick = {1,3,ln(7-5)+4,ln(20-5)+4},%,ln(30-5)+3},ymax=5.5,
         ylabel style={yshift=6pt},
@@
 \else
-\ifnum\pgfmathresult>20
-\pgfmathparse{ceil((exp(\tick-3)+5))*100}
+\ifnum\pgfmathresult>30
+\pgfmathparse{ceil((exp(\tick-4)+5))*100}
 $\pgfmathprintnumber{\pgfmathresult}$\%
@@
 \node[rotate=0,fill=white,inner sep=0pt,anchor=west] at (rel axis cs:1.02, 0.5) {$b=6$};
-% \node[font=\tiny,rotate=45,fill=white,inner sep=0pt,anchor=west] at (axis cs:0.725, 1.2) {\pgfmathprintnumber\storeresultI~GFLOP/s};
-% \node[font=\tiny,rotate=45,fill=white,inner sep=0pt,anchor=west] at (axis cs:0.95, 1.2) {\pgfmathprintnumber\storeresultII~GFLOP/s};
-% \node[font=\tiny,rotate=45,fill=white,inner sep=0pt,anchor=west] at (axis cs:1.2, 1.2) {\pgfmathprintnumber\storeresultIII~GFLOP/s};
-          \addplot+[draw=black,fill, area legend] table [x expr=\thisrow{N}>7?\thisrow{N}/2:\thisrow{N},y expr=\thisrow{aij-10-F}/\thisrow{aijmkl-10-T}/(\storeresultIII*1000000000.0)]{\loadedMatProduct};\label{pgfplots:aijmkl10}
-          \addplot+[nodes near coords style={xshift=0.55cm,yshift=-0.1cm,above}, nodes near coords*={\rotatebox{45}{\tiny\color{black}{\pgfmathfloattofixed{\pgfplotspointmeta}\pgfmathparse{int(round(\pgfmathresult))}\ifnum\pgfmathresult=1\pgfmathprintnumber\storeresultIII~GFLOP/s\fi}}},draw=black,fill, area legend] table [x expr=\thisrow{N}>7?\thisrow{N}/2:\thisrow{N},y expr=\thisrow{baij-10-F}/\thisrow{baijmkl-10-T}/(\storeresultIII*1000000000.0)]{\loadedMatProduct};\label{pgfplots:baijmkl10}
+          \addplot+[nodes near coords style={xshift=0.55cm,yshift=-0.1cm,above}, nodes near coords*={\rotatebox{45}{\tiny\color{black}{\pgfmathfloattofixed{\pgfplotspointmeta}\pgfmathparse{int(round(\pgfmathresult))}\ifnum\pgfmathresult=1\pgfmathprintnumber\storeresultIII~GFLOP/s\fi}}},draw=black,fill, area legend] table [x expr=\thisrow{N}>7?\thisrow{N}/2:\thisrow{N},y expr=\thisrow{aij-10-F}/\thisrow{aijmkl-10-T}/(\storeresultIII*1000000000.0)]{\loadedMatProduct};\label{pgfplots:aijmkl10}
+          \addplot+[draw=black,fill, area legend] table [x expr=\thisrow{N}>7?\thisrow{N}/2:\thisrow{N},y expr=\thisrow{baij-10-F}/\thisrow{baijmkl-10-T}/(\storeresultIII*1000000000.0)]{\loadedMatProduct};\label{pgfplots:baijmkl10}
           \addplot+[draw=black,fill, area legend] table [x expr=\thisrow{N}>7?\thisrow{N}/2:\thisrow{N},y expr=\thisrow{sbaij-10-F}/\thisrow{sbaijmkl-10-T}/(\storeresultIII*1000000000.0)]{\loadedMatProduct};\label{pgfplots:sbaijmkl10}
-          %\addplot+[draw=black,fill, area legend] table [x expr=\thisrow{N}>7?\thisrow{N}/2:\thisrow{N},y expr=\thisrow{aij-10-F}/\thisrow{cusparse-10-T}/(\storeresultIII*1000000000.0)]{\loadedMatProduct};\label{pgfplots:cusparse10}
-          \addplot+[draw=black,fill, area legend] table [x expr=\thisrow{N}>7?\thisrow{N}/2:\thisrow{N},y expr=\thisrow{N}>1.5?ln(\thisrow{aij-10-F}/\thisrow{cusparse-10-T}/(\storeresultIII*1000000000.0)-5.0)+3:\thisrow{aij-10-F}/\thisrow{cusparse-10-T}/(\storeresultIII*1000000000.0)]{\loadedMatProduct};\label{pgfplots:cusparse10}
+          % \addplot+[draw=black,fill, area legend] table [x expr=\thisrow{N}>7?\thisrow{N}/2:\thisrow{N},y expr=\thisrow{aij-10-F}/\thisrow{cusparse-10-T}/(\storeresultIII*1000000000.0)]{\loadedMatProduct};\label{pgfplots:cusparse10}
+\begin{pgfonlayer}{foreground}
+          \addplot+[draw=black,fill, area legend] table [x expr=\thisrow{N}>7?\thisrow{N}/2:\thisrow{N},y expr=ln(\thisrow{aij-10-F}/\thisrow{cusparse-10-T}/(\storeresultIII*1000000000.0)-5.0)+4]{\loadedMatProduct};\label{pgfplots:cusparse10}
+\end{pgfonlayer}{foreground}
           \begin{pgfonlayer}{foreground}
@@
           \path[shape=coordinate]
-(rel axis cs:-0.001,0.49) coordinate(b1) (rel axis cs: 1.001,0.49) coordinate(b2)
-(rel axis cs: 1.001,0.51) coordinate(b3) (rel axis cs:-0.001,0.51) coordinate(b4);
+(rel axis cs:-0.001,0.54) coordinate(b1) (rel axis cs: 1.001,0.54) coordinate(b2)
+(rel axis cs: 1.001,0.56) coordinate(b3) (rel axis cs:-0.001,0.56) coordinate(b4);
   \path[fill=white] (b2) -- (b3) 
@@
     \end{tikzpicture}
-   \caption{Performance of the matrix--matrix multiplication as implemented in PETSc for three sparse matrix types.
-As baseline values with $N=1$, for $b=1$, performance of \pk{MatAIJMKL} matrix--vector multiplication is used~\ref{pgfplots:aijmkl1}, while for $b=\{3,6\}$, it is the performance of \pk{MatBAIJMKL} matrix--vector multiplication~\ref{pgfplots:baijmkl3}.\label{fig:perfAXpar}}
+   \caption{Performance of the matrix--matrix multiplication as implemented in PETSc using intranode parallelism.
+As baseline values with $N=1$, performance of \pk{MatAIJMKL} matrix--vector multiplication is used~\ref{pgfplots:aijmkl1}.
+Results are averages of 5 consecutives multiplications.\label{fig:perfAXpar}}
 \end{figure}
@@
 modern large-scale applications dealing with partial differential equations.
-In PETSc, preconditioners inherit from the \pk{PC} base class.
+% In PETSc, preconditioners inherit from the \pk{PC} base class.
 Historically, the block Jacobi and overlapping Schwarz methods were implemented
@@
 
-In PETSc, there is also an implementation of two-level overlapping Schwarz
-methods using a generalized Dryja, Smith, Widlund
-preconditioner~\cite{dohrmann2008domain}. Its type is \pk{PCEXOTIC}. When it comes to
-nonoverlapping preconditioners, \pk{PCNN} implements a basic balancing
+% In PETSc, there is also an implementation of two-level overlapping Schwarz
+% methods using a generalized Dryja, Smith, Widlund
+% preconditioner~\cite{dohrmann2008domain}. Its type is \pk{PCEXOTIC}.
+When it comes to nonoverlapping preconditioners, \pk{PCNN} implements a basic balancing
 Neumann--Neumann method~\cite{mandel1993balancing}. Much more advanced
@@
 registered since PETSc version 3.12 when one configures PETSc with the extra
-flag -{}-download-hpddm -{}-download-slepc. It is then
+flag \pk{-{}-download-hpddm -{}-download-slepc}. It is then
 possible to select the corresponding \pk{PC}
-using the command line option -pc\_type hpddm, or the routine \pk{PCSetType}(pc,
-\pk{PCHPDDM}) and its Fortran equivalent. In this section, it is assumed that one
-wants to solve 
+using the command line option \pk{-pc\_type hpddm}, or the routine \pk{PCSetType}(pc,
+\pk{PCHPDDM}).% and its Fortran equivalent.
+In this section, it is assumed that one wants to solve 
 \begin{equation}
@@
 the local operators $\{R_i A R_i^T\}_{i=1}^N$, GenEO needs users to supply
-the local unassembled matrices $\{\mathring{A}_i\}_{i=1}^N$, also known as
+the local unassembled matrices~$\{\mathring{A}_i\}_{i=1}^N$, also known as
 Neumann matrices. These are yielded by the
@@
 notice is that the size of the coarse operator $P^TAP$ is much lower
-than $n$, the size of $A$. It is thus common to have the coarse operator distributed on
+than $n$, the size of $A$. Furthermore, coarsening is much more aggressive than
+algebraic or geometric multigrids, see~\cref{sec:linear-elasticity} for
+evidence. It is thus common to have the coarse operator distributed on
 fewer processes than for $A$, see for example the concept of telescoping later
@@
 
-Of course, users do not have to implement the three previous steps themselves.
+% Of course, users do not have to implement the three previous steps themselves.
 The auxiliary operators $\{\mathring{A}_i\}_{i=1}^N$ have simply to be provided
@@
 overlapping operators and the correspondence are automatically assembled by
-PETSc for users. Other \pk{DMType} may be supported by implementing the proper
-\pk{DMCreateNeumannOverlap} callback. With the local matrix pencils available,
-new options will be automatically registered such as
--pc\_hpddm\_levels\_1\_eps\_nev or -pc\_hpddm\_levels\_1\_eps\_threshold, which
+PETSc for users, see~\cref{fig:p4est}. Other \pk{DMType} may be supported by
+implementing the proper \pk{DMCreateNeumannOverlap} callback, e.g., \pk{DMDA}
+support could be easily added upon request. With the local
+matrix pencils available, new options will be automatically registered such as
+\pk{-pc\_hpddm\_levels\_1\_eps\_nev} or \pk{-pc\_hpddm\_levels\_1\_eps\_threshold}, which
 may be used to set how many eigenpairs SLEPc should compute in step \#2.  The
-default -eps\_nev value is 20, i.e., the dimension of the coarse space is
+default \pk{-eps\_nev} value is 20, i.e., the dimension of the coarse space is
 $20\times N$ under the assumption that all concurrent SLEPc solves converge.
-Another important option is -pc\_hpddm\_coarse\_p, which sets the size of the
+It is usually best to keep the same number of eigenvectors on each process for multiple reasons:
+\begin{itemize}
+    \item with standard eigensolvers such as ARPACK or \pk{EPSKRYLOVSCHUR}, it is not possible to compute all eigenpairs in a given interval of $\mathbb{R}^+$.
+Instead, an upper bound of the number of eigenpairs has to be provided, and then it is possible in \pk{PCHPDDM} to remove eigenvectors associated to eigenvalues above the given \pk{-eps\_threshold}; extra eigenvectors are thus computed, but immediately thrown away;
+    \item if the number of eigenvectors is the same, coarse operators take advantage of block structures and are assembled as \pk{MatMPISBAIJ} or \pk{MatMPIBAIJ} matrices, with usually large block sizes, by default 20; this can have serious memory and performance implications, see~\cref{fig:perfAX};
+    \item during solution phases, it is possible to use the functions \texttt{MPI\_Gather} and \texttt{MPI\_Scatter} to go from one to another level and vice versa, instead of the variable size alternatives \texttt{MPI\_Gatherv} and \texttt{MPI\_Scatterv}, which are more costly.
+\end{itemize}
+Another important option is \pk{-pc\_hpddm\_coarse\_p}, which sets the size of the
 communicator on which the coarse operator should be distributed. The default
@@
 The action of both inverses from~\cref{eq:two} can be parameterized through the
--pc\_hpddm\_coarse\_ksp\_type and -pc\_hpddm\_levels\_1\_ksp\_type options
+\pk{-pc\_hpddm\_coarse\_ksp\_type} and \pk{-pc\_hpddm\_levels\_1\_ksp\_type} options
 respectively.
+\begin{figure}[htbp]
+  \centering
+    \subfloat[Initial nonoverlapping decomposition.]{
+        \includegraphics[width=0.45\textwidth]{Data/p4est_nonoverlapping.pdf}
+    }
+    \subfloat[Overlapping decomposition on which auxiliary operators $\{\mathring{A}_i\}_{i=1}^N$ are assembled.]{
+        \includegraphics[width=0.45\textwidth]{Data/p4est_overlapping.pdf}
+    }
+    \caption{Automatic generation of overlapping subdomains when using a~\pk{DMP4EST}. For clarity, only four subdomains, one of which is disconnected, are colored.\label{fig:p4est}}
+\end{figure}
+
     \subsubsection{Coarse corrections and eigenvalue problems}
@@
 very efficient numerically~\cite{tang2009comparison}, and thus not used as
-default in \pk{PCHPDDM}. Users can choose between
-$M'^{-1}_{\text{additive}}$ and:
+default in \pk{PCHPDDM}. This observation is not specific to domain
+decomposition methods and also applies to multigrid~\cite{bastian1998additive}.
+Users can choose between $M'^{-1}_{\text{additive}}$ and:
 \begin{align*}
@@
 \end{align*}
-with $Q = P \left(P^T A P\right)^{-1} P^T$. These preconditioners may be selected using the additional
-options -pc\_hpddm\_coarse\_correction (deflated$|$additive$|$balanced) or the
+with $Q = P \left(P^T A P\right)^{-1} P^T$. These preconditioners may be selected using the 
+options \pk{-pc\_hpddm\_coarse\_correction} \pk{(deflated$|$additive$|$balanced)} or the
 routine \pk{PCHPDDMSetCoarseCorrectionType}(pc, type) and its Fortran equivalent.
@@
 The local generalized eigenvalue problems~\cref{eq:gevp} solved by SLEPc can also be adjusted.
-For example, in addition to the local unassembled Neumann matrices
-$\{\mathring{A}_i\}_{i=1}^N$, if one has access to local matrices with optimized
-boundary conditions, e.g., Robin transmission conditions, $\{B_i\}_{i=1}^N$, it
+For example, in addition to the local unassembled Neumann matrices~$\{\mathring{A}_i\}_{i=1}^N$,
+if one has access to local matrices with optimized
+boundary conditions $\{B_i\}_{i=1}^N$, e.g., Robin transmission conditions, it
 is possible to pass these using the routine \pk{PCHPDDMSetRHSMat}(pc, $B_i$).
@@
 of solving coarse systems. \pk{PCHPDDM} follows the same numbering of
-\pk{PCBDDC}\footnote{which itself follows the opposite numbering of \pk{PCMG}}: the
-finest level is always numbered 1, the level index increases, up until the
-coarsest level whose options are always prefixed by coarse\_. In order to
+\pk{PCBDDC}%\footnote{which itself follows the opposite numbering of \pk{PCMG}}
+: the finest level is always numbered 1, the level index increases, up until
+the coarsest level whose options are always prefixed by \pk{coarse\_}. In order to
 register an additional level $l' = l+1$, the following conditions must be met at
@@
 \begin{itemize}
-    \item -pc\_hpddm\_levels\_1\_eps\_threshold 0.4 -pc\_hpddm\_coarse\_p 8 would
+    \item \pk{-pc\_hpddm\_levels\_1\_eps\_threshold 0.4 -pc\_hpddm\_coarse\_p 8} would
         define a two-level method with the coarse operator being distributed
         among 8 processes, assuming that there is globally enough $\lambda_i$ in~\cref{eq:gevp} lower than 0.4;
-    \item -pc\_hpddm\_levels\_1\_eps\_nev 10 -pc\_hpddm\_levels\_2\_p 8
-        -pc\_hpddm\_levels\_2\_eps\_nev 10 -pc\_hpddm\_coarse\_p 2 would define
+    \item \pk{-pc\_hpddm\_levels\_1\_eps\_nev 10 -pc\_hpddm\_levels\_2\_p 8} \linebreak 
+        \pk{-pc\_hpddm\_levels\_2\_eps\_nev 10 -pc\_hpddm\_coarse\_p 2} would define
         a three-level method with the second level (resp.~ coarse operator)
         being distributed among 8 (resp.~2) processes;
-    \item -pc\_hpddm\_levels\_1\_eps\_nev 10 -pc\_hpddm\_levels\_2\_p 8 is not
-        valid because there are only two levels, so the options levels\_2\_ are
+    \item \pk{-pc\_hpddm\_levels\_1\_eps\_nev 10 -pc\_hpddm\_levels\_2\_p 8} is not
+        valid because there are only two levels, so the options \pk{levels\_2\_} are
         not registered and coarse\_ has to be used instead;
-    \item -pc\_hpddm\_levels\_1\_eps\_nev 10 -pc\_hpddm\_levels\_2\_eps\_nev 10 -pc\_hpddm\_coarse\_p 2 
-        is not valid because there is no option levels\_2\_p, so the default
+    \item \pk{-pc\_hpddm\_levels\_1\_eps\_nev 10 -pc\_hpddm\_levels\_2\_eps\_nev 10} \linebreak \pk{-pc\_hpddm\_coarse\_p 2} 
+        is not valid because there is no option \pk{levels\_2\_p}, so the default
         value of 1 is used for the second level, and it is thus not possible to
@@
 overlap, see for example $\Omega_1$ and $\Omega_2$ from~\cref{fig:poisson} in a
-finite element context. Using a local to global mapping, it is possible to
-reconstruct the local assembled matrices, equivalent to the $\{R_i A
-R_i^T\}_{i=1}^N$ from~\cref{sec:interface-pc}. All tools needed by the GenEO
-framework are thus readily available. Note however that one-level overlapping
-Schwarz methods are known for converging slowly when there is in fact no
-overlap. It is common in the field of BDD to see the coarse problem as an
-unassembled operator as well. In fact, this is mandatory to define efficient
-multilevel variants, as implemented in \pk{PCBDDC}. In~\cref{sec:bddc}, it will be
-shown how a coarse problem obtained in \pk{PCBDDC} may be solved using
+finite element context. Using a local to global mapping and \pk{MatConvert}, it
+is possible to reconstruct the local assembled matrices, equivalent to the
+$\{R_i A R_i^T\}_{i=1}^N$ from~\cref{sec:interface-pc}. All tools needed by the
+GenEO framework are thus readily available. Note however that one-level
+overlapping Schwarz methods are known for converging slowly when there is in
+fact no overlap. It is common in the field of BDD to see the coarse problem as
+an unassembled operator as well. In fact, this is mandatory to define efficient
+multilevel variants, as implemented in \pk{PCBDDC}. In~\cref{sec:bddc}, it will
+be shown how a coarse problem obtained in \pk{PCBDDC} may be solved using
 \pk{PCHPDDM}.
@@
 \end{enumerate}
-The options -ksp\_gmres\_modifiedgramschmidt
+The options \pk{-ksp\_gmres\_modifiedgramschmidt
 -ksp\_gmres\_restart 50
--ksp\_type fgmres are used in conjunction with all \pk{PC}. The options specific to each solver described above are given next. \\[-4pt]
+-ksp\_type fgmres} are used in conjunction with all \pk{PC}. The options specific to each solver described above are given next. \\[-4pt]
 \begin{minipage}[t]{0.34\textwidth}
@@
 \begin{semilogyaxis}[xmin=0,xmax=50,width=0.46\textwidth,
-    legend style={at={(1.05,0.95)},cells={anchor=west}},ymajorgrids,
+    ymajorgrids,
     xlabel={Iteration number},
@@
 respectively 64 and 36 seconds. Out of these 64 seconds, 31 are spent in SLEPc
-\pk{EPSSolve} for solving GenEO at each level but the coarsest. The code used to
-produce these results was borrowed from the original paper about the multilevel
-extension of GenEO~\cite{aldaas2019multi} and is available
+\pk{EPSSolve} for solving GenEO at each level but the coarsest. With respect to
+coarsening, \pk{PCGAMG} has an operator complexity of 1.501, while for \pk{PCHPDDM},
+it is 1.015. The code used to produce these results was borrowed from
+the original paper about the multilevel extension of
+GenEO~\cite{aldaas2019multi} and is available
 at~\url{https://github.com/prj-/aldaas2019multi}.
@@
 combustion models. Here, it is merely used to test \pk{PCHPDDM} in the context
-of solving successive linearized equations, but there are many other techniques
-for solving such a nonlinear equation, such as first decomposing the domain and
-then linearizing the equation~\cite{dryja1997nonlinear}.
-The physical domain $\Omega$ is the unit cube. It is decomposed in $10{,}272$ subdomains.
+of solving successive linearized equations%, but there are many other techniques for solving such a nonlinear equation, such as first decomposing the domain and then linearizing the equation~\cite{dryja1997nonlinear}
+. The physical domain $\Omega$ is the unit cube. It is decomposed in $10{,}272$ subdomains.
 \Cref{eq:bratu} is discretized using $\mathbb{P}_2$ finite elements.
@@
 convergence history, which is displayed in~\cref{fig:bratu-conv}. The complete set of
-options is given in~\cref{fig:bratu-options}.
+options is given in~\cref{fig:bratu-options}. Results can be reproduced using the following FreeFEM script: \url{http://jolivet.perso.enseeiht.fr/bratu-hpddm.edp}
 \pgfplotstableread{
@@
 CISS~\cite{sakurai2003projection} implementations in SLEPc are
-currently\footnote{in version 3.13.2} not using blocking when applying 
+currently %\footnote{in version 3.13.2}
+not using blocking when applying 
 spectral transformations, but this is being rewritten to use the new routine
@@
 Roman, and B. Smith for reviewing the successive merge requests in PETSc and
-SLEPc repositories. This work was granted access to the HPC resources of
-TGCC@CEA under the allocations A0070607519 made by GENCI.
+SLEPc repositories. This work was granted access to the GENCI-sponsored HPC resources of:
+\begin{itemize}
+    \item TGCC@CEA under allocation A0070607519;
+    \item IDRIS@CNRS under allocation AP010611780.
+\end{itemize}
 % We would like to acknowledge the assistance of volunteers in putting
@@
 
+\bibitem{bastian1998additive}
+{\sc P.~Bastian, G.~Wittum, and W.~Hackbusch}, {\em Additive and multiplicative
+  multi-grid—a comparison}, Computing, 60 (1998), pp.~345--364.
+
 \bibitem{bavier2012amesos2}
@@
 
-\bibitem{dohrmann2008domain}
-{\sc C.~R. Dohrmann, A.~Klawonn, and O.~B. Widlund}, {\em {Domain decomposition
-  for less regular subdomains: Overlapping Schwarz in two dimensions}}, SIAM
-  Journal on Numerical Analysis, 46 (2008), pp.~2153--2168.
-
 \bibitem{dohrmann2016bddc}
@@
   Communications on Pure and Applied Mathematics, 69 (2016), pp.~745--770.
-
-\bibitem{dolean2014ddm}
-{\sc V.~Dolean, P.~Jolivet, and F.~Nataf}, {\em An Introduction to Domain
-  Decomposition Methods\col\ Algorithms, Theory and Parallel Implementation},
-  SIAM, 2015.
-
-\bibitem{dryja1997nonlinear}
-{\sc M.~Dryja and W.~Hackbusch}, {\em On the nonlinear domain decomposition
-  method}, BIT Numerical Mathematics, 37 (1997), pp.~296--311.
 
@@
 
+\bibitem{heinecke2016libxsmm}
+{\sc A.~Heinecke, G.~Henry, M.~Hutchinson, and H.~Pabst}, {\em {LIBXSMM:
+  accelerating small matrix multiplications by runtime code generation}}, in
+  Proceedings of the 2016 International Conference for High Performance
+  Computing, Networking, Storage and Analysis, SC16, IEEE, 2016,
+  \url{https://github.com/hfp/libxsmm}.
+
 \bibitem{Heinlein:2016:PIT}
@@
 
+\bibitem{mkl-web-page}
+{\sc Intel}, {\em {MKL} web page}.
+\newblock
+  \url{https://software.intel.com/content/www/us/en/develop/tools/math-kernel-library.html},
+  2020.
+
 \bibitem{ji2017breakdown}
@@
 
+\bibitem{saad2000deflated}
+{\sc Y.~Saad, M.~Yeung, J.~Erhel, and F.~Guyomarc'h}, {\em A deflated version
+  of the conjugate gradient algorithm}, SIAM Journal on Scientific Computing,
+  21 (2000), pp.~1909--1926.
+
 \bibitem{sakurai2003projection}
@@
   and Applied Mathematics, 159 (2003), pp.~119--128.
+
+\bibitem{sakurai2010application}
+{\sc T.~Sakurai, H.~Tadano, and Y.~Kuramashi}, {\em {Application of block
+  Krylov subspace algorithms to the Wilson--Dirac equation with multiple
+  right-hand sides in lattice QCD}}, Computer Physics Communications, 181
+  (2010), pp.~113--117.
 
@@
 
-\bibitem{smith2004domain}
-{\sc B.~F. Smith, P.~Bj{\o}rstad, and B.~Gropp}, {\em Domain decomposition:
-  parallel multilevel methods for elliptic partial differential equations},
-  Cambridge University Press, 2004.
+\bibitem{soodhalter2014krylov}
+{\sc K.~M. Soodhalter, D.~B. Szyld, and F.~Xue}, {\em Krylov subspace recycling
+  for sequences of shifted linear systems}, Applied Numerical Mathematics, 81
+  (2014), pp.~105--118.
 

### 2020-06-20 MatProduct.c

@@ -244,7 +244,7 @@ int main(int argc, char** argv) {
 
           ierr = MatDenseGetColumnVecRead(C, 0, &cC);CHKERRQ(ierr);
           ierr = MatDenseGetColumnVecWrite(D, 0, &cD);CHKERRQ(ierr);
-
+          ierr = MatMult(A, cC, cD);CHKERRQ(ierr);
           ierr = PetscPrintf(PETSC_COMM_WORLD,"Benchmarking MatMult: with A %s %Dx%D\n",Atype,AM,AN);
           ierr = PetscLogStagePush(stage);CHKERRQ(ierr);
           for (mm = 0; mm < trial; ++mm) {

### 2020-06-21 article.tex
Figures
start to work on the paper (up to subsection 2.2.1 included)

@@
 \hypersetup{
-  pdftitle={KSPHPDDM and PCHPDDM: Two Classes of Adaptive Krylov and Overlapping Schwarz Methods in PETSc},
+  pdftitle={KSPHPDDM and PCHPDDM: extending PETSc with robust Overlapping Schwarz Methods and advanced Krylov methods},
   pdfauthor={P. Jolivet and S. Zampini}
@@
     libraries, PETSc and HPDDM, have been interfaced together in order to offer
-    to end-users adaptive Krylov and overlapping Schwarz methods.  The
-    flexibility of the implementation %, which for example may deal with various matrix formats, 
-    is showcased and explained.  Diverse examples are
-    displayed, ranging from fracture mechanics to linear stability analysis, and
-    the sets of options are thoroughly documented, so that they may be readily
-    tried on other applications and reproduce results.
+    end-users robust overlapping Schwarz methods and advanced Krylov methods featurin
+    recycling and block solution techniques.
+    The flexibility of the implementation %, which for example may deal with various matrix formats,
+    is showcased and explained with minimalist, easy-to-run and reproducible examples
+    ranging from {\color{red}fracture mechanics to linear stability analysis}
+    to ease the integration of these algorithms into more advanced frameworks.
 \end{abstract}
@@
 Implementing and maintaining robust and scalable applications is highly relevant
-in the field of computational science and engineering. Some tools exist to help
-end-users achieve this goal. One the one hand, PETSc \cite{balay1997cient}, the
+in the field of computational science and engineering. The efficient production of
+such applications requires a rich ecosystem of state-of-the-art reusable libraries
+\cite{xsdk}. In this paper, we focus on the interoperability of two such libraries.
+
+One the one hand, PETSc \cite{balay1997cient}, the
 Portable and Extensible Toolkit for Scientific computation, is a
-well-established software from the community. It may be used to efficiently
-discretize partial differential equations, or solve algebraic linear or
-nonlinear systems of equations. It comes with many built-in features such as
-advanced preconditioning or tailored matrix formats. Also, it can % be used to compile and 
+well-established, actively developed software from the community. It may be used to efficiently
+discretize partial differential equations and solve algebraic linear or
+nonlinear systems of time-dependent equations. Among its strenghts, it offers many
+advanced features regarding preconditioning and tailored matrix formats, and it can
 interact with third-party libraries, such as
 \emph{hypre}~\cite{falgout2002hypre} for linear solvers, TetGen~\cite{si2013tet}
-for mesh generation, and so on. These appealing features convinced computational
-scientists to use PETSc as one of the discretization and/or algebraic backend in
-many different higher-level projects, such as finite element or finite volume
-packages. On the other hand, HPDDM is a much smaller project purely focused on
-domain decomposition preconditioners~\cite{jolivet2013scalable} % , a small subclass of preconditioners available in PETSc as well, 
-and advanced iterative
-methods~\cite{jolivet2016block}. Such methods deal efficiently with linear
-systems when multiple right-hand sides are available simultaneously, and are
-then referred to as block iterative methods, or when there is a recurrence of
-varying coefficient matrices and right-hand sides, and are then referred to as
-recycling iterative methods.
+for mesh generation, p4est~\cite{BursteddeWilcoxGhattas11} for adaptive mesh
+refinement, to cite a few. The extensibility and robustness of the framework
+convinced computational scientists to use PETSc as one of the discretization and/or algebraic backend in
+many different higher-level projects {\color{red} ADD CIT TO DEAL.II, FIREDRAKE, MOOSE, HIPPYLIB, MFEM}.
+
+On the other hand, the High-Performance unified framework for Domain Decomposition Methods (HPDDM)
+is a much smaller project focusing on robust and scalable domain decomposition
+preconditioners~\cite{jolivet2013scalable} and advanced iterative
+methods~\cite{jolivet2016block} which can efficiently deal with linear
+systems with multiple right-hand sides (also known and as block iterative methods)
+and when there is a recurrence of varying coefficient matrices and right-hand
+sides (recycling iterative methods). For a introduction to domain decomposition methods, see
+~\cite{smith2004domain,dolean2014ddm,olofbook,tarekmathewbook} {\color{red} FIX CITATIONS}.
+
 % Htool ?
-In this paper, the integration of both aspects of HPDDM, its advanced iterative
-methods and its domain decomposition preconditioners, into PETSc is explained.
-The design principles of the interface are showcased, with % alongside
-minimalist and easy-to-run examples. The paper is divided in two main parts:
-\cref{sec:KSPHPDDM} which introduces \pk{KSPHPDDM}, and \cref{sec:PCHPDDM} which
-introduces \pk{PCHPDDM}. Eventually, concluding remarks may be found
+Here we describe the integration of the entire HPDDM suite, its advanced iterative
+methods and its domain decomposition preconditioners, into PETSc.
+We showcase the design principles of the interface considering % alongside
+minimalist and easy-to-run examples, equipped with the necessary command line options to
+fullfill reproducibility requirements {\color{red} FIND GOOD CIT, ADD 1 more sentences?,
+https://www.acm.org/publications/policies/artifact-review-badging} and ease the
+integration of these algorithms into more advanced frameworks.
+
+The paper is divided in two main parts {\color{red} COMPLETE WITH MORE DETAILS WHEN PAPER IS COMPLETED}:
+\cref{sec:KSPHPDDM} introduces \pk{KSPHPDDM}, provides applications within an eigensolver
+framework and discuss the readiness of the PETSc library in delivering highly efficient block Krylov methods.
+\cref{sec:PCHPDDM} introduces \pk{PCHPDDM}. Eventually, concluding remarks may be found
 \cref{sec:conclusion}. All PETSc keywords and options are typeset in a \texttt{typewriter} front, these
 are public and documented at \url{https://www.mcs.anl.gov/petsc/petsc-master/docs}.
+
 \section{KSPHPDDM}\label{sec:KSPHPDDM}
@@
 Krylov subspace methods are widely used in numerical linear algebra for solving
-systems of equations. Some of their strengths are memory efficiency and the fact
-that they are mostly format agnostic. Indeed, they mainly rely on matrix--vector
-operations such as multiplications or transpose multiplications. For
-challenging high-dimensional systems, finding appropriate preconditioners is
-mandatory to make these methods robust. %, but this goes beyond the scope of this section.
-In PETSc, Krylov methods belong to the \pk{KSP} base class. Some of the
+linear systems of equations because of their memory efficiency~\cite{templateBook}.
+They mainly rely on matrix--vector
+operations such as multiplications or transpose multiplications and,
+for robust convergence when dealing with challenging high-dimensional systems,
+on the application of appropriate preconditioners.
+PETSc (as of version 3.13.2) already offers 44 different types of Krylov methods within the \pk{KSP} base class. Some of the
 most used types are \pk{KSPGMRES}, \pk{KSPCG}, or \pk{KSPBCGS}, which respectively implement
@@
 
-Because these methods may need long-term recurrences to mitigate round-off
-errors during the generation of Krylov subspaces, it is common to introduce a
-restart parameter in order to control their memory consumption and the amount of
-global communication, e.g., when orthonormalizing a candidate basis vector. Such
+Because Krylov methods may need long-term recurrences to mitigate round-off
+errors during the generation of subspaces, it is common to introduce a
+restart parameter in order to control their memory consumption and the volume of
+global communications needed when orthonormalizing a candidate basis vector. Such
 restarts may hinder convergence of iterative methods, and even introduce
@@
 Deflated GMRES~\cite{erhel1996restarted,wakam2013158}~(DGMRES), implemented as
-\pk{KSPLGMRES} and \pk{KSPDGMRES}, are available. Neither handle variable
-preconditioning, and the latter does not support complex arithmetic. Also note
+\pk{KSPLGMRES} and \pk{KSPDGMRES}, are available. However, neither handle variable
+preconditioning, and the DGMRES does not support complex arithmetic. {\color{red} Also note
 that when solving a sequence of linear systems $A_i x_i = b_i, i = 1, 2,
@@
 restarting~\cite{parks2006recycling}~(GCRODR) does not suffer from this
-limitation.
+limitation.}
 
@@
 Another important aspect of Krylov methods is their ability
-to deal with multiple right-hand sides simultaneously. Block Krylov methods~\cite{gutknecht2006block} are
+to deal with multiple right-hand sides simultaneously.
+Block Krylov methods~\cite{gutknecht2006block} are
 designed for solving linear systems $A X = B$, where
-$X$ and $B$ are tall-and-skinny matrices with $k \geq 1$ columns. These methods are
+$X$ and $B$ are tall-and-skinny matrices with $k \geq 1$ columns.
+Such methods, while having higher arithmetic
+intensities, generate larger subspaces and typically converge in fewer iterates.
+
+These methods are not currently offered in PETSc and they are
 less frequently implemented in general purpose libraries because they require
 somehow more involved kernels such as matrix--matrix multiplication, instead of
-matrix--vector. Still, systems with multiple right-hand sides are ubiquitous,
+matrix--vector, and may involve different inner-products realizations~\cite{Frommer}.
+Still, systems with multiple right-hand sides are ubiquitous,
 e.g., in tomography~\cite{tournier2016micro}, data
-analytics~\cite{KALANTZIS2018136}, eigensolvers~\cite{knyazev2001toward}, geophysics~\cite{calandra2012flexible}, quantum chromodynamics~\cite{sakurai2010application}, etc.
+analytics~\cite{KALANTZIS2018136}, eigensolvers~\cite{knyazev2001toward}, geophysics~\cite{calandra2012flexible}, quantum chromodynamics~\cite{sakurai2010application}, and optimization with time-dependent partial differential equations as constraints.
 % Of course, it is possible to cast a system with multiple right-hand sides into a sequence of systems with the same coefficient matrix but changing right-hand sides $A x_i = b_i, i = 1,2,\ldots,k$.
 However,
-Block Krylov methods, while having higher arithmetic
-intensities, generate larger subspaces and typically converge in fewer iterates.
-There is none implemented in PETSc. It is nevertheless possible to solve
+{\color{red} REMOVE? It is nevertheless possible to solve
 systems with multiple right-hand sides using a specialized type of matrices,
-but with a very limited choice of preconditioners, cf.~\cref{sec:matkaij}.
-In the context of optimization, it is also
-frequent to solve a sequence of linear systems $A_i X_i = B_i, i = 1, 2,
-\ldots$
+but with a very limited choice of preconditioners, cf.~\cref{sec:matkaij}.}
 
 Trilinos~\cite{heroux2005overview} is another well-known software for scientific
-computing. Through its Belos package~\cite{bavier2012amesos2}, it provides most
-of the advanced iterative methods aforementioned. Although having PETSc and
+computing and it provides iterative methods through its Belos package~\cite{bavier2012amesos2},
+Although having PETSc and
 Trilinos interoperate is possible, there is currently no \pk{KSP} interface to Belos. % This limits its composability inside complete PETSc applications.
-Furthermore,
-as displayed in~\cref{sec:interface-ksp}, some Krylov methods implemented in
-HPDDM are not available in Belos. In the following section, the details of the
-interface between HPDDM Krylov methods and PETSc matrix and preconditioner
-classes will be explained. It is assumed that the right-hand sides, and
+Furthermore, some of the Krylov methods implemented in
+HPDDM and discussed in ~\cref{sec:interface-ksp}, are not available in Belos.
+
+  \subsection{Interfacing HPDDM Krylov methods in PETSc\label{sec:interface-ksp}}\
+In this section we provide details of the
+interface between HPDDM Krylov methods and the various PETSc
+classes. It is assumed that the right-hand sides, and
 consequently the solutions, are always dense vectors or matrices.
-  \subsection{Interfacing HPDDM Krylov methods in PETSc\label{sec:interface-ksp}}\
+
 \pk{KSPHPDDM}, the interface between HPDDM Krylov methods and PETSc, is automatically
 registered since PETSc version 3.12 when configuring PETSc with the extra
-flag \pk{-{}-download-hpddm}. It is then possible to select the corresponding \pk{KSP}
-using the command line option \pk{-ksp\_type hpddm}, or the routine \pk{KSPSetType}(ksp,
-\pk{KSPHPDDM}).% and its Fortran equivalent.
-When \pk{KSPHPDDM} is being used, it is then possible to select between the following
-Krylov methods:
+flag \pk{-{}-download-hpddm}. HPDDM linear solvers can be selected using the command
+line option \pk{-ksp\_type hpddm}, or the routine \pk{KSPSetType}(ksp,
+\pk{KSPHPDDM}), and the following Krylov methods can be accessed:
 \begin{itemize}
@@
 \end{itemize}
-In pseudo-block variants, multiple similar operations like matrix--vector
-products are fused to achieve higher arithmetic intensity, or to decrease the
-number of global synchronizations, e.g., reductions for scalar products. But they are mathematically equivalent to
-their ``standard'' counterpart. These methods may be selected using the 
-options \pk{-ksp\_hpddm\_type} \pk{(gmres$|$cg$|$gcrodr$|$bgmres$|$bcg$|$bfbcg$|$bgcrodr$|$preonly)} and
-\pk{-ksp\_hpddm\_variant} \pk{(left$|$right$|$flexible)} to select whether preconditioning
-is applied on the left, on the right, or variable. It is also possible to set
-this option through the more common PETSc option \pk{-ksp\_pc\_side (left$|$right)}. 
+While being mathematically equivalent to
+their ``standard'' counterparts, the pseudo-block variants try to fuse together multiple similar
+operations like matrix--vector
+products to achieve higher arithmetic intensity ({\color{red} which method does this?}), or to decrease the
+number of global synchronizations needed for scalar products ({\color{red} which method does this?}).
+HPDDM Krylov methods may be selected {\color{red} TODO API FOR KSPHPDDMSetType? KSPHPDDMSetVariant?} using the
+options \pk{-ksp\_hpddm\_type} \pk{(gmres$|$cg$|$gcrodr$|$bgmres$|$bcg$|$bfbcg$|$bgcrodr$|$preonly)}.
+Preconditioning variants can be specified via
+\pk{-ksp\_hpddm\_variant} \pk{(left$|$right$|$flexible)} to select whether the preconditioner
+has to applied on the left, on the right, or if it cannot be represented as a linear operator.
+{\color{red} out of curiosity. how can you choose the side if felixble is specified?}
+It is also possible to set the preconditioning side through the more common PETSc option \pk{-ksp\_pc\_side (left$|$right)}.
 Note that the value of the latter option is useless for BCG or BFBCG, and HPDDM CG only handles left and
-variable preconditioning. According to the standard PETSc notation, ksp($A$, $P_B$)
-defines a \pk{KSP} used to solve a system $Ax=b$ using a preconditioner $P_B$ built
-using an operator $B$ which is usually $A$ but not necessarily. If the
-type of ksp is \pk{KSPHPDDM}, the solution phase, invoked by \pk{KSPSolve}(ksp, $b$, $x$), will be carried out by HPDDM. To do
-so, HPDDM will repeatedly call the following PETSc routines:
+variable preconditioning {\color{red} can you elaborate more?}.
+Finally, convergence monitoring with the \pk{KSPMonitor} interface is fully supported, as well as the specification
+of customized convergence testing via the \pk{KSPSetConvergenceTest} callback.
+
+According to the standard PETSc notation, ksp($A$, $P_{A'}$)
+defines a \pk{KSP} used to solve a system $Ax=b$ using a preconditioner $P_{A'}$ built
+using an operator $A'$ which is usually, but not necessarily, equal to $A$. While solving
+the linear system via \pk{KSPSolve}(ksp, $b$, $x$) or \pk{KSPMatSolve}(ksp, $B$, $X$),
+HPDDM will repeatedly call the following PETSc routines:
 \begin{itemize}
@@
         AX$;
-    \item \pk{PCApply}($P_B$, $y$, $x$) for $x = P_By$.
+    \item \pk{PCApply}($P_{A'}$, $y$, $x$) for $x = P_{A'}y$.
 \end{itemize}
-Currently, PETSc does not have a general routine for applying $P_B$ to a
+As of version 3.13.2, PETSc does not have a general routine for applying $P_{A'}$ to a
 tall-and-skinny dense matrix $Y$. In HPDDM, a specialized routine has been implemented
-to compute $X = P_BY$ if the preconditioner is of one of the following types:
+to compute $X = P_{A'}Y$ if the preconditioner is of one of the following types:
 \begin{itemize}
-    \item exact $LU$ or Cholesky factorization, i.e., $P_B = B^{-1}$, \pk{PCLU} or
-        \pk{PCCHOLESKY};
-    \item incomplete $LU$ or Cholesky factorization, \pk{PCILU} or \pk{PCICC};
-    \item block Jacobi with % a single subdomain per process and
-        exact or incomplete subdomain solvers, \pk{PCBJACOBI};
-    \item overlapping Schwarz methods with % a single subdomain per process and
-        exact or incomplete subdomain solvers, \pk{PCASM}.
+    \item \pk{PCLU} or \pk{PCCHOLESKY}: exact $LU$ or Cholesky factorization, i.e., $P_{A'} = {A'}^{-1}$;
+    \item \pk{PCILU} or \pk{PCICC}: incomplete $LU$ or Cholesky factorization;
+    \item \pk{PCBJACOBI}: block Jacobi with exact or incomplete subdomain solvers as above;
+    \item \pk{PCASM}: overlapping Schwarz methods with exact or incomplete subdomain solvers as above.
 \end{itemize}
-Indeed, in these cases, it is possible to get access to a factored matrix $F$
-and call \pk{MatMatSolve}($F$, $\tilde{Y}$, $\tilde{X}$).
-% With the two first sets of preconditioners, $\tilde{Y} = Y$ and $\tilde{X} = X$.
-With block Jacobi, $\tilde{Y}$, resp.
-$\tilde{X}$, represents only the part of the matrix $Y$, resp.
-$X$, local to a given process.
-With overlapping Schwarz methods, there are furthermore overlapping ``ghost'' unknowns, interested readers are referred to monographs on practical aspects of domain decomposition methods~\cite{smith2004domain,dolean2014ddm}.
-For all other types of preconditioners, HPDDM will loop over the columns of $Y$
-and apply $P_B$ successively to get all columns of $X$.
-
-The rest of the operations are then performed directly inside HPDDM. For
-example, in all Krylov methods using the Arnoldi process, it is mandatory to orthonormalize
-candidate basis vectors.
-% It would be impractical to try to reuse PETSc routines like \pk{KSPGMRESModifiedGramSchmidtOrthogonalization} for two reasons: 1) it does not handle the orthonormalization of multiple vectors at once which is needed in pseudo-block methods 2) HPDDM may be used without PETSc as a stand-alone library.
-To give some more technical details to curious readers, in HPDDM,
-$QR$ factorizations of tall-and-skinny dense matrices, which are needed by
-block Krylov methods and for recycling Krylov methods are computed using
-the CholQR algorithm~\cite{stathopoulos2002block}, or the block modified or classical
-Gram--Schmidt method. This may be adjusted with the option \pk{-ksp\_hpddm\_qr (cholqr$|$mgs$|$cgs)}.
+In the aforementioned cases, it is indeed possible to access the so-called factored matrix $F$
+%via \pk{PCFactorGetMatrix} and then call \pk{MatMatSolve}($F$, $\tilde{Y}$, $\tilde{X}$).
+via \pk{PCFactorGetMatrix} and then call \pk{MatMatSolve}($F$, $Y$, $X$) to take advantage
+of the blocking solving strategy from the various factorization packages interfaced with PETSc
+such as MUMPS~\cite{amestoy2001fully}, SuiteSparse, MKL\_PARDISO and SUPERLU {\color{red} ADD REFERENCES}.
+For the types of preconditioners not implementing \pk{PCFactorGetMatrix},
+HPDDM will loop over the columns of $Y$ and successively apply $P_{A'}$ to get all columns of $X$.
+Future work will consider extending the various implementations of the PETSc
+preconditioning infrastructure to increase the arithmetic intensity of the
+block preconditioner application phase.
+
+The rest of the operations are then performed directly inside HPDDM {\color{red} what about VecAXPY, VecNorm etc?}.
+Specifically, within Krylov methods using the Arnoldi process, or when recycling is requested,
+we need $QR$ factorizations of tall-and-skinny dense matrices to orthonormalize
+candidate basis vectors. Such operations are performed using the CholQR algorithm~\cite{stathopoulos2002block}, or
+via the block variants of the (modified) Gram--Schmidt method.
 For Hessenberg matrices generated by the Arnoldi
@@
 used instead~\cite{gutknecht2008updating}.
-
-Monitoring convergence with the \pk{KSPMonitor} and \pk{KSPSetConvergenceTest} callbacks is fully supported.
+These orthonormalization variants can selected at runtime with the option \pk{-ksp\_hpddm\_qr (cholqr$|$mgs$|$cgs)}. {\color{red} API?}
 
     \subsubsection{User-defined deflation space} Recycling Krylov methods try to
-extract information at the end of each cycle or when convergence is reached by
-solving small standard or generalized dense eigenproblems. Then, during a
-subsequent solve, this information is reused appropriately~\cite{saad2000deflated,stathopoulos2002block,soodhalter2014krylov}. Two new routines were
-added in PETSc to either get the information generated by such recycling
-methods, or set the information in case users know what is a good initial
-deflation subspace. These routines are respectively
-\pk{KSPHPDDMGetDeflationSpace} and \pk{KSPHPDDMSetDeflationSpace}. The subspaces
-are once again stored as distributed
-dense tall-and-skinny matrices.
+extract convergence information by solving a small standard or a generalized dense eigenproblem
+at the end of each cycle or when convergence is reached and then reuse it appropriately for a
+subsequent solve~\cite{saad2000deflated,stathopoulos2002block,soodhalter2014krylov}.
+The routine \pk{KSPHPDDMGetDeflationSpace} can be used to inquire about the details of such deflation subspaces, stored as distributed
+dense tall-and-skinny matrices. It is also possible to provide this information
+via \pk{KSPHPDDMSetDeflationSpace} in case end users already know a good initial
+deflation subspace.
+
     \subsubsection{MatKAIJ and systems with multiple right-hand sides\label{sec:matkaij}}
+{\color{red} TODO: move KSPMatSolve discussion out, and clean up the section}
 There is a wide range of available matrix types in PETSc. One of them is
@@
 
+\bibitem{templateBook}
+{\sc R.~Barrett, M.~Berry, T.~F. Chan, J.~Demmel, J.~Donato, J.~Dongarra,
+ V.~Eijkhout, R.~Pozo, C.~Romine, and H.~Van der Vorst}
+ {\em Templates for the solution of linear systems: building blocks for iterative methods},
+  Society for Industrial and Applied Mathematics, (1994).
+
+\bibitem{xsdk}
+{\sc R.~Bartlett, I.~Demeshko, T.~Gamblin, G.~Hammond, M.~A. Heroux, J.~Johnson, A.~Klinvex, X.~Li, L.~Curfman~McInnes, J.~D. Moulton, et~al.},
+  {\em xSDK Foundations: Toward an Extreme-scale Scientific Software Development Kit},
+  Supercomputing Frontiers and Innovations, 4 (2017), pp.~69--82.
+
 \bibitem{bastian1998additive}
@@
   \url{https://www.llnl.gov/casc/hypre}.
+
+\bibitem{Frommer}
+{\sc A.~Frommer, K.~Lund, D.~B. Szyld},
+ {\em Block Krylov subspace methods for funcions of matrices},
+ Electronic Transactions on Numerical Analysis, 47 (2017), pp.~100--126.
 

### 2020-06-21 article.tex
Finish working on KSPHPDDM (some TODO left)

@@
 % SIAM Article Template
-\documentclass[draft,review,onefignum,onetabnum]{siamart190516}
+\documentclass[review,onefignum,onetabnum]{siamart190516}
 
@@
     al.~\cite{parks2006recycling}}
-Alongside the paper introducing GCRODR, a MATLAB implementation was provided and
-is since then
+Alongside the paper introducing GCRODR~\cite{parks2006recycling},
+a MATLAB implementation was provided and is since then
 available\footnote{\url{https://www.sandia.gov/~mlparks/GCRODR.zip}}. It comes
@@
 this paragraph is to explain how the results have been reproduced in PETSc.
-% Thanks to \pk{KSPHPDDM}, it is quite straightforward.
-First, the sequence of MATLAB systems had to be converted to PETSc \pk{Mat} and \pk{Vec}.
-Then, it is just a matter of calling \pk{KSPSolve} in sequence.
 The dataset and the source code are available at
@@
         GCRODR(40, 20), and Jacobi right-preconditioned GCRODR(40, 20);
-    \item in PETSc with \pk{KSPGMRES}, unpreconditioned unrestarted GMRES($\infty$);
+    \item in PETSc with unpreconditioned and unrestarted GMRES($\infty$) \pk{KSPGMRES};
     \item in PETSc with \pk{KSPHPDDM}, all of the above.
 \end{itemize}
-All PETSc tests are run on four MPI
-processes. They can be launched using the following command line. \\[-4pt]
+The notation GCRODR($n$, $m$) indicates a restart after $n$ iterations and a recycling subspace of dimension $m$.
+In all the cases, convergence is declared when the initial unpreconditioned residual is reduced by 10 orders of magnitude.
+Except for the right-preconditioned GCRODR, these tests are the same as the
+ones from the original GCRODR paper~\cite{parks2006recycling}.
+All PETSc tests are performed using four MPI
+processes and they can be launched using the following command line: \\[-4pt]
 \begin{Verbatim}[fontsize=\footnotesize,frame=single,framerule=0.1mm,commandchars=&\[\]]
@@
 \end{Verbatim}
-The various number of iterations needed to reach the convergence criterion as
-originally set to $10^{-10}$ are gathered in~\cref{fig:comparison}.
-Except for the right-preconditioned GCRODR, these tests are the same as the
-ones from the original GCRODR paper~\cite{parks2006recycling}. Furthermore,
-\pk{PCICC} and ichol from MATLAB are not consistent.  Apart from
-the tests \ref{pgfplots:r2} and \ref{pgfplots:r3} using $ICC(0)$ and \ref{pgfplots:r4} and \ref{pgfplots:r5}
-using the Jacobi method, one may notice that \ref{pgfplots:r0} and
-\ref{pgfplots:r1} (resp. \ref{pgfplots:r6} and \ref{pgfplots:r7}) are almost a
-perfect match and reproduce partially Figure 4.2 from Parks et
-al.~\cite{parks2006recycling}.
+The iteration numbers needed to reach convergence are gathered in~\cref{fig:comparison};
+the cases \ref{pgfplots:r0} and
+\ref{pgfplots:r1} (resp. \ref{pgfplots:r6} and \ref{pgfplots:r7}) reproduce Figure 4.2 in~\cite{parks2006recycling}.
+We note here that the results obtained with
+\pk{PCICC} from PETSc (\ref{pgfplots:r3}) and ichol from MATLAB (\ref{pgfplots:r2}, which uses SuiteSparse) are not consistent.
+%Apart from
+%the tests \ref{pgfplots:r2} and \ref{pgfplots:r3} using $ICC(0)$ and \ref{pgfplots:r4} and \ref{pgfplots:r5}
+%using the Jacobi method, one may notice that \ref{pgfplots:r0} and
+%\ref{pgfplots:r1} (resp. \ref{pgfplots:r6} and \ref{pgfplots:r7}) are almost a
+%perfect match and reproduce partially Figure 4.2 from Parks et
+%al.~\cite{parks2006recycling}.
 \pgfplotstableread{
@@
         \textnormal{ichol} from MATLAB are not consistent, hence the discrepancy
-        between \ref{pgfplots:r2} and \ref{pgfplots:r3}. All other \pk{KPSHPDDM}
-        numbers of iterations match with the respective MATLAB or PETSc
-        reference implementation.\label{fig:comparison}}
+        between \ref{pgfplots:r2} and \ref{pgfplots:r3}. In the other cases,
+        the numbers of iterations needed by KSPHPDDM match with the respective
+        MATLAB or PETSc reference implementation.\label{fig:comparison}}
     \end{figure}
     \subsubsection{Linear stability analysis}
-SLEPc~\cite{hernandez2005ssf}, the Scalable Library for Eigenvalue Problem
-computations, is built on top of PETSc. Thus, it is possible to reuse and
-compose \pk{KSP} or preconditioners for solving eigenproblems efficiently, e.g.,
-through spectral transformations. Recently, it has been observed that in the
-field of computational fluid dynamics, the role of recycling can be quite
-significant when performing linear stability analysis~\cite{moulin2018al}.
-The following generalized eigenvalue problem is considered next:
+In this section we apply recycling Krylov methods in the context linear
+stability analysis and consider the solution the following generalized eigenvalue problem:
 \begin{equation}\label{eq:lsa}
@@
 where $J$ is the Jacobian of the incompressible steady-state Navier--Stokes equation
-evaluated using a given base flow $q_b = \begin{bmatrix}u_b & p_b\end{bmatrix}^T$ and $M$ the discretization of the mass matrix
-on the space of velocities. There are many ways to compute the eigenvalues
-of~\cref{eq:lsa} near a complex-valued shift $\sigma$, but emphasis will be put
-in this paragraph on Krylov--Schur methods~\cite{Stewart2002}. These rely on the successive
-solution of linear systems such as:
+evaluated using a given base flow $q_b = \begin{bmatrix}u_b & p_b\end{bmatrix}^T$ and
+$M$ the discretization of the mass matrix on the space of velocities. Interested
+readers are referred to~\cite{moulin2018al} for more details and for larger
+runs.
+
+The mini-app employed for the numerical results is built on top of FreeFEM~\cite{hecht2012new} and it is
+available at \url{https://github.com/prj-/moulin2019al}. As eigenvalue solver, we consider the Scalable Library for Eigenvalue Problem
+computations (SLEPc)~\cite{hernandez2005ssf}, which is built on top of PETSc,
+and thus reuse its linear solver infrastructure.
+
+There are different methods to compute the eigenvalues
+of~\cref{eq:lsa} near a complex-valued shift $\sigma$; in this paragraph, emphasis will be put
+on Krylov--Schur methods~\cite{Stewart2002}, which rely on spectral tranformation steps
+and on the solution of successive linear systems such as:
 \begin{equation}\label{eq:ks}
-    \left(J(q_b) - \sigma \begin{bmatrix}M&0\\0&0\end{bmatrix}\right)x_i = b_i, i = 1,2,\ldots
+    \left(J(q_b) - \sigma \begin{bmatrix}M&0\\0&0\end{bmatrix}\right)x_i = b_i,
 \end{equation}
-In SLEPc, these systems are parameterized using the \pk{-st\_} prefix, e.g.,
-\pk{-st\_pc\_type~lu}. If they are solved iteratively, it is possible to
-use a \pk{KSP} of type \pk{KSPHPDDM} to recycle information between
-successive linear systems yielded by the convergence of the Krylov--Schur
-algorithm. Using the FreeFEM~\cite{hecht2012new} mini-app from \url{https://github.com/prj-/moulin2019al}
-which precondition~\cref{eq:ks} using a modified augmented Lagrangian
-approach~\cite{Benzi2011a}, the effect of such recycling is shown next for
+which, in SLEPc, are parameterized using the \pk{-st\_} prefix.
+In this work, the above linear system is preconditioned using a modified augmented Lagrangian
+approach~\cite{Benzi2011a,moulin2018al}
+
+The effectiveness of subspace recycling is shown for
 finding the 5 eigenpairs closest to the shift $\sigma = 10^{-6} + 0.6\mathfrak{i}$ for a
-flow past a cylinder at Reynolds 100, interested readers are again referred to
-the previously cited paper~\cite{moulin2018al} for more details and for larger
-runs. Results can be reproduced using the following four commands, the first two
-being merely for generating the base flow $q_b$ with \pk{SNESSolve} using a
+flow past a cylinder at Reynolds 100. Results can be reproduced using the following four commands;
+we note here that the first two commands are merely used to generate the base flow $q_b$ with \pk{SNESSolve} using a
 continuation method on the Reynolds number. \\[-4pt]
@@
 \end{Verbatim}
-The Krylov--Schur algorithm converges in six outer iterations. In each of these
-iterations, multiple linear systems having the structure of~\cref{eq:ks} are
-solved. In~\cref{fig:comparison-slepc}, numbers of iterates per inner solves
-are reported. After the first inner solve, once recycling kicks in, one may
-clearly see the benefits of switching from \pk{KSPFGMRES} to \pk{KSPHPDDM}.
-Indeed, through the 46 inner solves, these numbers are roughly decreased by 25.
-In terms of runtime, \SI[round-mode          = places,round-precision=1]{1.5098e+03}{\sec} (resp.~\SI[round-mode          = places,round-precision=1]{1.0524e+03}{\sec}) is spent in \pk{EPSSolve} when using \pk{KSPFGMRES} (resp.~\pk{KSPHPDDM}).
-These timings, and all which follow except in~\cref{sec:primitives}, have been obtained on {Ir\`ene}, a system composed of \pgfmathprintnumber[assume math mode=true]{1656} nodes with two 24-core Intel Xeon Platinum 8168 clocked at \SI{2.7}{\giga\hertz}.
+
+The Krylov--Schur algorithm converges in six outer iterations.
+Recycling of Krylov subspaces with \pk{KSPHPDDM} provides algorithmic speedup by lowering the number of iterations per inner solve,
+as reported in~\cref{fig:comparison-slepc}, from 46 with \pk{KSPFGMRES} to 25.
+A similar speedup can be observed in terms of runtime, with the time spent for the solution of the eigenvalue problem being reduced from \SI[round-mode = places,round-precision=1]{1509}{\sec} to \SI[round-mode = places,round-precision=1]{1052}{\sec}.
+{\color{red} FIXME LATER: These timings, and all which follow except in~\cref{sec:primitives}, have been obtained on {Ir\`ene}, a system composed of \pgfmathprintnumber[assume math mode=true]{1656} nodes with two 24-core Intel Xeon Platinum 8168 clocked at \SI{2.7}{\giga\hertz}}.
 \pgfplotstableread{
@@
     \subsubsection{Performance of primitives for block Krylov methods}\label{sec:primitives}
-As detailed in~\cref{sec:interface-ksp}, it is currently not possible with most
-PETSc preconditioners to apply them to a block of vectors. Even though there
-are efforts to enable such an operation for an increasing number of \pk{PC}, in
-this section, the focus is instead put on the matrix--matrix multiplication
-needed in block Krylov methods or when restarting recycling Krylov methods.
-In particular, the performance of three PETSc matrix types are evaluated:
+We close this section discussing the readiness of PETSc (as of version 3.13.2)
+in delivering efficient implementations of the core matrix--matrix multiplication
+needed by block Krylov methods and by recycling Krylov methods when restart is performed.
+
+In particular, we consider three PETSc matrix types:
 \begin{itemize}
@@
     \item \pk{MatSeqBAIJ}, block sparse matrices, based on block compressed sparse row format;
-    \item \pk{MatSeqSBAIJ}, symmetric block sparse matrices.
+    \item \pk{MatSeqSBAIJ}, symmetric block sparse matrices stored in upper triangular form.
 \end{itemize}
-Parallel matrices in PETSc heavily rely on sequential matrices.
-For example, \pk{MatMPIAIJ}, standard parallel sparse matrices, store on each rank two \pk{MatSeqAIJ}, one with local rows and columns, another with local rows and ``nonlocal'' columns.
-To keep this study succinct, only nondistributed sparse matrices will be evaluated next, which can help draw conclusions for the intraprocess performance of parallel sparse matrices.
-The Seq substring is now dropped, since in this case, \pk{MatSeqAIJ} is equivalent to \pk{MatAIJ} and so on.
-The performance of the multiplication primitives of the three aforementioned types will also be compared to MKL inspector--executor sparse BLAS routines~\cite{mkl-web-page}, using a single OpenMP thread.
-There is an ongoing effort to better integrate these routines in PETSc respective types \pk{MatAIJMKL} and \pk{MatBAIJMKL}, but at the time of writing, direct calls to the MKL were used.
-
-The Poisson equation on a cube is discretized with $\mathbb{P}_1$ finite elements in double precision with 32-bit integers.
-This yields a matrix $A$ of dimension one million.
-Though the numerical values of $A$ are of no interest here, the sparsity pattern is rather common and frequently encountered, e.g., when discretizing partial differential equations.
-Then, a random symmetric dense matrix $T$ of dimension $b$ is generated, for $b \in \{ 1,3,6 \}$, and the following matrix is evaluated $\mathbb{A} = A \otimes T$.
-The matrix $\mathbb{A}$ is then assembled into all three formats described above.
-For block matrices, a block size of $b$ is used.
-The performance of the \pk{MatProductNumeric} operation is evaluated for tall-and-skinny dense matrices with a varying number of columns $N \in \{1,2,8,16,32,64\}$.
-For $N=1$, the \pk{MatMult} operation for matrix--vector multiplication is used instead.
-All results are gathered in~\cref{fig:perfAX}.
-They can be reproduced using the following mini-app: \url{http://jolivet.perso.enseeiht.fr/MatProduct.c} and any input \pk{MatAIJ} stored in binary format, here using the default name \pk{binaryoutput}.
+Their distributed memory counterparts rely on the sequential implementations, with each process
+storing two such matrices, one with local rows and columns, and another with local rows and ``nonlocal'' columns.
+To keep this study succinct, we only evaluate the performances of the sequential implementations, which can help
+in drawing conclusions for the intraprocess performance of the parallel formats.
+For this reason, from now on we drop the \pk{Seq} substring.
+The performances of the multiplication primitives of the three aforementioned types will also be compared to
+those obtained with the MKL inspector--executor sparse BLAS routines~\cite{mkl-web-page}, using a single OpenMP thread.
+We note that there is an ongoing effort to better integrate these routines in PETSc, but at the time of writing, direct calls to the MKL were used.
+
+To perform the benchmarking, we first discretize the Poisson equation on a cube with $\mathbb{P}_1$ finite elements
+and obtain a matrix $A$ of dimension one million. Though the numerical
+values of $A$ are of no interest here, the sparsity pattern is rather common and frequently encountered when
+discretizing partial differential equations. We then generate a random symmetric dense matrix $T$ of dimension $b$,
+for $b \in \{ 1,3,6 \}$, and we construct the matrix $\mathbb{A} = A \otimes T$ into all three formats described above, using
+a block size of $b$ for the block storage formats.
+The performance of the \pk{MatProductNumeric} operation is evaluated for tall-and-skinny dense matrices with a varying number
+of columns $N \in \{2,8,16,32,64\}$. For the single column case $N=1$, the \pk{MatMult} operation is used.
+All results have been obtained using double precision arithmetic and 32-bit integers; we report the scaled efficiency (measured in GFLOP per second)
+with respect to the baseline \pk{MatAIJ} implementation with $N=1$ in~\cref{fig:perfAX}, where performances have been averaged over five consecutive
+product operations.
+They can be reproduced by using the mini-app \url{http://jolivet.perso.enseeiht.fr/MatProduct.c} and any input \pk{MatAIJ} stored in binary format,
+here using the default name \pk{binaryoutput} {\color{red} Do we want to provide the binary of the matrix used?}.
 \begin{Verbatim}[fontsize=\footnotesize,frame=single,framerule=0.1mm,commandchars=&\[\]]
@@
     \end{tikzpicture}
-   \caption{Performance of the matrix--matrix multiplication as implemented in PETSc for three sparse matrix types.
-As baseline values with $N=1$, performance of \pk{MatAIJ} matrix--vector multiplication is used~\ref{pgfplots:aij1}.
-Results are averages of 5 consecutives multiplications.\label{fig:perfAX}}
+   \caption{Performance of the matrix--matrix multiplication as implemented in PETSc for three sparse matrix types.\label{fig:perfAX}}
 \end{figure}
@@
 \begin{itemize}
-    \item with a block size of 1, top plot $b=1$ in~\cref{fig:perfAX}, using any type but \pk{MatAIJ} is counterproductive, see~\ref{pgfplots:baij1} and~\ref{pgfplots:sbaij1}, even with the MKL, see~\ref{pgfplots:mklbaijseq1} and~\ref{pgfplots:mklsbaijseq1}.
-However, with this format, as soon as the number of columns is greater or equal than 8, the efficiency reaches approximately 200\%, see~\ref{pgfplots:aij1}; PETSc implementation seems slightly better than MKL except for the single-column case;
+    \item with a block size of 1, top plot $b=1$ in~\cref{fig:perfAX}, using any type but \pk{MatAIJ}~\ref{pgfplots:aij1} is counterproductive when $N>1$, see~\ref{pgfplots:baij1} and~\ref{pgfplots:sbaij1}, even when using MKL, see~\ref{pgfplots:mklbaijseq1} and~\ref{pgfplots:mklsbaijseq1}.
+With 8 or more columns, the performance of \pk{MatAIJ} plateaus and the efficiency reaches approximately 200\%, while the perfomances of the MKL primitives stagnate at around 175\% see~\ref{pgfplots:aij1};
+for the single column case, \pk{MatSBAIJ} and \pk{MatAIJMKL} deliver slightly better performances;
     \item for block sizes lower or equal than 5, loops for matrix--vector and matrix--matrix multiplications are unrolled by hand for block formats.
@@
 For block formats \ref{pgfplots:baij3} and~\ref{pgfplots:sbaij3}, given the small value of $b$, it could be beneficial to switch to optimized libraries for small matrices, e.g., LIBXSMM~\cite{heinecke2016libxsmm}; in fact, MKL implementation for block formats is here clearly outperforming PETSc, see~\ref{pgfplots:mklbaijseq3} and~\ref{pgfplots:mklsbaijseq3};
-    \item for greater block sizes, \texttt{?gemv} or \texttt{?gemm} are used instead for block formats in PETSc.
-For \pk{MatAIJ}~\ref{pgfplots:aij10} in bottom plot $b=10$, the efficiency quickly caps near 400\% for $N\in\{8,16,32,64\}$, but for block formats, this reaches 500\% for large number of columns, see~\ref{pgfplots:baij10} and~\ref{pgfplots:sbaij10} for $N=64$.
+    \item for block sizes larger than 5, block entry multiplication with PETSc block formats is performed using \texttt{?gemv} or \texttt{?gemm} operations.
+For \pk{MatAIJ}~\ref{pgfplots:aij10} in the bottom plot corresponding to $b=6$, the efficiency quickly caps near 400\% for $N\in\{8,16,32,64\}$, while reaching 500\% for large number of columns for the block formats, see~\ref{pgfplots:baij10} and~\ref{pgfplots:sbaij10} for $N=64$.
 With these numbers of columns, PETSc and MKL perform similarly.
@@
 For example, if the matrices were to be offloaded to a GPU, one could instead benchmark \pk{MatAIJcuSPARSE}.
-The mini-app may also be used to benchmark intranode performance of matrix--matrix multiplications, either with multiple OpenMP threads, or using CUDA.
+
+We then reuse the mini-app to benchmark intranode performance of matrix--matrix multiplications, either with multiple OpenMP threads, or by offloading the operation to a GPU and use the cuSPARSE implementation~\cite{cusparse-web-page}.
 \begin{Verbatim}[fontsize=\footnotesize,frame=single,framerule=0.1mm,commandchars=&\[\]]
@@
 \end{Verbatim}
-Results are reported in~\cref{fig:perfAXpar}.
+Results reported in~\cref{fig:perfAXpar} have been scaled with respect to the baseline implementation \pk{MatAIJMKL} with $N=1$.
 They have been obtained on a single node of {Jean Zay}, a system composed of \pgfmathprintnumber[assume math mode=true]{261} nodes with two 20-core Intel Xeon Gold 6248 clocked at \SI{2.5}{\giga\hertz} and four NVIDIA Tesla V100 SXM2.
-Neither transfers between host and device nor time spent in \texttt{mkl\_sparse\_optimize} are taken into account.
-Once again, it is clear that the larger the number of right-hand sides $N$, the better the efficiency, no matter the block size.
+Neither transfers between host and device nor time spent in \texttt{mkl\_sparse\_optimize} have been taken into account.
+{\color{red} FEW MORE SENTENCES, percentage of peak, wait for results with different number of threads}
+
 \begin{figure}[tbp]
@@
     \end{tikzpicture}
-   \caption{Performance of the matrix--matrix multiplication as implemented in PETSc using intranode parallelism.
-As baseline values with $N=1$, performance of \pk{MatAIJMKL} matrix--vector multiplication is used~\ref{pgfplots:aijmkl1}.
-Results are averages of 5 consecutives multiplications.\label{fig:perfAXpar}}
+   \caption{Performance of the matrix--matrix multiplication as implemented in PETSc using intranode parallelism. \label{fig:perfAXpar}}
 \end{figure}
@@
 
+\bibitem{cusparse-web-page}
+{\sc NVIDIA}, {\em {cuSPARSE} web page}.
+\newblock
+  \url{https://docs.nvidia.com/cuda/cusparse/index.html},
+  2020.
+
 \bibitem{notay2000flexible}

### 2020-06-22 article.tex
WOrk on PCHPDDM section

@@
 and when there is a recurrence of varying coefficient matrices and right-hand
-sides (recycling iterative methods). For a introduction to domain decomposition methods, see
+sides (recycling iterative methods). For an introduction to domain decomposition methods, see
 ~\cite{smith2004domain,dolean2014ddm,olofbook,tarekmathewbook} {\color{red} FIX CITATIONS}.
@@
     \subsubsection{MatKAIJ and systems with multiple right-hand sides\label{sec:matkaij}}
-{\color{red} TODO: move KSPMatSolve discussion out, and clean up the section}
+{\color{red} TODO: move KSPMatSolve discussion out, and clean up the section, add results for irk?}
 There is a wide range of available matrix types in PETSc. One of them is
@@
 modern large-scale applications dealing with partial differential equations.
+For an introduction to domain decomposition methods, see
+~\cite{smith2004domain,dolean2014ddm,olofbook,tarekmathewbook} {\color{red} FIX CITATIONS}.
 % In PETSc, preconditioners inherit from the \pk{PC} base class.
-Historically, the block Jacobi and overlapping Schwarz methods were implemented
-as \pk{PCBJACOBI} and \pk{PCASM} respectively. ASM stands for additive Schwarz
-method, as originally introduced by Schwarz~\cite{schwarz1870alternating}, but
-the PETSc implementation is more versatile in the sense that one can easily
-switch between this method and other variants such as the restricted additive
-Schwarz method~\cite{cai1999restricted}. More details about overlapping domain decomposition
-preconditioners will be given in~\cref{sec:interface-pc}, but first the other
-related preconditioners will be quickly presented.
-
+
+Basic one-level methods such as the block Jacobi and the (overlapping) additive Schwarz method are implemented
+in PETSc as \pk{PCBJACOBI} and \pk{PCASM} respectively, with some additional customization~\cite{cai1999restricted}
+available for \pk{PCASM}. However, none of these methods provide automatic support for a coarse level solver,
+which is mandatory to obtain algorithmic scalability with a large number of processes.
+%ASM stands for additive Schwarz
+%method, as originally introduced by Schwarz~\cite{schwarz1870alternating}, but
+%the PETSc implementation is more versatile in the sense that one can easily
+%switch between this method and other variants such as the restricted additive
+%Schwarz method~\cite{cai1999restricted}. More details about overlapping domain decomposition
+%preconditioners will be given in~\cref{sec:interface-pc}, but first the other
+%related preconditioners will be quickly presented.
 % In PETSc, there is also an implementation of two-level overlapping Schwarz
@@
 % preconditioner~\cite{dohrmann2008domain}. Its type is \pk{PCEXOTIC}.
-When it comes to nonoverlapping preconditioners, \pk{PCNN} implements a basic balancing
-Neumann--Neumann method~\cite{mandel1993balancing}. Much more advanced
-techniques are available in \pk{PCBDDC}~\cite{zampini2016pcbddc}: adaptive selection
-of primal constraints~\cite{pechstein2017unified}, change of basis for problems in
-$H(\text{curl})$~\cite{dohrmann2016bddc}$\ldots$ For multigrid,
-\pk{PCMG} exposes a machinery for geometric or algebraic preconditioners. In
-particular, it is used for the implementation of the smoothed-aggregation
+Furthermore, PETSc provides domain decomposition methods of the non-overlapping type;
+\pk{PCNN} implements a basic balancing
+Neumann--Neumann method~\cite{mandel1993balancing}, while more advanced
+preconditioning techniques are available with \pk{PCBDDC}~\cite{zampini2016pcbddc}, including guaranteed robust convergence
+by means of adaptive selection
+of primal constraints~\cite{pechstein2017unified} and support for $H(\text{curl})$ and $H(\text{div})$
+conforming finite elements~\cite{zampinihcurl1,zampinihcurl2,zampinihdiv,zampinidarcy}, as well as for IsoGeometric Analysys \cite{zampiniiga1,zampiniiga2,zampiniiga3}.
+For multigrid,
+\pk{PCMG} provides a unified framework for geometric or algebraic preconditioners, which is used for the implementation of the smoothed-aggregation
 \pk{PCGAMG}~\cite{adams2004ultra}. There are also interfaces to other well-known
-multigrid packages such as \pk{PCHYPRE} or \pk{PCML}~\cite{gee2006ml} from Trilinos. Of
+multigrid packages such as \pk{PCHYPRE}~\cite{falgout2002hypre} or \pk{PCML}~\cite{gee2006ml} from Trilinos. Of
 course, there are many other available domain decomposition preconditioners,
@@
 
-HPDDM implements the same one-level overlapping Schwarz methods as PETSc, albeit
-emphasis is also put on optimized Schwarz methods~\cite{gander2006optimized}.
-The main theoretical framework for defining robust two-level methods is called GenEO,
-for generalized eigenvalue problem on the overlap~\cite{spillane2011robust,spillane2013abstract}.
-GenEO was also implemented in other libraries~\cite{butler2020high}, but to the best of our knowledge,
-the interface between HPDDM and PETSc proposes the first high-level access to
-GenEO so that it can, for example, be used as a block-decomposed
-solver~\cite{brown2012composable}.
+HPDDM implements the same one-level overlapping Schwarz methods as PETSc as well as
+optimized Schwarz methods~\cite{gander2006optimized} {\color{red} Are you planning to expose them too?}.
+It also provides support for defining robust two-level methods using the GENeralized Eigenvalue problem on the Overlap (GenEO) framework~\cite{spillane2011robust,spillane2013abstract}; the results presented in this work presents the first high-level access to
+GenEO from PETSc and its composable solver infrastructrure~\cite{brown2012composable}.
+We note that GenEO has been recently implemented in other libraries as well~\cite{butler2020high}
 
   \subsection{Interfacing HPDDM overlapping Schwarz methods in PETSc\label{sec:interface-pc}}\
-\pk{PCHPDDM}, the interface between HPDDM overlapping Schwarz methods and PETSc, is automatically
+The interface between HPDDM overlapping Schwarz methods and PETSc, \pk{PCHPDDM}, is automatically
 registered since PETSc version 3.12 when one configures PETSc with the extra
@@
 using the command line option \pk{-pc\_type hpddm}, or the routine \pk{PCSetType}(pc,
-\pk{PCHPDDM}).% and its Fortran equivalent.
-In this section, it is assumed that one wants to solve
+\pk{PCHPDDM}).
+ % and its Fortran equivalent.
+In this section, we are interested in solving
 \begin{equation}
@@
 assembled on $N$ processes, using a ksp($A$, $P_A$), i.e., the matrix used to
-define the preconditioner is $A$ itself. Without anything more from the user,
+define the preconditioner is $A$ itself {\color{red} WHY IS THAT? You do not support a different preconditioning matrix?}. Without anything more from the user,
 \pk{PCHPDDM} is strictly equivalent to \pk{PCASM}. That is, if $A$ is distributed among
-$N$ processes the preconditioner $M^{-1}$ is:
+$N$ processes, the action of the preconditioner $M^{-1}$ is:
 \begin{equation}\label{eq:one}
@@
 where $\{R_i\}_{i=1}^N$ are restriction operators that act on a global vector and
-return a vector local to each process, possibly with some overlap.
+return a local vector on each process, possibly with some overlap.
 $\{\tilde{R}_i\}_{i=1}^N$ are the same operators except that coefficients on the
-overlap are set to 0. The action of the inverse of all $\{R_i A R_i^T\}_{i=1}^N$
-can be parameterized through a local \pk{KSP}. If it is assumed that $A$ is the
-discretization of a linear operator $\mathcal{L}$, e.g., $\mathcal{L}=-\Delta$, on a domain $\Omega$, it is
+overlap are set to 0~\cite{cai1999restricted}. The action of each $\{R_i A R_i^T\}^{-1}_{i}$
+can be parameterized through a local \pk{KSP}. When $A$ is the
+discretization of a linear partial differential operator operator $\mathcal{L}$ on a domain $\Omega$, it is
 extremely common to exhibit parallelism by distributing $\Omega$ on $N$
 processes, possibly with some overlap, and to construct the $\{R_i\}_{i=1}^N$ so
-that they map global unknowns to unknowns local to each process. In addition to
+that they map global unknowns to unknowns local to each process.
+
+In addition to
 the local operators $\{R_i A R_i^T\}_{i=1}^N$, GenEO needs users to supply
@@
 domain $\Omega$. In the case of the 1D Poisson
-equation with homogeneous Dirichlet boundary conditions discretized using $\mathbb{P}_1$ finite elements, an explicit representation of these various operators is given~\cref{fig:poisson} for a
+equation with homogeneous Dirichlet boundary conditions discretized using $\mathbb{P}_1$ finite elements, an explicit representation of these  operators is given~\cref{fig:poisson} for a
 two-domain decomposition.
@@
 \end{figure}
-Then, it is possible to enrich the original one-level preconditioner
+
+With extra Neumann information at hand, it is then possible to enrich the original one-level preconditioner
 from~\cref{eq:one} using a spectral coarse grid built in the following way:
@@
 \end{enumerate}
-Note that $\{\tilde{R}_i R_i^T\}_{i=1}^N$ define a partition of unity, i.e.,
+Note that $\{\tilde{R}_i R_i^T\}_{i=1}^N$ defines a partition of unity, i.e.,
 \begin{equation*}
@@
 \end{equation*}
-When used without PETSc and thus SLEPc, step \#1 is solved in HPDDM with
-ARPACK~\cite{lehoucq1998arpack}. Step \#3 is common to most domain
-decomposition or multigrid preconditioners. However, in HPDDM, the special
-structure of the deflation matrix~$P$ which is sparse with dense blocks is
-exploited. Interested readers are referred to the paper describing the technical
-details of the implementation~\cite{jolivet2013scalable}. One important point to
-notice is that the size of the coarse operator $P^TAP$ is much lower
-than $n$, the size of $A$. Furthermore, coarsening is much more aggressive than
-algebraic or geometric multigrids, see~\cref{sec:linear-elasticity} for
-evidence. It is thus common to have the coarse operator distributed on
-fewer processes than for $A$, see for example the concept of telescoping later
-introduced in PETSc~\cite{may2016extreme}. In HPDDM, the coarse operator is
-computed and redistributed simultaneously. This could be technical to implement
-in PETSc as the result of the routine \pk{MatPtAP} is distributed on the same communicator as the one of the input matrix $A$.
+{\color{red} Should we mention the connections with the condition number here? and thus discuss the guaranteed convergence property of the method?}
+
+%When used without PETSc and thus SLEPc, step \#1 is solved in HPDDM with
+%ARPACK~\cite{lehoucq1998arpack}.
+While Step \#3 is common to most domain
+decomposition or multigrid preconditioners, HPDDM takes advantage of the special
+block sparse structure of the deflation matrix~$P$.
+Moreover,
+the size of the coarse operator $P^TAP$ is much lower
+than $n$, the size of $A$, and coarsening is much more aggressive than
+when using algebraic or geometric multigrids methods.
+%, see~\cref{sec:linear-elasticity} for
+%evidence.
+It is thus natural to remap the coarse operator on
+a subset of the processes used to decompose the original matrix $A$; in HPDDM, the coarse operator is
+computed and redistributed simultaneously.
+Interested readers are referred to the paper describing the technical
+details of the implementation~\cite{jolivet2013scalable}.
+We finally note here that similar techniques have been proven beneficial for other types
+of solvers in PETSc~\cite{may2016extreme}.
 
 % Of course, users do not have to implement the three previous steps themselves.
-The auxiliary operators $\{\mathring{A}_i\}_{i=1}^N$ have simply to be provided
-as well as the correspondence from local to global numbering,
-see~\cref{fig:poisson}, using the routine \pk{PCHPDDMSetAuxiliaryMat}.  If there
-is a \pk{DM} attached to the \pk{PC} of type \pk{DMPLEX}~\cite{knepley2009mesh}
-or \pk{DMP4EST}~\cite{BursteddeWilcoxGhattas11}, then these auxiliary
-overlapping operators and the correspondence are automatically assembled by
-PETSc for users, see~\cref{fig:p4est}. Other \pk{DMType} may be supported by
-implementing the proper \pk{DMCreateNeumannOverlap} callback, e.g., \pk{DMDA}
-support could be easily added upon request. With the local
-matrix pencils available, new options will be automatically registered such as
-\pk{-pc\_hpddm\_levels\_1\_eps\_nev} or \pk{-pc\_hpddm\_levels\_1\_eps\_threshold}, which
-may be used to set how many eigenpairs SLEPc should compute in step \#2.  The
-default \pk{-eps\_nev} value is 20, i.e., the dimension of the coarse space is
-$20\times N$ under the assumption that all concurrent SLEPc solves converge.
-It is usually best to keep the same number of eigenvectors on each process for multiple reasons:
-\begin{itemize}
-    \item with standard eigensolvers such as ARPACK or \pk{EPSKRYLOVSCHUR}, it is not possible to compute all eigenpairs in a given interval of $\mathbb{R}^+$.
-Instead, an upper bound of the number of eigenpairs has to be provided, and then it is possible in \pk{PCHPDDM} to remove eigenvectors associated to eigenvalues above the given \pk{-eps\_threshold}; extra eigenvectors are thus computed, but immediately thrown away;
-    \item if the number of eigenvectors is the same, coarse operators take advantage of block structures and are assembled as \pk{MatMPISBAIJ} or \pk{MatMPIBAIJ} matrices, with usually large block sizes, by default 20; this can have serious memory and performance implications, see~\cref{fig:perfAX};
-    \item during solution phases, it is possible to use the functions \texttt{MPI\_Gather} and \texttt{MPI\_Scatter} to go from one to another level and vice versa, instead of the variable size alternatives \texttt{MPI\_Gatherv} and \texttt{MPI\_Scatterv}, which are more costly.
-\end{itemize}
-Another important option is \pk{-pc\_hpddm\_coarse\_p}, which sets the size of the
-communicator on which the coarse operator should be distributed. The default
-value is 1, i.e., the matrix $P^TAP$ is centralized on a single process.
-The action of both inverses from~\cref{eq:two} can be parameterized through the
-\pk{-pc\_hpddm\_coarse\_ksp\_type} and \pk{-pc\_hpddm\_levels\_1\_ksp\_type} options
-respectively.
+In order to use GenEO from PETSc, the auxiliary operators $\{\mathring{A}_i\}_{i=1}^N$
+and the correspondending local to global map of degrees of freedom (see~\cref{fig:poisson})
+can be provided using the routine \pk{PCHPDDMSetAuxiliaryMat}.
+In special cases where the matrix has been constructed using the PETSc discretization
+instrastucture, the relevant information is automatically computed
+by PETSc. In \cref{fig:p4est} we provide an example of non-overlapping (left) and overlapping (right)
+decomposition using a non-conforming grid with \pk{DMPLEX}~\cite{knepley2009mesh}. The
+mesh has been constructed internally by \pk{DMPLEX} using a random pattern of non-conforming 2:1 refinements
+and a space-filling curve distribution by p4est~\cite{BursteddeWilcoxGhattas11}.
+The required elements needed to properly assemble the Neumann operators in a conforming way are pictured with a lighter color.
+Other \pk{DMType}s, including user-defined, may be supported by
+implementing the proper \pk{DMCreateNeumannOverlap} callback; support for \pk{DMDA}
+could be easily added upon request.
 \begin{figure}[htbp]
@@
 
-    \subsubsection{Coarse corrections and eigenvalue problems}
+With the local
+matrix pencils available, new options will be automatically registered such as
+\pk{-pc\_hpddm\_levels\_1\_eps\_nev} or \pk{-pc\_hpddm\_levels\_1\_eps\_threshold}, which
+may be used to customize the eigenpairs SLEPc will compute in step \#2.  The
+default \pk{-eps\_nev} value is 20, i.e., the dimension of the coarse space is
+$20\times N$ under the assumption that all concurrent SLEPc solves converge.
+
+For performance reasons, it is usually better to compute the same number of eigenvectors
+on each process, since in this case the coarse operator will be assembled using the \pk{MatMPISBAIJ} (or \pk{MatMPIBAIJ}) matrix format,
+with usually large block sizes; this can have serious memory and performance implications, see~\cref{fig:perfAX}. During solution phases,
+we can use \texttt{MPI\_Gather} and \texttt{MPI\_Scatter} operations to move between levels,
+instead of the variable size alternatives \texttt{MPI\_Gatherv} and \texttt{MPI\_Scatterv} that could exhibit worse performances.
+In addition, standard eigensolvers such as ARPACK~\cite{lehoucq1998arpack} or \pk{EPSKRYLOVSCHUR} from SLEPc,
+cannot compute all eigenpairs in a given interval of $\mathbb{R}^+$. Instead, an upper bound of the number of eigenpairs has to be provided.
+
+The option \pk{-pc\_hpddm\_coarse\_p} can be used to provide the size of the
+subcommunicator that will be used to remap the coarse operator. With the default
+value of 1, the coarse matrix $P^TAP$ is centralized on a single process.
+The action of the inverses in~\cref{eq:two} can be parameterized through the
+\pk{-pc\_hpddm\_coarse\_ksp\_type} and \pk{-pc\_hpddm\_levels\_1\_ksp\_type} options
+respectively.
+
+    \subsubsection{Coarse corrections customization}
 Some other parameters may be adjusted to define how the one-level preconditioner
-and the coarse grid operator interact with each other. In fact, the additive
-correction from~\cref{eq:two} is the simplest to write down, but most often not
-very efficient numerically~\cite{tang2009comparison}, and thus not used as
-default in \pk{PCHPDDM}. This observation is not specific to domain
-decomposition methods and also applies to multigrid~\cite{bastian1998additive}.
-Users can choose between $M'^{-1}_{\text{additive}}$ and:
+and the coarse grid operator interact. Besided simple the additive
+correction shown in~\cref{eq:two}, users can choose more numerically efficient
+corrections as:~\cite{tang2009comparison,bastian1998additive}
 \begin{align*}
@@
 options \pk{-pc\_hpddm\_coarse\_correction} \pk{(deflated$|$additive$|$balanced)} or the
-routine \pk{PCHPDDMSetCoarseCorrectionType}(pc, type) and its Fortran equivalent.
-
-The local generalized eigenvalue problems~\cref{eq:gevp} solved by SLEPc can also be adjusted.
-For example, in addition to the local unassembled Neumann matrices~$\{\mathring{A}_i\}_{i=1}^N$,
-if one has access to local matrices with optimized
-boundary conditions $\{B_i\}_{i=1}^N$, e.g., Robin transmission conditions, it
-is possible to pass these using the routine \pk{PCHPDDMSetRHSMat}(pc, $B_i$).
-The following generalized eigenvalue problems, called GenEO-2 in the
+routine \pk{PCHPDDMSetCoarseCorrectionType}.
+
+    \subsubsection{Extension of the GenEO framework}
+The local generalized eigenvalue problems~\cref{eq:gevp} can also be adjusted.
+In addition to the local unassembled Neumann matrices~$\{\mathring{A}_i\}_{i=1}^N$,
+users can prescribe additional local operators defined on the overlapped subdomain and
+with optimized, Robin
+boundary conditions $\{B_i\}_{i=1}^N$ via
+the routine \pk{PCHPDDMSetRHSMat}(pc, $B_i$).
+In such a case, the following generalized eigenvalue problems, called GenEO-2 in the
 literature~\cite{haferssas2017soras}, are then solved in each subdomain:
@@
 \end{equation*}
+{\color{red} One sentence or two on why this is needed? what are the cases where this is used in practice?}
+
     \subsubsection{Multilevel extension}\label{sec:ml}
@@
 \pk{PCHPDDM} in order to build robust two-level overlapping domain decomposition preconditioners.
-A multilevel extension of the GenEO framework is also
-possible~\cite{aldaas2019multi}. It can be used without any additional
-information provided by users. Since the dimension of the
-coarse operator of the second level grows with the number of subdomains of the
-fine level, switching to a multilevel scheme may alleviate the increasing cost
-of solving coarse systems. \pk{PCHPDDM} follows the same numbering of
-\pk{PCBDDC}%\footnote{which itself follows the opposite numbering of \pk{PCMG}}
-: the finest level is always numbered 1, the level index increases, up until
-the coarsest level whose options are always prefixed by \pk{coarse\_}. In order to
+Since the dimension of the
+coarse operator linearly depends on the number of subdomains of the
+fine level decomposition, switching to a multilevel scheme may alleviate the increasing cost
+of solving the coarse linear system. A multilevel extension of the GenEO framework has been recently
+proposed~\cite{aldaas2019multi}, with the advantage that the multilevel hierarchy can be
+automatically constructed without any additional information from the user.
+
+\pk{PCHPDDM} follows the same numbering of \pk{PCBDDC}:
+the finest level is always numbered with 1, and the level index increases as traversing the hierarchy, up until
+the coarsest level, whose solver options are prefixed by \pk{coarse\_}. In order to
 register an additional level $l' = l+1$, the following conditions must be met at
@@
     \item there must be more than one subdomain;
-    \item at least one local eigenvector must be computed.
+    \item at least one local eigenvector must be computed per coarse subdomain.
 \end{itemize}
-Here are some examples with $N = 16$ subdomains on the finest level:
-\begin{itemize}
-    \item \pk{-pc\_hpddm\_levels\_1\_eps\_threshold 0.4 -pc\_hpddm\_coarse\_p 8} would
-        define a two-level method with the coarse operator being distributed
-        among 8 processes, assuming that there is globally enough $\lambda_i$ in~\cref{eq:gevp} lower than 0.4;
-    \item \pk{-pc\_hpddm\_levels\_1\_eps\_nev 10 -pc\_hpddm\_levels\_2\_p 8} \linebreak
-        \pk{-pc\_hpddm\_levels\_2\_eps\_nev 10 -pc\_hpddm\_coarse\_p 2} would define
-        a three-level method with the second level (resp.~ coarse operator)
-        being distributed among 8 (resp.~2) processes;
-    \item \pk{-pc\_hpddm\_levels\_1\_eps\_nev 10 -pc\_hpddm\_levels\_2\_p 8} is not
-        valid because there are only two levels, so the options \pk{levels\_2\_} are
-        not registered and coarse\_ has to be used instead;
-    \item \pk{-pc\_hpddm\_levels\_1\_eps\_nev 10 -pc\_hpddm\_levels\_2\_eps\_nev 10} \linebreak \pk{-pc\_hpddm\_coarse\_p 2}
-        is not valid because there is no option \pk{levels\_2\_p}, so the default
-        value of 1 is used for the second level, and it is thus not possible to
-        define a third level with a single second-level subdomain.
-\end{itemize}
-    \subsubsection{MatIS}
+For example, using $N = 16$ subdomains on the finest level, the options
+
+\pk{-pc\_hpddm\_levels\_1\_eps\_threshold 0.4 -pc\_hpddm\_coarse\_p 8}
+
+\noindent will define a two-level method with the coarse operator being distributed
+among 8 processes, assuming that there is globally enough $\lambda_i$ in~\cref{eq:gevp} smaller than the prescribed threshold 0.4.
+Similarly, the options
+
+\pk{-pc\_hpddm\_levels\_1\_eps\_nev 10 -pc\_hpddm\_levels\_2\_p 8}
+
+\pk{-pc\_hpddm\_levels\_2\_eps\_nev 10 -pc\_hpddm\_coarse\_p 2}
+
+\noindent will define
+a three-level method with the second level distributed among 8 processes, while the coarsest level
+will be built from 10 eigenvalues from the second level and remapped onto 2 processes.
+Additional examples for the solver customization are provided in~\cref{sec:pchpddm-examples}
+%    \item \pk{-pc\_hpddm\_levels\_1\_eps\_nev 10 -pc\_hpddm\_levels\_2\_p 8} is not
+%        valid because there are only two levels, so the options \pk{levels\_2\_} are
+%        not registered and coarse\_ has to be used instead;
+%    \item \pk{-pc\_hpddm\_levels\_1\_eps\_nev 10 -pc\_hpddm\_levels\_2\_eps\_nev 10} \linebreak \pk{-pc\_hpddm\_coarse\_p 2}
+%        is not valid because there is no option \pk{levels\_2\_p}, so the default
+%        value of 1 is used for the second level, and it is thus not possible to
+%        define a third level with a single second-level subdomain.
+%\end{itemize}
+
+    \subsubsection{PCHPDDM for non-overlapping domain decomposition}
 Because of the intrinsic nature of balancing domain
@@
 overlap, see for example $\Omega_1$ and $\Omega_2$ from~\cref{fig:poisson} in a
-finite element context. Using a local to global mapping and \pk{MatConvert}, it
-is possible to reconstruct the local assembled matrices, equivalent to the
+finite element context. Since the assembled form of these operators can be reconstructed
+it is possible to construct the Dirichlet operators, e.g.
 $\{R_i A R_i^T\}_{i=1}^N$ from~\cref{sec:interface-pc}. All tools needed by the
@@
 overlapping Schwarz methods are known for converging slowly when there is in
-fact no overlap. It is common in the field of BDD to see the coarse problem as
-an unassembled operator as well. In fact, this is mandatory to define efficient
-multilevel variants, as implemented in \pk{PCBDDC}. In~\cref{sec:bddc}, it will
-be shown how a coarse problem obtained in \pk{PCBDDC} may be solved using
-\pk{PCHPDDM}.
-  \subsection{Applications and numerical results}
+fact no overlap. {\color{red} add one sentence with reference to Nicole's work on GenEO for BDD?}
+%It is common in the field of BDD to see the coarse problem as
+%an unassembled operator as well. In fact, this is mandatory to define efficient
+%multilevel variants, as implemented in \pk{PCBDDC}. In~\cref{sec:bddc}, it will
+%be shown how a coarse problem obtained in \pk{PCBDDC} may be solved using
+%\pk{PCHPDDM}.
+  \subsection{Applications and numerical results}\label{sec:pchpddm-examples}
     \subsubsection{System of elasticity}\label{sec:linear-elasticity}
-Considering only small displacements, this linear problem with highly heterogeneous elastic moduli is solved in 3D.
+We first report on solving the three-dimensional system of linear elasticy with highly heterogeneous elastic modulus.
 Its strong formulation is given by:
@@
 \Cref{eq:elasticity} is discretized using
-$(\mathbb{P}_1, \mathbb{P}_1, \mathbb{P}_1)$ finite elements.
+$(\mathbb{P}_1, \mathbb{P}_1, \mathbb{P}_1)$ finite elements resulting in $\pgfmathprintnumber{593} \times 10^6$ number of uknowns
+with approximately $\pgfmathprintnumber{45}$ nonzero coefficients per row.
 The physical domain $\Omega$ is decomposed in $13{,}824$ subdomains using the automatic graph partitioner ParMETIS~\cite{karypis1998fast}.
-The number of unknowns is $\pgfmathprintnumber{593} \times 10^6$, with approximately $\pgfmathprintnumber{45}$ nonzero coefficients per row.
 There are heterogeneities due to jumps in $E$ and $\nu$.
-The following discontinuous piecewise constant values are considered: $(E_1, \nu_1) = (2\times 10^{11}, 0.25)$, $(E_2, \nu_2) = (10^{7}, 0.45)$, see~\cref{fig:elas}.
+The following discontinuous piecewise constant values are considered: $(E_1, \nu_1) = (2\times 10^{11}, 0.25)$ in blue regions, while $(E_2, \nu_2) = (10^{7}, 0.45)$ in red regions, see~\cref{fig:elas}.
+{\color{red} Pierre, check the blue/red here and in the figure caption}
+The code used to produce these results was borrowed from
+the original paper about the multilevel extension of
+GenEO~\cite{aldaas2019multi} and it is available
+at~\url{https://github.com/prj-/aldaas2019multi}.
 \begin{figure}[htbp]
@@
 \end{figure}
+
 The following solvers are compared:
@@
     \item \pk{PCHPDDM} with an exact coarse solver;\label{enum:exact}
-    \item \pk{PCHPDDM} with an inexact coarse solver using \pk{PCHYPRE};\label{enum:inexacthypre}
-    \item \pk{PCHPDDM} with inexact coarse solvers using GenEO multilevel extension, cf.~\cref{sec:ml}.\label{enum:inexact}
+    \item \pk{PCHPDDM} with an inexact coarse solver using \pk{PCHYPRE};\label{enum:inexacthypre} {\color{red} I think we should not include this case}
+    \item \pk{PCHPDDM} with inexact coarse solvers using GenEO multilevel extension.\label{enum:inexact} %, cf.~\cref{sec:ml}.\label{enum:inexact}
 \end{enumerate}
-The options \pk{-ksp\_gmres\_modifiedgramschmidt
+As a linear solver, we use flexible GMRES, customized with the options \pk{-ksp\_gmres\_modifiedgramschmidt
 -ksp\_gmres\_restart 50
--ksp\_type fgmres} are used in conjunction with all \pk{PC}. The options specific to each solver described above are given next. \\[-4pt]
+-ksp\_type fgmres}. The options specific to each solver described above are given below. \\[-4pt]
 \begin{minipage}[t]{0.34\textwidth}
@@
 \end{minipage} \\
-By default, the coarse problem in \pk{PCHPDDM} is solved using an exact $LU$ or Cholesky factorization, in this case using 24 processes and MKL CPARDISO.
+By default, the coarse problem in \pk{PCHPDDM} is solved using an exact $LU$ or Cholesky factorization, in this case using 24 processes and \pk{MKL\_CPARDISO}.
 Switching to an inexact solver using \pk{PCHYPRE} (resp.~GenEO multilevel extension), cf.~\cref{enum:inexacthypre} (resp.~\cref{enum:inexact}), is straightforward.
@@
 \end{minipage} \\
-$M$ is the number of subdomains used to define the second-level domain decomposition, which is an aggregation of first-level subdomain.
+For the multivel extension of GenEO, $M$ is the number of subdomains used to define the second-level domain decomposition, which is an aggregation of first-level subdomain.
 When it comes to convergence, first note that BoomerAMG from \emph{hypre} has a really hard time reaching the stopping criterion of $10^{-2}$ when solving GenEO coarse problems.
@@
 
-Second, when properly parametrized, \cref{enum:exact} and \cref{enum:inexact} yield the same number of outer iterates.
+Considering the parametrization used for the \pk{PCHPDDM} solvers, \cref{enum:exact} and \cref{enum:inexact} yield the same number of outer iterates.
 In~\cref{fig:gamg-geneo}, the convergence history of \pk{PCGAMG} and \pk{PCHPDDM} are reported.
@@
 Note that in this test case, \pk{PCGAMG} is faster than all \pk{PCHPDDM}
-alternatives. Its setup (resp.~solution) phase takes 70 (resp.~19) seconds while
-for the best parameters of \pk{PCHPDDM}, which is
-configuration~\ref{pgfplots:p-256} with 256 second-level subdomains, its
-respectively 64 and 36 seconds. Out of these 64 seconds, 31 are spent in SLEPc
+alternatives. For \pk{PCGAMG}, the setup phase takes 70 seconds, while the solution phase requires 19 seconds;
+on the other hand, the fastest \pk{PCHPDDM} configuration, which corresponds to
+configuration~\ref{pgfplots:p-256} with 256 second-level subdomains, requires 64 seconds to setup and 31 for the solution of the linear system.
+Out of the 64 seconds needed for \pk{PCHPDDM} setup, 31 are spent in SLEPc
 \pk{EPSSolve} for solving GenEO at each level but the coarsest. With respect to
 coarsening, \pk{PCGAMG} has an operator complexity of 1.501, while for \pk{PCHPDDM},
-it is 1.015. The code used to produce these results was borrowed from
-the original paper about the multilevel extension of
-GenEO~\cite{aldaas2019multi} and is available
-at~\url{https://github.com/prj-/aldaas2019multi}.
+it is 1.015.
     \subsubsection{Liouville--Bratu--Gelfand equation} The strong formulation of this nonlinear problem is given by:
@@
     \subsubsection{Using PCHPDDM as a coarse grid solver for PCBDDC\label{sec:bddc}}
-A last toy problem is explained to show how one may switch between a multilevel
+A last toy problem is here proposed to show how one may switch between a multilevel
 BDDC solver to a two-level BDDC solver using \pk{PCHPDDM} for solving the
@@
 \end{align}
-It is discretized using order 1 N\'ed\'elec finite elements. There is a command
-line option -{}-nonoverlapping to activate \pk{PCBDDC}. The following command line
-is thus used. \\[-4pt]
+It is discretized using first order N\'ed\'elec finite elements and the \pk{PCBDDC} solver
+is automatically customized to obtain a stable method with these elements \cite{zampinihcurl2}.
+The following command line options are used. \\[-4pt]
 \begin{Verbatim}[fontsize=\footnotesize,frame=single,framerule=0.1mm,commandchars=&\[\]]
@@
  -dirichlet_pc_factor_mat_solver_type mumps
-&fvtextcolor[mygreen][# continue on the left column]
+&fvtextcolor[mygreen][# continue on the right column]
 \end{Verbatim}
@@
 \begin{Verbatim}[fontsize=\footnotesize,frame=single,framerule=0.1mm,commandchars=&\[\]]
-&fvtextcolor[mygreen][# continued from the right column]
+&fvtextcolor[mygreen][# continued from the left column]
  -prefix_push coarse_
@@
 Deluxe scaling is used to build an adaptive second level with \pk{PCBDDC}. On this
-second level, new subdomains are built by aggregating the constraints
-of 8 fine-level subdomains. The BDDC coarse problem composed of $\frac{512}{8} =
+second level, new subdomains are built by aggregating the coarse element matrices of
+8 fine-level subdomains. The BDDC coarse problem composed of $\frac{512}{8} =
 64$ subdomains is then solved using \pk{PCHPDDM}, which itself builds another adaptive
@@
 
+\bibitem{zampiniiga1}
+{\sc L.~Beirao da Veiga, L.~F. Pavarino, S.~Scacchi, O.~B. Widlund and  S.~Zampini},
+  {\em Isogeometric {BDDC} preconditioners with deluxe scaling},
+  SIAM Journal on Scientific Computing,
+  36 (2014), pp.~A1118--A1139.
+
+\bibitem{zampiniiga2}
+{\sc L.~Beirao da Veiga, L.~F. Pavarino, S.~Scacchi, O.~B. Widlund and  S.~Zampini},
+  {\em Adaptive selection of primal constraints for isogeometric {BDDC} deluxe preconditioners},
+  SIAM Journal on Scientific Computing, 39 (2017), pp.~A281--A302.
+
 \bibitem{Benzi2011a}
@@
 
-\bibitem{dohrmann2016bddc}
-{\sc C.~R. Dohrmann and O.~B. Widlund}, {\em {A BDDC Algorithm with Deluxe
-  Scaling for Three-Dimensional $H(\textnormal{curl})$ Problems}},
-  Communications on Pure and Applied Mathematics, 69 (2016), pp.~745--770.
+%\bibitem{dohrmann2016bddc}
+%{\sc C.~R. Dohrmann and O.~B. Widlund}, {\em {A BDDC Algorithm with Deluxe
+%  Scaling for Three-Dimensional $H(\textnormal{curl})$ Problems}},
+%  Communications on Pure and Applied Mathematics, 69 (2016), pp.~745--770.
 
@@
 
+\bibitem{tarekmathewbook}
+{\sc T.~Mathew},
+  {\em Domain decomposition methods for the numerical solution of partial differential equations},
+  Springer Science \& Business Media, 2008.
+
 \bibitem{may2016extreme}
@@
 
+\bibitem{zampinihdiv}
+{\sc D.~S. Oh, O.~B. Widlund, S.~Zampini and C.~Dohrmann},
+  {\em {BDDC} algorithms with deluxe scaling and adaptive selection of primal constraints for {R}aviart-{T}homas vector fields},
+  Mathematics of Computation, 87 (2018), pp.~659--692.
+
 \bibitem{o1980block}
@@
   on Scientific Computing, 28 (2006), pp.~1651--1674.
+
+\bibitem{zampiniiga3}
+{\sc L.~F. Pavarino, S.~Scacchi, O.~B. Widlund and  S.~Zampini},
+  {\em Isogeometric {BDDC} deluxe preconditioners for linear elasticity},
+  Mathematical Models and Methods in Applied Sciences, 28 (2018), pp.~1337--1370.
 
@@
   multigrid methods}, Journal of Scientific Computing, 39 (2009), pp.~340--370.
+
+\bibitem{olofbook}
+{\sc A.~Toselli and O.~B. Widlund},
+  {\em Domain decomposition methods: algorithms and theory},
+  Springer Science \& Business Media, 2006.
 
@@
 \bibitem{zampini2016pcbddc}
-{\sc S.~Zampini}, {\em {PCBDDC: a class of robust dual--primal methods in
+{\sc S.~Zampini}, {\em {{PCBDDC}: a class of robust dual--primal methods in
   PETSc}}, SIAM Journal on Scientific Computing, 38 (2016), pp.~S282--S306,
@@
 
+\bibitem{zampinihcurl1}
+{\sc S.~Zampini},
+  {\em Adaptive {BDDC} deluxe methods for {H}(curl)},
+  Domain Decomposition Methods in Science and Engineering XXIII, 2017, pp.~285--292.
+
+\bibitem{zampinidarcy}
+{\sc S.~Zampini and X.~Tu},
+  {\em Multilevel Balancing Domain Decomposition by Constraints Deluxe Algorithms with Adaptive Coarse Spaces for Flow in Porous Media},
+  SIAM Journal on Scientific Computing, 39 (2107), pp.~A1389--A1415.
+
+\bibitem{zampinihcurl2}
+{\sc S.~Zampini, P.~Vassilevski, V.~Dobrev and T.~Kolev},
+  {\em Balancing Domain Decomposition by Constraints Algorithms for Curl-Conforming Spaces of Arbitrary Order},
+  Domain Decomposition Methods in Science and Engineering XXIV, (2019).
+
 \end{thebibliography}

### 2020-07-08 article.tex
Update with LOBPCG + multithread results

@@
 \hypersetup{
-  pdftitle={KSPHPDDM and PCHPDDM: extending PETSc with robust Overlapping Schwarz Methods and advanced Krylov methods},
+  pdftitle={KSPHPDDM and PCHPDDM: Extending PETSc with Robust Overlapping Schwarz Preconditioners and Advanced Krylov Methods},
   pdfauthor={P. Jolivet and S. Zampini}
@@
 \newcommand*{\fvtextcolor}[2]{\textcolor{#1}{#2}}
+\usepackage{colortbl,booktabs}
 \definecolor{last}{RGB}{44,123,182}
@@
     libraries, PETSc and HPDDM, have been interfaced together in order to offer
-    end-users robust overlapping Schwarz methods and advanced Krylov methods featurin
-    recycling and block solution techniques.
+    end-users robust overlapping Schwarz preconditioners and advanced Krylov
+    methods featuring recycling and block solution techniques.
     The flexibility of the implementation %, which for example may deal with various matrix formats,
@@
 convinced computational scientists to use PETSc as one of the discretization and/or algebraic backend in
-many different higher-level projects {\color{red} ADD CIT TO DEAL.II, FIREDRAKE, MOOSE, HIPPYLIB, MFEM}.
-
-On the other hand, the High-Performance unified framework for Domain Decomposition Methods (HPDDM)
+many different higher-level projects, see \url{https://www.mcs.anl.gov/petsc} for a comprehensive list.
+
+On the other hand, HPDDM~\cite{jolivet2013scalable}, the High-Performance unified framework for Domain Decomposition Methods (HPDDM)
 is a much smaller project focusing on robust and scalable domain decomposition
-preconditioners~\cite{jolivet2013scalable} and advanced iterative
+preconditioners and advanced iterative
 methods~\cite{jolivet2016block} which can efficiently deal with linear
-systems with multiple right-hand sides (also known and as block iterative methods)
+systems with multiple right-hand sides, also known and as block iterative methods,
 and when there is a recurrence of varying coefficient matrices and right-hand
-sides (recycling iterative methods). For an introduction to domain decomposition methods, see
-~\cite{smith2004domain,dolean2014ddm,olofbook,tarekmathewbook} {\color{red} FIX CITATIONS}.
+sides, recycling iterative methods.
 
 % Htool ?
-Here we describe the integration of the entire HPDDM suite, its advanced iterative
+Here, we describe the integration of the entire HPDDM suite, its advanced iterative
 methods and its domain decomposition preconditioners, into PETSc.
-We showcase the design principles of the interface considering % alongside
-minimalist and easy-to-run examples, equipped with the necessary command line options to
-fullfill reproducibility requirements {\color{red} FIND GOOD CIT, ADD 1 more sentences?,
-https://www.acm.org/publications/policies/artifact-review-badging} and ease the
-integration of these algorithms into more advanced frameworks.
+The design principles of the interface are showcased considering % alongside
+minimalist and easy-to-run examples. Necessary command line
+options are provided to:
+\begin{itemize}
+    \item fulfill reproducibility requirements as advocated by the
+Association for Computing Machinery
+\url{https://www.acm.org/publications/policies/artifact-review-badging};
+    \item ease the integration of these algorithms into more advanced frameworks;
+    \item analyze performance results~\cite{hoefler2015scientific}.
+\end{itemize}
 
@@
 Krylov subspace methods are widely used in numerical linear algebra for solving
-linear systems of equations because of their memory efficiency~\cite{templateBook}.
+linear systems of equations because of their memory efficiency~\cite{barrett1994templates}.
 They mainly rely on matrix--vector
@@
 \pk{KSPLGMRES} and \pk{KSPDGMRES}, are available. However, neither handle variable
-preconditioning, and the DGMRES does not support complex arithmetic. {\color{red} Also note
+preconditioning, and \pk{KSPDGMRES} does not support complex arithmetic. {\color{red} Also note
 that when solving a sequence of linear systems $A_i x_i = b_i, i = 1, 2,
@@
 somehow more involved kernels such as matrix--matrix multiplication, instead of
-matrix--vector, and may involve different inner-products realizations~\cite{Frommer}.
+matrix--vector, and may involve different inner-product realizations~\cite{frommer2017block}.
 Still, systems with multiple right-hand sides are ubiquitous,
@@
 However,
-{\color{red} REMOVE? It is nevertheless possible to solve
-systems with multiple right-hand sides using a specialized type of matrices,
-but with a very limited choice of preconditioners, cf.~\cref{sec:matkaij}.}
 
 Trilinos~\cite{heroux2005overview} is another well-known software for scientific
-computing and it provides iterative methods through its Belos package~\cite{bavier2012amesos2},
+computing and it provides iterative methods through its Belos package~\cite{bavier2012amesos2}.
 Although having PETSc and
@@
 operations like matrix--vector
-products to achieve higher arithmetic intensity ({\color{red} which method does this?}), or to decrease the
-number of global synchronizations needed for scalar products ({\color{red} which method does this?}).
-HPDDM Krylov methods may be selected {\color{red} TODO API FOR KSPHPDDMSetType? KSPHPDDMSetVariant?} using the
-options \pk{-ksp\_hpddm\_type} \pk{(gmres$|$cg$|$gcrodr$|$bgmres$|$bcg$|$bfbcg$|$bgcrodr$|$preonly)}.
+products to achieve higher arithmetic intensity, or to decrease the
+number of global synchronizations needed for scalar products.
+HPDDM Krylov methods may be selected {\color{red} TODO API FOR KSPHPDDMSetVariant?} using the
+options \pk{-ksp\_hpddm\_type}
+\pk{(gmres$|$cg$|$gcrodr$|$bgmres$|$bcg$|$bfbcg$|$bgcrodr$|$preonly)}, or the
+routine \pk{KSPHPDDMSetType}(ksp, \pk{KSPHPDDMType}).
 Preconditioning variants can be specified via
-\pk{-ksp\_hpddm\_variant} \pk{(left$|$right$|$flexible)} to select whether the preconditioner
-has to applied on the left, on the right, or if it cannot be represented as a linear operator.
-{\color{red} out of curiosity. how can you choose the side if felixble is specified?}
-It is also possible to set the preconditioning side through the more common PETSc option \pk{-ksp\_pc\_side (left$|$right)}.
-Note that the value of the latter option is useless for BCG or BFBCG, and HPDDM CG only handles left and
-variable preconditioning {\color{red} can you elaborate more?}.
-Finally, convergence monitoring with the \pk{KSPMonitor} interface is fully supported, as well as the specification
-of customized convergence testing via the \pk{KSPSetConvergenceTest} callback.
+\pk{-ksp\_hpddm\_variant} \pk{(left$|$right$|$flexible)} to select whether the
+preconditioner is applied on the left, on the right, or if it cannot be
+represented as a linear operator. In this latter case, the preconditioner is
+always applied on the right, except for FCG. HPDDM CG
+and its variants BCG and BFBCG only handle left-preconditioning.  It is also
+possible to set the preconditioning side through the more common PETSc option
+\pk{-ksp\_pc\_side (left$|$right)}.  Finally, convergence monitoring with the
+\pk{KSPMonitor} interface is fully supported, as well as the specification of
+customized convergence testing via the \pk{KSPSetConvergenceTest} callback.
 
@@
     \item \pk{MatMult}($A$, $x$, $y$) for $y = Ax$;
-    \item \pk{MatMatMult}($A$, $X$, {\scriptsize \pk{MAT\_REUSE\_MATRIX}}, {\scriptsize
+    \item \pk{MatMatMult}($A$, $X$, {\small \pk{MAT\_REUSE\_MATRIX}}, {\small
         \pk{PETSC\_DEFAULT}}, $Y$)\footnote{or \pk{MatProductNumeric}($Y$) with PETSc 3.14.0 and above} for $Y =
         AX$;
-    \item \pk{PCApply}($P_{A'}$, $y$, $x$) for $x = P_{A'}y$.
+    \item \pk{PCApply}($P_{A'}$, $y$, $x$) for $x = P_{A'}y$;
+    \item \pk{PCMatApply}($P_{A'}$, $Y$, $X$) for $X = P_{A'}Y$;
 \end{itemize}
-As of version 3.13.2, PETSc does not have a general routine for applying $P_{A'}$ to a
-tall-and-skinny dense matrix $Y$. In HPDDM, a specialized routine has been implemented
-to compute $X = P_{A'}Y$ if the preconditioner is of one of the following types:
+As of version 3.14.0, PETSc provides unified routines for applying $A$ or $P_{A'}$
+to a tall-and-skinny dense matrix $Y$. However, when no specialized implementation
+is available, PETSc will loop over the columns of $Y$ and successively apply $A$
+or $P_{A'}$ to get all columns of $X$. For matrix--matrix products, there are
+currently specialized implementations for the following \pk{MatType}:
 \begin{itemize}
-    \item \pk{PCLU} or \pk{PCCHOLESKY}: exact $LU$ or Cholesky factorization, i.e., $P_{A'} = {A'}^{-1}$;
-    \item \pk{PCILU} or \pk{PCICC}: incomplete $LU$ or Cholesky factorization;
-    \item \pk{PCBJACOBI}: block Jacobi with exact or incomplete subdomain solvers as above;
-    \item \pk{PCASM}: overlapping Schwarz methods with exact or incomplete subdomain solvers as above.
+    \item \pk{MatMPIAIJ};
+    \item \pk{MatShell};
+    \item \pk{MatNest};
+    \item \pk{MatSeqAIJ}, \pk{MatSeqBAIJ}, and \pk{MatSeqSBAIJ};
+    \item \pk{MatSeqAIJcuSPARSE}.
 \end{itemize}
-In the aforementioned cases, it is indeed possible to access the so-called factored matrix $F$
+The following preconditioners have specialized implementations for dealing
+efficiently with multiple vectors:
+\begin{enumerate}
+    \item \pk{PCNONE}: $P_{A'} = I$, identity matrix;
+    \item \pk{PCKSP}: embedded Krylov method inside $P_{A'}$;
+    \item \pk{PCMAT}: multiplication by $A'$, i.e., $P_{A'} = {A'}$;
+    \item \pk{PCSHELL}: user-defined preconditioner, e.g., matrix-free;
+    \item \pk{PCHARA}~\cite{10.1145/3232850};
+    \item \pk{PCHPDDM}: see~\cref{sec:PCHPDDM};\label{it:hpddm}
+    \item \pk{PCASM}: overlapping Schwarz methods;\label{it:asm}
+    \item \pk{PCBJACOBI}: block Jacobi;\label{it:bjacobi}
+    \item \pk{PCLU} or \pk{PCCHOLESKY}: exact $LU$ or Cholesky factorization,
+        i.e., $P_{A'} = {A'}^{-1}$;\label{it:lu}
+    \item \pk{PCILU} or \pk{PCICC}: incomplete $LU$ or Cholesky
+        factorization.\label{it:ilu}
+\end{enumerate}
+One appealing feature of domain decomposition-like preconditioners 
+\cref{it:hpddm,it:asm,it:bjacobi} is that they most often rely on exact or inexact
+factorizations~\cref{it:lu,it:ilu} as subdomain solvers.
+In these cases, it is possible to access the so-called local or distributed factored matrix $F$
 %via \pk{PCFactorGetMatrix} and then call \pk{MatMatSolve}($F$, $\tilde{Y}$, $\tilde{X}$).
-via \pk{PCFactorGetMatrix} and then call \pk{MatMatSolve}($F$, $Y$, $X$) to take advantage
-of the blocking solving strategy from the various factorization packages interfaced with PETSc
-such as MUMPS~\cite{amestoy2001fully}, SuiteSparse, MKL\_PARDISO and SUPERLU {\color{red} ADD REFERENCES}.
-For the types of preconditioners not implementing \pk{PCFactorGetMatrix},
-HPDDM will loop over the columns of $Y$ and successively apply $P_{A'}$ to get all columns of $X$.
-Future work will consider extending the various implementations of the PETSc
+via \pk{PCFactorGetMatrix}, and then call \pk{MatMatSolve}($F$, $Y$, $X$) to take advantage
+of blocked forward eliminitations and backward substitutions from the various factorization packages
+interfaced with PETSc. In the case of a sparse coefficient matrix $A'$,
+this strategy is possible with: MUMPS~\cite{amestoy2001fully},
+SuiteSparse~\cite{davis2004algorithm}, MKL
+PARDISO or CPARDISO~\cite{mkl-web-page}, and SuperLU~\cite{li2005overview} or
+SuperLU\_DIST~\cite{li2003superlu_dist}. For the the case of a dense coefficient
+matrix $A'$: ScaLAPACK and Elemental~\cite{poulson2013elemental}. 
+Future work will consider extending the various implementations of PETSc
 preconditioning infrastructure to increase the arithmetic intensity of the
@@
 
-The rest of the operations are then performed directly inside HPDDM {\color{red} what about VecAXPY, VecNorm etc?}.
+The rest of the operations are then performed directly inside HPDDM.
 Specifically, within Krylov methods using the Arnoldi process, or when recycling is requested,
-we need $QR$ factorizations of tall-and-skinny dense matrices to orthonormalize
+one has to compute $QR$ factorizations of tall-and-skinny dense matrices to orthonormalize
 candidate basis vectors. Such operations are performed using the CholQR algorithm~\cite{stathopoulos2002block}, or
@@
     \subsubsection{User-defined deflation space} Recycling Krylov methods try to
-extract convergence information by solving a small standard or a generalized dense eigenproblem
-at the end of each cycle or when convergence is reached and then reuse it appropriately for a
-subsequent solve~\cite{saad2000deflated,stathopoulos2002block,soodhalter2014krylov}.
-The routine \pk{KSPHPDDMGetDeflationSpace} can be used to inquire about the details of such deflation subspaces, stored as distributed
+extract convergence information by solving a small standard or a generalized
+dense eigenproblem at the end of each cycle or when convergence is reached, and
+then reuse it appropriately for subsequent
+solves~\cite{saad2000deflated,stathopoulos2002block,soodhalter2014krylov}.
+The routine \pk{KSPHPDDMGetDeflationSpace} can be used to retrieve such deflation subspaces, stored as distributed
 dense tall-and-skinny matrices. It is also possible to provide this information
-via \pk{KSPHPDDMSetDeflationSpace} in case end users already know a good initial
+via \pk{KSPHPDDMSetDeflationSpace} in case end-users already know a good initial
 deflation subspace.
 
-    \subsubsection{MatKAIJ and systems with multiple right-hand sides\label{sec:matkaij}}
-{\color{red} TODO: move KSPMatSolve discussion out, and clean up the section, add results for irk?}
-There is a wide range of available matrix types in PETSc. One of them is
-\pk{MatKAIJ}, for which it is assumed that an operator may be written as $I \otimes S
+%     \subsubsection{MatKAIJ and systems with multiple right-hand sides\label{sec:matkaij}}
+% {\color{red} TODO: move KSPMatSolve discussion out, and clean up the section, add results for irk?}
+% There is a wide range of available matrix types in PETSc. One of them is
+% \pk{MatKAIJ}, for which it is assumed that an operator may be written as $I \otimes S
+% + A \otimes T$, with $S$ and $T$ elements of $\mathbb{K}^{p\times q}$ and
+% $A$ a standard sparse matrix of dimension $n$. If $p=q=k$, $S = \lambda I$, and $T =
+% \mu I$, then solving $(I \otimes S + A \otimes T) x = b$ may be cast into a
+% system with multiple right-hand sides $(\lambda I + \mu A)X = B$ where $B$ is
+% the row-major representation of $b$, i.e., $b_{k \cdot i + j} = B_{i, j},
+% \forall (i, j) \in \llbracket 1;n\rrbracket \times \llbracket 1;k\rrbracket$.
+% On the one hand, due to the structure of the operator, it is challenging to
+% define appropriate preconditioners for $(I \otimes S + A \otimes T)$: only
+% \pk{PCPBJACOBI} is currently implemented for \pk{MatKAIJ}. On the other hand,
+% it may be simpler to efficiently approximate the inverse of $A$. In
+% \pk{KSPHPDDM}, the type of the operators is checked and if it is \pk{MatKAIJ},
+% the single right-hand side system is automatically cast into a system with
+% multiple right-hand sides instead. % The standard function \pk{KSPSolve} takes as input argument a \pk{KSP} and two \pk{Vec}.
+% A new function \pk{KSPMatSolve} was added in PETSc 3.14.0, which takes as input argument
+% a \pk{KSP} and two dense \pk{Mat}, for solving true
+% multiple right-hand side systems. If the input \pk{KSP} is not of type
+% \pk{KSPHPDDM}, the solution phase is done in a column by column fashion.
+% Otherwise, it is possible to use any \pk{KSPHPDDM} \mbox{(pseudo-)block} Krylov methods.
+% Furthermore, a function \pk{KSPSetMatSolveBlockSize} is also provided to decompose
+% a single large block of column vectors into multiple sub-blocks.
+% \pk{KSPMatSolve} is then called repeatedly until all sub-blocks are traversed.
+% This was inspired by MUMPS~\cite{amestoy2001fully} option
+% ICNTL(27)\footnote{\url{http://mumps.enseeiht.fr/doc/userguide_5.3.1.pdf}, section 6.1}.
+    \subsubsection{Systems with multiple right-hand sides\label{sec:kspmatsolve}}
+Along with \pk{PCMatApply} introduced in PETSc 3.14.0, a new function
+\pk{KSPMatSolve} was added for solving systems with multiple right-hand sides.
+It takes as input argument a \pk{KSP} and, just as for \pk{PCMatApply}, two
+dense \pk{Mat}. If the input \pk{KSP} is not of type \pk{KSPHPDDM}, the
+solution phase is done in a column by column fashion.  Otherwise, it is
+possible to use any \pk{KSPHPDDM} \mbox{(pseudo-)block} Krylov methods.
+Furthermore, a function \pk{KSPSetMatSolveBlockSize} is also provided to
+decompose a single large block of column vectors into multiple sub-blocks.
+\pk{KSPMatSolve} is then called repeatedly until all sub-blocks are traversed.
+This was inspired by MUMPS~\cite{amestoy2001fully} option
+ICNTL(27)\footnote{\url{http://mumps.enseeiht.fr/doc/userguide_5.3.3.pdf},
+section 6.1}.
+
+With these added functions, it is now possible to do code refactorization for
+applications that deal with multiple right-hand sides.  As a striking example,
+LOBPCG~\cite{knyazev2001toward} or CISS~\cite{sakurai2003projection}
+implementations in SLEPc~\cite{hernandez2005ssf}, the Scalable Library for
+Eigenvalue Problem computations built on top of PETSc, were previously not
+using blocking.  It has been rewritten for version 3.14.0 and the benefits of
+this new implementation is displayed~\cref{sec:lobpcg}.
+
+As a side note, when dealing with
+\pk{MatKAIJ} matrices, for which it is assumed that an operator may be written as $I \otimes S
 + A \otimes T$, with $S$ and $T$ elements of $\mathbb{K}^{p\times q}$ and
-$A$ a standard sparse matrix of dimension $n$. If $p=q=k$, $S = \lambda I$, and $T =
+$A$ a standard sparse matrix of dimension $n$, it is also possible to use block Krylov methods. 
+Indeed, if $p=q=k$, $S = \lambda I$, and $T =
 \mu I$, then solving $(I \otimes S + A \otimes T) x = b$ may be cast into a
@@
 it may be simpler to efficiently approximate the inverse of $A$. In
-\pk{KSPHPDDM}, the type of the operators is checked and if it is \pk{MatKAIJ},
-the single right-hand side system is automatically cast into a system with
-multiple right-hand sides instead. % The standard function \pk{KSPSolve} takes as input argument a \pk{KSP} and two \pk{Vec}.
-A new function \pk{KSPMatSolve} was added in PETSc 3.14.0, which takes as input argument
-a \pk{KSP} and two dense \pk{Mat}, for solving true
-multiple right-hand side systems. If the input \pk{KSP} is not of type
-\pk{KSPHPDDM}, the solution phase is done in a column by column fashion.
-Otherwise, it is possible to use any \pk{KSPHPDDM} \mbox{(pseudo-)block} Krylov methods.
-Furthermore, a function \pk{KSPSetMatSolveBlockSize} is also provided to decompose
-a single large block of column vectors into multiple sub-blocks.
-\pk{KSPMatSolve} is then called repeatedly until all sub-blocks are traversed.
-This was inspired by MUMPS~\cite{amestoy2001fully} option
-ICNTL(27)\footnote{\url{http://mumps.enseeiht.fr/doc/userguide_5.3.1.pdf}, section 6.1}.
+\pk{KSPHPDDM}, the type of the operators is checked: if it is \pk{MatKAIJ} and
+the previous condition on $p, q, S$, and $T$ is fulfilled, the single
+right-hand side system is automatically cast into a system with multiple
+right-hand sides instead.
+
   \subsection{Applications and numerical results}
@@
 We note here that the results obtained with
-\pk{PCICC} from PETSc (\ref{pgfplots:r3}) and ichol from MATLAB (\ref{pgfplots:r2}, which uses SuiteSparse) are not consistent.
+\pk{PCICC} from PETSc \ref{pgfplots:r3} and ichol from MATLAB \ref{pgfplots:r2}, which uses SuiteSparse, are not consistent.
 %Apart from
@@
     \subsubsection{Linear stability analysis}
-In this section we apply recycling Krylov methods in the context linear
+In this section, we apply recycling Krylov methods in the context linear
 stability analysis and consider the solution the following generalized eigenvalue problem:
@@
 
-The mini-app employed for the numerical results is built on top of FreeFEM~\cite{hecht2012new} and it is
-available at \url{https://github.com/prj-/moulin2019al}. As eigenvalue solver, we consider the Scalable Library for Eigenvalue Problem
-computations (SLEPc)~\cite{hernandez2005ssf}, which is built on top of PETSc,
-and thus reuse its linear solver infrastructure.
+The mini-app employed for the numerical results is built on top of
+FreeFEM~\cite{hecht2012new} and it is available at
+\url{https://github.com/prj-/moulin2019al}. As eigenvalue solver, it uses
+SLEPc, and thus reuses PETSc linear solver infrastructure.
 
 There are different methods to compute the eigenvalues
-of~\cref{eq:lsa} near a complex-valued shift $\sigma$; in this paragraph, emphasis will be put
-on Krylov--Schur methods~\cite{Stewart2002}, which rely on spectral tranformation steps
+of~\cref{eq:lsa} near a complex-valued shift $\sigma$. In this paragraph, emphasis will be put
+on Krylov--Schur methods~\cite{Stewart2002}, which rely on spectral transformations
 and on the solution of successive linear systems such as:
@@
 finding the 5 eigenpairs closest to the shift $\sigma = 10^{-6} + 0.6\mathfrak{i}$ for a
-flow past a cylinder at Reynolds 100. Results can be reproduced using the following four commands;
-we note here that the first two commands are merely used to generate the base flow $q_b$ with \pk{SNESSolve} using a
+flow past a cylinder at Reynolds 100. Results can be reproduced using the following four commands.
+We note here that the first two commands are merely used to generate the base flow $q_b$ with \pk{SNESSolve} using a
 continuation method on the Reynolds number. \\[-4pt]
@@
 Recycling of Krylov subspaces with \pk{KSPHPDDM} provides algorithmic speedup by lowering the number of iterations per inner solve,
-as reported in~\cref{fig:comparison-slepc}, from 46 with \pk{KSPFGMRES} to 25.
-A similar speedup can be observed in terms of runtime, with the time spent for the solution of the eigenvalue problem being reduced from \SI[round-mode = places,round-precision=1]{1509}{\sec} to \SI[round-mode = places,round-precision=1]{1052}{\sec}.
+as reported in~\cref{fig:comparison-slepc}, from 80 with \pk{KSPFGMRES} to 45.
+A similar speedup can be observed in terms of runtime, with the time spent for the solution of the eigenvalue problem being reduced from \SI[round-mode = places,round-precision=1,group-minimum-digits=3,group-separator = {,}]{1509}{\sec} to \SI[round-mode = places,round-precision=1,group-minimum-digits=3,group-separator = {,}]{1052}{\sec}.
 {\color{red} FIXME LATER: These timings, and all which follow except in~\cref{sec:primitives}, have been obtained on {Ir\`ene}, a system composed of \pgfmathprintnumber[assume math mode=true]{1656} nodes with two 24-core Intel Xeon Platinum 8168 clocked at \SI{2.7}{\giga\hertz}}.
@@
 }\loadedtableSLEPc
-    \begin{figure}[!b]
+    \begin{figure}[!t]
 \centering \begin{tikzpicture}
@@
     \subsubsection{Performance of primitives for block Krylov methods}\label{sec:primitives}
-We close this section discussing the readiness of PETSc (as of version 3.13.2)
-in delivering efficient implementations of the core matrix--matrix multiplication
-needed by block Krylov methods and by recycling Krylov methods when restart is performed.
+The readiness of PETSc, as of version 3.13.2, is now discussed
+when it comes to delivering efficient implementations of the core matrix--matrix multiplication
+needed by block Krylov methods and when a restart occurs in recycling Krylov methods.
 
@@
 storing two such matrices, one with local rows and columns, and another with local rows and ``nonlocal'' columns.
-To keep this study succinct, we only evaluate the performances of the sequential implementations, which can help
+To keep this study succinct, we only evaluate the performance of the sequential implementations, which can help
 in drawing conclusions for the intraprocess performance of the parallel formats.
-For this reason, from now on we drop the \pk{Seq} substring.
-The performances of the multiplication primitives of the three aforementioned types will also be compared to
+For this reason, from now on, the \pk{Seq} substring is dropped.
+The performance of the multiplication primitives of the three aforementioned types will also be compared to
 those obtained with the MKL inspector--executor sparse BLAS routines~\cite{mkl-web-page}, using a single OpenMP thread.
-We note that there is an ongoing effort to better integrate these routines in PETSc, but at the time of writing, direct calls to the MKL were used.
-
-To perform the benchmarking, we first discretize the Poisson equation on a cube with $\mathbb{P}_1$ finite elements
+There is an ongoing effort to better integrate these routines in PETSc, but at the time of writing, direct calls to the MKL were used.
+
+For benchmarking, we first discretize the Poisson equation on a cube with $\mathbb{P}_1$ finite elements
 and obtain a matrix $A$ of dimension one million. Though the numerical
@@
 discretizing partial differential equations. We then generate a random symmetric dense matrix $T$ of dimension $b$,
-for $b \in \{ 1,3,6 \}$, and we construct the matrix $\mathbb{A} = A \otimes T$ into all three formats described above, using
-a block size of $b$ for the block storage formats.
+for $b \in \{ 1,3,6 \}$, and the matrix $\mathbb{A} = A \otimes T$ is assembled into all three formats described above, using
+a block size of $b$ for the block formats.
 The performance of the \pk{MatProductNumeric} operation is evaluated for tall-and-skinny dense matrices with a varying number
 of columns $N \in \{2,8,16,32,64\}$. For the single column case $N=1$, the \pk{MatMult} operation is used.
-All results have been obtained using double precision arithmetic and 32-bit integers; we report the scaled efficiency (measured in GFLOP per second)
+All results have been obtained using double precision arithmetic and 32-bit integers. The scaled efficiency measured in GFLOP per second is reported
 with respect to the baseline \pk{MatAIJ} implementation with $N=1$ in~\cref{fig:perfAX}, where performances have been averaged over five consecutive
@@
 They can be reproduced by using the mini-app \url{http://jolivet.perso.enseeiht.fr/MatProduct.c} and any input \pk{MatAIJ} stored in binary format,
-here using the default name \pk{binaryoutput} {\color{red} Do we want to provide the binary of the matrix used?}.
+here using the default name \pk{binaryoutput}, available at \url{http://jolivet.perso.enseeiht.fr/binaryoutput}.
 \begin{Verbatim}[fontsize=\footnotesize,frame=single,framerule=0.1mm,commandchars=&\[\]]
@@
     \item with a block size of 1, top plot $b=1$ in~\cref{fig:perfAX}, using any type but \pk{MatAIJ}~\ref{pgfplots:aij1} is counterproductive when $N>1$, see~\ref{pgfplots:baij1} and~\ref{pgfplots:sbaij1}, even when using MKL, see~\ref{pgfplots:mklbaijseq1} and~\ref{pgfplots:mklsbaijseq1}.
-With 8 or more columns, the performance of \pk{MatAIJ} plateaus and the efficiency reaches approximately 200\%, while the perfomances of the MKL primitives stagnate at around 175\% see~\ref{pgfplots:aij1};
+With 8 or more columns, the performance of \pk{MatAIJ} plateaus and the efficiency reaches approximately 200\%, while the performance of the MKL primitives stagnate at around 175\% see~\ref{pgfplots:aij1};
 for the single column case, \pk{MatSBAIJ} and \pk{MatAIJMKL} deliver slightly better performances;
@@
 This yields rather disappointing performance, except for \pk{MatAIJ}, see~\ref{pgfplots:aij3} in middle plot $b=3$, and its 350\% efficiency, even with a moderate number of columns.
-For block formats \ref{pgfplots:baij3} and~\ref{pgfplots:sbaij3}, given the small value of $b$, it could be beneficial to switch to optimized libraries for small matrices, e.g., LIBXSMM~\cite{heinecke2016libxsmm}; in fact, MKL implementation for block formats is here clearly outperforming PETSc, see~\ref{pgfplots:mklbaijseq3} and~\ref{pgfplots:mklsbaijseq3};
+For block formats \ref{pgfplots:baij3} and~\ref{pgfplots:sbaij3}, given the small value of $b$, it could be beneficial to switch to optimized libraries for small matrices, e.g., LIBXSMM~\cite{heinecke2016libxsmm}.
+In fact, MKL implementation for block formats is here clearly outperforming PETSc, see~\ref{pgfplots:mklbaijseq3} and~\ref{pgfplots:mklsbaijseq3};
     \item for block sizes larger than 5, block entry multiplication with PETSc block formats is performed using \texttt{?gemv} or \texttt{?gemm} operations.
@@
 
-We then reuse the mini-app to benchmark intranode performance of matrix--matrix multiplications, either with multiple OpenMP threads, or by offloading the operation to a GPU and use the cuSPARSE implementation~\cite{cusparse-web-page}.
+The mini-app is then used to benchmark intranode performance of matrix--matrix multiplications, either with multiple OpenMP threads, or by offloading the operation to a GPU and use cuSPARSE~\cite{cusparse-web-page} as interfaced in PETSc.
 \begin{Verbatim}[fontsize=\footnotesize,frame=single,framerule=0.1mm,commandchars=&\[\]]
@@
 \end{Verbatim}
-Results reported in~\cref{fig:perfAXpar} have been scaled with respect to the baseline implementation \pk{MatAIJMKL} with $N=1$.
-They have been obtained on a single node of {Jean Zay}, a system composed of \pgfmathprintnumber[assume math mode=true]{261} nodes with two 20-core Intel Xeon Gold 6248 clocked at \SI{2.5}{\giga\hertz} and four NVIDIA Tesla V100 SXM2.
-Neither transfers between host and device nor time spent in \texttt{mkl\_sparse\_optimize} have been taken into account.
-{\color{red} FEW MORE SENTENCES, percentage of peak, wait for results with different number of threads}
+Results reported in~\cref{fig:perfAXpar} have been scaled with respect to the
+baseline implementation \pk{MatAIJMKL} with $N=1$. They have been obtained on a
+single node of {Jean Zay}, a system composed of \pgfmathprintnumber[assume math
+mode=true]{261} nodes with two 20-core Intel Xeon Gold 6248 clocked at
+\SI{2.5}{\giga\hertz} and four NVIDIA Tesla V100 SXM2. Neither transfers
+between host and device nor time spent in \texttt{mkl\_sparse\_optimize} have
+been taken into account. For the sake of completeness, the performance of
+\pk{MatAIJMKL} with $N=1$ but different numbers of threads is also reported
+\cref{fig:scaling-Ax}. Clearly, it is not advised to use the full socket to
+perform this type of workload for such small sparse matrices and skinny dense
+matrices. But this helps in drawing a fair comparison between a full CPU socket
+and a full GPU device. All in all, the CPU configuration which reaches the
+highest percentage of peak is MatSBAIJMKL~\ref{pgfplots:sbaijmkl3} with $b=3$
+and $N=64$, right-most middle plot in~\cref{fig:perfAXpar}, with approximately
+63 GFLOP/s. It is still less than 1\% of peak for an Intel Xeon Gold 6248.
+For the GPU configuration, \pk{MatAIJcuSPARSE}~\ref{pgfplots:cusparse10} with $b=6$
+and $N=64$, right-most bottom plot in~\cref{fig:perfAXpar}, performs at around
+306 GFLOP/s, about 4\% of peak for an NVIDIA Tesla V100 SXM2.
+\pgfplotstableread{
+P       aij1        aij3        aij6
+1       6.6857e-02  7.4649e-01  3.1128e+00       
+2       3.4206e-02  3.7820e-01  1.5723e+00
+4       1.6565e-02  1.8634e-01  7.7238e-01
+8       9.1370e-03  9.7549e-02  3.9884e-01
+16      7.2334e-03  7.9871e-02  3.2154e-01
+20      7.8399e-03  8.6487e-02  3.5536e-01
+}\loadedOMP
+\pgfplotstablegetelem{0}{aij1}\of{\loadedOMP}
+\let\refI\pgfplotsretval
+\pgfplotstablecreatecol[create col/expr={\refI/\thisrow{aij1}}]{speedupI}\loadedOMP
+\pgfplotstablegetelem{0}{aij3}\of{\loadedOMP}
+\let\refIII\pgfplotsretval
+\pgfplotstablecreatecol[create col/expr={\refIII/\thisrow{aij3}}]{speedupIII}\loadedOMP
+\pgfplotstablegetelem{0}{aij6}\of{\loadedOMP}
+\let\refVI\pgfplotsretval
+\pgfplotstablecreatecol[create col/expr={\refVI/\thisrow{aij6}}]{speedupVI}\loadedOMP
+\begin{figure}[tbp]
+  \centering
+    \begin{tikzpicture}
+      \begin{loglogaxis}[
+              ytick={1,2,4,8,20},
+        yticklabel={
+            \pgfmathparse{100*round(exp(\tick))}
+$\pgfmathprintnumber{\pgfmathresult}$\%
+},
+        xtick = {\empty},
+        extra x ticks = {1,2,4,8,16,20},height=6cm,
+          extra x tick labels = {1,2,4,8,16,\hspace*{0.3cm}20},
+          legend style={at={(0.02,0.98)}, legend cell align=left,anchor=north west},
+      ]
+          \addplot[black, dashed] table [x expr=\thisrow{P},y expr=\thisrow{P}]{\loadedOMP};\label{pgfplots:linear}
+          \addplot[blueaccent,mark=square*] table [x expr=\thisrow{P},y expr=\thisrow{speedupI}]{\loadedOMP};\label{pgfplots:speedupI}
+          \addplot[darkgray,mark=*] table [x expr=\thisrow{P},y expr=\thisrow{speedupIII}]{\loadedOMP};\label{pgfplots:speedupIII}
+          \addplot[last,mark=triangle*] table [x expr=\thisrow{P},y expr=\thisrow{speedupVI}]{\loadedOMP};\label{pgfplots:speedupVI}
+          \addlegendentry{Linear speedup}
+          \addlegendentry{$b = 1$}
+          \addlegendentry{$b = 3$}
+          \addlegendentry{$b = 6$}
+    \end{loglogaxis}
+    \end{tikzpicture}
+    \caption{\label{fig:scaling-Ax}}
+\end{figure}
 
@@
 \end{figure}
+\subsubsection{Blocking inside SLEPc\label{sec:lobpcg}}
+Finally, the performance of LOBPCG~\cite{knyazev2001toward} as implemented in SLEPc is studied using three different preconditioners:
+\begin{itemize}
+    \item \pk{PCASM}: one-level overlapping Schwarz method with one level of overlap and exact Cholesky factorizations in each subdomain;
+    \item \pk{PCHPDDM}: multilevel overlapping Schwarz method;
+    \item \pk{PCGAMG}: algebraic multigrid method~\cite{adams2004ultra}.
+\end{itemize}
+More details are given in the next section where focus is put on preconditioning rather than Krylov methods. The two main points are that:
+\begin{itemize}
+    \item \pk{PCASM} and \pk{PCHPDDM} handle blocking, see the list~\cref{sec:interface-ksp};
+    \item \pk{PCHPDDM} and \pk{PCGAMG} have low contraction factors.
+\end{itemize}
+The following generalized eigenvalue problem is solved in a three-dimensional cube:
+\begin{equation*}
+    -\nabla \cdot (\kappa \nabla u) = \lambda u.\label{eq:lobpcg}
+\end{equation*}
+This continuous equation is discretized with $\mathbb{P}_3$ finite elements using FreeFEM.
+The script is available at \url{http://jolivet.perso.enseeiht.fr/lobpcg.edp}.
+Following timings and convergence histories have been obtained on $2{,}304$ processes of Ir\`ene, for solving a problem of dimension \pgfmathprintnumber{74618461}.
+By default, SLEPc applies at most 5 iterations of GMRES on each column of a so-called set of active columns.
+Since the option \pk{-eps\_lobpcg\_blocksize 40} is used for computing 20 eigenpairs, the size of the set varies between 40 to 1.
+This size corresponds to the number of right-hand sides in systems that will be solved throughout LOBPCG iterations.
+Instead of using \pk{KSPGMRES}, which does not handle blocking, \pk{KSPHPDDM} is used, with the pseudo-block GMRES.
+That means that, assuming the size of the set of active columns is 40, one LOBPCG iterations implies doing at most 5 GMRES iterations with 40 vectors simultaneously, using \pk{MatProductNumeric} and \pk{PCMatApply}, instead of doing at most $40\times 5$ successive GMRES iterations, using \pk{MatMult} and \pk{PCApply}.
+Results are gathered in~\cref{fig:lobpcg}.
+In the first line, the numbers of outer LOBPCG iterations are reported with the three different inner preconditioners mentioned above.
+\pk{PCASM} is known to yield a preconditioned operator whose condition number grows as the number of subdomains, here processes, increases.
+Thus, the number of outer iterations is higher than with \pk{PCHPDDM} or \pk{PCGAMG} since the inner solves are not converging to meaningful solutions in just 5 iterations.
+For both multilevel preconditioners, however, it seems that they perform an equivalent job since the difference in number of outer iterations is only 1.
+The second and third lines represent the number of times \pk{PCApply} and \pk{PCMatApply} are called.
+Since \pk{PCGAMG} does not handle blocking, at each pseudo-block GMRES iteration, \pk{PCApply} is called on all active columns.
+\pk{PCASM} and \pk{PCHPDDM} handle blocking, so only rely on \pk{PCMatApply}.
+In the fourth line, the time spent setting up the preconditioners is given.
+\pk{PCASM} is extremely cheap but not very robust numerically, while both multilevel preconditioners have a similar setup time.
+Eventually, in the last line, the time spent in the full eigensolve is given.
+Clearly, having both a \pk{KSP} and a \pk{PC} which can deal with multiple vectors efficiently is mandatory to achieve reasonable performance with such an outer solver that heavily relies on blocking.
+\pgfplotstableread{
+PC       outer PCApply PCMatApply PCSetUp       EPSSolve
+PCASM    54    ---     378        2.8536e+00    265.6564
+PCHPDDM  14    ---     98         2.1566e+01    121.584
+PCGAMG   15    4{,}049    ---     2.6528e+01    391.822
+}\mytable
+\pgfplotstabletranspose[
+    colnames from=PC,
+    input colnames to=PC
+]\mytablenew{\mytable}
+\begin{figure}[h!]
+    \centering
+\pgfplotstabletypeset[
+columns/PCHPDDM/.style={
+postproc cell content/.append code={%
+\ifnum\pgfplotstablerow>2
+\pgfkeyssetvalue{/pgfplots/table/@cell content}{\pgfmathparse{##1<100?"0":{}}\phantom{\pgfmathresult}\pgfmathparse{##1<10?"0":{}}\phantom{\pgfmathresult}\pgfmathprintnumber[precision=1,fixed]{##1}}%
+\fi
+}
+},
+columns/PCGAMG/.style={
+postproc cell content/.append code={%
+\ifnum\pgfplotstablerow>2
+\pgfkeyssetvalue{/pgfplots/table/@cell content}{\pgfmathparse{##1<100?"0":{}}\phantom{\pgfmathresult}\pgfmathparse{##1<10?"0":{}}\phantom{\pgfmathresult}\pgfmathprintnumber[precision=1,fixed]{##1}}%
+\fi
+}
+},
+columns/PCASM/.style={
+postproc cell content/.append code={%
+\ifnum\pgfplotstablerow>2
+\pgfkeyssetvalue{/pgfplots/table/@cell content}{\pgfmathparse{##1<100?"0":{}}\phantom{\pgfmathresult}\pgfmathparse{##1<10?"0":{}}\phantom{\pgfmathresult}\pgfmathprintnumber[precision=1,fixed]{##1}}%
+\fi
+}
+},
+columns/PC/.style={
+    column name={},
+% stupid example for multiple postprocessors:
+postproc cell content/.append style={
+/pgfplots/table/@cell content/.add={\tt}{},
+},
+postproc cell content/.append code={%
+\ifnum\pgfplotstablerow=0
+\pgfkeyssetvalue{/pgfplots/table/@cell content}{\# of outer iterations}%
+\else
+\ifnum\pgfplotstablerow>2
+\pgfkeyssetvalue{/pgfplots/table/@cell content}{\pk{##1} (sec)}%
+\fi
+\fi
+}
+},
+    every even row/.style={
+before row={\rowcolor[gray]{0.9}}},
+        assign column name/.style={/pgfplots/table/column name={\pk{#1}}},
+    every head row/.style={
+        before row=\toprule,
+        after row=\midrule
+    },
+    every last row/.style={
+        after row=\bottomrule
+    },
+string type]{\mytablenew}
+\caption{Performance of \pk{EPSLOBPCG} with three different inner preconditioners: \pk{PCASM} and \pk{PCHPDDM}, which utilize \pk{PCMatApply}, and \pk{PCGAMG}, which utilizes \pk{PCApply}.\label{fig:lobpcg}}
+\end{figure}
 \section{PCHPDDM}\label{sec:PCHPDDM}
@@
 modern large-scale applications dealing with partial differential equations.
-For an introduction to domain decomposition methods, see
-~\cite{smith2004domain,dolean2014ddm,olofbook,tarekmathewbook} {\color{red} FIX CITATIONS}.
-% In PETSc, preconditioners inherit from the \pk{PC} base class.
+There are many monographs on domain decomposition methods \cite{smith2004domain,dolean2014ddm,toselli2005domain,pechstein2012finite,quarteroni1999domain,mathew2008domain}.
 
@@
 % preconditioner~\cite{dohrmann2008domain}. Its type is \pk{PCEXOTIC}.
-Furthermore, PETSc provides domain decomposition methods of the non-overlapping type;
-\pk{PCNN} implements a basic balancing
-Neumann--Neumann method~\cite{mandel1993balancing}, while more advanced
-preconditioning techniques are available with \pk{PCBDDC}~\cite{zampini2016pcbddc}, including guaranteed robust convergence
-by means of adaptive selection
-of primal constraints~\cite{pechstein2017unified} and support for $H(\text{curl})$ and $H(\text{div})$
-conforming finite elements~\cite{zampinihcurl1,zampinihcurl2,zampinihdiv,zampinidarcy}, as well as for IsoGeometric Analysys \cite{zampiniiga1,zampiniiga2,zampiniiga3}.
+Furthermore, PETSc provides domain decomposition methods of the non-overlapping
+type; \pk{PCNN} implements a basic balancing Neumann--Neumann
+method~\cite{mandel1993balancing}, while more advanced preconditioning
+techniques are available with \pk{PCBDDC}~\cite{zampini2016pcbddc}, including
+adaptive selection of primal constraints~\cite{pechstein2017unified}, support
+for $H(\text{curl})$~\cite{zampini2017balancing} and
+$H(\text{div})$~\cite{oh2018bddc} conforming finite elements, and isogeometric
+analysis \cite{da2014isogeometric}.
 For multigrid,
@@
 
-HPDDM implements the same one-level overlapping Schwarz methods as PETSc as well as
-optimized Schwarz methods~\cite{gander2006optimized} {\color{red} Are you planning to expose them too?}.
-It also provides support for defining robust two-level methods using the GENeralized Eigenvalue problem on the Overlap (GenEO) framework~\cite{spillane2011robust,spillane2013abstract}; the results presented in this work presents the first high-level access to
-GenEO from PETSc and its composable solver infrastructrure~\cite{brown2012composable}.
-We note that GenEO has been recently implemented in other libraries as well~\cite{butler2020high}
+HPDDM implements the same one-level overlapping Schwarz methods as PETSc as
+well as optimized Schwarz methods~\cite{gander2006optimized}, which may be used
+in PETSc with \pk{PCSetModifySubMatrices}.  It also provides support for
+defining robust two-level methods using the Generalized Eigenvalue problem on
+the Overlap (GenEO) framework, for finite
+elements~\cite{spillane2011robust,spillane2013abstract} and boundary
+elements~\cite{marchand2019two}. Results in this work present the first
+high-level access to GenEO from PETSc and its composable solver
+infrastructrure~\cite{brown2012composable}. We note that GenEO has been
+recently implemented in other libraries as well~\cite{butler2020high}.
 
@@
 for a given coefficient matrix $A$ and right-hand side $b$ of dimension $n$
-assembled on $N$ processes, using a ksp($A$, $P_A$), i.e., the matrix used to
-define the preconditioner is $A$ itself {\color{red} WHY IS THAT? You do not support a different preconditioning matrix?}. Without anything more from the user,
+assembled on $N$ processes. For simplicity, a ksp($A$, $P_A$) is used next, i.e., the matrix used to
+define the preconditioner is $A$ itself. Without anything more from the user,
 \pk{PCHPDDM} is strictly equivalent to \pk{PCASM}. That is, if $A$ is distributed among
@@
 \end{equation*}
-{\color{red} Should we mention the connections with the condition number here? and thus discuss the guaranteed convergence property of the method?}
+Such a construction of a two-level method yields a preconditioner with which it is possible to bound the condition number of the preconditioned operator.
+Thus, the convergence speed of a preconditioned Krylov method such as the conjugate gradient can be guaranteed a priori.
 
@@
 %ARPACK~\cite{lehoucq1998arpack}.
-While Step \#3 is common to most domain
+While step \#3 is common to most domain
 decomposition or multigrid preconditioners, HPDDM takes advantage of the special
@@
 It is thus natural to remap the coarse operator on
-a subset of the processes used to decompose the original matrix $A$; in HPDDM, the coarse operator is
+a subset of the processes used to decompose the original matrix $A$. In HPDDM, the coarse operator is
 computed and redistributed simultaneously.
@@
 details of the implementation~\cite{jolivet2013scalable}.
-We finally note here that similar techniques have been proven beneficial for other types
+Note that similar techniques have been proven beneficial for other types
 of solvers in PETSc~\cite{may2016extreme}.
@@
 Other \pk{DMType}s, including user-defined, may be supported by
-implementing the proper \pk{DMCreateNeumannOverlap} callback; support for \pk{DMDA}
+implementing the proper \pk{DMCreateNeumannOverlap} callback. Support for \pk{DMDA}
 could be easily added upon request.
@@
 
-For performance reasons, it is usually better to compute the same number of eigenvectors
-on each process, since in this case the coarse operator will be assembled using the \pk{MatMPISBAIJ} (or \pk{MatMPIBAIJ}) matrix format,
-with usually large block sizes; this can have serious memory and performance implications, see~\cref{fig:perfAX}. During solution phases,
-we can use \texttt{MPI\_Gather} and \texttt{MPI\_Scatter} operations to move between levels,
-instead of the variable size alternatives \texttt{MPI\_Gatherv} and \texttt{MPI\_Scatterv} that could exhibit worse performances.
-In addition, standard eigensolvers such as ARPACK~\cite{lehoucq1998arpack} or \pk{EPSKRYLOVSCHUR} from SLEPc,
-cannot compute all eigenpairs in a given interval of $\mathbb{R}^+$. Instead, an upper bound of the number of eigenpairs has to be provided.
+For performance reasons, it is better to compute the same number of
+eigenvectors on each process, since in this case the coarse operator will be
+assembled using the \pk{MatMPISBAIJ} or \pk{MatMPIBAIJ} matrix type, with
+usually large block sizes. This can have serious memory and performance
+implications, see~\cref{fig:perfAX}. When applying the preconditioner,
+\texttt{MPI\_Gather} and \texttt{MPI\_Scatter} operations are used to move
+between levels, instead of the variable-size alternatives \texttt{MPI\_Gatherv}
+and \texttt{MPI\_Scatterv}, that could exhibit worse performance. In addition,
+standard eigensolvers such as ARPACK~\cite{lehoucq1998arpack} or
+\pk{EPSKRYLOVSCHUR} from SLEPc, cannot compute all eigenpairs in a given
+interval of $\mathbb{R}^+$. Instead, an upper bound of the number of eigenpairs
+has to be provided.
 
@@
 Some other parameters may be adjusted to define how the one-level preconditioner
-and the coarse grid operator interact. Besided simple the additive
-correction shown in~\cref{eq:two}, users can choose more numerically efficient
-corrections as:~\cite{tang2009comparison,bastian1998additive}
+and the coarse grid operator interact. Besides the additive
+correction shown in~\cref{eq:two}, end-users may select more numerically efficient
+corrections~\cite{tang2009comparison,bastian1998additive} as:
 \begin{align*}
@@
 \end{align*}
-with $Q = P \left(P^T A P\right)^{-1} P^T$. These preconditioners may be selected using the
-options \pk{-pc\_hpddm\_coarse\_correction} \pk{(deflated$|$additive$|$balanced)} or the
-routine \pk{PCHPDDMSetCoarseCorrectionType}.
+with $Q = P \left(P^T A P\right)^{-1} P^T$. This is parameterized using the
+options \pk{-pc\_hpddm\_coarse\_correction}
+\pk{(deflated$|$additive$|$balanced)} or the routine
+\pk{PCHPDDMSetCoarseCorrectionType}. It is important to keep in mind that all
+these correction formulae may be applied to either a single vector or a block
+of multiple vectors optimally. Indeed, the one-level preconditioner $M^{-1}$ is
+applied with either \pk{PCApply} or \pk{PCMatApply}, and the
+restriction-correction-interpolation operator $Q$ is applied in a block fashion
+in HPDDM, and not column by column. See the effect of this strategy~\cref{sec:lobpcg}.
 
@@
 In addition to the local unassembled Neumann matrices~$\{\mathring{A}_i\}_{i=1}^N$,
-users can prescribe additional local operators defined on the overlapped subdomain and
-with optimized, Robin
+users can prescribe additional local operators defined on the overlapping subdomain and
+with optimized Robin
 boundary conditions $\{B_i\}_{i=1}^N$ via
@@
 \end{equation*}
-{\color{red} One sentence or two on why this is needed? what are the cases where this is used in practice?}
+This may be useful when dealing with nearly indefinite systems, e.g., nearly incompressible elasticity or the system of Stokes.
+Indeed, for such equations, one needs to compute a lot of eigenvectors from the classical GenEO eigenproblem~\cref{eq:gevp} to generate a robust preconditioner. GenEO-2 alleviates this phenomenon.
 
@@
 \pk{PCHPDDM} follows the same numbering of \pk{PCBDDC}:
-the finest level is always numbered with 1, and the level index increases as traversing the hierarchy, up until
+the finest level is always numbered with 1, and the level index increases as the hierarchy is traversed, up until
 the coarsest level, whose solver options are prefixed by \pk{coarse\_}. In order to
@@
 a three-level method with the second level distributed among 8 processes, while the coarsest level
-will be built from 10 eigenvalues from the second level and remapped onto 2 processes.
+will be built using 10 eigenvalues per subdomain from the second level and remapped onto 2 processes.
 Additional examples for the solver customization are provided in~\cref{sec:pchpddm-examples}
@@
 overlap, see for example $\Omega_1$ and $\Omega_2$ from~\cref{fig:poisson} in a
-finite element context. Since the assembled form of these operators can be reconstructed
-it is possible to construct the Dirichlet operators, e.g.
-$\{R_i A R_i^T\}_{i=1}^N$ from~\cref{sec:interface-pc}. All tools needed by the
+finite element context. Since the assembled form of these operators can be reconstructed,
+it is possible to get the Dirichlet operators
+$\{R_i A R_i^T\}_{i=1}^N$ from~\cref{sec:interface-pc} as well. All tools needed by the
 GenEO framework are thus readily available. Note however that one-level
 overlapping Schwarz methods are known for converging slowly when there is in
-fact no overlap. {\color{red} add one sentence with reference to Nicole's work on GenEO for BDD?}
+fact no overlap. The proposed methodology is not strictly equivalent to the
+BDD-GenEO method~\cite{spillane2013automatic}, in which local Schur complements
+on subdomain interfaces are computed explicitly, and then used in concurrent
+dense generalized eigenvalue problems.
 %It is common in the field of BDD to see the coarse problem as
@@
 There are heterogeneities due to jumps in $E$ and $\nu$.
-The following discontinuous piecewise constant values are considered: $(E_1, \nu_1) = (2\times 10^{11}, 0.25)$ in blue regions, while $(E_2, \nu_2) = (10^{7}, 0.45)$ in red regions, see~\cref{fig:elas}.
-{\color{red} Pierre, check the blue/red here and in the figure caption}
+The following discontinuous piecewise constant values are considered: $(E_1, \nu_1) = ({\color{paraviewred}2\times 10^{11}}, 0.25)$ in blue regions, while $(E_2, \nu_2) = ({\color{paraviewblue}10^7}, 0.45)$ in red regions, see~\cref{fig:elas}.
 The code used to produce these results was borrowed from
@@
 The striped plane in the background shows a cut of the resting configuration with the jumps in the coefficient $E_1={\color{paraviewblue}10^7}$ and $E_2={\color{paraviewred}2\times 10^{11}}$, aligned with those of $\nu$.\label{fig:elas}}
-\end{figure}
-
-The following solvers are compared:
+\end{figure} \\
+The following preconditioners are compared:
 \begin{enumerate}
@@
     \item \pk{PCHPDDM} with an exact coarse solver;\label{enum:exact}
-    \item \pk{PCHPDDM} with an inexact coarse solver using \pk{PCHYPRE};\label{enum:inexacthypre} {\color{red} I think we should not include this case}
     \item \pk{PCHPDDM} with inexact coarse solvers using GenEO multilevel extension.\label{enum:inexact} %, cf.~\cref{sec:ml}.\label{enum:inexact}
 \end{enumerate}
-As a linear solver, we use flexible GMRES, customized with the options \pk{-ksp\_gmres\_modifiedgramschmidt
+The flexible GMRES is used and customized with the options \pk{-ksp\_gmres\_modifiedgramschmidt
 -ksp\_gmres\_restart 50
@@
 \end{minipage} \\
-By default, the coarse problem in \pk{PCHPDDM} is solved using an exact $LU$ or Cholesky factorization, in this case using 24 processes and \pk{MKL\_CPARDISO}.
-Switching to an inexact solver using \pk{PCHYPRE} (resp.~GenEO multilevel extension), cf.~\cref{enum:inexacthypre} (resp.~\cref{enum:inexact}), is straightforward.
+By default, the coarse problem in \pk{PCHPDDM} is solved using an exact $LU$ or Cholesky factorization, in this case using 24 processes and MKL CPARDISO.
+Switching to an inexact solver using the multilevel extension of GenEO, cf.~\cref{enum:inexact}, is straightforward.
 \\[-4pt]
-\begin{minipage}[t]{0.4\textwidth}
-\begin{Verbatim}[fontsize=\footnotesize,frame=single,framerule=0.1mm,commandchars=\\\{\}, codes={\catcode`$=3\catcode`^=7}]
-\fvtextcolor{mygreen}{# cf. \cref{enum:inexacthypre}}
--prefix_push pc_hpddm_coarse_
- -ksp_type    gmres
- -ksp_rtol    1.0e-2
- -ksp_max_it  100
- -ksp_pc_side right
- -pc_type     hypre
- -prefix_push pc_hypre_
-  -boomeramg_interp_type  ext+i
-  -boomeramg_coarsen_type hmis
- -prefix_pop
--prefix_pop
-\end{Verbatim}
-\end{minipage}
-\hfill
 \begin{minipage}[t]{0.525\textwidth}
@@
 \end{Verbatim}
-\vspace*{-0.5cm}
 \end{minipage} \\
-For the multivel extension of GenEO, $M$ is the number of subdomains used to define the second-level domain decomposition, which is an aggregation of first-level subdomain.
-When it comes to convergence, first note that BoomerAMG from \emph{hypre} has a really hard time reaching the stopping criterion of $10^{-2}$ when solving GenEO coarse problems.
-This is emphasized in~\cref{fig:hypre-conv}, where the lack of convergence at the coarse level with \pk{PCHYPRE} greatly deteriorates the outer solver convergence, this configuration is thus discarded next and motivates the need for GenEO multilevel extension.
-\pgfplotstableread{
-it res inner
-  0  3.059298477541e+01 nan
-  1  3.059296797779e+01 100
-  2  3.059296195928e+01 15
-  3  3.059296184099e+01 15
-  4  3.059295742636e+01 13
-  5  3.059295742508e+01 15
-  6  3.059294966626e+01 14
-  7  3.059293719161e+01 95
-  8  3.059292563008e+01 100
-  9  3.059286757287e+01 100
- 10  3.059276348161e+01 100
- 11  3.059259253443e+01 100
- 12  3.059200963623e+01 100
- 13  3.059116858285e+01 100
- 14  3.059008706025e+01 100
- 15  3.058899295395e+01 100
- 16  3.058765538548e+01 100
- 17  3.058718339920e+01 100
- 18  3.058690573439e+01 100
- 19  3.058631705473e+01 100
- 20  3.058529856313e+01 100
- 21  3.058458303471e+01 100
- 22  3.058441527393e+01 100
- 23  3.058404813350e+01 100
- 24  3.058350395496e+01 100
- 25  3.058297671639e+01 100
- 26  3.058267730942e+01 100
- 27  3.058267032565e+01 100
- 28  3.058260186627e+01 19
- 29  3.058233416052e+01 100
- 30  3.058177883143e+01 100
- 31  3.058107492291e+01 100
- 32  3.058053304435e+01 100
- 33  3.058021917750e+01 100
- 34  3.057991067536e+01 100
- 35  3.057984123224e+01 100
- 36  3.057981166461e+01 100
- 37  3.057969981702e+01 100
- 38  3.057951130340e+01 100
- 39  3.057909738295e+01 100
- 40  3.057879065735e+01 100
- 41  3.057844662403e+01 100
- 42  3.057813110973e+01 100
- 43  3.057805844313e+01 100
- 44  3.057748148817e+01 100
- 45  3.057716298347e+01 100
- 46  3.057695365582e+01 100
- 47  3.057687162665e+01 100
- 48  3.057685976587e+01 100
- 49  3.057676977855e+01 100
- 50  3.057666112196e+01 100
- 51  3.057666112157e+01 100
- 52  3.057666066500e+01 15
- 53  3.057666020080e+01 15
- 54  3.057666000522e+01 12
- 55  3.057666000081e+01 15
- 56  3.057665991326e+01 15
- 57  3.057665913485e+01 95
- 58  3.057665848806e+01 100
- 59  3.057665837181e+01 100
- 60  3.057665830553e+01 100
- 61  3.057665733071e+01 100
- 62  3.057665722344e+01 100
- 63  3.057665608783e+01 100
- 64  3.057664613215e+01 100
- 65  3.057661541445e+01 100
- 66  3.057659606149e+01 100
- 67  3.057657581806e+01 100
- 68  3.057648665398e+01 100
- 69  3.057648651148e+01 100
- 70  3.057647274070e+01 100
- 71  3.057646787321e+01 100
- 72  3.057646700351e+01 100
- 73  3.057646583987e+01 100
- 74  3.057640958561e+01 100
- 75  3.057640655402e+01 100
- 76  3.057640464884e+01 100
- 77  3.057634384959e+01 100
- 78  3.057630002417e+01 100
- 79  3.057628670010e+01 100
- 80  3.057628640121e+01 100
- 81  3.057628480757e+01 100
- 82  3.057625930943e+01 100
- 83  3.057619950478e+01 100
- 84  3.057615004023e+01 100
- 85  3.057605590648e+01 100
- 86  3.057603257752e+01 100
- 87  3.057575042666e+01 100
- 88  3.057537924704e+01 100
- 89  3.057516167721e+01 100
- 90  3.057445320162e+01 100
- 91  3.057434150644e+01 100
- 92  3.057433943630e+01 100
- 93  3.057431618884e+01 100
- 94  3.057419597979e+01 100
- 95  3.057412470447e+01 100
- 96  3.057409280981e+01 100
- 97  3.057407988562e+01 100
- 98  3.057397192115e+01 100
- 99  3.057365742468e+01 100
-100  3.057298339599e+01 100
-}\loadedtableHYPRE
-\begin{figure}[htbp]
-    \centering
-        \begin{tikzpicture}[font=\scriptsize]
-\begin{axis}[%title=\framebox{{\upcase diffusion 2D, $\mathbb{P}_4$ FE, $\varepsilon = \num[exponent-product=\cdot]{1e-06}$}},
-        axis y line*=left,ymajorgrids, unbounded coords=jump, xmin=0,xmax=100,extra y ticks={100}, width=0.5\textwidth,extra y tick labels=\empty, extra y tick style={grid=major}, ylabel = {\# of inner iterations~\ref{pgfplots:inner-hypre}}, xlabel = {Outer iteration index}]
-\addplot[last, very thick] table[x=it, y=inner] {\loadedtableHYPRE};\label{pgfplots:inner-hypre}
-% \node[rectangle, draw, font=\scriptsize, text width = 4.6cm] at (axis cs:12.9, 18) {\texttt{-hpddm\_master\_p 32 \\ -hpddm\_master\_aggregate\_size 16}};
-\end{axis}
-\begin{axis}[axis x line=none,ymax=31,ymin=30, axis y line*=right, xmin=0,xmax=100,width=0.5\textwidth, ylabel near ticks=right, ylabel = {Unpreconditioned residual~\ref{pgfplots:res-hypre}}]
-\addplot[myred, densely dashed, very thick] table[x=it, y=res] {\loadedtableHYPRE};\label{pgfplots:res-hypre}
-\end{axis}
-\end{tikzpicture}
-    \caption{Convergence history of \pk{PCHPDDM} using \pk{PCHYPRE} on 24 processes
-    as a coarse solver. The maximum number of iterations of both the inner coarse solver and the outer solver is set to 100.\label{fig:hypre-conv}}
-\end{figure}
+In this third configuration, $M$ is the number of subdomains used to define the second-level domain decomposition, which is an aggregation of first-level subdomain.
 
@@
 It is discretized using first order N\'ed\'elec finite elements and the \pk{PCBDDC} solver
-is automatically customized to obtain a stable method with these elements \cite{zampinihcurl2}.
+is automatically customized to obtain a stable method with these elements~\cite{zampini2017balancing}.
 The following command line options are used. \\[-4pt]
@@
 
-Concerning PETSc, one may notice that it is now possible to do some code
-refactorization for applications that had to deal with multiple right-hand sides.
-As a striking example, LOBPCG~\cite{knyazev2001toward} or
-CISS~\cite{sakurai2003projection} implementations in SLEPc are
-currently %\footnote{in version 3.13.2}
-not using blocking when applying
-spectral transformations, but this is being rewritten to use the new routine
-\pk{KSPMatSolve}\footnote{\url{https://gitlab.com/slepc/slepc/-/merge_requests/63}} and should be available in version 3.14.0.
+Concerning PETSc, the infrastructure for dealing with blocks of vectors has
+been laid out. Still, only a little fraction of the built-in matrix types and
+preconditioners may efficiently handle such a format.
 
@@
 \bibitem{balay1997cient}
-{\sc S.~Balay, W.~Gropp, L.~Curfman~McInnes, and B.~Smith}, {\em Efficient
+{\sc S.~Balay, W.~Gropp, L.~Curfman~McInnes, and B.~F. Smith}, {\em Efficient
   management of parallelism in object-oriented numerical software libraries},
@@
 
-\bibitem{templateBook}
+\bibitem{barrett1994templates}
 {\sc R.~Barrett, M.~Berry, T.~F. Chan, J.~Demmel, J.~Donato, J.~Dongarra,
- V.~Eijkhout, R.~Pozo, C.~Romine, and H.~Van der Vorst}
- {\em Templates for the solution of linear systems: building blocks for iterative methods},
-  Society for Industrial and Applied Mathematics, (1994).
-
-\bibitem{xsdk}
-{\sc R.~Bartlett, I.~Demeshko, T.~Gamblin, G.~Hammond, M.~A. Heroux, J.~Johnson, A.~Klinvex, X.~Li, L.~Curfman~McInnes, J.~D. Moulton, et~al.},
-  {\em xSDK Foundations: Toward an Extreme-scale Scientific Software Development Kit},
-  Supercomputing Frontiers and Innovations, 4 (2017), pp.~69--82.
+  V.~Eijkhout, R.~Pozo, C.~Romine, and H.~Van~der Vorst}, {\em Templates for
+  the solution of linear systems: building blocks for iterative methods}, SIAM,
+  1994.
 
@@
   Scientific Programming, 20 (2012), pp.~241--255.
-
-\bibitem{zampiniiga1}
-{\sc L.~Beirao da Veiga, L.~F. Pavarino, S.~Scacchi, O.~B. Widlund and  S.~Zampini},
-  {\em Isogeometric {BDDC} preconditioners with deluxe scaling},
-  SIAM Journal on Scientific Computing,
-  36 (2014), pp.~A1118--A1139.
-
-\bibitem{zampiniiga2}
-{\sc L.~Beirao da Veiga, L.~F. Pavarino, S.~Scacchi, O.~B. Widlund and  S.~Zampini},
-  {\em Adaptive selection of primal constraints for isogeometric {BDDC} deluxe preconditioners},
-  SIAM Journal on Scientific Computing, 39 (2017), pp.~A281--A302.
 
@@
 
+\bibitem{10.1145/3232850}
+{\sc W.~Boukaram, G.~Turkiyyah, and D.~Keyes}, {\em {Hierarchical Matrix
+  Operations on GPUs: Matrix-Vector Multiplication and Compression}}, ACM
+  Transactions on Mathematical Software, 45 (2019).
+
 \bibitem{brown2012composable}
-{\sc J.~Brown, M.~G. Knepley, D.~A. May, L.~C. McInnes, and B.~F. Smith}, {\em
-  Composable linear solvers for multiphysics}, in 2012 11th International
+{\sc J.~Brown, M.~G. Knepley, D.~A. May, L.~Curfman~McInnes, and B.~F. Smith},
+  {\em Composable linear solvers for multiphysics}, in 2012 11th International
   Symposium on Parallel and Distributed Computing, IEEE, 2012, pp.~55--62.
@@
 
-%\bibitem{dohrmann2016bddc}
-%{\sc C.~R. Dohrmann and O.~B. Widlund}, {\em {A BDDC Algorithm with Deluxe
-%  Scaling for Three-Dimensional $H(\textnormal{curl})$ Problems}},
-%  Communications on Pure and Applied Mathematics, 69 (2016), pp.~745--770.
+\bibitem{da2014isogeometric}
+{\sc L.~B. Da~Veiga, L.~F. Pavarino, S.~Scacchi, O.~B. Widlund, and
+  S.~Zampini}, {\em {Isogeometric BDDC preconditioners with deluxe scaling}},
+  SIAM Journal on Scientific Computing, 36 (2014), pp.~A1118--A1139.
+
+\bibitem{davis2004algorithm}
+{\sc T.~A. Davis}, {\em Algorithm 832: {UMFPACK}—an unsymmetric-pattern
+  multifrontal method}, ACM Transactions on Mathematical Software, 30 (2004),
+  pp.~196--199.
+
+\bibitem{dolean2014ddm}
+{\sc V.~Dolean, P.~Jolivet, and F.~Nataf}, {\em An Introduction to Domain
+  Decomposition Methods\col\ Algorithms, Theory and Parallel Implementation},
+  SIAM, 2015.
 
@@
 
-\bibitem{Frommer}
-{\sc A.~Frommer, K.~Lund, D.~B. Szyld},
- {\em Block Krylov subspace methods for funcions of matrices},
- Electronic Transactions on Numerical Analysis, 47 (2017), pp.~100--126.
+\bibitem{frommer2017block}
+{\sc A.~Frommer, K.~Lund, and D.~B. Szyld}, {\em {Block Krylov Subspace Methods
+  for Functions of Matrices}}, Electronic Transactions on Numerical Analysis,
+  47 (2017), pp.~100--126.
 
@@
 \bibitem{gee2006ml}
-{\sc M.~Gee, C.~Siefert, J.~Hu, R.~Tuminaro, and M.~Sala}, {\em {ML} 5.0
-  smoothed aggregation user's guide}, Tech. Report SAND2006-2649, Sandia
-  National Laboratories, 2006, \url{http://trilinos.sandia.gov/packages/ml}.
+{\sc M.~W. Gee, C.~M. Siefert, J.~J. Hu, R.~S. Tuminaro, and M.~G. Sala}, {\em
+  {ML} 5.0 smoothed aggregation user's guide}, Tech. Report SAND2006-2649,
+  Sandia National Laboratories, 2006, \url{https://trilinos.github.io/ml.html}.
 
@@
 
+\bibitem{hoefler2015scientific}
+{\sc T.~Hoefler and R.~Belli}, {\em Scientific benchmarking of parallel
+  computing systems: twelve ways to tell the masses when reporting performance
+  results}, in Proceedings of the 2015 International Conference for High
+  Performance Computing, Networking, Storage and Analysis, SC15, 2015.
+
 \bibitem{mkl-web-page}
@@
 
+\bibitem{li2005overview}
+{\sc X.~S. Li}, {\em {An overview of SuperLU: Algorithms, implementation, and
+  user interface}}, ACM Transactions on Mathematical Software, 31 (2005),
+  pp.~302--325.
+
+\bibitem{li2003superlu_dist}
+{\sc X.~S. Li and J.~Demmel}, {\em {SuperLU\_DIST: A scalable
+  distributed-memory sparse direct solver for unsymmetric linear systems}}, ACM
+  Transactions on Mathematical Software, 29 (2003), pp.~110--140.
+
 \bibitem{mandel1993balancing}
@@
 
-\bibitem{tarekmathewbook}
-{\sc T.~Mathew},
-  {\em Domain decomposition methods for the numerical solution of partial differential equations},
-  Springer Science \& Business Media, 2008.
+\bibitem{marchand2019two}
+{\sc P.~Marchand, X.~Claeys, P.~Jolivet, F.~Nataf, and P.-H. Tournier}, {\em
+  {Two-level preconditioning for $h$-version boundary element approximation of
+  hypersingular operator with GenEO}}, Numerische Mathematik,  (2019),
+  p.~submitted for publication.
+
+\bibitem{mathew2008domain}
+{\sc T.~Mathew}, {\em Domain decomposition methods for the numerical solution
+  of partial differential equations}, vol.~61, Springer Science \& Business
+  Media, 2008.
 
@@
 
-\bibitem{cusparse-web-page}
-{\sc NVIDIA}, {\em {cuSPARSE} web page}.
-\newblock
-  \url{https://docs.nvidia.com/cuda/cusparse/index.html},
-  2020.
-
 \bibitem{notay2000flexible}
@@
 
-\bibitem{zampinihdiv}
-{\sc D.~S. Oh, O.~B. Widlund, S.~Zampini and C.~Dohrmann},
-  {\em {BDDC} algorithms with deluxe scaling and adaptive selection of primal constraints for {R}aviart-{T}homas vector fields},
-  Mathematics of Computation, 87 (2018), pp.~659--692.
+\bibitem{cusparse-web-page}
+{\sc NVIDA}, {\em {cuSPARSE} web page}.
+\newblock \url{https://docs.nvidia.com/cuda/cusparse/}, 2020.
+
+\bibitem{oh2018bddc}
+{\sc D.-S. Oh, O.~B. Widlund, S.~Zampini, and C.~Dohrmann}, {\em {BDDC
+  algorithms with deluxe scaling and adaptive selection of primal constraints
+  for Raviart--Thomas vector fields}}, Mathematics of Computation, 87 (2018),
+  pp.~659--692.
 
@@
 
-\bibitem{zampiniiga3}
-{\sc L.~F. Pavarino, S.~Scacchi, O.~B. Widlund and  S.~Zampini},
-  {\em Isogeometric {BDDC} deluxe preconditioners for linear elasticity},
-  Mathematical Models and Methods in Applied Sciences, 28 (2018), pp.~1337--1370.
+\bibitem{pechstein2012finite}
+{\sc C.~Pechstein}, {\em {Finite and Boundary Element Tearing and
+  Interconnecting Solvers for Multiscale Problems}}, vol.~90 of Lecture Notes
+  in Computational Science and Engineering, Springer, 2012.
 
@@
   pp.~273--336.
+
+\bibitem{poulson2013elemental}
+{\sc J.~Poulson, B.~Marker, R.~A. Van~de Geijn, J.~R. Hammond, and N.~A.
+  Romero}, {\em {Elemental: A new framework for distributed memory dense matrix
+  computations}}, ACM Transactions on Mathematical Software, 39 (2013).
+
+\bibitem{quarteroni1999domain}
+{\sc A.~Quarteroni and A.~Valli}, {\em Domain decomposition methods for partial
+  differential equations}, vol.~10, Clarendon Press, 1999.
 
@@
 
-\bibitem{schwarz1870alternating}
-{\sc H.~Schwarz}, {\em {\"Uber einen Grenz\"ubergang durch alternierendes
-  Verfahren}}, Vierteljahrsschrift der Naturforschenden Gesellschaft in
-  Z\"urich, 15 (1870), pp.~272--286.
-
 \bibitem{si2013tet}
@@
 
+\bibitem{smith2004domain}
+{\sc B.~F. Smith, P.~Bj{\o}rstad, and W.~Gropp}, {\em Domain decomposition:
+  parallel multilevel methods for elliptic partial differential equations},
+  Cambridge University Press, 2004.
+
 \bibitem{soodhalter2014krylov}
@@
 
+\bibitem{spillane2013automatic}
+{\sc N.~Spillane and D.~J. Rixen}, {\em Automatic spectral coarse spaces for
+  robust finite element tearing and interconnecting and balanced domain
+  decomposition algorithms}, International Journal for Numerical Methods in
+  Engineering, 95 (2013), pp.~953--990.
+
 \bibitem{stathopoulos2002block}
@@
 
-\bibitem{olofbook}
-{\sc A.~Toselli and O.~B. Widlund},
-  {\em Domain decomposition methods: algorithms and theory},
-  Springer Science \& Business Media, 2006.
+\bibitem{toselli2005domain}
+{\sc A.~Toselli and O.~B. Widlund}, {\em Domain decomposition methods:
+  algorithms and theory}, vol.~34 of Series in Computational Mathematics,
+  Springer, 2005.
 
@@
 \bibitem{zampini2016pcbddc}
-{\sc S.~Zampini}, {\em {{PCBDDC}: a class of robust dual--primal methods in
+{\sc S.~Zampini}, {\em {PCBDDC: a class of robust dual--primal methods in
   PETSc}}, SIAM Journal on Scientific Computing, 38 (2016), pp.~S282--S306,
@@
 
-\bibitem{zampinihcurl1}
-{\sc S.~Zampini},
-  {\em Adaptive {BDDC} deluxe methods for {H}(curl)},
-  Domain Decomposition Methods in Science and Engineering XXIII, 2017, pp.~285--292.
-
-\bibitem{zampinidarcy}
-{\sc S.~Zampini and X.~Tu},
-  {\em Multilevel Balancing Domain Decomposition by Constraints Deluxe Algorithms with Adaptive Coarse Spaces for Flow in Porous Media},
-  SIAM Journal on Scientific Computing, 39 (2107), pp.~A1389--A1415.
-
-\bibitem{zampinihcurl2}
-{\sc S.~Zampini, P.~Vassilevski, V.~Dobrev and T.~Kolev},
-  {\em Balancing Domain Decomposition by Constraints Algorithms for Curl-Conforming Spaces of Arbitrary Order},
-  Domain Decomposition Methods in Science and Engineering XXIV, (2019).
+\bibitem{zampini2017balancing}
+{\sc S.~Zampini, P.~Vassilevski, V.~Dobrev, and T.~Kolev}, {\em {Balancing
+  Domain Decomposition by Constraints Algorithms for Curl-Conforming Spaces of
+  Arbitrary Order}}, in International Conference on Domain Decomposition
+  Methods, Springer, 2017, pp.~103--116.
 

### 2020-07-08 article.tex
Missing caption + scripts

@@
   \subsection{Interfacing HPDDM Krylov methods in PETSc\label{sec:interface-ksp}}\
-In this section we provide details of the
+In this section, we provide details of the
 interface between HPDDM Krylov methods and the various PETSc
@@
     \end{tikzpicture}
-    \caption{\label{fig:scaling-Ax}}
+    \caption{Scalability of \pk{MatMult} with \pk{MatAIJMKL} and different number of processes.\label{fig:scaling-Ax}}
 \end{figure}
@@
 instrastucture, the relevant information is automatically computed
-by PETSc. In \cref{fig:p4est} we provide an example of non-overlapping (left) and overlapping (right)
-decomposition using a non-conforming grid with \pk{DMPLEX}~\cite{knepley2009mesh}. The
+by PETSc. \Cref{fig:nonoverlapping} (resp.~\cref{fig:overlapping}) is an
+example of a non-overlapping (resp. overlapping) decomposition using a
+non-conforming grid with \pk{DMPLEX}~\cite{knepley2009mesh}. The
 mesh has been constructed internally by \pk{DMPLEX} using a random pattern of non-conforming 2:1 refinements
@@
   \centering
-    \subfloat[Initial nonoverlapping decomposition.]{
+    \subfloat[Initial non-overlapping decomposition.\label{fig:nonoverlapping}]{
         \includegraphics[width=0.45\textwidth]{Data/p4est_nonoverlapping.pdf}
     }
-    \subfloat[Overlapping decomposition on which auxiliary operators $\{\mathring{A}_i\}_{i=1}^N$ are assembled.]{
+    \subfloat[Overlapping decomposition on which auxiliary operators $\{\mathring{A}_i\}_{i=1}^N$ are assembled.\label{fig:overlapping}]{
         \includegraphics[width=0.45\textwidth]{Data/p4est_overlapping.pdf}
@@
 \bibliographystyle{siamplain}
-\begin{thebibliography}{10}
-
-\bibitem{adams2004ultra}
-{\sc M.~Adams, H.~H. Bayraktar, T.~M. Keaveny, and P.~Papadopoulos}, {\em
-  {Ultrascalable Implicit Finite Element Analyses in Solid Mechanics with over
-  a Half a Billion Degrees of Freedom}}, in Proceedings of the 2004 ACM/IEEE
-  Conference on Supercomputing, SC04, IEEE Computer Society, 2004,
-  pp.~\mbox{34:1--34:15}.
-
-\bibitem{aldaas2019multi}
-{\sc H.~Al~Daas, L.~Grigori, P.~Jolivet, and P.-H. Tournier}, {\em {A
-  Multilevel Schwarz Preconditioner Based on a Hierarchy of Robust Coarse
-  Spaces}}, Journal of Scientific Computing,  (2019), p.~submitted for
-  publication, \url{https://github.com/prj-/aldaas2019multi}.
-
-\bibitem{amestoy2001fully}
-{\sc P.~Amestoy, I.~Duff, J.-Y. L'Excellent, and J.~Koster}, {\em A fully
-  asynchronous multifrontal solver using distributed dynamic scheduling}, SIAM
-  Journal on Matrix Analysis and Applications, 23 (2001), pp.~15--41,
-  \url{http://mumps.enseeiht.fr}.
-
-\bibitem{anderson2019mfem}
-{\sc R.~Anderson, J.~Andrej, A.~Barker, J.~Bramwell, J.-S. Camier, J.~Cerveny,
-  V.~Dobrev, Y.~Dudouit, A.~Fisher, T.~Kolev, et~al.}, {\em {MFEM: a modular
-  finite element methods library}}, arXiv:1911.09220,  (2019),
-  \url{http://mfem.org}.
-
-\bibitem{badia2014highly}
-{\sc S.~Badia, A.~F. Mart{\'\i}n, and J.~Principe}, {\em A highly scalable
-  parallel implementation of balancing domain decomposition by constraints},
-  SIAM Journal on Scientific Computing, 36 (2014), pp.~C190--C218.
-
-\bibitem{baker2005technique}
-{\sc A.~H. Baker, E.~R. Jessup, and T.~Manteuffel}, {\em {A Technique for
-  Accelerating the Convergence of Restarted GMRES}}, SIAM Journal on Matrix
-  Analysis and Applications, 26 (2005), pp.~962--984.
-
-\bibitem{balay1997cient}
-{\sc S.~Balay, W.~Gropp, L.~Curfman~McInnes, and B.~F. Smith}, {\em Efficient
-  management of parallelism in object-oriented numerical software libraries},
-  Modern Software Tools in Scientific Computing,  (1997), pp.~163--202.
-
-\bibitem{barrett1994templates}
-{\sc R.~Barrett, M.~Berry, T.~F. Chan, J.~Demmel, J.~Donato, J.~Dongarra,
-  V.~Eijkhout, R.~Pozo, C.~Romine, and H.~Van~der Vorst}, {\em Templates for
-  the solution of linear systems: building blocks for iterative methods}, SIAM,
-  1994.
-
-\bibitem{bastian1998additive}
-{\sc P.~Bastian, G.~Wittum, and W.~Hackbusch}, {\em Additive and multiplicative
-  multi-grid—a comparison}, Computing, 60 (1998), pp.~345--364.
-
-\bibitem{bavier2012amesos2}
-{\sc E.~Bavier, M.~Hoemmen, S.~Rajamanickam, and H.~Thornquist}, {\em {Amesos2
-  and Belos: Direct and Iterative Solvers for Large Sparse Linear Systems}},
-  Scientific Programming, 20 (2012), pp.~241--255.
-
-\bibitem{Benzi2011a}
-{\sc M.~Benzi, M.~A. Olshanskii, and Z.~Wang}, {\em {Modified augmented
-  Lagrangian preconditioners for the incompressible Navier--Stokes equations}},
-  International Journal for Numerical Methods in Fluids, 66 (2011),
-  pp.~486--508.
-
-\bibitem{10.1145/3232850}
-{\sc W.~Boukaram, G.~Turkiyyah, and D.~Keyes}, {\em {Hierarchical Matrix
-  Operations on GPUs: Matrix-Vector Multiplication and Compression}}, ACM
-  Transactions on Mathematical Software, 45 (2019).
-
-\bibitem{brown2012composable}
-{\sc J.~Brown, M.~G. Knepley, D.~A. May, L.~Curfman~McInnes, and B.~F. Smith},
-  {\em Composable linear solvers for multiphysics}, in 2012 11th International
-  Symposium on Parallel and Distributed Computing, IEEE, 2012, pp.~55--62.
-
-\bibitem{BursteddeWilcoxGhattas11}
-{\sc C.~Burstedde, L.~C. Wilcox, and O.~Ghattas}, {\em {\texttt{p4est}}:
-  Scalable algorithms for parallel adaptive mesh refinement on forests of
-  octrees}, SIAM Journal on Scientific Computing, 33 (2011), pp.~1103--1133,
-  \url{http://www.p4est.org}.
-
-\bibitem{butler2020high}
-{\sc R.~Butler, T.~Dodwell, A.~Reinarz, A.~Sandhu, R.~Scheichl, and
-  L.~Seelinger}, {\em High-performance dune modules for solving large-scale,
-  strongly anisotropic elliptic problems with applications to aerospace
-  composites}, Computer Physics Communications, 249 (2020).
-
-\bibitem{cai1999restricted}
-{\sc X.-C. Cai and M.~Sarkis}, {\em {A restricted additive Schwarz
-  preconditioner for general sparse linear systems}}, SIAM Journal on
-  Scientific Computing, 21 (1999), pp.~792--797.
-
-\bibitem{calandra2012flexible}
-{\sc H.~Calandra, S.~Gratton, J.~Langou, X.~Pinel, and X.~Vasseur}, {\em
-  {Flexible Variants of Block Restarted GMRES Methods with Application to
-  Geophysics}}, SIAM Journal on Scientific Computing, 34 (2012),
-  pp.~A714--A736.
-
-\bibitem{carvalho2011flexible}
-{\sc L.~M. Carvalho, S.~Gratton, R.~Lago, and X.~Vasseur}, {\em {A Flexible
-  Generalized Conjugate Residual Method With Inner Orthogonalization and
-  Deflated Restarting}}, SIAM Journal on Matrix Analysis and Applications, 32
-  (2011), pp.~1212--1235.
-
-\bibitem{da2014isogeometric}
-{\sc L.~B. Da~Veiga, L.~F. Pavarino, S.~Scacchi, O.~B. Widlund, and
-  S.~Zampini}, {\em {Isogeometric BDDC preconditioners with deluxe scaling}},
-  SIAM Journal on Scientific Computing, 36 (2014), pp.~A1118--A1139.
-
-\bibitem{davis2004algorithm}
-{\sc T.~A. Davis}, {\em Algorithm 832: {UMFPACK}—an unsymmetric-pattern
-  multifrontal method}, ACM Transactions on Mathematical Software, 30 (2004),
-  pp.~196--199.
-
-\bibitem{dolean2014ddm}
-{\sc V.~Dolean, P.~Jolivet, and F.~Nataf}, {\em An Introduction to Domain
-  Decomposition Methods\col\ Algorithms, Theory and Parallel Implementation},
-  SIAM, 2015.
-
-\bibitem{erhel1996restarted}
-{\sc J.~Erhel, K.~Burrage, and B.~Pohl}, {\em {Restarted GMRES preconditioned
-  by deflation}}, Journal of Computational and Applied Mathematics, 69 (1996),
-  pp.~303--318.
-
-\bibitem{falgout2002hypre}
-{\sc R.~Falgout and U.~Yang}, {\em \emph{hypre}: a library of high performance
-  preconditioners}, Computational Science---ICCS 2002,  (2002), pp.~632--641,
-  \url{https://www.llnl.gov/casc/hypre}.
-
-\bibitem{frommer2017block}
-{\sc A.~Frommer, K.~Lund, and D.~B. Szyld}, {\em {Block Krylov Subspace Methods
-  for Functions of Matrices}}, Electronic Transactions on Numerical Analysis,
-  47 (2017), pp.~100--126.
-
-\bibitem{gander2006optimized}
-{\sc M.~J. Gander}, {\em Optimized {S}chwarz {M}ethods}, SIAM Journal on
-  Numerical Analysis, 44 (2006), pp.~699--731.
-
-\bibitem{gee2006ml}
-{\sc M.~W. Gee, C.~M. Siefert, J.~J. Hu, R.~S. Tuminaro, and M.~G. Sala}, {\em
-  {ML} 5.0 smoothed aggregation user's guide}, Tech. Report SAND2006-2649,
-  Sandia National Laboratories, 2006, \url{https://trilinos.github.io/ml.html}.
-
-\bibitem{gutknecht2006block}
-{\sc M.~H. Gutknecht}, {\em Block {K}rylov space methods for linear systems
-  with multiple right-hand sides: an introduction}, in Modern Mathematical
-  Models, Methods and Algorithms for Real World Systems, A.~Siddiqui, I.~Duff,
-  and O.~Christensen, eds., 2006, pp.~420--447.
-
-\bibitem{gutknecht2008updating}
-{\sc M.~H. Gutknecht and T.~Schmelzer}, {\em {Updating the $QR$ decomposition
-  of block tridiagonal and block Hessenberg matrices}}, {Applied Numerical
-  Mathematics}, 58 (2008), pp.~871--883.
-
-\bibitem{haferssas2017soras}
-{\sc R.~Haferssas, P.~Jolivet, and F.~Nataf}, {\em {An additive Schwarz method
-  type theory for Lions's algorithm and a Symmetrized Optimized Restricted
-  Additive Schwarz Method}}, Journal on Scientific Computing, 39 (2017),
-  pp.~A1345--A1365.
-
-\bibitem{hecht2012new}
-{\sc F.~Hecht}, {\em New development in {FreeFem++}}, Journal of Numerical
-  Mathematics, 20 (2012), pp.~251--266, \url{http://freefem.org}.
-
-\bibitem{heinecke2016libxsmm}
-{\sc A.~Heinecke, G.~Henry, M.~Hutchinson, and H.~Pabst}, {\em {LIBXSMM:
-  accelerating small matrix multiplications by runtime code generation}}, in
-  Proceedings of the 2016 International Conference for High Performance
-  Computing, Networking, Storage and Analysis, SC16, IEEE, 2016,
-  \url{https://github.com/hfp/libxsmm}.
-
-\bibitem{Heinlein:2016:PIT}
-{\sc A.~Heinlein, A.~Klawonn, and O.~Rheinbach}, {\em A parallel implementation
-  of a two-level overlapping {S}chwarz method with energy-minimizing coarse
-  space based on {T}rilinos}, SIAM Journal on Scientific Computing, 38 (2016),
-  pp.~C713--C747.
-
-\bibitem{hernandez2005ssf}
-{\sc V.~Hernandez, J.~E. Roman, and V.~Vidal}, {\em {SLEP}c: A scalable and
-  flexible toolkit for the solution of eigenvalue problems}, ACM Transactions
-  on Mathematical Software, 31 (2005), pp.~351--362,
-  \url{https://slepc.upv.es}.
-
-\bibitem{heroux2005overview}
-{\sc M.~A. Heroux, R.~A. Bartlett, V.~E. Howle, R.~J. Hoekstra, J.~J. Hu, T.~G.
-  Kolda, R.~B. Lehoucq, K.~R. Long, R.~P. Pawlowski, E.~T. Phipps, et~al.},
-  {\em {An overview of the Trilinos project}}, ACM Transactions on Mathematical
-  Software (TOMS), 31 (2005), pp.~397--423, \url{https://trilinos.github.io}.
-
-\bibitem{hestenes1952methods}
-{\sc M.~R. Hestenes and E.~Stiefel}, {\em {Methods of Conjugate Gradients for
-  Solving Linear Systems}}, Journal of Research of the National Bureau of
-  Standards, 49 (1952), pp.~409--436.
-
-\bibitem{hoefler2015scientific}
-{\sc T.~Hoefler and R.~Belli}, {\em Scientific benchmarking of parallel
-  computing systems: twelve ways to tell the masses when reporting performance
-  results}, in Proceedings of the 2015 International Conference for High
-  Performance Computing, Networking, Storage and Analysis, SC15, 2015.
-
-\bibitem{mkl-web-page}
-{\sc Intel}, {\em {MKL} web page}.
-\newblock
-  \url{https://software.intel.com/content/www/us/en/develop/tools/math-kernel-library.html},
-  2020.
-
-\bibitem{ji2017breakdown}
-{\sc H.~Ji and Y.~Li}, {\em A breakdown-free block conjugate gradient method},
-  BIT Numerical Mathematics, 57 (2017), pp.~379--403.
-
-\bibitem{jolivet2013scalable}
-{\sc P.~Jolivet, F.~Hecht, F.~Nataf, and C.~Prud'homme}, {\em {Scalable Domain
-  Decomposition Preconditioners for Heterogeneous Elliptic Problems}}, in
-  Proceedings of the International Conference on High Performance Computing,
-  Networking, Storage and Analysis, SC13, ACM, 2013.
-
-\bibitem{jolivet2016block}
-{\sc P.~Jolivet and P.-H. Tournier}, {\em {Block Iterative Methods and
-  Recycling for Improved Scalability of Linear Solvers}}, in Proceedings of the
-  2016 International Conference for High Performance Computing, Networking,
-  Storage and Analysis, SC16, IEEE, 2016.
-
-\bibitem{KALANTZIS2018136}
-{\sc V.~Kalantzis, A.~C.~I. Malossi, C.~Bekas, A.~Curioni, E.~Gallopoulos, and
-  Y.~Saad}, {\em A scalable iterative dense linear system solver for multiple
-  right-hand sides in data analytics}, Parallel Computing, 74 (2018),
-  pp.~136--153.
-
-\bibitem{karypis1998fast}
-{\sc G.~Karypis and V.~Kumar}, {\em A fast and high quality multilevel scheme
-  for partitioning irregular graphs}, SIAM Journal on Scientific Computing, 20
-  (1998), pp.~359--392,
-  \url{http://glaros.dtc.umn.edu/gkhome/metis/parmetis/overview}.
-
-\bibitem{knepley2009mesh}
-{\sc M.~G. Knepley and D.~A. Karpeev}, {\em {Mesh algorithms for PDE with Sieve
-  I: Mesh distribution}}, Scientific Programming, 17 (2009), pp.~215--230.
-
-\bibitem{knyazev2001toward}
-{\sc A.~V. Knyazev}, {\em {Toward the optimal preconditioned eigensolver:
-  Locally optimal block preconditioned conjugate gradient method}}, SIAM
-  journal on scientific computing, 23 (2001), pp.~517--541.
-
-\bibitem{lehoucq1998arpack}
-{\sc R.~Lehoucq, D.~Sorensen, and C.~Yang}, {\em {ARPACK} users’ guide\col\
-  solution of large-scale eigenvalue problems with implicitly restarted Arnoldi
-  methods}, vol.~6, Society for Industrial and Applied Mathematics, 1998,
-  \url{http://www.caam.rice.edu/software/ARPACK}.
-
-\bibitem{li2005overview}
-{\sc X.~S. Li}, {\em {An overview of SuperLU: Algorithms, implementation, and
-  user interface}}, ACM Transactions on Mathematical Software, 31 (2005),
-  pp.~302--325.
-
-\bibitem{li2003superlu_dist}
-{\sc X.~S. Li and J.~Demmel}, {\em {SuperLU\_DIST: A scalable
-  distributed-memory sparse direct solver for unsymmetric linear systems}}, ACM
-  Transactions on Mathematical Software, 29 (2003), pp.~110--140.
-
-\bibitem{mandel1993balancing}
-{\sc J.~Mandel}, {\em Balancing domain decomposition}, Communications in
-  Numerical Methods in Engineering, 9 (1993), pp.~233--241.
-
-\bibitem{marchand2019two}
-{\sc P.~Marchand, X.~Claeys, P.~Jolivet, F.~Nataf, and P.-H. Tournier}, {\em
-  {Two-level preconditioning for $h$-version boundary element approximation of
-  hypersingular operator with GenEO}}, Numerische Mathematik,  (2019),
-  p.~submitted for publication.
-
-\bibitem{mathew2008domain}
-{\sc T.~Mathew}, {\em Domain decomposition methods for the numerical solution
-  of partial differential equations}, vol.~61, Springer Science \& Business
-  Media, 2008.
-
-\bibitem{may2016extreme}
-{\sc D.~A. May, P.~Sanan, K.~Rupp, M.~G. Knepley, and B.~F. Smith}, {\em
-  {Extreme-scale multigrid components within PETSc}}, in Proceedings of the
-  Platform for Advanced Scientific Computing Conference, 2016.
-
-\bibitem{moulin2018al}
-{\sc J.~Moulin, P.~Jolivet, and O.~Marquet}, {\em {Augmented Lagrangian
-  Preconditioner for Large-Scale Hydrodynamic Stability Analysis}}, Computer
-  Methods in Applied Mechanics and Engineering, 351 (2019), pp.~718--743,
-  \url{https://github.com/prj-/moulin2019al}.
-
-\bibitem{notay2000flexible}
-{\sc Y.~Notay}, {\em Flexible conjugate gradients}, SIAM Journal on Scientific
-  Computing, 22 (2000), pp.~1444--1460.
-
-\bibitem{cusparse-web-page}
-{\sc NVIDA}, {\em {cuSPARSE} web page}.
-\newblock \url{https://docs.nvidia.com/cuda/cusparse/}, 2020.
-
-\bibitem{oh2018bddc}
-{\sc D.-S. Oh, O.~B. Widlund, S.~Zampini, and C.~Dohrmann}, {\em {BDDC
-  algorithms with deluxe scaling and adaptive selection of primal constraints
-  for Raviart--Thomas vector fields}}, Mathematics of Computation, 87 (2018),
-  pp.~659--692.
-
-\bibitem{o1980block}
-{\sc D.~P. O'Leary}, {\em {The Block Conjugate Gradient Algorithm and Related
-  Methods}}, Linear Algebra and its Applications, 29 (1980), pp.~293--322.
-
-\bibitem{parks2006recycling}
-{\sc M.~L. Parks, E.~de~Sturler, G.~Mackey, D.~D. Johnson, and S.~Maiti}, {\em
-  {Recycling Krylov Subspaces for Sequences of Linear Systems}}, SIAM Journal
-  on Scientific Computing, 28 (2006), pp.~1651--1674.
-
-\bibitem{pechstein2012finite}
-{\sc C.~Pechstein}, {\em {Finite and Boundary Element Tearing and
-  Interconnecting Solvers for Multiscale Problems}}, vol.~90 of Lecture Notes
-  in Computational Science and Engineering, Springer, 2012.
-
-\bibitem{pechstein2017unified}
-{\sc C.~Pechstein and C.~R. Dohrmann}, {\em {A unified framework for adaptive
-  BDDC}}, Electronic Transactions on Numerical Analysis, 46 (2017),
-  pp.~273--336.
-
-\bibitem{poulson2013elemental}
-{\sc J.~Poulson, B.~Marker, R.~A. Van~de Geijn, J.~R. Hammond, and N.~A.
-  Romero}, {\em {Elemental: A new framework for distributed memory dense matrix
-  computations}}, ACM Transactions on Mathematical Software, 39 (2013).
-
-\bibitem{quarteroni1999domain}
-{\sc A.~Quarteroni and A.~Valli}, {\em Domain decomposition methods for partial
-  differential equations}, vol.~10, Clarendon Press, 1999.
-
-\bibitem{saad1993flexible}
-{\sc Y.~Saad}, {\em {A Flexible Inner--Outer Preconditioned GMRES Algorithm}},
-  SIAM Journal on Scientific Computing, 14 (1993), pp.~461--469.
-
-\bibitem{saad2003iterative}
-{\sc Y.~Saad}, {\em {Iterative Methods for Sparse Linear Systems}}, SIAM, 2003.
-
-\bibitem{saad1986gmres}
-{\sc Y.~Saad and M.~H. Schultz}, {\em {GMRES: a generalized minimal residual
-  algorithm for solving nonsymmetric linear systems}}, SIAM Journal on
-  Scientific and Statistical Computing, 7 (1986), pp.~856--869.
-
-\bibitem{saad2000deflated}
-{\sc Y.~Saad, M.~Yeung, J.~Erhel, and F.~Guyomarc'h}, {\em A deflated version
-  of the conjugate gradient algorithm}, SIAM Journal on Scientific Computing,
-  21 (2000), pp.~1909--1926.
-
-\bibitem{sakurai2003projection}
-{\sc T.~Sakurai and H.~Sugiura}, {\em A projection method for generalized
-  eigenvalue problems using numerical integration}, Journal of Computational
-  and Applied Mathematics, 159 (2003), pp.~119--128.
-
-\bibitem{sakurai2010application}
-{\sc T.~Sakurai, H.~Tadano, and Y.~Kuramashi}, {\em {Application of block
-  Krylov subspace algorithms to the Wilson--Dirac equation with multiple
-  right-hand sides in lattice QCD}}, Computer Physics Communications, 181
-  (2010), pp.~113--117.
-
-\bibitem{si2013tet}
-{\sc H.~Si}, {\em {TetGen: A Quality Tetrahedral Mesh Generator and 3D Delaunay
-  Triangulator}}, Tech. Report~13, 2013,
-  \url{http://wias-berlin.de/software/tetgen}.
-
-\bibitem{vsistek2013parallel}
-{\sc J.~{\v{S}}{\'\i}stek, J.~Mandel, B.~Soused{\'\i}k, and P.~Burda}, {\em
-  Parallel implementation of multilevel {BDDC}}, in Numerical Mathematics and
-  Advanced Applications 2011, Springer, 2013, pp.~681--689,
-  \url{http://users.math.cas.cz/~sistek/software/bddcml.html}.
-
-\bibitem{smith2004domain}
-{\sc B.~F. Smith, P.~Bj{\o}rstad, and W.~Gropp}, {\em Domain decomposition:
-  parallel multilevel methods for elliptic partial differential equations},
-  Cambridge University Press, 2004.
-
-\bibitem{soodhalter2014krylov}
-{\sc K.~M. Soodhalter, D.~B. Szyld, and F.~Xue}, {\em Krylov subspace recycling
-  for sequences of shifted linear systems}, Applied Numerical Mathematics, 81
-  (2014), pp.~105--118.
-
-\bibitem{spillane2011robust}
-{\sc N.~Spillane, V.~Dolean, P.~Hauret, F.~Nataf, C.~Pechstein, and
-  R.~Scheichl}, {\em A robust two-level domain decomposition preconditioner for
-  systems of {PDE}s}, Comptes Rendus Mathématique, 349 (2011), pp.~1255--1259.
-
-\bibitem{spillane2013abstract}
-{\sc N.~Spillane, V.~Dolean, P.~Hauret, F.~Nataf, C.~Pechstein, and
-  R.~Scheichl}, {\em Abstract robust coarse spaces for systems of {PDEs} via
-  generalized eigenproblems in the overlaps}, Numerische Mathematik, 126
-  (2013), pp.~741--700.
-
-\bibitem{spillane2013automatic}
-{\sc N.~Spillane and D.~J. Rixen}, {\em Automatic spectral coarse spaces for
-  robust finite element tearing and interconnecting and balanced domain
-  decomposition algorithms}, International Journal for Numerical Methods in
-  Engineering, 95 (2013), pp.~953--990.
-
-\bibitem{stathopoulos2002block}
-{\sc A.~Stathopoulos and K.~Wu}, {\em {A Block Orthogonalization Procedure with
-  Constant Synchronization Requirements}}, SIAM Journal on Scientific
-  Computing, 23 (2002), pp.~2165--2182.
-
-\bibitem{Stewart2002}
-{\sc G.~W. Stewart}, {\em {A Krylov--Schur Algorithm for Large Eigenproblems}},
-  SIAM Journal on Matrix Analysis and Applications, 23 (2002), pp.~601--614.
-
-\bibitem{tang2009comparison}
-{\sc J.~Tang, R.~Nabben, C.~Vuik, and Y.~Erlangga}, {\em Comparison of
-  two-level preconditioners derived from deflation, domain decomposition and
-  multigrid methods}, Journal of Scientific Computing, 39 (2009), pp.~340--370.
-
-\bibitem{toselli2005domain}
-{\sc A.~Toselli and O.~B. Widlund}, {\em Domain decomposition methods:
-  algorithms and theory}, vol.~34 of Series in Computational Mathematics,
-  Springer, 2005.
-
-\bibitem{tournier2016micro}
-{\sc P.-H. Tournier, I.~Aliferis, M.~Bonazzoli, M.~de~Buhan, M.~Darbas,
-  V.~Dolean, F.~Hecht, P.~Jolivet, I.~El~Kanfoud, C.~Migliaccio, F.~Nataf,
-  C.~Pichot, and S.~Semenov}, {\em {Microwave Tomographic Imaging of
-  Cerebrovascular Accidents by Using High-Performance Computing}}, Parallel
-  Computing, 85 (2019), pp.~88--97.
-
-\bibitem{van1992bi}
-{\sc H.~A. Van~der Vorst}, {\em {Bi-CGSTAB: a fast and smoothly converging
-  variant of Bi-CG for the solution of nonsymmetric linear systems}}, SIAM
-  Journal on Scientific and Statistical Computing, 13 (1992), pp.~631--644.
-
-\bibitem{wakam2013158}
-{\sc D.~N. Wakam and F.~Pacull}, {\em Memory efficient hybrid algebraic solvers
-  for linear systems arising from compressible flows}, Computers \& Fluids, 80
-  (2013), pp.~158--167.
-
-\bibitem{zampini2016pcbddc}
-{\sc S.~Zampini}, {\em {PCBDDC: a class of robust dual--primal methods in
-  PETSc}}, SIAM Journal on Scientific Computing, 38 (2016), pp.~S282--S306,
-  \url{https://www.mcs.anl.gov/petsc/petsc-master/docs/manualpages/PC/PCBDDC.html}.
-
-\bibitem{zampini2017balancing}
-{\sc S.~Zampini, P.~Vassilevski, V.~Dobrev, and T.~Kolev}, {\em {Balancing
-  Domain Decomposition by Constraints Algorithms for Curl-Conforming Spaces of
-  Arbitrary Order}}, in International Conference on Domain Decomposition
-  Methods, Springer, 2017, pp.~103--116.
-
-\end{thebibliography}
+\bibliography{article.bib , ../../2018/SC18/bare_conf.bib , ../../2016/SC16/bare_conf.bib , ../../../Courses/2019/Tokyo/main.bib , ../../../Bibliography/main.bib , ../../../CV/cv.bib}
 \end{document}

### 2020-07-09 article.tex
Missing bibliography

--- 2020-07-08 article.tex
@@
 \bibitem{cusparse-web-page}
-{\sc NVIDA}, {\em {cuSPARSE} web page}.
+{\sc NVIDIA}, {\em {cuSPARSE} web page}.
 \newblock \url{https://docs.nvidia.com/cuda/cusparse/}, 2020.

### 2020-07-10 MatProduct.c
Cleanup code

@@ -1,16 +1,36 @@
+/*
+   Author(s): Pierre Jolivet <pierre.jolivet@enseeiht.fr>
+              Jose Roman <jroman@dsic.upv.es>
+              Stefano Zampini <stefano.zampini@kaust.edu.sa>
+        Date: 2020-07-10
+
+   Copyright (C) 2020-     Centre National de la Recherche Scientifique
+                 2020-     Universitat Politècnica de València
+                 2020-     King Abdullah University of Science and Technology
+
+   This script is distributed in the hope that it will be useful,
+   but WITHOUT ANY WARRANTY; without even the implied warranty of
+   MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.
+
+   If you use this script, you are kindly asked to cite the following article:
+
+   "KSPHPDDM and PCHPDDM: Extending PETSc with Robust Overlapping Schwarz Preconditioners and Advanced Krylov Methods",
+   P. Jolivet, J. E. Roman, and S. Zampini (2020).
+ */
+
 #include <petsc.h>
 
 #if defined(PETSC_HAVE_MKL)
 #include <mkl.h>
-#define PetscStackCallMKLSparse(func,args) do {                                                           \
-    sparse_status_t __ierr;                                                                               \
-    PetscStackPush(#func);                                                                                \
-    __ierr = func args;                                                                                   \
-    PetscStackPop;                                                                                        \
+#define PetscStackCallMKLSparse(func,args) do {              \
+    sparse_status_t __ierr;                                  \
+    PetscStackPush(#func);                                   \
+    __ierr = func args;                                      \
+    PetscStackPop;                                           \
     if (__ierr != SPARSE_STATUS_SUCCESS) SETERRQ2(PETSC_COMM_SELF,PETSC_ERR_LIB,"Error in %s(): error code %d",#func,(int)__ierr); \
   } while (0)
 #else
-#define PetscStackCallMKLSparse(func,args) do {                                                           \
+#define PetscStackCallMKLSparse(func,args) do {              \
     SETERRQ(PETSC_COMM_SELF,PETSC_ERR_SUP,"No MKL support"); \
   } while (0)
 #endif

### 2020-07-10 article.tex
Add myself as a co-author
Add xSDK reference

@@
   pdftitle={KSPHPDDM and PCHPDDM: Extending PETSc with Robust Overlapping Schwarz Preconditioners and Advanced Krylov Methods},
-  pdfauthor={P. Jolivet and S. Zampini}
+  pdfauthor={P. Jolivet, J. E. Roman, and S. Zampini}
 }
@@
 such applications requires a rich ecosystem of state-of-the-art reusable libraries
-\cite{xsdk}. In this paper, we focus on the interoperability of two such libraries.
+\cite{Bartlett:2017:XFT}. In this paper, we focus on the interoperability of two such libraries.
 
@@
 \section*{Acknowledgments}
-The authors would like to thank S. Balay, J. Brown, V. Hapla, M. Knepley, J.
-Roman, and B. Smith for reviewing the successive merge requests in PETSc and
-SLEPc repositories. This work was granted access to the GENCI-sponsored HPC resources of:
+The authors would like to thank S. Balay, J. Brown, V. Hapla, M. Knepley,
+and B. Smith for reviewing the successive merge requests in the PETSc
+repository. This work was granted access to the GENCI-sponsored HPC resources of:
 \begin{itemize}
@@
 
+\bibitem{Bartlett:2017:XFT}
+{\sc R.~Bartlett, I.~Demeshko, T.~Gamblin, G.~Hammond, M.~A. Heroux,
+  J.~Johnson, A.~Klinvex, X.~Li, L.~C. McInnes, J.~D. Moulton, D.~Osei-Kuffuor,
+  J.~Sarich, B.~Smith, J.~Willenbring, and U.~M. Yang}, {\em {xSDK}
+  foundations: Toward an extreme-scale scientific software development kit},
+  Supercomputing Frontiers and Innovations, 4 (2017),
+  \url{https://doi.org/10.14529/jsfi170104}.
+
 \bibitem{bastian1998additive}

### 2020-07-12 article.tex
Add SPAI and GASM + README.md

@@
     \item \pk{PCSHELL}: user-defined preconditioner, e.g., matrix-free;
+    \item \pk{PCSPAI}: sparse approximate inverse~\cite{grote1997parallel};
     \item \pk{PCHARA}~\cite{10.1145/3232850};
     \item \pk{PCHPDDM}: see~\cref{sec:PCHPDDM};\label{it:hpddm}
-    \item \pk{PCASM}: overlapping Schwarz methods;\label{it:asm}
+    \item \pk{PCASM} and \pk{PCGASM}: overlapping Schwarz methods;\label{it:asm}
     \item \pk{PCBJACOBI}: block Jacobi;\label{it:bjacobi}
@@
 This continuous equation is discretized with $\mathbb{P}_3$ finite elements using FreeFEM.
-The script is available at \url{http://jolivet.perso.enseeiht.fr/lobpcg.edp}.
+The script is available at \url{https://github.com/prj-/jolivet2020sisc}.
 Following timings and convergence histories have been obtained on $2{,}304$ processes of Ir\`ene, for solving a problem of dimension \pgfmathprintnumber{74618461}.
@@
 convergence history, which is displayed in~\cref{fig:bratu-conv}. The complete set of
-options is given in~\cref{fig:bratu-options}. Results can be reproduced using the following FreeFEM script: \url{http://jolivet.perso.enseeiht.fr/bratu-hpddm.edp}
+options is given in~\cref{fig:bratu-options}. Results can be reproduced using the FreeFEM script at \url{https://github.com/prj-/jolivet2020sisc}
 \pgfplotstableread{
@@
   J.~Johnson, A.~Klinvex, X.~Li, L.~C. McInnes, J.~D. Moulton, D.~Osei-Kuffuor,
-  J.~Sarich, B.~Smith, J.~Willenbring, and U.~M. Yang}, {\em {xSDK}
-  foundations: Toward an extreme-scale scientific software development kit},
-  Supercomputing Frontiers and Innovations, 4 (2017),
-  \url{https://doi.org/10.14529/jsfi170104}.
+  J.~Sarich, B.~Smith, J.~Willenbring, and U.~M. Yang}, {\em {xSDK Foundations:
+  Toward an Extreme-scale Scientific Software Development Kit}}, Supercomputing
+  Frontiers and Innovations, 4 (2017).
 
@@
   Sandia National Laboratories, 2006, \url{https://trilinos.github.io/ml.html}.
+
+\bibitem{grote1997parallel}
+{\sc M.~J. Grote and T.~Huckle}, {\em Parallel preconditioning with sparse
+  approximate inverses}, SIAM Journal on Scientific Computing, 18 (1997),
+  pp.~838--853.
 

### 2020-07-15 article.tex
Fix a few typos

@@
 
-One the one hand, PETSc \cite{balay1997cient}, the
+On the one hand, PETSc \cite{balay1997cient}, the
 Portable and Extensible Toolkit for Scientific computation, is a
@@
 discretize partial differential equations and solve algebraic linear or
-nonlinear systems of time-dependent equations. Among its strenghts, it offers many
+nonlinear systems of time-dependent equations. Among its strengths, it offers many
 advanced features regarding preconditioning and tailored matrix formats, and it can
@@
 \cref{sec:PCHPDDM} introduces \pk{PCHPDDM}. Eventually, concluding remarks may be found
-\cref{sec:conclusion}. All PETSc keywords and options are typeset in a \texttt{typewriter} front, these
+\cref{sec:conclusion}. All PETSc keywords and options are typeset in a \texttt{typewriter} font, these
 are public and documented at \url{https://www.mcs.anl.gov/petsc/petsc-master/docs}.
@@
 via \pk{PCFactorGetMatrix}, and then call \pk{MatMatSolve}($F$, $Y$, $X$) to take advantage
-of blocked forward eliminitations and backward substitutions from the various factorization packages
+of blocked forward eliminations and backward substitutions from the various factorization packages
 interfaced with PETSc. In the case of a sparse coefficient matrix $A'$,
@@
 PARDISO or CPARDISO~\cite{mkl-web-page}, and SuperLU~\cite{li2005overview} or
-SuperLU\_DIST~\cite{li2003superlu_dist}. For the the case of a dense coefficient
+SuperLU\_DIST~\cite{li2003superlu_dist}. For the case of a dense coefficient
 matrix $A'$: ScaLAPACK and Elemental~\cite{poulson2013elemental}. 
@@
 high-level access to GenEO from PETSc and its composable solver
-infrastructrure~\cite{brown2012composable}. We note that GenEO has been
+infrastructure~\cite{brown2012composable}. We note that GenEO has been
 recently implemented in other libraries as well~\cite{butler2020high}.
@@
 than $n$, the size of $A$, and coarsening is much more aggressive than
-when using algebraic or geometric multigrids methods.
+when using algebraic or geometric multigrid methods.
 %, see~\cref{sec:linear-elasticity} for
@@
 In order to use GenEO from PETSc, the auxiliary operators $\{\mathring{A}_i\}_{i=1}^N$
-and the correspondending local to global map of degrees of freedom (see~\cref{fig:poisson})
+and the corresponding local to global map of degrees of freedom (see~\cref{fig:poisson})
 can be provided using the routine \pk{PCHPDDMSetAuxiliaryMat}.
 In special cases where the matrix has been constructed using the PETSc discretization
-instrastucture, the relevant information is automatically computed
+infrastructure, the relevant information is automatically computed
 by PETSc. \Cref{fig:nonoverlapping} (resp.~\cref{fig:overlapping}) is an
@@
     \subsubsection{System of elasticity}\label{sec:linear-elasticity}
-We first report on solving the three-dimensional system of linear elasticy with highly heterogeneous elastic modulus.
+We first report on solving the three-dimensional system of linear elasticity with highly heterogeneous elastic modulus.
 Its strong formulation is given by:
@@
 \Cref{eq:elasticity} is discretized using
-$(\mathbb{P}_1, \mathbb{P}_1, \mathbb{P}_1)$ finite elements resulting in $\pgfmathprintnumber{593} \times 10^6$ number of uknowns
+$(\mathbb{P}_1, \mathbb{P}_1, \mathbb{P}_1)$ finite elements resulting in $\pgfmathprintnumber{593} \times 10^6$ number of unknowns
 with approximately $\pgfmathprintnumber{45}$ nonzero coefficients per row.

### 2020-07-16 article.tex
Three suggested changes, fix broken URL, and several typos

@@
 used instead~\cite{gutknecht2008updating}.
-These orthonormalization variants can selected at runtime with the option \pk{-ksp\_hpddm\_qr (cholqr$|$mgs$|$cgs)}. {\color{red} API?}
+These orthonormalization variants can be selected at runtime with the option \pk{-ksp\_hpddm\_qr (cholqr$|$mgs$|$cgs)}. {\color{red} API?}
 
@@
 Eigenvalue Problem computations built on top of PETSc, were previously not
-using blocking.  It has been rewritten for version 3.14.0 and the benefits of
-this new implementation is displayed~\cref{sec:lobpcg}.
+using blocking when applying the preconditioner or solving linear systems. This has been rewritten for version 3.14.0 and the benefit of
+this new implementation is displayed in~\cref{sec:lobpcg}.
 
@@
     \subsubsection{Linear stability analysis}
-In this section, we apply recycling Krylov methods in the context linear
-stability analysis and consider the solution the following generalized eigenvalue problem:
+In this section, we apply recycling Krylov methods in the context of linear
+stability analysis and consider the solution of the following generalized eigenvalue problem:
 \begin{equation}\label{eq:lsa}
@@
 of~\cref{eq:lsa} near a complex-valued shift $\sigma$. In this paragraph, emphasis will be put
-on Krylov--Schur methods~\cite{Stewart2002}, which rely on spectral transformations
+on the Krylov--Schur method~\cite{Stewart2002}, which for interior eigenvalues relies on spectral transformations
 and on the solution of successive linear systems such as:
@@
 
-The Krylov--Schur algorithm converges in six outer iterations.
+The Krylov--Schur algorithm converges in six outer iterations (restarts), with a total of 46 inner solves~\cref{eq:ks}.
 Recycling of Krylov subspaces with \pk{KSPHPDDM} provides algorithmic speedup by lowering the number of iterations per inner solve,
@@
 product operations.
-They can be reproduced by using the mini-app \url{http://jolivet.perso.enseeiht.fr/MatProduct.c} and any input \pk{MatAIJ} stored in binary format,
+They can be reproduced by using the mini-app \url{https://github.com/prj-/jolivet2020petsc/blob/master/MatProduct.c} and any input \pk{MatAIJ} stored in binary format,
 here using the default name \pk{binaryoutput}, available at \url{http://jolivet.perso.enseeiht.fr/binaryoutput}.
@@
     \item with a block size of 1, top plot $b=1$ in~\cref{fig:perfAX}, using any type but \pk{MatAIJ}~\ref{pgfplots:aij1} is counterproductive when $N>1$, see~\ref{pgfplots:baij1} and~\ref{pgfplots:sbaij1}, even when using MKL, see~\ref{pgfplots:mklbaijseq1} and~\ref{pgfplots:mklsbaijseq1}.
-With 8 or more columns, the performance of \pk{MatAIJ} plateaus and the efficiency reaches approximately 200\%, while the performance of the MKL primitives stagnate at around 175\% see~\ref{pgfplots:aij1};
+With 8 or more columns, the performance of \pk{MatAIJ} plateaus and the efficiency reaches approximately 200\%, while the performance of the MKL primitives stagnate at around 175\%, see~\ref{pgfplots:aij1};
 for the single column case, \pk{MatSBAIJ} and \pk{MatAIJMKL} deliver slightly better performances;
@@
 
-The mini-app is then used to benchmark intranode performance of matrix--matrix multiplications, either with multiple OpenMP threads, or by offloading the operation to a GPU and use cuSPARSE~\cite{cusparse-web-page} as interfaced in PETSc.
+The mini-app is then used to benchmark intranode performance of matrix--matrix multiplications, either with multiple OpenMP threads, or by offloading the operation to a GPU and using cuSPARSE~\cite{cusparse-web-page} as interfaced in PETSc.
 \begin{Verbatim}[fontsize=\footnotesize,frame=single,framerule=0.1mm,commandchars=&\[\]]
@@
 been taken into account. For the sake of completeness, the performance of
-\pk{MatAIJMKL} with $N=1$ but different numbers of threads is also reported
+\pk{MatAIJMKL} with $N=1$ but different numbers of threads is also reported in
 \cref{fig:scaling-Ax}. Clearly, it is not advised to use the full socket to
@@
 highest percentage of peak is MatSBAIJMKL~\ref{pgfplots:sbaijmkl3} with $b=3$
-and $N=64$, right-most middle plot in~\cref{fig:perfAXpar}, with approximately
+and $N=64$, rightmost middle plot in~\cref{fig:perfAXpar}, with approximately
 63 GFLOP/s. It is still less than 1\% of peak for an Intel Xeon Gold 6248.
 For the GPU configuration, \pk{MatAIJcuSPARSE}~\ref{pgfplots:cusparse10} with $b=6$
-and $N=64$, right-most bottom plot in~\cref{fig:perfAXpar}, performs at around
+and $N=64$, rightmost bottom plot in~\cref{fig:perfAXpar}, performs at around
 306 GFLOP/s, about 4\% of peak for an NVIDIA Tesla V100 SXM2.

### 2020-07-16 article.tex
Several fixes, including URL

@@
 This continuous equation is discretized with $\mathbb{P}_3$ finite elements using FreeFEM.
-The script is available at \url{https://github.com/prj-/jolivet2020sisc}.
-Following timings and convergence histories have been obtained on $2{,}304$ processes of Ir\`ene, for solving a problem of dimension \pgfmathprintnumber{74618461}.
+The script is available at \url{https://github.com/prj-/jolivet2020petsc}.
+The following timings and convergence histories have been obtained on $2{,}304$ processes of Ir\`ene, for solving a problem of dimension \pgfmathprintnumber{74618461}.
 By default, SLEPc applies at most 5 iterations of GMRES on each column of a so-called set of active columns.
@@
 restriction-correction-interpolation operator $Q$ is applied in a block fashion
-in HPDDM, and not column by column. See the effect of this strategy~\cref{sec:lobpcg}.
+in HPDDM, and not column by column. See the effect of this strategy in~\cref{sec:lobpcg}.
 
@@
 The physical domain $\Omega$ is a beam of dimensions $[0,6] \times [0,1] \times [0,1]$.
-The Cauchy stress tensor $\sigma$ is given by Hooke’s law: it can be expressed in terms of Young’s modulus $E$ and Poisson’s ratio $\nu$.
+The Cauchy stress tensor $\sigma$ is given by Hooke’s law: it can be expressed in terms of Young’s modulus $E$ and Poisson’s ratio $\nu$,
 \[
@@
 \end{minipage} \\
-In this third configuration, $M$ is the number of subdomains used to define the second-level domain decomposition, which is an aggregation of first-level subdomain.
+In this third configuration, $M$ is the number of subdomains used to define the second-level domain decomposition, which is an aggregation of first-level subdomains.
 
@@
 convergence history, which is displayed in~\cref{fig:bratu-conv}. The complete set of
-options is given in~\cref{fig:bratu-options}. Results can be reproduced using the FreeFEM script at \url{https://github.com/prj-/jolivet2020sisc}
+options is given in~\cref{fig:bratu-options}. Results can be reproduced using the FreeFEM script at \url{https://github.com/prj-/jolivet2020petsc}
 \pgfplotstableread{
@@
 }
-    \subfloat[\parbox{5cm}{\pk{SNES}, \pk{KSP}, and \pk{PC} options for the
+    \subfloat[\parbox{5cm}{\pk{SNES}, \pk{KSP}, and \pk{PC} options for
     solving~\cref{eq:bratu}}\label{fig:bratu-options}]{
@@
 64$ subdomains is then solved using \pk{PCHPDDM}, which itself builds another adaptive
-level using $10$ local eigenvectors per subdomains. The coarsest level is
+level using $10$ local eigenvectors per subdomain. The coarsest level is
 aggregated on a single process and solved using an exact Cholesky factorization.
-The magnitude of the electric field is plotted~\cref{fig:E}. For the
-convergence history of the solver, it is~\cref{fig:BDDC}. The outer solver
-reaches the prescribed convergence criterion in 17 iterations~\ref{pgfplots:res}. The number of
+The magnitude of the electric field is plotted in~\cref{fig:E}. The
+convergence history of the solver is shown in~\cref{fig:BDDC}. The outer solver
+reaches the prescribed convergence tolerance in 17 iterations~\ref{pgfplots:res}. The number of
 inner iterations for solving the BDDC coarse problem is reported as well~\ref{pgfplots:inner}.

### 2020-07-16 article.tex
Remarks of Jose

@@
 on the application of appropriate preconditioners.
-PETSc (as of version 3.13.2) already offers 44 different types of Krylov methods within the \pk{KSP} base class. Some of the
+PETSc, as of version 3.14.0, already offers 44 different types of
+Krylov methods within the \pk{KSP} base class. Some of the
 most used types are \pk{KSPGMRES}, \pk{KSPCG}, or \pk{KSPBCGS}, which respectively implement
@@
         AX$;
-    \item \pk{PCApply}($P_{A'}$, $y$, $x$) for $x = P_{A'}y$;
-    \item \pk{PCMatApply}($P_{A'}$, $Y$, $X$) for $X = P_{A'}Y$;
+    \item \pk{PCApply}($P_{A'}$, $x$, $y$) for $y = P_{A'}x$;
+    \item \pk{PCMatApply}($P_{A'}$, $X$, $Y$) for $Y = P_{A'}X$;
 \end{itemize}
 As of version 3.14.0, PETSc provides unified routines for applying $A$ or $P_{A'}$
-to a tall-and-skinny dense matrix $Y$. However, when no specialized implementation
-is available, PETSc will loop over the columns of $Y$ and successively apply $A$
-or $P_{A'}$ to get all columns of $X$. For matrix--matrix products, there are
+to a tall-and-skinny dense matrix $X$. However, when no specialized implementation
+is available, PETSc will loop over the columns of $X$ and successively apply $A$
+or $P_{A'}$ to get all columns of $Y$. For matrix--matrix products, there are
 currently specialized implementations for the following \pk{MatType}:
 \begin{itemize}
-    \item \pk{MatMPIAIJ};
-    \item \pk{MatShell};
-    \item \pk{MatNest};
-    \item \pk{MatSeqAIJ}, \pk{MatSeqBAIJ}, and \pk{MatSeqSBAIJ};
-    \item \pk{MatSeqAIJcuSPARSE}.
+    \item \pk{MatSeqAIJ}, standard sequential sparse matrix, based on compressed sparse row format;
+    \item \pk{MatSeqBAIJ}, block sparse matrix, based on block compressed sparse row format;
+    \item \pk{MatSeqSBAIJ}, symmetric block sparse matrix stored in upper triangular form.
+    \item \pk{MatMPIAIJ}, standard parallel sparse matrix;
+    \item \pk{MatShell}, user-defined matrix;
+    \item \pk{MatNest}, block-defined matrix with nested submatrices;
+    \item \pk{MatSeqAIJCUSPARSE}, standard sequential sparse matrix, offloaded to a GPU using cuSPARSE~\cite{cusparse-web-page}.
 \end{itemize}
@@
     \item \pk{PCSHELL}: user-defined preconditioner, e.g., matrix-free;
-    \item \pk{PCSPAI}: sparse approximate inverse~\cite{grote1997parallel};
+    % \item \pk{PCSPAI}: sparse approximate inverse~\cite{grote1997parallel};
     \item \pk{PCHARA}~\cite{10.1145/3232850};
@@
     \item \pk{PCLU} or \pk{PCCHOLESKY}: exact $LU$ or Cholesky factorization,
-        i.e., $P_{A'} = {A'}^{-1}$;\label{it:lu}
+        e.g., $P_{A'} = {L}^{-1}{U}^{-1}$;\label{it:lu}
     \item \pk{PCILU} or \pk{PCICC}: incomplete $LU$ or Cholesky
@@
 %via \pk{PCFactorGetMatrix} and then call \pk{MatMatSolve}($F$, $\tilde{Y}$, $\tilde{X}$).
-via \pk{PCFactorGetMatrix}, and then call \pk{MatMatSolve}($F$, $Y$, $X$) to take advantage
+via \pk{PCFactorGetMatrix}, and then call \pk{MatMatSolve}($F$, $X$, $Y$) to take advantage
 of blocked forward eliminations and backward substitutions from the various factorization packages
@@
 candidate basis vectors. Such operations are performed using the CholQR algorithm~\cite{stathopoulos2002block}, or
-via the block variants of the (modified) Gram--Schmidt method.
+via the (modified) Gram--Schmidt method.
 For Hessenberg matrices generated by the Arnoldi
@@
 then reuse it appropriately for subsequent
-solves~\cite{saad2000deflated,stathopoulos2002block,soodhalter2014krylov}.
+solves~\cite{saad2000deflated,stathopoulos2009deflation,soodhalter2014krylov}.
 The routine \pk{KSPHPDDMGetDeflationSpace} can be used to retrieve such deflation subspaces, stored as distributed
@@
 \mu I$, then solving $(I \otimes S + A \otimes T) x = b$ may be cast into a
-system with multiple right-hand sides $(\lambda I + \mu A)X = B$ where $B$ is
+system with $k$ right-hand sides $(\lambda I + \mu A)X = B$ where $B$ is
 the row-major representation of $b$, i.e., $b_{k \cdot i + j} = B_{i, j},
@@
 All PETSc tests are performed using four MPI
-processes and they can be launched using the following command line: \\[-4pt]
+processes and they can be launched using the following command lines: \\[-4pt]
 \begin{Verbatim}[fontsize=\footnotesize,frame=single,framerule=0.1mm,commandchars=&\[\]]
@@
   -load_dir &fvtextcolor[myyellow][${DATAFILESPATH}]/matrices/hpddm/GCRODR
-&fvtextcolor[mygreen][$] &fvtextcolor[myyellow][mpirun] -n 4 ./ex75 -ksp_converged_reason -sub_pc_type icc -ksp_rtol 1e-10          &fvtextcolor[myred][\]
+&fvtextcolor[mygreen][$] &fvtextcolor[myyellow][mpirun] -n 4 ./ex75 -ksp_converged_reason -redundant_pc_type icc -ksp_rtol 1e-10    &fvtextcolor[myred][\]
   -ksp_type hpddm -ksp_hpddm_type gcrodr -ksp_gmres_restart 40 -ksp_hpddm_recycle 20 &fvtextcolor[myred][\]
-  -load_dir &fvtextcolor[myyellow][${DATAFILESPATH}]/matrices/hpddm/GCRODR
+  -load_dir &fvtextcolor[myyellow][${DATAFILESPATH}]/matrices/hpddm/GCRODR -pc_type redundant
 &fvtextcolor[mygreen][$] &fvtextcolor[myyellow][mpirun] -n 4 ./ex75 -ksp_converged_reason -pc_type jacobi -ksp_rtol 1e-10           &fvtextcolor[myred][\]
@@
 \end{Verbatim}
-The iteration numbers needed to reach convergence are gathered in~\cref{fig:comparison};
-the cases \ref{pgfplots:r0} and
+For the case of ICC(0), a \pk{PCREDUNDANT} is needed as well since
+PETSc does not provide a parallel ICC solver. The iteration numbers
+needed to reach convergence are gathered in~\cref{fig:comparison}.  The
+cases \ref{pgfplots:r0} and
 \ref{pgfplots:r1} (resp. \ref{pgfplots:r6} and \ref{pgfplots:r7}) reproduce Figure 4.2 in~\cite{parks2006recycling}.
@@
 sys none-matlab icc-matlab jacobi-matlab none-hpddm icc-hpddm jacobi-hpddm icc-gmres none-gmres icc-petsc none-petsc
-401 498         293        457           497        131       460          127       438        127       438
-402 228         125        187           231        53        184          127       449        127       449
-403 209         113        182           206        47        177          127       450        127       450
-404 199         112        175           198        47        174          127       450        127       450
-405 198         110        174           198        47        172          127       451        127       451
-406 198         111        174           199        47        174          127       452        127       452
-407 206         114        180           206        49        179          126       458        126       458
-408 207         115        181           208        49        182          128       458        135       458
-409 207         110        181           207        49        181          128       458        128       458
-410 206         110        179           206        49        179          128       457        128       457
+401 498         293        457           497        94        460          127       438        127       438
+402 228         125        187           231        37        184          127       449        127       449
+403 209         113        182           206        34        177          127       450        127       450
+404 199         112        175           198        33        174          127       450        127       450
+405 198         110        174           198        34        172          127       451        127       451
+406 198         111        174           199        34        174          127       452        127       452
+407 206         114        180           206        33        179          126       458        126       458
+408 207         115        181           208        34        182          128       458        135       458
+409 207         110        181           207        34        181          128       458        128       458
+410 206         110        179           206        34        179          128       457        128       457
 }\loadedtable
@@
 where $J$ is the Jacobian of the incompressible steady-state Navier--Stokes equation
-evaluated using a given base flow $q_b = \begin{bmatrix}u_b & p_b\end{bmatrix}^T$ and
+evaluated using a given base flow $q_b = \begin{bmatrix}u_b \\ p_b\end{bmatrix}$ and
 $M$ the discretization of the mass matrix on the space of velocities. Interested
@@
 flow past a cylinder at Reynolds 100. Results can be reproduced using the following four commands.
-We note here that the first two commands are merely used to generate the base flow $q_b$ with \pk{SNESSolve} using a
-continuation method on the Reynolds number. \\[-4pt]
+We note here that the first two commands are merely used to generate
+the base flow $q_b$ with a \pk{SNES}, a PETSc object used to solve
+nonlinear problems, using a continuation method on the Reynolds number. \\[-4pt]
 \begin{Verbatim}[fontsize=\footnotesize,frame=single,framerule=0.1mm,commandchars=&\[\]]
@@
     \subsubsection{Performance of primitives for block Krylov methods}\label{sec:primitives}
-The readiness of PETSc, as of version 3.13.2, is now discussed
+The readiness of PETSc, as of version 3.14.0, is now discussed
 when it comes to delivering efficient implementations of the core matrix--matrix multiplication
@@
 
-In particular, we consider three PETSc matrix types:
-\begin{itemize}
-    \item \pk{MatSeqAIJ}, standard sequential sparse matrices, based on compressed sparse row format;
-    \item \pk{MatSeqBAIJ}, block sparse matrices, based on block compressed sparse row format;
-    \item \pk{MatSeqSBAIJ}, symmetric block sparse matrices stored in upper triangular form.
-\end{itemize}
-Their distributed memory counterparts rely on the sequential implementations, with each process
+In particular, the first three matrix types from~\cref{sec:interface-ksp}, \pk{MatSeqAIJ}, \pk{MatSeqBAIJ}, and \pk{MatSeqSBAIJ} are first considered.
+Their distributed memory counterparts, e.g., \pk{MatMPIAIJ}, rely on the sequential implementations, with each process
 storing two such matrices, one with local rows and columns, and another with local rows and ``nonlocal'' columns.
@@
 values of $A$ are of no interest here, the sparsity pattern is rather common and frequently encountered when
-discretizing partial differential equations. We then generate a random symmetric dense matrix $T$ of dimension $b$,
+discretizing partial differential equations. We then generate a random symmetric dense $b$-by-$b$ matrix $T$,
 for $b \in \{ 1,3,6 \}$, and the matrix $\mathbb{A} = A \otimes T$ is assembled into all three formats described above, using
@@
 product operations.
-They can be reproduced by using the mini-app \url{https://github.com/prj-/jolivet2020petsc/blob/master/MatProduct.c} and any input \pk{MatAIJ} stored in binary format,
+They can be reproduced by using the mini-app \texttt{MatProduct.c} from \url{https://github.com/prj-/jolivet2020petsc} and any input \pk{MatAIJ} stored in binary format,
 here using the default name \pk{binaryoutput}, available at \url{http://jolivet.perso.enseeiht.fr/binaryoutput}.
@@
 between host and device nor time spent in \texttt{mkl\_sparse\_optimize} have
-been taken into account. For the sake of completeness, the performance of
-\pk{MatAIJMKL} with $N=1$ but different numbers of threads is also reported in
-\cref{fig:scaling-Ax}. Clearly, it is not advised to use the full socket to
-perform this type of workload for such small sparse matrices and skinny dense
-matrices. But this helps in drawing a fair comparison between a full CPU socket
-and a full GPU device. All in all, the CPU configuration which reaches the
-highest percentage of peak is MatSBAIJMKL~\ref{pgfplots:sbaijmkl3} with $b=3$
-and $N=64$, rightmost middle plot in~\cref{fig:perfAXpar}, with approximately
-63 GFLOP/s. It is still less than 1\% of peak for an Intel Xeon Gold 6248.
-For the GPU configuration, \pk{MatAIJcuSPARSE}~\ref{pgfplots:cusparse10} with $b=6$
-and $N=64$, rightmost bottom plot in~\cref{fig:perfAXpar}, performs at around
-306 GFLOP/s, about 4\% of peak for an NVIDIA Tesla V100 SXM2.
-\pgfplotstableread{
-P       aij1        aij3        aij6
-1       6.6857e-02  7.4649e-01  3.1128e+00       
-2       3.4206e-02  3.7820e-01  1.5723e+00
-4       1.6565e-02  1.8634e-01  7.7238e-01
-8       9.1370e-03  9.7549e-02  3.9884e-01
-16      7.2334e-03  7.9871e-02  3.2154e-01
-20      7.8399e-03  8.6487e-02  3.5536e-01
-}\loadedOMP
-\pgfplotstablegetelem{0}{aij1}\of{\loadedOMP}
-\let\refI\pgfplotsretval
-\pgfplotstablecreatecol[create col/expr={\refI/\thisrow{aij1}}]{speedupI}\loadedOMP
-\pgfplotstablegetelem{0}{aij3}\of{\loadedOMP}
-\let\refIII\pgfplotsretval
-\pgfplotstablecreatecol[create col/expr={\refIII/\thisrow{aij3}}]{speedupIII}\loadedOMP
-\pgfplotstablegetelem{0}{aij6}\of{\loadedOMP}
-\let\refVI\pgfplotsretval
-\pgfplotstablecreatecol[create col/expr={\refVI/\thisrow{aij6}}]{speedupVI}\loadedOMP
-\begin{figure}[tbp]
-  \centering
-    \begin{tikzpicture}
-      \begin{loglogaxis}[
-              ytick={1,2,4,8,20},
-        yticklabel={
-            \pgfmathparse{100*round(exp(\tick))}
-$\pgfmathprintnumber{\pgfmathresult}$\%
-},
-        xtick = {\empty},
-        extra x ticks = {1,2,4,8,16,20},height=6cm,
-          extra x tick labels = {1,2,4,8,16,\hspace*{0.3cm}20},
-          legend style={at={(0.02,0.98)}, legend cell align=left,anchor=north west},
-      ]
-          \addplot[black, dashed] table [x expr=\thisrow{P},y expr=\thisrow{P}]{\loadedOMP};\label{pgfplots:linear}
-          \addplot[blueaccent,mark=square*] table [x expr=\thisrow{P},y expr=\thisrow{speedupI}]{\loadedOMP};\label{pgfplots:speedupI}
-          \addplot[darkgray,mark=*] table [x expr=\thisrow{P},y expr=\thisrow{speedupIII}]{\loadedOMP};\label{pgfplots:speedupIII}
-          \addplot[last,mark=triangle*] table [x expr=\thisrow{P},y expr=\thisrow{speedupVI}]{\loadedOMP};\label{pgfplots:speedupVI}
-          \addlegendentry{Linear speedup}
-          \addlegendentry{$b = 1$}
-          \addlegendentry{$b = 3$}
-          \addlegendentry{$b = 6$}
-    \end{loglogaxis}
-    \end{tikzpicture}
-    \caption{Scalability of \pk{MatMult} with \pk{MatAIJMKL} and different number of processes.\label{fig:scaling-Ax}}
-\end{figure}
+been taken into account.
 
@@
            \addlegendentry{MatSBAIJMKL}
-           \addlegendentry{\pk{MatAIJcuSPARSE}}
+           \addlegendentry{\pk{MatAIJCUSPARSE}}
 \node[rotate=0,fill=white,inner sep=0pt,anchor=west] at (rel axis cs:1.02, 0.5) {$b=1$};
@@
 \end{figure}
+
+For the sake of completeness, the performance of
+\pk{MatAIJMKL} with $N=1$ but different numbers of threads is also reported in
+\cref{fig:scaling-Ax}. Clearly, it is not advised to use the full socket to
+perform this type of workload for such small sparse matrices and skinny dense
+matrices. But this helps in drawing a fair comparison between a full CPU socket
+and a full GPU device. All in all, the CPU configuration which reaches the
+highest percentage of peak is MatSBAIJMKL~\ref{pgfplots:sbaijmkl3} with $b=3$
+and $N=64$, rightmost middle plot in~\cref{fig:perfAXpar}, with approximately
+63 GFLOP/s. It is still less than 1\% of peak for an Intel Xeon Gold 6248.
+For the GPU configuration, \pk{MatAIJCUSPARSE}~\ref{pgfplots:cusparse10} with $b=6$
+and $N=64$, rightmost bottom plot in~\cref{fig:perfAXpar}, performs at around
+306 GFLOP/s, about 4\% of peak for an NVIDIA Tesla V100 SXM2.
+\pgfplotstableread{
+P       aij1        aij3        aij6
+1       6.6857e-02  7.4649e-01  3.1128e+00       
+2       3.4206e-02  3.7820e-01  1.5723e+00
+4       1.6565e-02  1.8634e-01  7.7238e-01
+8       9.1370e-03  9.7549e-02  3.9884e-01
+16      7.2334e-03  7.9871e-02  3.2154e-01
+20      7.8399e-03  8.6487e-02  3.5536e-01
+}\loadedOMP
+\pgfplotstablegetelem{0}{aij1}\of{\loadedOMP}
+\let\refI\pgfplotsretval
+\pgfplotstablecreatecol[create col/expr={\refI/\thisrow{aij1}}]{speedupI}\loadedOMP
+\pgfplotstablegetelem{0}{aij3}\of{\loadedOMP}
+\let\refIII\pgfplotsretval
+\pgfplotstablecreatecol[create col/expr={\refIII/\thisrow{aij3}}]{speedupIII}\loadedOMP
+\pgfplotstablegetelem{0}{aij6}\of{\loadedOMP}
+\let\refVI\pgfplotsretval
+\pgfplotstablecreatecol[create col/expr={\refVI/\thisrow{aij6}}]{speedupVI}\loadedOMP
+\begin{figure}[tbp]
+  \centering
+    \begin{tikzpicture}
+      \begin{loglogaxis}[
+              ytick={1,2,4,8,20},
+        yticklabel={
+            \pgfmathparse{100*round(exp(\tick))}
+$\pgfmathprintnumber{\pgfmathresult}$\%
+},
+        xtick = {\empty},
+        extra x ticks = {1,2,4,8,16,20},height=6cm,
+          extra x tick labels = {1,2,4,8,16,\hspace*{0.3cm}20},
+          legend style={at={(0.02,0.98)}, legend cell align=left,anchor=north west},
+      ]
+          \addplot[black, dashed] table [x expr=\thisrow{P},y expr=\thisrow{P}]{\loadedOMP};\label{pgfplots:linear}
+          \addplot[blueaccent,mark=square*] table [x expr=\thisrow{P},y expr=\thisrow{speedupI}]{\loadedOMP};\label{pgfplots:speedupI}
+          \addplot[darkgray,mark=*] table [x expr=\thisrow{P},y expr=\thisrow{speedupIII}]{\loadedOMP};\label{pgfplots:speedupIII}
+          \addplot[last,mark=triangle*] table [x expr=\thisrow{P},y expr=\thisrow{speedupVI}]{\loadedOMP};\label{pgfplots:speedupVI}
+          \addlegendentry{Linear speedup}
+          \addlegendentry{$b = 1$}
+          \addlegendentry{$b = 3$}
+          \addlegendentry{$b = 6$}
+    \end{loglogaxis}
+    \end{tikzpicture}
+    \caption{Scalability of \pk{MatMult} with \pk{MatAIJMKL} and different number of processes.\label{fig:scaling-Ax}}
+\end{figure}
+
 \subsubsection{Blocking inside SLEPc\label{sec:lobpcg}}
@@
 Furthermore, PETSc provides domain decomposition methods of the non-overlapping
-type; \pk{PCNN} implements a basic balancing Neumann--Neumann
+type. \pk{PCNN} implements a basic balancing Neumann--Neumann
 method~\cite{mandel1993balancing}, while more advanced preconditioning
@@
 course, there are many other available domain decomposition preconditioners,
-either accessible through high-level libraries, e.g., FROSch~\cite{Heinlein:2016:PIT} in
+either accessible through high-level libraries other than PETSc, e.g., FROSch~\cite{Heinlein:2016:PIT} in
 Trilinos, or as stand-alone packages, e.g., BDDCML~\cite{vsistek2013parallel} or
@@
 can be parameterized through a local \pk{KSP}. When $A$ is the
-discretization of a linear partial differential operator operator $\mathcal{L}$ on a domain $\Omega$, it is
+discretization of a linear partial differential operator $\mathcal{L}$ on a domain $\Omega$, it is
 extremely common to exhibit parallelism by distributing $\Omega$ on $N$
@@
         \begin{equation}\label{eq:gevp}
-            \mathring{A}_i \Lambda_i = \lambda_i \tilde{R}_i R_i^T (R_i A R_i^T) \tilde{R}_i R_i^T \Lambda_i ;
+            \mathring{A}_i y_i = \lambda_i \tilde{R}_i R_i^T (R_i A R_i^T) \tilde{R}_i R_i^T y_i ;
         \end{equation}
-    \item retrieve the $\nu_i$ smallest eigenpairs $\{\Lambda_{i_j}, \lambda_{i_j}\}_{j=1}^{\nu_i}$ and assemble a local deflation dense matrix $W_i = \begin{bmatrix} \tilde{R}_i R_i^T \Lambda_{i_1} & \cdots & \tilde{R}_i R_i^T \Lambda_{i_{\nu_j}} \end{bmatrix}$;
+    \item retrieve the $\nu_i$ smallest eigenpairs $\{y_{i_j}, \lambda_{i_j}\}_{j=1}^{\nu_i}$ and assemble a local deflation dense matrix $W_i = \begin{bmatrix} \tilde{R}_i R_i^T y_{i_1} & \cdots & \tilde{R}_i R_i^T y_{i_{\nu_i}} \end{bmatrix}$;
     \item define a global deflation matrix $P = \begin{bmatrix} R_1^T W_1 & \cdots & R_N^T W_N\end{bmatrix}$ and a new two-level preconditioner using the Galerkin product of $A$ and $P$:
@@
     }
-    \caption{Automatic generation of overlapping subdomains when using a~\pk{DMP4EST}. For clarity, only four subdomains, one of which is disconnected, are colored.\label{fig:p4est}}
+    \caption{Automatic generation of overlapping subdomains when using a~\pk{DMPLEX}. For clarity, only four subdomains, one of which is disconnected, are colored.\label{fig:p4est}}
 \end{figure}
@@
 standard eigensolvers such as ARPACK~\cite{lehoucq1998arpack} or
-\pk{EPSKRYLOVSCHUR} from SLEPc, cannot compute all eigenpairs in a given
-interval of $\mathbb{R}^+$. Instead, an upper bound of the number of eigenpairs
-has to be provided.
+\pk{EPSKRYLOVSCHUR} from SLEPc, cannot cheaply evaluate the exact number of eigenpairs in a given
+interval of $\mathbb{R}^+$. Instead, it is often more efficient to provide an upper bound of the number of eigenpairs.
 
@@
 \end{align*}
-with $Q = P \left(P^T A P\right)^{-1} P^T$. This is parameterized using the
+with $Q = P \left(P^T A P\right)^{-1} P^T$ and $M^{-1}$ defined in~\cref{eq:one}. This is parameterized using the
 options \pk{-pc\_hpddm\_coarse\_correction}
@@
 \begin{equation*}
-    \mathring{A}_i \Lambda_i = \lambda_i B_i \Lambda_i.
+    \mathring{A}_i y_i = \lambda_i B_i y_i.
 \end{equation*}
@@
 Because of the intrinsic nature of balancing domain
-decomposition methods, \pk{PCNN} and \pk{PCBDDC} must be supplied local unassembled
-matrices. These matrices are equivalent to the $\{\mathring{A}_i\}_{i=1}^N$
+decomposition methods, \pk{PCNN} and \pk{PCBDDC} must be supplied a \pk{MatIS}
+which stores local unassembled matrices and local to global mappings.
+These local matrices are equivalent to the $\{\mathring{A}_i\}_{i=1}^N$
 defined \cref{sec:interface-pc}, assuming the domain decomposition is without
@@
 \begin{itemize}
-\item supply whenever the Jacobian is being updated in the \pk{SNESSetJacobian}
-function new unassembled Neumann matrices at the current linearization point with the previously
-explained \pk{PCHPDDMSetAuxiliaryMat} function, this is done automatically with a
+\item whenever the Jacobian is being updated in the \pk{SNESSetJacobian}
+function, new unassembled Neumann matrices at the current linearization point are updated as well and supplied to the previously
+explained \pk{PCHPDDMSetAuxiliaryMat} function, this is done automatically if the nonliear solver is attached to a
 \pk{DMPLEX} or a \pk{DMP4EST};
@@
 \bibitem{balay1997cient}
-{\sc S.~Balay, W.~Gropp, L.~Curfman~McInnes, and B.~F. Smith}, {\em Efficient
-  management of parallelism in object-oriented numerical software libraries},
-  Modern Software Tools in Scientific Computing,  (1997), pp.~163--202.
+{\sc S.~Balay, W.~D. Gropp, L.~Curfman~McInnes, and B.~F. Smith}, {\em
+  Efficient management of parallelism in object-oriented numerical software
+  libraries}, Modern Software Tools in Scientific Computing,  (1997),
+  pp.~163--202.
 
@@
 
-\bibitem{grote1997parallel}
-{\sc M.~J. Grote and T.~Huckle}, {\em Parallel preconditioning with sparse
-  approximate inverses}, SIAM Journal on Scientific Computing, 18 (1997),
-  pp.~838--853.
-
 \bibitem{gutknecht2006block}
@@
 \bibitem{saad2003iterative}
-{\sc Y.~Saad}, {\em {Iterative Methods for Sparse Linear Systems}}, SIAM, 2003.
+{\sc Y.~Saad}, {\em {Iterative Methods for Sparse Linear Systems, Second
+  Edition}}, SIAM, 2003.
 
@@
   Engineering, 95 (2013), pp.~953--990.
+
+\bibitem{stathopoulos2009deflation}
+{\sc A.~Stathopoulos, A.~Abdel-Rehim, and K.~Orginos}, {\em {Deflation for
+  inversion with multiple right-hand sides in QCD}}, in Journal of Physics:
+  Conference Series, vol.~180, IOP Publishing, 2009.
 

### 2020-07-16 article.tex
Small note about DtN

@@
 In addition to the local unassembled Neumann matrices~$\{\mathring{A}_i\}_{i=1}^N$,
-users can prescribe additional local operators defined on the overlapping subdomain and
-with optimized Robin
-boundary conditions $\{B_i\}_{i=1}^N$ via
-the routine \pk{PCHPDDMSetRHSMat}(pc, $B_i$).
-In such a case, the following generalized eigenvalue problems, called GenEO-2 in the
+users can prescribe additional local operators $\{B_i\}_{i=1}^N$, defined on
+the overlapping decomposition, via the routine \pk{PCHPDDMSetRHSMat}(pc, $B_i$).
+In the case of local matrices with optimized Robin transmission conditions, the
+following generalized eigenvalue problems, called GenEO-2 in the
 literature~\cite{haferssas2017soras}, are then solved in each subdomain:
@@
 \end{equation*}
-This may be useful when dealing with nearly indefinite systems, e.g., nearly incompressible elasticity or the system of Stokes.
-Indeed, for such equations, one needs to compute a lot of eigenvectors from the classical GenEO eigenproblem~\cref{eq:gevp} to generate a robust preconditioner. GenEO-2 alleviates this phenomenon.
+This may be useful when dealing with nearly indefinite systems, e.g., nearly
+incompressible elasticity or the system of Stokes. Indeed, for such equations,
+one needs to compute a lot of eigenvectors from the classical GenEO
+eigenproblem~\cref{eq:gevp} to generate a robust preconditioner. GenEO-2
+alleviates this phenomenon. In the case where the local operators
+$\{B_i\}_{i=1}^N$ are the discrete mass matrices on the subdomains interfaces,
+the so-called Dirichlet-to-Neumann coarse operator is constructed. This has
+proven to be efficient for solving the heterogeneous Helmholtz
+equation~\cite{conen2014coarse}.
 
@@
 
+\bibitem{conen2014coarse}
+{\sc L.~Conen, V.~Dolean, R.~Krause, and F.~Nataf}, {\em {A coarse space for
+  heterogeneous Helmholtz problems based on the Dirichlet-to-Neumann
+  operator}}, Journal of Computational and Applied Mathematics, 271 (2014),
+  pp.~83--99.
+
 \bibitem{da2014isogeometric}

### 2020-07-17 article.tex
Some more typo
Some more remarks

@@
     \item pseudo-block GCRODR or flexible GCRODR~\cite{carvalho2011flexible};
-    \item block GMRES or flexible GMRES~\cite{calandra2012flexible};
+    \item block GMRES or flexible GMRES~\cite{calandra2012flexible}, with deflation at each restart;
     \item block CG~\cite{o1980block};
-    \item breakdown-free block CG~\cite{ji2017breakdown};
-    \item block GCRODR or flexible GCRODR.
+    \item breakdown-free block CG~\cite{ji2017breakdown}, with deflation at each iteration;
+    \item block GCRODR or flexible GCRODR, with deflation at each restart.
 \end{itemize}
@@
     \item \pk{PCLU} or \pk{PCCHOLESKY}: exact $LU$ or Cholesky factorization,
-        e.g., $P_{A'} = {L}^{-1}{U}^{-1}$;\label{it:lu}
+        e.g., $P_{A'} = {U}^{-1}{L}^{-1}$;\label{it:lu}
     \item \pk{PCILU} or \pk{PCICC}: incomplete $LU$ or Cholesky
@@
 and \texttt{MPI\_Scatterv}, that could exhibit worse performance. In addition,
-standard eigensolvers such as ARPACK~\cite{lehoucq1998arpack} or
-\pk{EPSKRYLOVSCHUR} from SLEPc, cannot cheaply evaluate the exact number of eigenpairs in a given
-interval of $\mathbb{R}^+$. Instead, it is often more efficient to provide an upper bound of the number of eigenpairs.
+standard eigensolvers such as \pk{EPSKRYLOVSCHUR} from SLEPc, cannot cheaply
+evaluate the exact number of eigenpairs in a given interval of $\mathbb{R}^+$.
+Instead, it is often more efficient to provide an upper bound of the number of
+eigenpairs.
 
@@
 function, new unassembled Neumann matrices at the current linearization point are updated as well and supplied to the previously
-explained \pk{PCHPDDMSetAuxiliaryMat} function, this is done automatically if the nonliear solver is attached to a
+explained \pk{PCHPDDMSetAuxiliaryMat} function, this is done automatically if the nonlinear solver is attached to a
 \pk{DMPLEX} or a \pk{DMP4EST};
@@
 
-\bibitem{lehoucq1998arpack}
-{\sc R.~Lehoucq, D.~Sorensen, and C.~Yang}, {\em {ARPACK} users’ guide\col\
-  solution of large-scale eigenvalue problems with implicitly restarted Arnoldi
-  methods}, vol.~6, Society for Industrial and Applied Mathematics, 1998,
-  \url{http://www.caam.rice.edu/software/ARPACK}.
-
 \bibitem{li2005overview}

### 2020-07-17 article.tex
PCBDDC unknowns

@@
 \end{align}
-It is discretized using first order N\'ed\'elec finite elements and the \pk{PCBDDC} solver
-is automatically customized to obtain a stable method with these elements~\cite{zampini2017balancing}.
+It is discretized using first order N\'ed\'elec finite elements and \pk{PCBDDC}
+is automatically customized to obtain a stable method with these
+elements~\cite{zampini2017balancing}. The number of unknowns is $\pgfmathprintnumber{5.6} \times 10^6$.
 The following command line options are used. \\[-4pt]
@@
 Deluxe scaling is used to build an adaptive second level with \pk{PCBDDC}. On this
-second level, new subdomains are built by aggregating the coarse element matrices of
+second level with 16${,}$284 unknowns, new subdomains are built by aggregating the coarse element matrices of
 8 fine-level subdomains. The BDDC coarse problem composed of $\frac{512}{8} =

### 2020-07-18 article.tex
A bit longer description of SLEPc

@@
 With these added functions, it is now possible to do code refactorization for
-applications that deal with multiple right-hand sides.  As a striking example,
-LOBPCG~\cite{knyazev2001toward} or CISS~\cite{sakurai2003projection}
-implementations in SLEPc~\cite{hernandez2005ssf}, the Scalable Library for
-Eigenvalue Problem computations built on top of PETSc, were previously not
-using blocking when applying the preconditioner or solving linear systems. This has been rewritten for version 3.14.0 and the benefit of
-this new implementation is displayed in~\cref{sec:lobpcg}.
+applications that deal with multiple right-hand sides. This is the case of some
+eigensolvers in SLEPc~\cite{hernandez2005ssf}, the Scalable Library for
+Eigenvalue Problem computations. SLEPc is an add-on library that extends PETSc
+with several classes for linear and nonlinear eigenvalue problems, as well as
+tools to, e.g., facilitate the use of shift-and-invert spectral transformations.
+The implementations of LOBPCG~\cite{knyazev2001toward} and CISS~\cite{sakurai2003projection}
+in SLEPc were previously not using blocking when applying the preconditioner or solving linear systems.
+In version 3.14.0 we have adapted these solvers to employ \pk{KSPMatSolve} and \pk{PCMatApply}.
+In the case of LOBPCG, which is a purely blocked method, all steps were already
+implemented as block operations, except the applications of the preconditioner,
+which ruined the benefits of blocking. With the new implementation, the gains
+are striking, as reported in~\cref{sec:lobpcg}.
 

### 2020-07-18 article.tex
Add sentence and project reference

@@
 
+Concerning SLEPc, as a future work we plan to implement new solvers (or extend existing
+ones) to further exploit the block solve primitives, for instance block Krylov-Schur.
+
 \section*{Acknowledgments}
@@
 \end{itemize}
+Jose E.~Roman was supported by the Spanish Agencia Estatal de Investigaci\'{o}n (AEI) under project SLEPc-DA (PID2019-107379RB-I00).
 % We would like to acknowledge the assistance of volunteers in putting

### 2020-07-18 article.tex
Fixing some red part

@@
 
-The paper is divided in two main parts {\color{red} COMPLETE WITH MORE DETAILS WHEN PAPER IS COMPLETED}:
-\cref{sec:KSPHPDDM} introduces \pk{KSPHPDDM}, provides applications within an eigensolver
-framework and discuss the readiness of the PETSc library in delivering highly efficient block Krylov methods.
-\cref{sec:PCHPDDM} introduces \pk{PCHPDDM}. Eventually, concluding remarks may be found
-\cref{sec:conclusion}. All PETSc keywords and options are typeset in a \texttt{typewriter} font, these
-are public and documented at \url{https://www.mcs.anl.gov/petsc/petsc-master/docs}.
+The paper is divided in two main parts. \Cref{sec:KSPHPDDM} introduces
+\pk{KSPHPDDM}. The readiness of PETSc for interacting with block Krylov methods
+is discussed.  Multiple applications are provided, two of which are within
+eigensolver frameworks, showing the interoperability of the proposed
+methodology.  \Cref{sec:PCHPDDM} introduces \pk{PCHPDDM}. Again, diverse
+problems are dealt with, for example in the context of a nonlinear solver.
+Eventually, concluding remarks may be found \cref{sec:conclusion}. All PETSc
+keywords and options are typeset in \texttt{typewriter} font, these are
+public and documented at \url{https://www.mcs.anl.gov/petsc/petsc-master/docs}.
+Codes not already available in other public repositories are available at 
+\url{https://github.com/prj-/jolivet2020petsc}.
 
@@
 number of global synchronizations needed for scalar products.
-HPDDM Krylov methods may be selected {\color{red} TODO API FOR KSPHPDDMSetVariant?} using the
+HPDDM Krylov methods may be selected using the % {\color{red} TODO API FOR KSPHPDDMSetVariant?}
 options \pk{-ksp\_hpddm\_type}
@@
     % \item \pk{PCSPAI}: sparse approximate inverse~\cite{grote1997parallel};
-    \item \pk{PCHARA}~\cite{10.1145/3232850};
+    \item \pk{PCHARA}~\cite{10.1145/3232850} {\color{red}more info?};
     \item \pk{PCHPDDM}: see~\cref{sec:PCHPDDM};\label{it:hpddm}
@@
 used instead~\cite{gutknecht2008updating}.
-These orthonormalization variants can be selected at runtime with the option \pk{-ksp\_hpddm\_qr (cholqr$|$mgs$|$cgs)}. {\color{red} API?}
+These orthonormalization variants can be selected at runtime with the option \pk{-ksp\_hpddm\_qr (cholqr$|$mgs$|$cgs)}. % {\color{red} API?}
 
@@
 in SLEPc were previously not using blocking when applying the preconditioner or solving linear systems.
-In version 3.14.0 we have adapted these solvers to employ \pk{KSPMatSolve} and \pk{PCMatApply}.
+In version 3.14.0, we have adapted these solvers to employ \pk{KSPMatSolve} and \pk{PCMatApply}.
 In the case of LOBPCG, which is a purely blocked method, all steps were already
@@
 A similar speedup can be observed in terms of runtime, with the time spent for the solution of the eigenvalue problem being reduced from \SI[round-mode = places,round-precision=1,group-minimum-digits=3,group-separator = {,}]{1509}{\sec} to \SI[round-mode = places,round-precision=1,group-minimum-digits=3,group-separator = {,}]{1052}{\sec}.
-{\color{red} FIXME LATER: These timings, and all which follow except in~\cref{sec:primitives}, have been obtained on {Ir\`ene}, a system composed of \pgfmathprintnumber[assume math mode=true]{1656} nodes with two 24-core Intel Xeon Platinum 8168 clocked at \SI{2.7}{\giga\hertz}}.
+These timings, and all which follow except in~\cref{sec:primitives}, have been obtained on {Ir\`ene}, a system composed of \pgfmathprintnumber[assume math mode=true]{1656} nodes with two 24-core Intel Xeon Platinum 8168 clocked at \SI{2.7}{\giga\hertz}.
 \pgfplotstableread{
@@
 convergence history, which is displayed in~\cref{fig:bratu-conv}. The complete set of
-options is given in~\cref{fig:bratu-options}. Results can be reproduced using the FreeFEM script at \url{https://github.com/prj-/jolivet2020petsc}
+options is given in~\cref{fig:bratu-options}. Results can be reproduced using the FreeFEM script at \url{https://github.com/prj-/jolivet2020petsc}.
 \pgfplotstableread{

### 2020-07-18 article.tex
Almost there

@@
 \hypersetup{
-  pdftitle={KSPHPDDM and PCHPDDM: Extending PETSc with Robust Overlapping Schwarz Preconditioners and Advanced Krylov Methods},
+  pdftitle={KSPHPDDM and PCHPDDM: Extending PETSc with Robust Overlapping Schwarz Preconditioners and Advanced Krylov Solvers},
   pdfauthor={P. Jolivet, J. E. Roman, and S. Zampini}
@@
 methods~\cite{jolivet2016block} which can efficiently deal with linear
-systems with multiple right-hand sides, also known and as block iterative methods,
+systems with multiple right-hand sides (i.e. block iterative methods),
 and when there is a recurrence of varying coefficient matrices and right-hand
-sides, recycling iterative methods.
+sides (recycling iterative methods).
 
 % Htool ?
-Here, we describe the integration of the entire HPDDM suite, its advanced iterative
-methods and its domain decomposition preconditioners, into PETSc.
+Here, we describe the integration of the HPDDM solvers suite consisting of advanced iterative
+methods and robust domain decomposition preconditioners into PETSc.
 The design principles of the interface are showcased considering % alongside
-minimalist and easy-to-run examples. Necessary command line
-options are provided to:
-\begin{itemize}
-    \item fulfill reproducibility requirements as advocated by the
+minimalist and easy-to-run examples, with a focus on fulfilling reproducibility requirements as advocated by the
 Association for Computing Machinery
-\url{https://www.acm.org/publications/policies/artifact-review-badging};
-    \item ease the integration of these algorithms into more advanced frameworks;
-    \item analyze performance results~\cite{hoefler2015scientific}.
-\end{itemize}
+\url{https://www.acm.org/publications/policies/artifact-review-badging}, as well as to 
+ease the integration of these algorithms into more advanced frameworks
+and facilitate the analysis and comparison of performance results~\cite{hoefler2015scientific}.
 
 The paper is divided in two main parts. \Cref{sec:KSPHPDDM} introduces
-\pk{KSPHPDDM}. The readiness of PETSc for interacting with block Krylov methods
-is discussed.  Multiple applications are provided, two of which are within
-eigensolver frameworks, showing the interoperability of the proposed
-methodology.  \Cref{sec:PCHPDDM} introduces \pk{PCHPDDM}. Again, diverse
-problems are dealt with, for example in the context of a nonlinear solver.
-Eventually, concluding remarks may be found \cref{sec:conclusion}. All PETSc
+\pk{KSPHPDDM} and discusses the readiness of PETSc with respect to block Krylov methods.
+Different applications from eigen-analysis are also provided.
+\Cref{sec:PCHPDDM} introduces \pk{PCHPDDM} and presents diverse
+applications for solving linear and nonlinear partial differential equations.
+Eventually, concluding remarks may be found in \cref{sec:conclusion}.
+
+All PETSc
 keywords and options are typeset in \texttt{typewriter} font, these are
@@
 
-\section{KSPHPDDM}\label{sec:KSPHPDDM}
+\section{Advanced Krylov methods in PETSc: KSPHPDDM}\label{sec:KSPHPDDM}
   \subsection{Related work}
@@
 \pk{KSPLGMRES} and \pk{KSPDGMRES}, are available. However, neither handle variable
-preconditioning, and \pk{KSPDGMRES} does not support complex arithmetic. {\color{red} Also note
-that when solving a sequence of linear systems $A_i x_i = b_i, i = 1, 2,
-\ldots$, it is not trivial to extend the theory of these methods to perform
-recycling throughout both the restarts and the different systems. The
+preconditioning, and \pk{KSPDGMRES} does not support complex arithmetic;
+moreover, extending the recycling capabilites of such methods to sequence of linear systems
+with smoothly varying coefficient matrices and right-hand sides is not trivial.
+On the other hand, HPDDM offers support for the
 generalized conjugate residual method with inner orthogonalization and deflated
-restarting~\cite{parks2006recycling}~(GCRODR) does not suffer from this
-limitation.}
+restarting~\cite{parks2006recycling}~(GCRODR), which does not suffer from the aforementioned
+limitations.
 
@@
 Such methods, while having higher arithmetic
-intensities, generate larger subspaces and typically converge in fewer iterates.
-
-These methods are not currently offered in PETSc and they are
+intensities, generate larger subspaces and typically converge in fewer iterates. These methods are not currently offered in PETSc and they are
 less frequently implemented in general purpose libraries because they require
 somehow more involved kernels such as matrix--matrix multiplication, instead of
-matrix--vector, and may involve different inner-product realizations~\cite{frommer2017block}.
+matrix--vector products, and may involve different inner-product realizations~\cite{frommer2017block}.
 Still, systems with multiple right-hand sides are ubiquitous,
@@
 flag \pk{-{}-download-hpddm}. HPDDM linear solvers can be selected using the command
-line option \pk{-ksp\_type hpddm}, or the routine \pk{KSPSetType}(ksp,
+line option \pk{-ksp\_type hpddm}, or programmatically via \pk{KSPSetType}(\pk{ksp},
 \pk{KSPHPDDM}), and the following Krylov methods can be accessed:
@@
 number of global synchronizations needed for scalar products.
+
 HPDDM Krylov methods may be selected using the % {\color{red} TODO API FOR KSPHPDDMSetVariant?}
 options \pk{-ksp\_hpddm\_type}
-\pk{(gmres$|$cg$|$gcrodr$|$bgmres$|$bcg$|$bfbcg$|$bgcrodr$|$preonly)}, or the
-routine \pk{KSPHPDDMSetType}(ksp, \pk{KSPHPDDMType}).
+\pk{(gmres$|$cg$|$gcrodr$|$bgmres$|$bcg$|$bfbcg$|$bgcrodr$|$preonly)}, or programmatically 
+by using \pk{KSPHPDDMSetType}(\pk{ksp}, \pk{KSPHPDDMType}).
 Preconditioning variants can be specified via
@@
 
-According to the standard PETSc notation, ksp($A$, $P_{A'}$)
-defines a \pk{KSP} used to solve a system $Ax=b$ using a preconditioner $P_{A'}$ built
-using an operator $A'$ which is usually, but not necessarily, equal to $A$. While solving
-the linear system via \pk{KSPSolve}(ksp, $b$, $x$) or \pk{KSPMatSolve}(ksp, $B$, $X$),
-HPDDM will repeatedly call the following PETSc routines:
+Recycling Krylov methods try to
+extract convergence information by solving a small standard or a generalized
+dense eigenproblem at the end of each cycle or when convergence is reached, and
+then reuse it appropriately for subsequent
+solves~\cite{saad2000deflated,stathopoulos2009deflation,soodhalter2014krylov}.
+Such deflation subspaces, stored as dense tall-and-skinny matrices, can be accessed with \pk{KSPHPDDMGetDeflationSpace}.
+User-defined deflation subspaces can also be specified via \pk{KSPHPDDMSetDeflationSpace}.
+
+To fully support HPDDM solvers, from version 3.14.0 of PETSc we added interface routines
+for solving systems with multiple right-hand sides, \pk{KSPMatSolve}(\pk{ksp},\pk{X},\pk{Y}),
+and to apply preconditioning methods, i.e. \pk{PCMatApply}(\pk{ksp},\pk{X},\pk{Y}).
+Both input $X$ and output $Y$ matrices are currently limited to be dense tall-and-skinny matrices.
+With \pk{KSPHPDDM}, \mbox{(pseudo-)block} methods Krylov methods will be used; instead, when no specialized implementation
+is available, PETSc will perform the solution phase
+in a column by column fashion.
+Furthermore, a function \pk{KSPSetMatSolveBlockSize} is also provided to
+decompose a single large block of column vectors into multiple sub-blocks.
+\pk{KSPMatSolve} is then called repeatedly until all sub-blocks are traversed.
+This was inspired by MUMPS~\cite{amestoy2001fully} option
+ICNTL(27)\footnote{\url{http://mumps.enseeiht.fr/doc/userguide_5.3.3.pdf},
+section 6.1}.
+
+While solving the (block) linear system, HPDDM will repeatedly call the following PETSc routines:
 \begin{itemize}
@@
         AX$;
-    \item \pk{PCApply}($P_{A'}$, $x$, $y$) for $y = P_{A'}x$;
-    \item \pk{PCMatApply}($P_{A'}$, $X$, $Y$) for $Y = P_{A'}X$;
+    \item \pk{PCApply}(M, $x$, $y$) to apply a preconditioner to $x$, i.e. for $y = M^{-1}x$;
+    \item \pk{PCMatApply}(M, $X$, $Y$) for $Y = M^{-1}X$;
 \end{itemize}
-As of version 3.14.0, PETSc provides unified routines for applying $A$ or $P_{A'}$
-to a tall-and-skinny dense matrix $X$. However, when no specialized implementation
-is available, PETSc will loop over the columns of $X$ and successively apply $A$
-or $P_{A'}$ to get all columns of $Y$. For matrix--matrix products, there are
-currently specialized implementations for the following \pk{MatType}:
-\begin{itemize}
-    \item \pk{MatSeqAIJ}, standard sequential sparse matrix, based on compressed sparse row format;
-    \item \pk{MatSeqBAIJ}, block sparse matrix, based on block compressed sparse row format;
-    \item \pk{MatSeqSBAIJ}, symmetric block sparse matrix stored in upper triangular form.
-    \item \pk{MatMPIAIJ}, standard parallel sparse matrix;
-    \item \pk{MatShell}, user-defined matrix;
-    \item \pk{MatNest}, block-defined matrix with nested submatrices;
-    \item \pk{MatSeqAIJCUSPARSE}, standard sequential sparse matrix, offloaded to a GPU using cuSPARSE~\cite{cusparse-web-page}.
-\end{itemize}
-The following preconditioners have specialized implementations for dealing
-efficiently with multiple vectors:
-\begin{enumerate}
-    \item \pk{PCNONE}: $P_{A'} = I$, identity matrix;
-    \item \pk{PCKSP}: embedded Krylov method inside $P_{A'}$;
-    \item \pk{PCMAT}: multiplication by $A'$, i.e., $P_{A'} = {A'}$;
-    \item \pk{PCSHELL}: user-defined preconditioner, e.g., matrix-free;
-    % \item \pk{PCSPAI}: sparse approximate inverse~\cite{grote1997parallel};
-    \item \pk{PCHARA}~\cite{10.1145/3232850} {\color{red}more info?};
-    \item \pk{PCHPDDM}: see~\cref{sec:PCHPDDM};\label{it:hpddm}
-    \item \pk{PCASM} and \pk{PCGASM}: overlapping Schwarz methods;\label{it:asm}
-    \item \pk{PCBJACOBI}: block Jacobi;\label{it:bjacobi}
-    \item \pk{PCLU} or \pk{PCCHOLESKY}: exact $LU$ or Cholesky factorization,
-        e.g., $P_{A'} = {U}^{-1}{L}^{-1}$;\label{it:lu}
-    \item \pk{PCILU} or \pk{PCICC}: incomplete $LU$ or Cholesky
-        factorization.\label{it:ilu}
-\end{enumerate}
-One appealing feature of domain decomposition-like preconditioners 
-\cref{it:hpddm,it:asm,it:bjacobi} is that they most often rely on exact or inexact
-factorizations~\cref{it:lu,it:ilu} as subdomain solvers.
-In these cases, it is possible to access the so-called local or distributed factored matrix $F$
-%via \pk{PCFactorGetMatrix} and then call \pk{MatMatSolve}($F$, $\tilde{Y}$, $\tilde{X}$).
-via \pk{PCFactorGetMatrix}, and then call \pk{MatMatSolve}($F$, $X$, $Y$) to take advantage
-of blocked forward eliminations and backward substitutions from the various factorization packages
-interfaced with PETSc. In the case of a sparse coefficient matrix $A'$,
-this strategy is possible with: MUMPS~\cite{amestoy2001fully},
-SuiteSparse~\cite{davis2004algorithm}, MKL
-PARDISO or CPARDISO~\cite{mkl-web-page}, and SuperLU~\cite{li2005overview} or
-SuperLU\_DIST~\cite{li2003superlu_dist}. For the case of a dense coefficient
-matrix $A'$: ScaLAPACK and Elemental~\cite{poulson2013elemental}. 
-Future work will consider extending the various implementations of PETSc
-preconditioning infrastructure to increase the arithmetic intensity of the
-block preconditioner application phase.
-
-The rest of the operations are then performed directly inside HPDDM.
+The rest of the operations are performed directly inside HPDDM.
 Specifically, within Krylov methods using the Arnoldi process, or when recycling is requested,
@@
 used instead~\cite{gutknecht2008updating}.
-These orthonormalization variants can be selected at runtime with the option \pk{-ksp\_hpddm\_qr (cholqr$|$mgs$|$cgs)}. % {\color{red} API?}
-
-    \subsubsection{User-defined deflation space} Recycling Krylov methods try to
-extract convergence information by solving a small standard or a generalized
-dense eigenproblem at the end of each cycle or when convergence is reached, and
-then reuse it appropriately for subsequent
-solves~\cite{saad2000deflated,stathopoulos2009deflation,soodhalter2014krylov}.
-The routine \pk{KSPHPDDMGetDeflationSpace} can be used to retrieve such deflation subspaces, stored as distributed
-dense tall-and-skinny matrices. It is also possible to provide this information
-via \pk{KSPHPDDMSetDeflationSpace} in case end-users already know a good initial
-deflation subspace.
+These orthonormalization variants can be selected at runtime with the option \pk{-ksp\_hpddm\_qr (cholqr$|$mgs$|$cgs)}.
+
+For matrix--matrix products, there are
+currently specialized implementations for the following \pk{MatType}:
+\begin{itemize}
+    \item \pk{MATSEQAIJ}, standard sequential sparse matrix, based on compressed sparse row format;
+    \item \pk{MATSEQBAIJ}, block sparse matrix, based on block compressed sparse row format;
+    \item \pk{MATSEQSBAIJ}, symmetric block sparse matrix stored in upper triangular form.
+    \item \pk{MATMPIAIJ}, standard parallel sparse matrix;
+    \item \pk{MATSHELL}, user-defined matrix;
+    \item \pk{MATNEST}, block-defined matrix with nested submatrices;
+    \item \pk{MATAIJCUSPARSE}, sequential or parallel sparse matrix, offloaded to a GPU using cuSPARSE~\cite{cusparse-web-page}.
+\end{itemize}
+
+The following preconditioners have specialized implementations for dealing
+efficiently with multiple vectors:
+\begin{enumerate}
+    %\item \pk{PCNONE}: $P_{A'} = I$, identity matrix;
+    \item \pk{PCKSP}: preconditionining via an embedded Krylov;
+    \item \pk{PCMAT}: preconditioning via matrix multiplication;
+    %\item \pk{PCSHELL}: user-defined preconditioner, e.g., matrix-free;
+    % \item \pk{PCSPAI}: sparse approximate inverse~\cite{grote1997parallel};
+    \item \pk{PCHARA}~\cite{10.1145/3232850,ambartsumyan2020hierarchical}: hierarchical matrix preconditioners; 
+    \item \pk{PCHPDDM}: see~\cref{sec:PCHPDDM};\label{it:hpddm}
+    \item \pk{PCASM} and \pk{PCGASM}: overlapping Schwarz methods;\label{it:asm}
+    \item \pk{PCBJACOBI}: block Jacobi;\label{it:bjacobi}
+    \item \pk{PCLU} or \pk{PCCHOLESKY}: exact $LU$ or Cholesky factorization;\label{it:lu}
+    \item \pk{PCILU} or \pk{PCICC}: incomplete $LU$ or Cholesky
+        factorization.\label{it:ilu}
+\end{enumerate}
+One appealing feature of domain decomposition-like preconditioners (i.e. 
+\cref{it:hpddm,it:asm,it:bjacobi}) is that they most often rely on exact or inexact
+factorizations~ (\cref{it:lu,it:ilu}) as subdomain solvers.
+In these cases, it is possible to access the so-called factored matrix $F$
+%via \pk{PCFactorGetMatrix} and then call \pk{MatMatSolve}($F$, $\tilde{Y}$, $\tilde{X}$).
+via \pk{PCFactorGetMatrix}, and then call \pk{MatMatSolve}($F$, $X$, $Y$) to take advantage
+of blocked forward eliminations and backward substitutions from the various factorization packages
+interfaced with PETSc. In the case of a sparse matrix,
+this strategy is possible with: MUMPS~\cite{amestoy2001fully},
+SuiteSparse~\cite{davis2004algorithm}, MKL
+PARDISO or CPARDISO~\cite{mkl-web-page}, and SuperLU~\cite{li2005overview} or
+SuperLU\_DIST~\cite{li2003superlu_dist}. For the case of a dense
+matrix: ScaLAPACK and Elemental~\cite{poulson2013elemental}. 
+Future work will consider extending the multigrid framework \pk{PCMG}, as well as other preconditioning classes to increase the arithmetic intensity of the
+block preconditioner application phase.
 
@@
 % ICNTL(27)\footnote{\url{http://mumps.enseeiht.fr/doc/userguide_5.3.1.pdf}, section 6.1}.
-    \subsubsection{Systems with multiple right-hand sides\label{sec:kspmatsolve}}
-Along with \pk{PCMatApply} introduced in PETSc 3.14.0, a new function
-\pk{KSPMatSolve} was added for solving systems with multiple right-hand sides.
-It takes as input argument a \pk{KSP} and, just as for \pk{PCMatApply}, two
-dense \pk{Mat}. If the input \pk{KSP} is not of type \pk{KSPHPDDM}, the
-solution phase is done in a column by column fashion.  Otherwise, it is
-possible to use any \pk{KSPHPDDM} \mbox{(pseudo-)block} Krylov methods.
-Furthermore, a function \pk{KSPSetMatSolveBlockSize} is also provided to
-decompose a single large block of column vectors into multiple sub-blocks.
-\pk{KSPMatSolve} is then called repeatedly until all sub-blocks are traversed.
-This was inspired by MUMPS~\cite{amestoy2001fully} option
-ICNTL(27)\footnote{\url{http://mumps.enseeiht.fr/doc/userguide_5.3.3.pdf},
-section 6.1}.
-
+
+{\color{red} THIS PARAGRAPH IS A BIT DETACHED FROM THE REST OF THE SECTION.
+WHAT ABOUT ADDING A SMALL PARAGRAPH IN THE INTRODUCTION (in the spirit of third party code that directly benefits from the PETSc-HPDDM interface) RIGHT BEFORE THE OUTLINE OF THE PAPER
+AND CONTINUE THE DISCUSSION IN ~\cref{sec:lobpcg}?}
 With these added functions, it is now possible to do code refactorization for
@@
 Eigenvalue Problem computations. SLEPc is an add-on library that extends PETSc
-with several classes for linear and nonlinear eigenvalue problems, as well as
-tools to, e.g., facilitate the use of shift-and-invert spectral transformations.
-The implementations of LOBPCG~\cite{knyazev2001toward} and CISS~\cite{sakurai2003projection}
+with several classes for linear and nonlinear eigenvalue problems, as well as with
+tools to facilitate the use of shift-and-invert spectral transformations.
+The implementations {\color{red} spell out LOBPGC and CISS} of LOBPCG~\cite{knyazev2001toward} and CISS~\cite{sakurai2003projection}
 in SLEPc were previously not using blocking when applying the preconditioner or solving linear systems.
@@
 
-As a side note, when dealing with
-\pk{MatKAIJ} matrices, for which it is assumed that an operator may be written as $I \otimes S
-+ A \otimes T$, with $S$ and $T$ elements of $\mathbb{K}^{p\times q}$ and
-$A$ a standard sparse matrix of dimension $n$, it is also possible to use block Krylov methods. 
-Indeed, if $p=q=k$, $S = \lambda I$, and $T =
-\mu I$, then solving $(I \otimes S + A \otimes T) x = b$ may be cast into a
-system with $k$ right-hand sides $(\lambda I + \mu A)X = B$ where $B$ is
-the row-major representation of $b$, i.e., $b_{k \cdot i + j} = B_{i, j},
-\forall (i, j) \in \llbracket 1;n\rrbracket \times \llbracket 1;k\rrbracket$.
-On the one hand, due to the structure of the operator, it is challenging to
-define appropriate preconditioners for $(I \otimes S + A \otimes T)$: only
-\pk{PCPBJACOBI} is currently implemented for \pk{MatKAIJ}. On the other hand,
-it may be simpler to efficiently approximate the inverse of $A$. In
-\pk{KSPHPDDM}, the type of the operators is checked: if it is \pk{MatKAIJ} and
-the previous condition on $p, q, S$, and $T$ is fulfilled, the single
-right-hand side system is automatically cast into a system with multiple
-right-hand sides instead.
+%As a side note, when dealing with
+%\pk{MatKAIJ} matrices, for which it is assumed that an operator may be written as $I \otimes S
+%+ A \otimes T$, with $S$ and $T$ elements of $\mathbb{K}^{p\times q}$ and
+%$A$ a standard sparse matrix of dimension $n$, it is also possible to use block Krylov methods. 
+%Indeed, if $p=q=k$, $S = \lambda I$, and $T =
+%\mu I$, then solving $(I \otimes S + A \otimes T) x = b$ may be cast into a
+%system with $k$ right-hand sides $(\lambda I + \mu A)X = B$ where $B$ is
+%the row-major representation of $b$, i.e., $b_{k \cdot i + j} = B_{i, j},
+%\forall (i, j) \in \llbracket 1;n\rrbracket \times \llbracket 1;k\rrbracket$.
+%On the one hand, due to the structure of the operator, it is challenging to
+%define appropriate preconditioners for $(I \otimes S + A \otimes T)$: only
+%\pk{PCPBJACOBI} is currently implemented for \pk{MatKAIJ}. On the other hand,
+%it may be simpler to efficiently approximate the inverse of $A$. In
+%\pk{KSPHPDDM}, the type of the operators is checked: if it is \pk{MatKAIJ} and
+%the previous condition on $p, q, S$, and $T$ is fulfilled, the single
+%right-hand side system is automatically cast into a system with multiple
+%right-hand sides instead.
 
@@
 problem constructed by Philippe H.\ Geubelle and Spandan Maiti." The goal of
-this paragraph is to explain how the results have been reproduced in PETSc.
-The dataset and the source code are available at
+this paragraph is to explain how the results can be reproduced with PETSc and \pk{KSPHPDDM}: the matrix files used are available at
 \url{https://gitlab.com/petsc/datafiles/-/tree/master/matrices/hpddm/GCRODR}
-and
-\url{https://www.mcs.anl.gov/petsc/petsc-master/src/ksp/ksp/tutorials/ex75.c.html}
-respectively. In order to check the correctness of the \pk{KSPHPDDM} interface,
+while the driver code is part of the PETSc testsuite and available at
+\url{https://www.mcs.anl.gov/petsc/petsc-master/src/ksp/ksp/tutorials/ex75.c.html}. In order to check the correctness of the \pk{KSPHPDDM} interface,
 the following tests are performed:
@@
 
-In particular, the first three matrix types from~\cref{sec:interface-ksp}, \pk{MatSeqAIJ}, \pk{MatSeqBAIJ}, and \pk{MatSeqSBAIJ} are first considered.
-Their distributed memory counterparts, e.g., \pk{MatMPIAIJ}, rely on the sequential implementations, with each process
+In particular, we consider the first three matrix types from~\cref{sec:interface-ksp}, namely \pk{MATSEQAIJ}, \pk{MATSEQBAIJ}, and \pk{MATSEQSBAIJ}.
+Their distributed memory counterparts, e.g., \pk{MATMPIAIJ}, rely on the sequential implementations, with each process
 storing two such matrices, one with local rows and columns, and another with local rows and ``nonlocal'' columns.
@@
 in drawing conclusions for the intraprocess performance of the parallel formats.
-For this reason, from now on, the \pk{Seq} substring is dropped.
+For this reason, from now on, we drop the \pk{SEQ} substring.
 The performance of the multiplication primitives of the three aforementioned types will also be compared to
@@
 
-For benchmarking, we first discretize the Poisson equation on a cube with $\mathbb{P}_1$ finite elements
+For benchmarking, we first discretize the Poisson equation on a cube with trilinear finite elements
 and obtain a matrix $A$ of dimension one million. Though the numerical
@@
 All results have been obtained using double precision arithmetic and 32-bit integers. The scaled efficiency measured in GFLOP per second is reported
-with respect to the baseline \pk{MatAIJ} implementation with $N=1$ in~\cref{fig:perfAX}, where performances have been averaged over five consecutive
+with respect to the baseline \pk{MATAIJ} implementation with $N=1$ in~\cref{fig:perfAX}, where performances have been averaged over five consecutive
 product operations.
-They can be reproduced by using the mini-app \texttt{MatProduct.c} from \url{https://github.com/prj-/jolivet2020petsc} and any input \pk{MatAIJ} stored in binary format,
+They can be reproduced by using the mini-app \texttt{MatProduct.c} from \url{https://github.com/prj-/jolivet2020petsc} and any input \pk{MATAIJ} stored in binary format,
 here using the default name \pk{binaryoutput}, available at \url{http://jolivet.perso.enseeiht.fr/binaryoutput}.
@@
            \addplot+[draw=black,fill, area legend,postaction={pattern=bricks}] table [x expr=\thisrow{N}>7?\thisrow{N}/2:\thisrow{N},y expr=\thisrow{sbaij-1-F}/\thisrow{sbaijmkl-seq-1-T}/(\storeresultI*1000000000.0)]{\loadedMatProduct};\label{pgfplots:mklsbaijseq1}
-           \addlegendentry{\pk{MatAIJ}}
-           \addlegendentry{\pk{MatBAIJ}}
-           \addlegendentry{\pk{MatSBAIJ}}
-           \addlegendentry{\pk{MatAIJMKL}}
-           \addlegendentry{\pk{MatBAIJMKL}}
-           \addlegendentry{MatSBAIJMKL}
+           \addlegendentry{\pk{MATAIJ}}
+           \addlegendentry{\pk{MATBAIJ}}
+           \addlegendentry{\pk{MATSBAIJ}}
+           \addlegendentry{\pk{MATAIJMKL}}
+           \addlegendentry{\pk{MATBAIJMKL}}
+           \addlegendentry{\pk{MATSBAIJMKL}}
 \node[rotate=0,fill=white,inner sep=0pt,anchor=west] at (rel axis cs:1.02, 0.5) {$b=1$};
@@
 \begin{itemize}
-    \item with a block size of 1, top plot $b=1$ in~\cref{fig:perfAX}, using any type but \pk{MatAIJ}~\ref{pgfplots:aij1} is counterproductive when $N>1$, see~\ref{pgfplots:baij1} and~\ref{pgfplots:sbaij1}, even when using MKL, see~\ref{pgfplots:mklbaijseq1} and~\ref{pgfplots:mklsbaijseq1}.
-With 8 or more columns, the performance of \pk{MatAIJ} plateaus and the efficiency reaches approximately 200\%, while the performance of the MKL primitives stagnate at around 175\%, see~\ref{pgfplots:aij1};
-for the single column case, \pk{MatSBAIJ} and \pk{MatAIJMKL} deliver slightly better performances;
+    \item with a block size of 1, top plot $b=1$ in~\cref{fig:perfAX}, using any type but \pk{MATAIJ}~\ref{pgfplots:aij1} is counterproductive when $N>1$, see~\ref{pgfplots:baij1} and~\ref{pgfplots:sbaij1}, even when using MKL, see~\ref{pgfplots:mklbaijseq1} and~\ref{pgfplots:mklsbaijseq1}.
+With 8 or more columns, the performance of \pk{MATAIJ} plateaus and the efficiency reaches approximately 200\%, while the performance of the MKL primitives stagnate at around 175\%, see~\ref{pgfplots:aij1};
+for the single column case, \pk{MATSBAIJ} and \pk{MATAIJMKL} deliver slightly better performances;
     \item for block sizes lower or equal than 5, loops for matrix--vector and matrix--matrix multiplications are unrolled by hand for block formats.
-This yields rather disappointing performance, except for \pk{MatAIJ}, see~\ref{pgfplots:aij3} in middle plot $b=3$, and its 350\% efficiency, even with a moderate number of columns.
+This yields rather disappointing performance, except for \pk{MATAIJ}, see~\ref{pgfplots:aij3} in middle plot $b=3$, and its 350\% efficiency, even with a moderate number of columns.
 For block formats \ref{pgfplots:baij3} and~\ref{pgfplots:sbaij3}, given the small value of $b$, it could be beneficial to switch to optimized libraries for small matrices, e.g., LIBXSMM~\cite{heinecke2016libxsmm}.
@@
     \item for block sizes larger than 5, block entry multiplication with PETSc block formats is performed using \texttt{?gemv} or \texttt{?gemm} operations.
-For \pk{MatAIJ}~\ref{pgfplots:aij10} in the bottom plot corresponding to $b=6$, the efficiency quickly caps near 400\% for $N\in\{8,16,32,64\}$, while reaching 500\% for large number of columns for the block formats, see~\ref{pgfplots:baij10} and~\ref{pgfplots:sbaij10} for $N=64$.
+For \pk{MATAIJ}~\ref{pgfplots:aij10} in the bottom plot corresponding to $b=6$, the efficiency quickly caps near 400\% for $N\in\{8,16,32,64\}$, while reaching 500\% for large number of columns for the block formats, see~\ref{pgfplots:baij10} and~\ref{pgfplots:sbaij10} for $N=64$.
 With these numbers of columns, PETSc and MKL perform similarly.
@@
 % These results are of course specific to the sparsity pattern of the initial matrix $A$, and thus $\mathbb{A} = A\otimes T$, and matrix types.
-For example, if the matrices were to be offloaded to a GPU, one could instead benchmark \pk{MatAIJcuSPARSE}.
+For example, if the matrices were to be offloaded to a GPU, one could instead benchmark \pk{MATAIJcuSPARSE}.
 
@@
 Results reported in~\cref{fig:perfAXpar} have been scaled with respect to the
-baseline implementation \pk{MatAIJMKL} with $N=1$. They have been obtained on a
+baseline implementation \pk{MATAIJMKL} with $N=1$. They have been obtained on a
 single node of {Jean Zay}, a system composed of \pgfmathprintnumber[assume math
@@
 \end{pgfonlayer}{foreground}
-           \addlegendentry{\pk{MatAIJMKL}}
-           \addlegendentry{\pk{MatBAIJMKL}}
-           \addlegendentry{MatSBAIJMKL}
-           \addlegendentry{\pk{MatAIJCUSPARSE}}
+           \addlegendentry{\pk{MATAIJMKL}}
+           \addlegendentry{\pk{MATBAIJMKL}}
+           \addlegendentry{\pk{MATSBAIJMKL}}
+           \addlegendentry{\pk{MATAIJCUSPARSE}}
 \node[rotate=0,fill=white,inner sep=0pt,anchor=west] at (rel axis cs:1.02, 0.5) {$b=1$};
@@
 For the sake of completeness, the performance of
-\pk{MatAIJMKL} with $N=1$ but different numbers of threads is also reported in
+\pk{MATAIJMKL} with $N=1$ but different numbers of threads is also reported in
 \cref{fig:scaling-Ax}. Clearly, it is not advised to use the full socket to
@@
 and a full GPU device. All in all, the CPU configuration which reaches the
-highest percentage of peak is MatSBAIJMKL~\ref{pgfplots:sbaijmkl3} with $b=3$
+highest percentage of peak is \pk{MATSBAIJMKL}~\ref{pgfplots:sbaijmkl3} with $b=3$
 and $N=64$, rightmost middle plot in~\cref{fig:perfAXpar}, with approximately
 63 GFLOP/s. It is still less than 1\% of peak for an Intel Xeon Gold 6248.
-For the GPU configuration, \pk{MatAIJCUSPARSE}~\ref{pgfplots:cusparse10} with $b=6$
+For the GPU configuration, \pk{MATAIJCUSPARSE}~\ref{pgfplots:cusparse10} with $b=6$
 and $N=64$, rightmost bottom plot in~\cref{fig:perfAXpar}, performs at around
@@
     \end{tikzpicture}
-    \caption{Scalability of \pk{MatMult} with \pk{MatAIJMKL} and different number of processes.\label{fig:scaling-Ax}}
+    \caption{Scalability of \pk{MatMult} with \pk{MATAIJMKL} and different number of threads.\label{fig:scaling-Ax}}
 \end{figure}
@@
 Thus, the number of outer iterations is higher than with \pk{PCHPDDM} or \pk{PCGAMG} since the inner solves are not converging to meaningful solutions in just 5 iterations.
-For both multilevel preconditioners, however, it seems that they perform an equivalent job since the difference in number of outer iterations is only 1.
+The multilevel preconditioners perform similarly, since the difference in the number of outer iterations is minimal (1).
 The second and third lines represent the number of times \pk{PCApply} and \pk{PCMatApply} are called.
-Since \pk{PCGAMG} does not handle blocking, at each pseudo-block GMRES iteration, \pk{PCApply} is called on all active columns.
+Since \pk{PCGAMG} does not handle blocking, at each pseudo-block GMRES iteration, \pk{PCApply} is called on each single active column.
 \pk{PCASM} and \pk{PCHPDDM} handle blocking, so only rely on \pk{PCMatApply}.
@@
 possible to select the corresponding \pk{PC}
-using the command line option \pk{-pc\_type hpddm}, or the routine \pk{PCSetType}(pc,
+using the command line option \pk{-pc\_type hpddm}, or the routine \pk{PCSetType}(\pk{pc},
 \pk{PCHPDDM}).
  % and its Fortran equivalent.
-In this section, we are interested in solving
-\begin{equation}
-    Ax = b,
-\end{equation}
-for a given coefficient matrix $A$ and right-hand side $b$ of dimension $n$
-assembled on $N$ processes. For simplicity, a ksp($A$, $P_A$) is used next, i.e., the matrix used to
-define the preconditioner is $A$ itself. Without anything more from the user,
+
+In this section, we are interested in constructing a preconditioner
+for a given coefficient matrix $A$. Without user intervention,
 \pk{PCHPDDM} is strictly equivalent to \pk{PCASM}. That is, if $A$ is distributed among
@@
 the local unassembled matrices~$\{\mathring{A}_i\}_{i=1}^N$, also known as
-Neumann matrices. These are yielded by the
-discretization of $\mathcal{L}$ on each subdomain, with added natural boundary
-conditions on the subdomain boundaries that are not boundaries of the initial
-domain $\Omega$. In the case of the 1D Poisson
-equation with homogeneous Dirichlet boundary conditions discretized using $\mathbb{P}_1$ finite elements, an explicit representation of these  operators is given~\cref{fig:poisson} for a
+Neumann matrices, representing the
+discretization of $\mathcal{L}$ on the extended local subdomain with overlap, endowed with natural boundary
+conditions. In the case of the 1D Poisson
+equation with homogeneous Dirichlet boundary conditions discretized using linear finite elements, an explicit representation of these  operators is given~\cref{fig:poisson} for a
 two-domain decomposition.
@@
 Other \pk{DMType}s, including user-defined, may be supported by
-implementing the proper \pk{DMCreateNeumannOverlap} callback. Support for \pk{DMDA}
+implementing the proper \pk{DMCreateNeumannOverlap} callback. Support for the structured grid infrastructure (\pk{DMDA})
 could be easily added upon request.
@@
 eigenvectors on each process, since in this case the coarse operator will be
-assembled using the \pk{MatMPISBAIJ} or \pk{MatMPIBAIJ} matrix type, with
-usually large block sizes. This can have serious memory and performance
-implications, see~\cref{fig:perfAX}. When applying the preconditioner,
-\texttt{MPI\_Gather} and \texttt{MPI\_Scatter} operations are used to move
-between levels, instead of the variable-size alternatives \texttt{MPI\_Gatherv}
-and \texttt{MPI\_Scatterv}, that could exhibit worse performance. In addition,
+assembled using the blocked matrix formats, with
+usually large block sizes, with consequent impact on memory, floating point and message passing performances.
+% When applying the preconditioner,
+%\texttt{MPI\_Gather} and \texttt{MPI\_Scatter} operations are used to move
+%between levels, instead of the variable-size alternatives \texttt{MPI\_Gatherv}
+%and \texttt{MPI\_Scatterv}, that could exhibit worse performance.
+In addition,
 standard eigensolvers such as \pk{EPSKRYLOVSCHUR} from SLEPc, cannot cheaply
@@
 \pk{(deflated$|$additive$|$balanced)} or the routine
-\pk{PCHPDDMSetCoarseCorrectionType}. It is important to keep in mind that all
-these correction formulae may be applied to either a single vector or a block
+\pk{PCHPDDMSetCoarseCorrectionType}. All of
+the correction formulas can be applied to either a single vector or a block
 of multiple vectors optimally. Indeed, the one-level preconditioner $M^{-1}$ is
@@
 restriction-correction-interpolation operator $Q$ is applied in a block fashion
-in HPDDM, and not column by column. See the effect of this strategy in~\cref{sec:lobpcg}.
+in HPDDM, and not column by column. % See the effect of this strategy in~\cref{sec:lobpcg}.
 
@@
 \section{Conclusion and perspectives}\label{sec:conclusion}
+{\color{red} I WOULD PREFER CONCLUSIONS in ACTIVE TENSE. ALSO, THE TONE IS 
+TOO MUCH COLLOQUIAL (i.e. using "and such"). THERE ARE FEW POINTS WE CAN ADD FOR
+FUTURE WORK: BDD-GENEO?, GPU AWARE HPDDM? SOME MORE SPECIFICS ON BLOCK KRYLOV-SCHUR?}
+
 In this paper, the interface between PETSc and HPDDM is presented. The new
@@
 examples are provided to show the applicability of these. Solving well-known
-problems, various options have been described. Some more unconventional test
-cases were commented as well, for example using \pk{MatKAIJ} or \pk{MatIS} matrix types.
+problems, various options have been described. %Some more unconventional test
+%cases were commented as well, for example using \pk{MatKAIJ} or \pk{MatIS} matrix types.
 Overall, the interface paves the way for having recycling and block Krylov
@@
 
+\bibitem{ambartsumyan2020hierarchical}
+{\sc I.~Ambartsumyan, W.~Boukaram, T.~Bui-Thanh, O.~Ghattas, D.~Keyes, G.~Stadler, G.~Turkiyyah and S.~Zampini},
+  {\em Hierarchical Matrix Approximations of Hessians Arising in Inverse Problems Governed by PDEs},
+  arXiv preprint arXiv:2003.10173 (2020).
+
 \bibitem{anderson2019mfem}

### 2020-07-18 article.tex
Style fixes

@@
 methods~\cite{jolivet2016block} which can efficiently deal with linear
-systems with multiple right-hand sides (i.e. block iterative methods),
+systems with multiple right-hand sides, i.e., block iterative methods,
 and when there is a recurrence of varying coefficient matrices and right-hand
-sides (recycling iterative methods).
+sides, i.e., recycling iterative methods.
 
@@
 Krylov subspace methods are widely used in numerical linear algebra for solving
-linear systems of equations because of their memory efficiency~\cite{barrett1994templates}.
+linear systems of equations because of their memory efficiency~\cite{barrett1994templates,saad2003iterative}.
 They mainly rely on matrix--vector
@@
 \end{equation*}
-This continuous equation is discretized with $\mathbb{P}_3$ finite elements using FreeFEM.
+This continuous equation is discretized with third-order Lagrange finite elements using FreeFEM.
 The script is available at \url{https://github.com/prj-/jolivet2020petsc}.
@@
 well as optimized Schwarz methods~\cite{gander2006optimized}, which may be used
-in PETSc with \pk{PCSetModifySubMatrices}.  It also provides support for
+in PETSc with \pk{PCSetModifySubMatrices}, see~\cite{gander2019new}.  It also provides support for
 defining robust two-level methods using the Generalized Eigenvalue problem on
@@
 In order to use GenEO from PETSc, the auxiliary operators $\{\mathring{A}_i\}_{i=1}^N$
-and the corresponding local to global map of degrees of freedom (see~\cref{fig:poisson})
+and the corresponding local to global map of degrees of freedom, see~\cref{fig:poisson},
 can be provided using the routine \pk{PCHPDDMSetAuxiliaryMat}.
@@
 $\Gamma_N$ is defined as the complementary of $\Gamma_D$ with respect to the boundary of $\Omega$.
-\Cref{eq:elasticity} is discretized using
-$(\mathbb{P}_1, \mathbb{P}_1, \mathbb{P}_1)$ finite elements resulting in $\pgfmathprintnumber{593} \times 10^6$ number of unknowns
+\Cref{eq:elasticity} is discretized using piecewise linear
+finite elements resulting in $\pgfmathprintnumber{593} \times 10^6$ number of unknowns
 with approximately $\pgfmathprintnumber{45}$ nonzero coefficients per row.
@@
 . The physical domain $\Omega$ is the unit cube. It is decomposed in $10{,}272$ subdomains.
-\Cref{eq:bratu} is discretized using $\mathbb{P}_2$ finite elements.
+\Cref{eq:bratu} is discretized using second-order Lagrange finite elements.
 The number of unknowns is $\pgfmathprintnumber{217} \times 10^6$, with
@@
 \section*{Acknowledgments}
-The authors would like to thank S. Balay, J. Brown, V. Hapla, M. Knepley,
-and B. Smith for reviewing the successive merge requests in the PETSc
+The authors would like to thank S.~Balay, J.~Brown, V.~Hapla, M.~Knepley,
+and B.~Smith for reviewing the successive merge requests in the PETSc
 repository. This work was granted access to the GENCI-sponsored HPC resources of:
@@
 \bibitem{falgout2002hypre}
-{\sc R.~Falgout and U.~Yang}, {\em \emph{hypre}: a library of high performance
-  preconditioners}, Computational Science---ICCS 2002,  (2002), pp.~632--641,
-  \url{https://www.llnl.gov/casc/hypre}.
+{\sc R.~Falgout and U.~M. Yang}, {\em \emph{hypre}: a library of high
+  performance preconditioners}, Computational Science---ICCS 2002,  (2002),
+  pp.~632--641, \url{https://www.llnl.gov/casc/hypre}.
 
@@
   Numerical Analysis, 44 (2006), pp.~699--731.
+
+\bibitem{gander2019new}
+{\sc M.~J. Gander and S.~Van~Criekingen}, {\em {New Coarse Corrections for
+  Optimized Restricted Additive Schwarz Using PETSc}}, in International
+  Conference on Domain Decomposition Methods, Springer, 2019.
 
@@
 \bibitem{smith2004domain}
-{\sc B.~F. Smith, P.~Bj{\o}rstad, and W.~Gropp}, {\em Domain decomposition:
+{\sc B.~F. Smith, P.~Bj{\o}rstad, and W.~D. Gropp}, {\em Domain decomposition:
   parallel multilevel methods for elliptic partial differential equations},

### 2020-07-18 article.tex
First attempt at fixing the last (?) issues

@@
     The flexibility of the implementation %, which for example may deal with various matrix formats,
-    is showcased and explained with minimalist, easy-to-run and reproducible examples
-    ranging from {\color{red}fracture mechanics to linear stability analysis}
+    is showcased and explained with minimalist, easy-to-run, and reproducible examples,
     to ease the integration of these algorithms into more advanced frameworks.
+    The examples simulate the linear or nonlinear solves of typical
+    applications in fracture mechanics, linear stability analysis or more
+    broadly-speaking eigenanalysis, elasticity, combustion, and
+    electromagnetism.
 \end{abstract}
@@
 \section{Introduction}
-Implementing and maintaining robust and scalable applications is highly relevant
-in the field of computational science and engineering. The efficient production of
-such applications requires a rich ecosystem of state-of-the-art reusable libraries
-\cite{Bartlett:2017:XFT}. In this paper, we focus on the interoperability of two such libraries.
+Implementing and maintaining robust and scalable applications is highly
+relevant in the field of computational science and engineering. The efficient
+production of such applications requires a rich ecosystem of state-of-the-art
+reusable libraries \cite{Bartlett:2017:XFT}. They can then be used by domain
+specialists who can decide what are the best tools available for their
+applications. In this paper, we focus on the interoperability of two such
+libraries.
 
@@
 and facilitate the analysis and comparison of performance results~\cite{hoefler2015scientific}.
+Furthermore, the benefits of code refactorization are shown for
+applications that deal with multiple right-hand sides. This is the case of some
+eigensolvers in SLEPc~\cite{hernandez2005ssf}, the Scalable Library for
+Eigenvalue Problem computations. SLEPc is an add-on library that extends PETSc
+with several classes for linear and nonlinear eigenvalue problems, as well as with
+tools to facilitate the use of shift-and-invert spectral transformations.
 
@@
 % ICNTL(27)\footnote{\url{http://mumps.enseeiht.fr/doc/userguide_5.3.1.pdf}, section 6.1}.
-
-{\color{red} THIS PARAGRAPH IS A BIT DETACHED FROM THE REST OF THE SECTION.
-WHAT ABOUT ADDING A SMALL PARAGRAPH IN THE INTRODUCTION (in the spirit of third party code that directly benefits from the PETSc-HPDDM interface) RIGHT BEFORE THE OUTLINE OF THE PAPER
-AND CONTINUE THE DISCUSSION IN ~\cref{sec:lobpcg}?}
-With these added functions, it is now possible to do code refactorization for
-applications that deal with multiple right-hand sides. This is the case of some
-eigensolvers in SLEPc~\cite{hernandez2005ssf}, the Scalable Library for
-Eigenvalue Problem computations. SLEPc is an add-on library that extends PETSc
-with several classes for linear and nonlinear eigenvalue problems, as well as with
-tools to facilitate the use of shift-and-invert spectral transformations.
-The implementations {\color{red} spell out LOBPGC and CISS} of LOBPCG~\cite{knyazev2001toward} and CISS~\cite{sakurai2003projection}
-in SLEPc were previously not using blocking when applying the preconditioner or solving linear systems.
-In version 3.14.0, we have adapted these solvers to employ \pk{KSPMatSolve} and \pk{PCMatApply}.
-In the case of LOBPCG, which is a purely blocked method, all steps were already
-implemented as block operations, except the applications of the preconditioner,
-which ruined the benefits of blocking. With the new implementation, the gains
-are striking, as reported in~\cref{sec:lobpcg}.
 
@@
 of columns $N \in \{2,8,16,32,64\}$. For the single column case $N=1$, the \pk{MatMult} operation is used.
-All results have been obtained using double precision arithmetic and 32-bit integers. The scaled efficiency measured in GFLOP per second is reported
+All results have been obtained using double-precision arithmetic and 32-bit integers. The scaled efficiency measured in GFLOP per second is reported
 with respect to the baseline \pk{MATAIJ} implementation with $N=1$ in~\cref{fig:perfAX}, where performances have been averaged over five consecutive
@@
 \subsubsection{Blocking inside SLEPc\label{sec:lobpcg}}
+The implementations of the Locally Optimal Block Preconditioned Conjugate
+Gradient~\cite{knyazev2001toward}~(LOBPCG) and the Contour Integral Spectrum
+Slicing method~\cite{sakurai2003projection}~(CISS) in SLEPc were previously not
+using blocking when applying the preconditioner or solving linear systems. In
+version 3.14.0, these solvers have been adapted to employ \pk{KSPMatSolve} and
+\pk{PCMatApply}.  In the case of LOBPCG, which is a purely blocked method, all
+steps were already implemented as block operations, except the applications of
+the preconditioner, which undermined the benefits of blocking. With the new
+implementation, the gains are striking, as reported in~\cref{sec:lobpcg}.
+
 Finally, the performance of LOBPCG~\cite{knyazev2001toward} as implemented in SLEPc is studied using three different preconditioners:
@@
 \section{Conclusion and perspectives}\label{sec:conclusion}
-{\color{red} I WOULD PREFER CONCLUSIONS in ACTIVE TENSE. ALSO, THE TONE IS 
-TOO MUCH COLLOQUIAL (i.e. using "and such"). THERE ARE FEW POINTS WE CAN ADD FOR
-FUTURE WORK: BDD-GENEO?, GPU AWARE HPDDM? SOME MORE SPECIFICS ON BLOCK KRYLOV-SCHUR?}
-
-In this paper, the interface between PETSc and HPDDM is presented. The new
-registered Krylov and overlapping Schwarz methods are explained and numerical
-examples are provided to show the applicability of these. Solving well-known
-problems, various options have been described. %Some more unconventional test
+
+In this paper, we presente the interface between PETSc and HPDDM. The new
+registered Krylov and overlapping Schwarz methods are explained and we show the
+applicability of these  with numerical examples. Solving well-known problems,
+we describe various runtime options. %Some more unconventional test
 %cases were commented as well, for example using \pk{MatKAIJ} or \pk{MatIS} matrix types.
@@
 
+Concerning SLEPc, we plan to implement new solvers or extend existing ones to
+further exploit block solve primitives, for instance, block
+Krylov--Schur~\cite{zhou2008block}. The current code for Arnoldi iterations
+would then handle block Arnoldi iterations.
+
 Concerning HPDDM, its integration inside PETSc offers a much-needed flexibility
@@
 runtime many parameters that are determined at compile-time in the standard implementation, e.g.,
-subdomain solvers, matrix types, and such.
-
-Concerning SLEPc, as a future work we plan to implement new solvers (or extend existing
-ones) to further exploit the block solve primitives, for instance block Krylov-Schur.
+subdomain solvers and matrix types.
+
+The interface between PETSc and HPDDM has still room for future improvements,
+and we discuss openly two of them next. First, HPDDM handles mixed-precision,
+e.g., using single-precision scalars for assembling coarse operators while
+using double-precision scalars at the fine level. The topic of mixed-precision
+is under scrutiny for the next major overhaul of PETSc. Second, HPDDM does not
+currently handle GPU efficiently, but the interface with PETSc provides a very
+thorough test bed with \pk{Mat} and \pk{PC} implementations that are exploiting
+GPU.
 
@@
 
+\bibitem{ambartsumyan2020hierarchical}
+{\sc I.~Ambartsumyan, W.~Boukaram, T.~Bui-Thanh, O.~Ghattas, D.~Keyes,
+  G.~Stadler, G.~Turkiyyah, and S.~Zampini}, {\em {Hierarchical Matrix
+  Approximations of Hessians Arising in Inverse Problems Governed by PDEs}},
+  arXiv:2003.10173,  (2020).
+
 \bibitem{amestoy2001fully}
@@
 
-\bibitem{ambartsumyan2020hierarchical}
-{\sc I.~Ambartsumyan, W.~Boukaram, T.~Bui-Thanh, O.~Ghattas, D.~Keyes, G.~Stadler, G.~Turkiyyah and S.~Zampini},
-  {\em Hierarchical Matrix Approximations of Hessians Arising in Inverse Problems Governed by PDEs},
-  arXiv preprint arXiv:2003.10173 (2020).
-
 \bibitem{anderson2019mfem}
 {\sc R.~Anderson, J.~Andrej, A.~Barker, J.~Bramwell, J.-S. Camier, J.~Cerveny,
-  V.~Dobrev, Y.~Dudouit, A.~Fisher, T.~Kolev, et~al.}, {\em {MFEM: a modular
-  finite element methods library}}, arXiv:1911.09220,  (2019),
-  \url{http://mfem.org}.
+  V.~Dobrev, Y.~Dudouit, A.~Fisher, T.~Kolev, W.~Pazner, M.~Stowell, V.~Tomov,
+  I.~Akkerman, J.~Dahm, D.~Medina, and S.~Zampini}, {\em {MFEM: A modular
+  finite element methods library}}, Computers \& Mathematics with Applications,
+   (2020), \url{http://mfem.org}.
 
@@
 
+\bibitem{zhou2008block}
+{\sc Y.~Zhou and Y.~Saad}, {\em {Block Krylov--Schur method for large symmetric
+  eigenvalue problems}}, Numerical Algorithms, 47 (2008), pp.~341--359.
+
 \end{thebibliography}

### 2020-07-19 article.tex
minor

@@
     to ease the integration of these algorithms into more advanced frameworks.
-    The examples simulate the linear or nonlinear solves of typical
-    applications in fracture mechanics, linear stability analysis or more
-    broadly-speaking eigenanalysis, elasticity, combustion, and
+    The examples provided cover applications from
+    %simulate the linear or nonlinear solves of typical
+    %applications in fracture mechanics, linear stability analysis or more
+    %broadly-speaking
+    eigenanalysis, elasticity, combustion, and
     electromagnetism.
@@
 steps were already implemented as block operations, except the applications of
-the preconditioner, which undermined the benefits of blocking. With the new
-implementation, the gains are striking, as reported in~\cref{sec:lobpcg}.
-
-Finally, the performance of LOBPCG~\cite{knyazev2001toward} as implemented in SLEPc is studied using three different preconditioners:
+the preconditioner, which undermined the benefits of blocking.
+%With the new
+%implementation, the gains are striking, as reported in~\cref{sec:lobpcg}.
+
+The performance of LOBPCG~\cite{knyazev2001toward} as implemented in SLEPc is studied using three different preconditioners:
 \begin{itemize}
@@
 
-In this paper, we presente the interface between PETSc and HPDDM. The new
-registered Krylov and overlapping Schwarz methods are explained and we show the
-applicability of these  with numerical examples. Solving well-known problems,
-we describe various runtime options. %Some more unconventional test
+In this paper, we presented the interface between PETSc and HPDDM and discussed the new Krylov and overlapping Schwarz methods, providing several
+numerical examples and describing various runtime options. %Some more unconventional test
 %cases were commented as well, for example using \pk{MatKAIJ} or \pk{MatIS} matrix types.

### 2020-07-19 article.tex
Title + clarification about MATSBAIJMKL.

@@
 \hypersetup{
-  pdftitle={KSPHPDDM and PCHPDDM: Extending PETSc with Robust Overlapping Schwarz Preconditioners and Advanced Krylov Solvers},
+  pdftitle={KSPHPDDM and PCHPDDM: Extending PETSc with Robust Overlapping Schwarz Preconditioners and Advanced Krylov Methods},
   pdfauthor={P. Jolivet, J. E. Roman, and S. Zampini}
@@
            \addlegendentry{\pk{MATBAIJMKL}}
-           \addlegendentry{\pk{MATSBAIJMKL}}
+           \addlegendentry{MATSBAIJMKL}
 \node[rotate=0,fill=white,inner sep=0pt,anchor=west] at (rel axis cs:1.02, 0.5) {$b=1$};
@@
            \addlegendentry{\pk{MATBAIJMKL}}
-           \addlegendentry{\pk{MATSBAIJMKL}}
+           \addlegendentry{MATSBAIJMKL}
            \addlegendentry{\pk{MATAIJCUSPARSE}}
@@
 and a full GPU device. All in all, the CPU configuration which reaches the
-highest percentage of peak is \pk{MATSBAIJMKL}~\ref{pgfplots:sbaijmkl3} with $b=3$
+highest percentage of peak is MATSBAIJMKL~\ref{pgfplots:sbaijmkl3} with $b=3$
 and $N=64$, rightmost middle plot in~\cref{fig:perfAXpar}, with approximately
-63 GFLOP/s. It is still less than 1\% of peak for an Intel Xeon Gold 6248.
+63 GFLOP/s. This format is not available yet in PETSc, but it is handled by the
+mini-app. Still, it is less than 1\% of peak for an Intel Xeon Gold 6248.
 For the GPU configuration, \pk{MATAIJCUSPARSE}~\ref{pgfplots:cusparse10} with $b=6$
@@
 \begin{equation*}
-    \hspace*{-0.5cm}  {\scriptsize \mathring{A}_i =
+    \hspace*{-0.5cm}  {\scriptsize \mathring{A}_2 =
     \begin{bmatrix}1&0&0&0\\0&2&-1&0\\0&-1&2&-1\\0&0&-1&1\end{bmatrix} \;
@@
 is under scrutiny for the next major overhaul of PETSc. Second, HPDDM does not
-currently handle GPU efficiently, but the interface with PETSc provides a very
+handle GPU efficiently currently, but the interface with PETSc provides a very
 thorough test bed with \pk{Mat} and \pk{PC} implementations that are exploiting

### 2020-07-19 article.tex
Change one keyword, change Fig 6 to Table 1, and modify one sentence in the conclusions

@@
 \begin{keywords}
-  Krylov methods, domain decomposition preconditioners, distributed computing
+  Krylov methods, domain decomposition preconditioners, distributed-memory parallel computing
 \end{keywords}
@@
 That means that, assuming the size of the set of active columns is 40, one LOBPCG iterations implies doing at most 5 GMRES iterations with 40 vectors simultaneously, using \pk{MatProductNumeric} and \pk{PCMatApply}, instead of doing at most $40\times 5$ successive GMRES iterations, using \pk{MatMult} and \pk{PCApply}.
-Results are gathered in~\cref{fig:lobpcg}.
+Results are gathered in~\cref{tab:lobpcg}.
 In the first line, the numbers of outer LOBPCG iterations are reported with the three different inner preconditioners mentioned above.
@@
 ]\mytablenew{\mytable}
-\begin{figure}[h!]
+\begin{table}[h!]
+\caption{Performance of \pk{EPSLOBPCG} with three different inner preconditioners: \pk{PCASM} and \pk{PCHPDDM}, which utilize \pk{PCMatApply}, and \pk{PCGAMG}, which utilizes \pk{PCApply}.\label{tab:lobpcg}}
     \centering
@@
 string type]{\mytablenew}
-\caption{Performance of \pk{EPSLOBPCG} with three different inner preconditioners: \pk{PCASM} and \pk{PCHPDDM}, which utilize \pk{PCMatApply}, and \pk{PCGAMG}, which utilizes \pk{PCApply}.\label{fig:lobpcg}}
-\end{figure}
+\end{table}
 \section{PCHPDDM}\label{sec:PCHPDDM}
@@
 Krylov--Schur~\cite{zhou2008block}. The current code for Arnoldi iterations
-would then handle block Arnoldi iterations.
+would then handle block Arnoldi iterations, and shift-and-invert solves would
+operate on the whole block instead of column by column.
 

### 2020-07-19 article.tex
New small paragraph about EPSCISS et al.

@@
 \url{https://gitlab.com/petsc/datafiles/-/tree/master/matrices/hpddm/GCRODR}
-while the driver code is part of the PETSc testsuite and available at
+while the driver code is part of the PETSc test suite and available at
 \url{https://www.mcs.anl.gov/petsc/petsc-master/src/ksp/ksp/tutorials/ex75.c.html}. In order to check the correctness of the \pk{KSPHPDDM} interface,
@@
 \end{equation*}
-This continuous equation is discretized with third-order Lagrange finite elements using FreeFEM.
-The script is available at \url{https://github.com/prj-/jolivet2020petsc}.
-The following timings and convergence histories have been obtained on $2{,}304$ processes of Ir\`ene, for solving a problem of dimension \pgfmathprintnumber{74618461}.
-By default, SLEPc applies at most 5 iterations of GMRES on each column of a so-called set of active columns.
-Since the option \pk{-eps\_lobpcg\_blocksize 40} is used for computing 20 eigenpairs, the size of the set varies between 40 to 1.
-This size corresponds to the number of right-hand sides in systems that will be solved throughout LOBPCG iterations.
-Instead of using \pk{KSPGMRES}, which does not handle blocking, \pk{KSPHPDDM} is used, with the pseudo-block GMRES.
-That means that, assuming the size of the set of active columns is 40, one LOBPCG iterations implies doing at most 5 GMRES iterations with 40 vectors simultaneously, using \pk{MatProductNumeric} and \pk{PCMatApply}, instead of doing at most $40\times 5$ successive GMRES iterations, using \pk{MatMult} and \pk{PCApply}.
-Results are gathered in~\cref{tab:lobpcg}.
-In the first line, the numbers of outer LOBPCG iterations are reported with the three different inner preconditioners mentioned above.
-\pk{PCASM} is known to yield a preconditioned operator whose condition number grows as the number of subdomains, here processes, increases.
-Thus, the number of outer iterations is higher than with \pk{PCHPDDM} or \pk{PCGAMG} since the inner solves are not converging to meaningful solutions in just 5 iterations.
-The multilevel preconditioners perform similarly, since the difference in the number of outer iterations is minimal (1).
-The second and third lines represent the number of times \pk{PCApply} and \pk{PCMatApply} are called.
-Since \pk{PCGAMG} does not handle blocking, at each pseudo-block GMRES iteration, \pk{PCApply} is called on each single active column.
-\pk{PCASM} and \pk{PCHPDDM} handle blocking, so only rely on \pk{PCMatApply}.
-In the fourth line, the time spent setting up the preconditioners is given.
-\pk{PCASM} is extremely cheap but not very robust numerically, while both multilevel preconditioners have a similar setup time.
-Eventually, in the last line, the time spent in the full eigensolve is given.
-Clearly, having both a \pk{KSP} and a \pk{PC} which can deal with multiple vectors efficiently is mandatory to achieve reasonable performance with such an outer solver that heavily relies on blocking.
+This continuous equation is discretized with third-order Lagrange finite
+elements using FreeFEM. The script \texttt{blocking-slepc.edp} available at
+\url{https://github.com/prj-/jolivet2020petsc} can be use for reproducibility.
+The following timings and convergence histories have been obtained on $2{,}304$
+processes of Ir\`ene, for solving a problem of dimension
+\pgfmathprintnumber{74618461}. By default, SLEPc applies at most 5 iterations
+of GMRES on each column of a so-called set of active columns. Since the option
+\pk{-eps\_lobpcg\_blocksize 40} is used for computing 20 eigenpairs, the size
+of the set varies between 40 to 1. This size corresponds to the number of
+right-hand sides in systems that will be solved throughout LOBPCG iterations.
+Instead of using \pk{KSPGMRES}, which does not handle blocking, \pk{KSPHPDDM}
+is used, with the pseudo-block GMRES. That means that, assuming the size of the
+set of active columns is 40, one LOBPCG iterations implies doing at most 5
+GMRES iterations with 40 vectors simultaneously, using \pk{MatProductNumeric}
+and \pk{PCMatApply}, instead of doing at most $40\times 5$ successive GMRES
+iterations, using \pk{MatMult} and \pk{PCApply}. Results are gathered
+in~\cref{tab:lobpcg}. In the first line, the numbers of outer LOBPCG iterations
+are reported with the three different inner preconditioners mentioned above.
+\pk{PCASM} is known to yield a preconditioned operator whose condition number
+grows as the number of subdomains, here processes, increases. Thus, the number
+of outer iterations is higher than with \pk{PCHPDDM} or \pk{PCGAMG} since the
+inner solves are not converging to meaningful solutions in just 5 iterations.
+The multilevel preconditioners perform similarly, since the difference in the
+number of outer iterations is minimal (1). The second and third lines represent
+the number of times \pk{PCApply} and \pk{PCMatApply} are called. Since
+\pk{PCGAMG} does not handle blocking, at each pseudo-block GMRES iteration,
+\pk{PCApply} is called on each single active column. \pk{PCASM} and
+\pk{PCHPDDM} handle blocking, so only rely on \pk{PCMatApply}. In the fourth
+line, the time spent setting up the preconditioners is given. \pk{PCASM} is
+extremely cheap but not very robust numerically, while both multilevel
+preconditioners have a similar setup time.  Eventually, in the last line, the
+time spent in the full eigensolve is given.  Clearly, having both a \pk{KSP}
+and a \pk{PC} which can deal with multiple vectors efficiently is mandatory to
+achieve reasonable performance with such an outer solver that heavily relies on
+blocking.
 \pgfplotstableread{
@@
 \end{table}
+The script available at \url{https://github.com/prj-/jolivet2020petsc} may be
+used to solve the same eigenproblem using \pk{EPSCISS} by providing the
+additional command line argument \pk{-eps\_type ciss}. Blocking is also
+available and tested in SLEPc test suite for \pk{EPSSUBSPACE} and
+\pk{NEPCISS} since version 3.14.0.
 \section{PCHPDDM}\label{sec:PCHPDDM}
@@
 convergence history, which is displayed in~\cref{fig:bratu-conv}. The complete set of
-options is given in~\cref{fig:bratu-options}. Results can be reproduced using the FreeFEM script at \url{https://github.com/prj-/jolivet2020petsc}.
+options is given in~\cref{fig:bratu-options}. Results can be reproduced using
+the FreeFEM script \texttt{bratu.edp} available at
+\url{https://github.com/prj-/jolivet2020petsc}.
 \pgfplotstableread{
@@
 The authors would like to thank S.~Balay, J.~Brown, V.~Hapla, M.~Knepley,
-and B.~Smith for reviewing the successive merge requests in the PETSc
+and B.~Smith for reviewing the successive merge requests in PETSc
 repository. This work was granted access to the GENCI-sponsored HPC resources of:

### 2020-07-20 article.tex
Additional comment about CISS

@@
 used to solve the same eigenproblem using \pk{EPSCISS} by providing the
-additional command line argument \pk{-eps\_type ciss}. Blocking is also
-available and tested in SLEPc test suite for \pk{EPSSUBSPACE} and
-\pk{NEPCISS} since version 3.14.0.
+additional command line argument \pk{-eps\_type ciss}.
+However, this eigensolver is mostly suited for finding interior eigenpairs.
+For fairness, its performance is not reported, but the options from the script
+can be readily used for problems where \pk{EPSCISS} is an appropriate solver.
+Blocking is also available and tested in SLEPc test suite for \pk{EPSSUBSPACE}
+and \pk{NEPCISS} since version 3.14.0.
 \section{PCHPDDM}\label{sec:PCHPDDM}

### 2020-07-20 article.tex
Last pass

--- 2020-07-19 article.tex
@@
 \section{Introduction}
-Implementing and maintaining robust and scalable applications is highly
-relevant in the field of computational science and engineering. The efficient
-production of such applications requires a rich ecosystem of state-of-the-art
-reusable libraries \cite{Bartlett:2017:XFT}. They can then be used by domain
-specialists who can decide what are the best tools available for their
-applications. In this paper, we focus on the interoperability of two such
+Computational science and engineering today enjoys unprecedented opportunities to transform
+the way society solves many of its most urgent technological problems through predictive simulations.
+At the heart of simulations are robust and scalable solution algorithms, and the efficient
+production of applications requires a rich ecosystem of state-of-the-art
+reusable libraries from which domain
+specialists can benefit \cite{Bartlett:2017:XFT}. In this paper, we focus on the interoperability of two such
 libraries.
@@
 and facilitate the analysis and comparison of performance results~\cite{hoefler2015scientific}.
-Furthermore, the benefits of code refactorization are shown for
-applications that deal with multiple right-hand sides. This is the case of some
-eigensolvers in SLEPc~\cite{hernandez2005ssf}, the Scalable Library for
-Eigenvalue Problem computations. SLEPc is an add-on library that extends PETSc
+
+Fruits of the proposed enhancement are discussed for 
+applications from
+the Scalable Library for
+Eigenvalue Problem computations (SLEPc)~\cite{hernandez2005ssf}, 
+an add-on library that extends PETSc
 with several classes for linear and nonlinear eigenvalue problems, as well as with
@@
 The paper is divided in two main parts. \Cref{sec:KSPHPDDM} introduces
-\pk{KSPHPDDM} and discusses the readiness of PETSc with respect to block Krylov methods.
+the interface to recycling and block Krylov methods
+(\pk{KSPHPDDM}) and discusses the readiness of the PETSc library with respect to block Krylov methods.
 Different applications from eigen-analysis are also provided.
-\Cref{sec:PCHPDDM} introduces \pk{PCHPDDM} and presents diverse
-applications for solving linear and nonlinear partial differential equations.
-Eventually, concluding remarks may be found in \cref{sec:conclusion}.
+\Cref{sec:PCHPDDM} introduces the suite of robust multilevel overlapping Schwarz methods (\pk{PCHPDDM})
+with examples from linear and nonlinear partial differential equations.
+Eventually, concluding remarks are provided in \cref{sec:conclusion}.
 
@@
 represented as a linear operator. In this latter case, the preconditioner is
-always applied on the right, except for FCG. HPDDM CG
-and its variants BCG and BFBCG only handle left-preconditioning.  It is also
+always applied on the right, except for the conjugate gradient methods BCG, 
+BFCG and BFBCG, that only handle left-preconditioning.  It is also
 possible to set the preconditioning side through the more common PETSc option
-\pk{-ksp\_pc\_side (left$|$right)}.  Finally, convergence monitoring with the
+\pk{-ksp\_pc\_side (left$|$right)}.  In addition, convergence monitoring with the
 \pk{KSPMonitor} interface is fully supported, as well as the specification of
@@
 for solving systems with multiple right-hand sides, \pk{KSPMatSolve}(\pk{ksp},\pk{X},\pk{Y}),
-and to apply preconditioning methods, i.e. \pk{PCMatApply}(\pk{ksp},\pk{X},\pk{Y}).
+and to apply preconditioning methods, i.e. \pk{PCMatApply}(\pk{pc},\pk{X},\pk{Y}).
 Both input $X$ and output $Y$ matrices are currently limited to be dense tall-and-skinny matrices.
@@
 ICNTL(27)\footnote{\url{http://mumps.enseeiht.fr/doc/userguide_5.3.3.pdf},
-section 6.1}.
+section 6.1}. Similar considerations apply to the block preconditioning application phase.
 
@@
 \end{Verbatim}
-For the case of ICC(0), a \pk{PCREDUNDANT} is needed as well since
-PETSc does not provide a parallel ICC solver. The iteration numbers
+%For the case of ICC(0), a \pk{PCREDUNDANT} is needed as well since
+%PETSc does not provide a parallel ICC solver.
+The iteration numbers
 needed to reach convergence are gathered in~\cref{fig:comparison}.  The
@@
         between \ref{pgfplots:r2} and \ref{pgfplots:r3}. In the other cases,
-        the numbers of iterations needed by KSPHPDDM match with the respective
+        the numbers of iterations needed by \pk{KSPHPDDM} match with the respective
         MATLAB or PETSc reference implementation.\label{fig:comparison}}
     \end{figure}
-    \subsubsection{Linear stability analysis}
-In this section, we apply recycling Krylov methods in the context of linear
-stability analysis and consider the solution of the following generalized eigenvalue problem:
-\begin{equation}\label{eq:lsa}
-    J(q_b)x = \lambda \begin{bmatrix}M & 0 \\ 0 & 0 \end{bmatrix} x,
-\end{equation}
-where $J$ is the Jacobian of the incompressible steady-state Navier--Stokes equation
-evaluated using a given base flow $q_b = \begin{bmatrix}u_b \\ p_b\end{bmatrix}$ and
-$M$ the discretization of the mass matrix on the space of velocities. Interested
-readers are referred to~\cite{moulin2018al} for more details and for larger
-runs.
-
-The mini-app employed for the numerical results is built on top of
-FreeFEM~\cite{hecht2012new} and it is available at
-\url{https://github.com/prj-/moulin2019al}. As eigenvalue solver, it uses
-SLEPc, and thus reuses PETSc linear solver infrastructure.
-
-There are different methods to compute the eigenvalues
-of~\cref{eq:lsa} near a complex-valued shift $\sigma$. In this paragraph, emphasis will be put
-on the Krylov--Schur method~\cite{Stewart2002}, which for interior eigenvalues relies on spectral transformations
-and on the solution of successive linear systems such as:
-\begin{equation}\label{eq:ks}
-    \left(J(q_b) - \sigma \begin{bmatrix}M&0\\0&0\end{bmatrix}\right)x_i = b_i,
-\end{equation}
-which, in SLEPc, are parameterized using the \pk{-st\_} prefix.
-In this work, the above linear system is preconditioned using a modified augmented Lagrangian
-approach~\cite{Benzi2011a,moulin2018al}
-
-The effectiveness of subspace recycling is shown for
-finding the 5 eigenpairs closest to the shift $\sigma = 10^{-6} + 0.6\mathfrak{i}$ for a
-flow past a cylinder at Reynolds 100. Results can be reproduced using the following four commands.
-We note here that the first two commands are merely used to generate
-the base flow $q_b$ with a \pk{SNES}, a PETSc object used to solve
-nonlinear problems, using a continuation method on the Reynolds number. \\[-4pt]
-\begin{Verbatim}[fontsize=\footnotesize,frame=single,framerule=0.1mm,commandchars=&\[\]]
-&fvtextcolor[mygreen][$] &fvtextcolor[myyellow][mpirun] -n 4 FreeFem++-mpi Nonlinear-solver.edp -Re 50 -v 0
-&fvtextcolor[mygreen][$] &fvtextcolor[myyellow][mpirun] -n 4 FreeFem++-mpi Nonlinear-solver.edp -Re 100 -v 0
-&fvtextcolor[mygreen][$] &fvtextcolor[myyellow][mpirun] -n 4 FreeFem++-mpi Eigensolver.edp -Re 100 -v 0 -st_ksp_rtol 1.0e-4         &fvtextcolor[myred][\]
-  -st_ksp_type fgmres -st_ksp_gmres_restart 200 -st_ksp_converged_reason
-&fvtextcolor[mygreen][$] &fvtextcolor[myyellow][mpirun] -n 4 FreeFem++-mpi Eigensolver.edp -Re 100 -v 0 -st_ksp_rtol 1.0e-4         &fvtextcolor[myred][\]
-  -st_ksp_type hpddm -st_ksp_gmres_restart 200 -st_ksp_converged_reason              &fvtextcolor[myred][\]
-  -st_ksp_hpddm_variant flexible -st_ksp_hpddm_recycle 10 -st_ksp_hpddm_type gcrodr
-\end{Verbatim}
-
-The Krylov--Schur algorithm converges in six outer iterations (restarts), with a total of 46 inner solves~\cref{eq:ks}.
-Recycling of Krylov subspaces with \pk{KSPHPDDM} provides algorithmic speedup by lowering the number of iterations per inner solve,
-as reported in~\cref{fig:comparison-slepc}, from 80 with \pk{KSPFGMRES} to 45.
-A similar speedup can be observed in terms of runtime, with the time spent for the solution of the eigenvalue problem being reduced from \SI[round-mode = places,round-precision=1,group-minimum-digits=3,group-separator = {,}]{1509}{\sec} to \SI[round-mode = places,round-precision=1,group-minimum-digits=3,group-separator = {,}]{1052}{\sec}.
-These timings, and all which follow except in~\cref{sec:primitives}, have been obtained on {Ir\`ene}, a system composed of \pgfmathprintnumber[assume math mode=true]{1656} nodes with two 24-core Intel Xeon Platinum 8168 clocked at \SI{2.7}{\giga\hertz}.
-\pgfplotstableread{
-sys hpddm fgmres
-1   77    77
-2   54    83
-3   53    83
-4   52    81
-5   52    81
-6   51    78
-7   51    79
-8   51    80
-9   51    80
-10  51    77
-11  51    79
-12  51    79
-13  51    79
-14  53    77
-15  52    79
-16  49    78
-17  51    77
-18  51    79
-19  52    79
-20  52    79
-21  51    77
-22  51    78
-23  51    79
-24  53    80
-25  54    79
-26  52    79
-27  49    75
-28  51    78
-29  51    78
-30  51    77
-31  50    78
-32  55    81
-33  56    79
-34  54    80
-35  53    80
-36  53    78
-37  50    78
-38  52    79
-39  53    77
-40  52    77
-41  54    80
-42  53    80
-43  55    80
-44  53    79
-45  52    77
-46  53    80
-}\loadedtableSLEPc
-    \begin{figure}[!t]
-\centering \begin{tikzpicture}
-    \begin{axis}[xlabel={Krylov--Schur system index},ylabel={\# of
-        iterations},xmin=1,xmax=46,ymin=0,ymajorgrids,
-        legend columns=1,enlarge x limits=0.05, height=5.5cm,
-        legend style={font=\small,at={(0.98,0.02)}, legend cell align=left,
-        anchor=south east}]
-\addplot[very thick,last]     table[x expr=\thisrowno{0},y expr=\thisrowno{2}, col sep=space] {\loadedtableSLEPc};\label{pgfplots:m0}
-\addplot[very thick,myred,densely dashed]       table[x expr=\thisrowno{0},y expr=\thisrowno{1}, col sep=space] {\loadedtableSLEPc};\label{pgfplots:m1}
-\legend{{\pk{KSPFGMRES}},{\pk{KSPHPDDM}}}
-\end{axis}
-\end{tikzpicture}
-        \caption{\# of inner iterations for systems~\cref{eq:ks} with~\ref{pgfplots:m0} or without~\ref{pgfplots:m1} recycling.\label{fig:comparison-slepc}}
-    \end{figure}
+
 \pgfplotstableread{
@@
 For this reason, from now on, we drop the \pk{SEQ} substring.
-The performance of the multiplication primitives of the three aforementioned types will also be compared to
+The performance of the multiplication primitives of the three aforementioned types will also be compared against
 those obtained with the MKL inspector--executor sparse BLAS routines~\cite{mkl-web-page}, using a single OpenMP thread.
@@
            \addlegendentry{\pk{MATBAIJMKL}}
-           \addlegendentry{MATSBAIJMKL}
+           \addlegendentry{\pk{MATSBAIJMKL}}
 \node[rotate=0,fill=white,inner sep=0pt,anchor=west] at (rel axis cs:1.02, 0.5) {$b=1$};
@@
     \end{tikzpicture}
-   \caption{Performance of the matrix--matrix multiplication as implemented in PETSc for three sparse matrix types.\label{fig:perfAX}}
+   \caption{Performance of the matrix--matrix multiplication different sparse matrix formats.\label{fig:perfAX}}
 \end{figure}
@@
 been taken into account. 
+For the sake of completeness, the performance of
+the matrix--vector multiplication with \pk{MATAIJMKL} and different numbers of threads is also reported in
+\cref{fig:scaling-Ax}. Clearly, it is not advised to use the full socket to
+perform this type of workload for such small sparse matrices and skinny dense
+matrices. However, these results help in drawing a fair comparison between a full CPU socket
+and a GPU device.
 
@@
            \addlegendentry{\pk{MATBAIJMKL}}
-           \addlegendentry{MATSBAIJMKL}
+           \addlegendentry{\pk{MATSBAIJMKL}}
            \addlegendentry{\pk{MATAIJCUSPARSE}}
@@
 
-For the sake of completeness, the performance of
-\pk{MATAIJMKL} with $N=1$ but different numbers of threads is also reported in
-\cref{fig:scaling-Ax}. Clearly, it is not advised to use the full socket to
-perform this type of workload for such small sparse matrices and skinny dense
-matrices. But this helps in drawing a fair comparison between a full CPU socket
-and a full GPU device. All in all, the CPU configuration which reaches the
-highest percentage of peak is MATSBAIJMKL~\ref{pgfplots:sbaijmkl3} with $b=3$
+The CPU configuration which reaches the
+highest percentage of peak is \pk{MATSBAIJMKL}~\ref{pgfplots:sbaijmkl3} with $b=3$
 and $N=64$, rightmost middle plot in~\cref{fig:perfAXpar}, with approximately
@@
 and $N=64$, rightmost bottom plot in~\cref{fig:perfAXpar}, performs at around
-306 GFLOP/s, about 4\% of peak for an NVIDIA Tesla V100 SXM2.
+306 GFLOP/s, about 4\% of peak for an NVIDIA Tesla V100 SXM2. These rather disappointing performances
+suggest to consider different, interlaced layouts for storing the dense right-hand sides and we plan to extend the
+current PETSc functionality in the future to maximize performances while maintaining interface flexibility
+and the user friendliness of the library.
+{\color{red} Pierre, can you do a quick run hacking MatProduct.cpp and using MAIJ MatMult instead of MatProductNumeric just to confirm this
+sentence and add preliminary numbers in terms of CPU peak performance? (forget about GPU, we need a new interface to CUSPARSE for that)}
+
 \pgfplotstableread{
@@
 
+    \subsubsection{Linear stability analysis}
+In this section, we apply recycling Krylov methods in the context of linear
+stability analysis and consider the solution of the following generalized eigenvalue problem:
+\begin{equation}\label{eq:lsa}
+    J(q_b)x = \lambda \begin{bmatrix}M & 0 \\ 0 & 0 \end{bmatrix} x,
+\end{equation}
+where $J$ is the Jacobian of the incompressible steady-state Navier--Stokes equation
+evaluated using a given base flow $q_b = \begin{bmatrix}u_b \\ p_b\end{bmatrix}$ and
+$M$ the discretization of the mass matrix on the space of velocities.
+The mini-app employed for the numerical results is built on top of
+FreeFEM~\cite{hecht2012new} and it is available at
+\url{https://github.com/prj-/moulin2019al}.
+Interested
+readers are referred to~\cite{moulin2018al} for more details and for larger
+runs.
+
+%As eigenvalue solver, it uses
+%SLEPc, and thus reuses PETSc linear solver infrastructure.
+
+There are different methods to compute the eigenvalues
+of~\cref{eq:lsa} near a complex-valued shift $\sigma$. In this paragraph we consider 
+a Krylov--Schur method~\cite{Stewart2002}, which, for interior eigenvalues, relies on spectral transformations
+and on the solution of successive linear systems such as:
+\begin{equation}\label{eq:ks}
+    \left(J(q_b) - \sigma \begin{bmatrix}M&0\\0&0\end{bmatrix}\right)x_i = b_i,
+\end{equation}
+which, in SLEPc, are parameterized using the \pk{-st\_} prefix.
+In this work, the above linear system is preconditioned using a modified augmented Lagrangian
+approach~\cite{Benzi2011a,moulin2018al}
+
+The effectiveness of subspace recycling is shown for
+finding the 5 eigenpairs closest to the shift $\sigma = 10^{-6} + 0.6\mathfrak{i}$ for a
+flow past a cylinder at Reynolds 100. Results can be reproduced using the following four commands.
+We note here that the first two commands are merely used to generate
+the base flow $q_b$ with a \pk{SNES}, a PETSc object used to solve
+nonlinear problems, using a continuation method on the Reynolds number. \\[-4pt]
+\begin{Verbatim}[fontsize=\footnotesize,frame=single,framerule=0.1mm,commandchars=&\[\]]
+&fvtextcolor[mygreen][$] &fvtextcolor[myyellow][mpirun] -n 4 FreeFem++-mpi Nonlinear-solver.edp -Re 50 -v 0
+&fvtextcolor[mygreen][$] &fvtextcolor[myyellow][mpirun] -n 4 FreeFem++-mpi Nonlinear-solver.edp -Re 100 -v 0
+&fvtextcolor[mygreen][$] &fvtextcolor[myyellow][mpirun] -n 4 FreeFem++-mpi Eigensolver.edp -Re 100 -v 0 -st_ksp_rtol 1.0e-4         &fvtextcolor[myred][\]
+  -st_ksp_type fgmres -st_ksp_gmres_restart 200 -st_ksp_converged_reason
+&fvtextcolor[mygreen][$] &fvtextcolor[myyellow][mpirun] -n 4 FreeFem++-mpi Eigensolver.edp -Re 100 -v 0 -st_ksp_rtol 1.0e-4         &fvtextcolor[myred][\]
+  -st_ksp_type hpddm -st_ksp_gmres_restart 200 -st_ksp_converged_reason              &fvtextcolor[myred][\]
+  -st_ksp_hpddm_variant flexible -st_ksp_hpddm_recycle 10 -st_ksp_hpddm_type gcrodr
+\end{Verbatim}
+
+The Krylov--Schur algorithm converges in six outer iterations (restarts), with a total of 46 inner solves~\cref{eq:ks}.
+Recycling of Krylov subspaces with \pk{KSPHPDDM} provides algorithmic speedup by lowering the number of iterations per inner solve,
+as reported in~\cref{fig:comparison-slepc}, from 80 with \pk{KSPFGMRES} to 45.
+A similar speedup can be observed in terms of runtime, with the time spent for the solution of the eigenvalue problem being reduced from \SI[round-mode = places,round-precision=1,group-minimum-digits=3,group-separator = {,}]{1509}{\sec} to \SI[round-mode = places,round-precision=1,group-minimum-digits=3,group-separator = {,}]{1052}{\sec}.
+These timings, and all that follow, have been obtained on {Ir\`ene}, a system composed of \pgfmathprintnumber[assume math mode=true]{1656} nodes with two 24-core Intel Xeon Platinum 8168 clocked at \SI{2.7}{\giga\hertz}.
+\pgfplotstableread{
+sys hpddm fgmres
+1   77    77
+2   54    83
+3   53    83
+4   52    81
+5   52    81
+6   51    78
+7   51    79
+8   51    80
+9   51    80
+10  51    77
+11  51    79
+12  51    79
+13  51    79
+14  53    77
+15  52    79
+16  49    78
+17  51    77
+18  51    79
+19  52    79
+20  52    79
+21  51    77
+22  51    78
+23  51    79
+24  53    80
+25  54    79
+26  52    79
+27  49    75
+28  51    78
+29  51    78
+30  51    77
+31  50    78
+32  55    81
+33  56    79
+34  54    80
+35  53    80
+36  53    78
+37  50    78
+38  52    79
+39  53    77
+40  52    77
+41  54    80
+42  53    80
+43  55    80
+44  53    79
+45  52    77
+46  53    80
+}\loadedtableSLEPc
+    \begin{figure}[!t]
+\centering \begin{tikzpicture}
+    \begin{axis}[xlabel={Krylov--Schur system index},ylabel={\# of
+        iterations},xmin=1,xmax=46,ymin=0,ymajorgrids,
+        legend columns=1,enlarge x limits=0.05, height=5.5cm,
+        legend style={font=\small,at={(0.98,0.02)}, legend cell align=left,
+        anchor=south east}]
+\addplot[very thick,last]     table[x expr=\thisrowno{0},y expr=\thisrowno{2}, col sep=space] {\loadedtableSLEPc};\label{pgfplots:m0}
+\addplot[very thick,myred,densely dashed]       table[x expr=\thisrowno{0},y expr=\thisrowno{1}, col sep=space] {\loadedtableSLEPc};\label{pgfplots:m1}
+\legend{{\pk{KSPFGMRES}},{\pk{KSPHPDDM}}}
+\end{axis}
+\end{tikzpicture}
+        \caption{\# of inner iterations for systems~\cref{eq:ks} with~\ref{pgfplots:m0} or without~\ref{pgfplots:m1} recycling.\label{fig:comparison-slepc}}
+    \end{figure}
+
 \subsubsection{Blocking inside SLEPc\label{sec:lobpcg}}
@@
 \end{itemize}
-More details are given in the next section where focus is put on preconditioning rather than Krylov methods. The two main points are that:
-\begin{itemize}
-    \item \pk{PCASM} and \pk{PCHPDDM} handle blocking, see the list~\cref{sec:interface-ksp};
-    \item \pk{PCHPDDM} and \pk{PCGAMG} have low contraction factors.
-\end{itemize}
-The following generalized eigenvalue problem is solved in a three-dimensional cube:
+More details on these solvers are given in the next section where focus is put on preconditioning rather than Krylov methods.
+We note here that only \pk{PCASM} and \pk{PCHPDDM} currently handle blocking (see the list~\cref{sec:interface-ksp}),
+and that \pk{PCHPDDM} and \pk{PCGAMG} have lower contraction factors than the cheaper alternative \pk{PCASM}.
+
+As a testbed, we consider the following generalized eigenvalue problem on a three-dimensional cube:
 \begin{equation*}
@@
 \end{equation*}
-This continuous equation is discretized with third-order Lagrange finite elements using FreeFEM.
-The script is available at \url{https://github.com/prj-/jolivet2020petsc}.
-The following timings and convergence histories have been obtained on $2{,}304$ processes of Ir\`ene, for solving a problem of dimension \pgfmathprintnumber{74618461}.
-By default, SLEPc applies at most 5 iterations of GMRES on each column of a so-called set of active columns.
-Since the option \pk{-eps\_lobpcg\_blocksize 40} is used for computing 20 eigenpairs, the size of the set varies between 40 to 1.
-This size corresponds to the number of right-hand sides in systems that will be solved throughout LOBPCG iterations.
-Instead of using \pk{KSPGMRES}, which does not handle blocking, \pk{KSPHPDDM} is used, with the pseudo-block GMRES.
-That means that, assuming the size of the set of active columns is 40, one LOBPCG iterations implies doing at most 5 GMRES iterations with 40 vectors simultaneously, using \pk{MatProductNumeric} and \pk{PCMatApply}, instead of doing at most $40\times 5$ successive GMRES iterations, using \pk{MatMult} and \pk{PCApply}.
-Results are gathered in~\cref{tab:lobpcg}.
-In the first line, the numbers of outer LOBPCG iterations are reported with the three different inner preconditioners mentioned above.
-\pk{PCASM} is known to yield a preconditioned operator whose condition number grows as the number of subdomains, here processes, increases.
-Thus, the number of outer iterations is higher than with \pk{PCHPDDM} or \pk{PCGAMG} since the inner solves are not converging to meaningful solutions in just 5 iterations.
-The multilevel preconditioners perform similarly, since the difference in the number of outer iterations is minimal (1).
-The second and third lines represent the number of times \pk{PCApply} and \pk{PCMatApply} are called.
-Since \pk{PCGAMG} does not handle blocking, at each pseudo-block GMRES iteration, \pk{PCApply} is called on each single active column.
-\pk{PCASM} and \pk{PCHPDDM} handle blocking, so only rely on \pk{PCMatApply}.
-In the fourth line, the time spent setting up the preconditioners is given.
-\pk{PCASM} is extremely cheap but not very robust numerically, while both multilevel preconditioners have a similar setup time.
-Eventually, in the last line, the time spent in the full eigensolve is given.
-Clearly, having both a \pk{KSP} and a \pk{PC} which can deal with multiple vectors efficiently is mandatory to achieve reasonable performance with such an outer solver that heavily relies on blocking.
+This continuous equation is discretized with third-order Lagrange finite
+elements using FreeFEM. The script \texttt{blocking-slepc.edp} available at
+\url{https://github.com/prj-/jolivet2020petsc} can be use for reproducibility.
+The following timings and convergence histories have been obtained on $2{,}304$
+processes of Ir\`ene, for solving a problem of dimension
+\pgfmathprintnumber{74618461}.
+
+By default, SLEPc applies at most 5 iterations
+of GMRES on each column of a so-called set of active columns. Since the option
+\pk{-eps\_lobpcg\_blocksize 40} is used for computing 20 eigenpairs, the size
+of the set varies between 40 to 1. This size corresponds to the number of
+right-hand sides that will be solved for throughout LOBPCG iterations.
+Instead of using \pk{KSPGMRES}, which does not handle blocking, \pk{KSPHPDDM}
+is used, with the pseudo-block GMRES. Thus, assuming the size of the
+set of active columns is 40, one LOBPCG iteration performs at most 5
+pseudo-block GMRES iterations with 40 vectors simultaneously, using \pk{MatProductNumeric}
+and \pk{PCMatApply}, instead of doing (at most) $40\times 5$ successive GMRES
+iterations, using \pk{MatMult} and \pk{PCApply}.
+
+Results are gathered
+in~\cref{tab:lobpcg}. In the first row, the numbers of outer LOBPCG iterations
+are reported with the three different inner preconditioners mentioned above.
+\pk{PCASM} is known to yield a preconditioned operator whose condition number
+grows as the number of subdomains, here processes, increases. Thus, the number
+of outer iterations is higher than with \pk{PCHPDDM} or \pk{PCGAMG} since the
+inner solves are not converging to meaningful solutions in just 5 iterations.
+On the other hand, both multilevel preconditioners perform similarly, with a minimal difference in the
+number of outer iterations. The second and third rows record
+the number of times \pk{PCApply} and \pk{PCMatApply} are called. Since
+\pk{PCGAMG} does not handle blocking, at each pseudo-block GMRES iteration,
+\pk{PCApply} is called for each active column separately. \pk{PCASM} and
+\pk{PCHPDDM} handle blocking, so the only rely on \pk{PCMatApply}. In the fourth
+row, the time spent setting up the preconditioners is given. \pk{PCASM} is
+extremely cheap but not very robust numerically, while both multilevel
+preconditioners have a similar setup time. The
+time spent in the full eigenvalue solver is reported in the last row. Clearly, having both a \pk{KSP}
+and a \pk{PC} which can efficiently deal with multiple vectors is mandatory to
+achieve reasonable performance with such an outer solver that heavily relies on
+blocking.
 \pgfplotstableread{
@@
 \end{table}
+
+The script available at \url{https://github.com/prj-/jolivet2020petsc} may be
+used to solve the same eigenproblem using \pk{EPSCISS} by providing the
+additional command line argument \pk{-eps\_type ciss}.
+However, this eigensolver is mostly suited for finding interior eigenpairs.
+For fairness, its performance is not reported, but the options from the script
+can be readily used for problems where \pk{EPSCISS} is an appropriate solver.
+We finally note that blocking support is also available and tested in SLEPc for \pk{EPSSUBSPACE}
+and \pk{NEPCISS} since version 3.14.0.
+
 \section{PCHPDDM}\label{sec:PCHPDDM}
@@
 from~\cref{eq:bratu} in few iterates.  There are two strategies to define the
-\pk{PC} here:
+preconditioner:
 \begin{itemize}
 \item whenever the Jacobian is being updated in the \pk{SNESSetJacobian}
-function, new unassembled Neumann matrices at the current linearization point are updated as well and supplied to the previously
-explained \pk{PCHPDDMSetAuxiliaryMat} function, this is done automatically if the nonlinear solver is attached to a
-\pk{DMPLEX} or a \pk{DMP4EST};
+function, new unassembled Neumann matrices at the current linearization point can be updated as well and supplied via
+\pk{PCHPDDMSetAuxiliaryMat}
 \item by reusing the \pk{PCHPDDM} hierarchy assembled when solving the first
@@
 Concerning PETSc, the infrastructure for dealing with blocks of vectors has
-been laid out. Still, only a little fraction of the built-in matrix types and
-preconditioners may efficiently handle such a format.
+been laid out. However, only a little fraction of the built-in matrix types and
+preconditioners can currently handle them efficiently, and additional
+storage formats may be considered to maximize performances.
 
@@
 Concerning HPDDM, its integration inside PETSc offers a much-needed flexibility
-compared to the standard implementation. Indeed, it is now possible to switch at
-runtime many parameters that are determined at compile-time in the standard implementation, e.g.,
-subdomain solvers and matrix types.
-
-The interface between PETSc and HPDDM has still room for future improvements,
-and we discuss openly two of them next. First, HPDDM handles mixed-precision,
+compared to the standard implementation. Indeed, it is now possible to decide at
+runtime many solvers parameters that are instead determined at compile-time in a
+stand-alone HPDDM implementation.
+
+The interface between PETSc and HPDDM has still room for future improvements.
+First, HPDDM handles mixed-precision,
 e.g., using single-precision scalars for assembling coarse operators while
@@
 handle GPU efficiently currently, but the interface with PETSc provides a very
-thorough test bed with \pk{Mat} and \pk{PC} implementations that are exploiting
+thorough testbed with \pk{Mat} and \pk{PC} implementations that are exploiting
 GPU.

### 2020-07-20 article.tex
Again, MATSBAIJMKL is not a PETSc keyword...

@@
 The CPU configuration which reaches the
-highest percentage of peak is \pk{MATSBAIJMKL}~\ref{pgfplots:sbaijmkl3} with $b=3$
+highest percentage of peak is MATSBAIJMKL~\ref{pgfplots:sbaijmkl3} with $b=3$
 and $N=64$, rightmost middle plot in~\cref{fig:perfAXpar}, with approximately

### 2020-07-20 article.tex
Two amendments

@@
     \end{tikzpicture}
-   \caption{Performance of the matrix--matrix multiplication different sparse matrix formats.\label{fig:perfAX}}
+   \caption{Performance of the matrix--matrix multiplication for different sparse matrix formats.\label{fig:perfAX}}
 \end{figure}
@@
 \pk{PCApply} is called for each active column separately. \pk{PCASM} and
-\pk{PCHPDDM} handle blocking, so the only rely on \pk{PCMatApply}. In the fourth
+\pk{PCHPDDM} handle blocking, so they only rely on \pk{PCMatApply}. In the fourth
 row, the time spent setting up the preconditioners is given. \pk{PCASM} is

### 2020-07-20 article.tex
MatProduct.c: handle MAIJ

@@
 always applied on the right, except for the conjugate gradient methods BCG, 
-BFCG and BFBCG, that only handle left-preconditioning.  It is also
+BFCG and BFBCG, that only handle left preconditioning.  It is also
 possible to set the preconditioning side through the more common PETSc option
@@
 \pk{KSPMatSolve} is then called repeatedly until all sub-blocks are traversed.
+Similar considerations apply to the underlying calls to \pk{PCMatApply} and \pk{MatProductNumeric}.
 This was inspired by MUMPS~\cite{amestoy2001fully} option
 ICNTL(27)\footnote{\url{http://mumps.enseeiht.fr/doc/userguide_5.3.3.pdf},
-section 6.1}. Similar considerations apply to the block preconditioning application phase.
+section 6.1}.
 
@@
 are reported with the three different inner preconditioners mentioned above.
-\pk{PCASM} is known to yield a preconditioned operator whose condition number
+On the one hand, \pk{PCASM} is known to yield a preconditioned operator whose condition number
 grows as the number of subdomains, here processes, increases. Thus, the number

### 2020-07-20 MatProduct.c

@@ -22,16 +22,16 @@
 
 #if defined(PETSC_HAVE_MKL)
 #include <mkl.h>
-#define PetscStackCallMKLSparse(func,args) do {              \
-    sparse_status_t __ierr;                                  \
-    PetscStackPush(#func);                                   \
-    __ierr = func args;                                      \
-    PetscStackPop;                                           \
-    if (__ierr != SPARSE_STATUS_SUCCESS) SETERRQ2(PETSC_COMM_SELF,PETSC_ERR_LIB,"Error in %s(): error code %d",#func,(int)__ierr); \
+#define PetscStackCallMKLSparse(func, args) do {               \
+    sparse_status_t __ierr;                                    \
+    PetscStackPush(#func);                                     \
+    __ierr = func args;                                        \
+    PetscStackPop;                                             \
+    if (__ierr != SPARSE_STATUS_SUCCESS) SETERRQ2(PETSC_COMM_SELF, PETSC_ERR_LIB, "Error in %s(): error code %d",#func,(int)__ierr); \
   } while (0)
 #else
-#define PetscStackCallMKLSparse(func,args) do {              \
-    SETERRQ(PETSC_COMM_SELF,PETSC_ERR_SUP,"No MKL support"); \
+#define PetscStackCallMKLSparse(func, args) do {               \
+    SETERRQ(PETSC_COMM_SELF, PETSC_ERR_SUP, "No MKL support"); \
   } while (0)
 #endif
 
@@ -42,7 +42,7 @@ int main(int argc, char** argv) {
   PetscInt       bs[10], N[8];
   char           *type[10];
   PetscMPIInt    size;
-  PetscBool      flg, iscuda, check = PETSC_FALSE, trans = PETSC_FALSE, mkl;
+  PetscBool      flg, cuda, maij, check = PETSC_FALSE, trans = PETSC_FALSE, mkl;
   char           file[PETSC_MAX_PATH_LEN];
   PetscErrorCode ierr;
 
[…]

### 2020-07-20 article.tex
tone down row-major storage comment

@@
 and $N=64$, rightmost bottom plot in~\cref{fig:perfAXpar}, performs at around
-306 GFLOP/s, about 4\% of peak for an NVIDIA Tesla V100 SXM2. These rather disappointing performances
-suggest to consider different, interlaced layouts for storing the dense right-hand sides and we plan to extend the
-current PETSc functionality in the future to maximize performances while maintaining interface flexibility
+306 GFLOP/s, about 4\% of peak for an NVIDIA Tesla V100 SXM2. Future work may
+consider using interlaced layouts for storing the dense right-hand sides and extend the
+current PETSc functionality in order to maximize performances while maintaining interface flexibility
 and the user friendliness of the library.
-{\color{red} Pierre, can you do a quick run hacking MatProduct.cpp and using MAIJ MatMult instead of MatProductNumeric just to confirm this
-sentence and add preliminary numbers in terms of CPU peak performance? (forget about GPU, we need a new interface to CUSPARSE for that)}
 

### 2020-07-20 article.tex
Again, not a PETSc keyword, but I prefer consistency in characters

@@
 The CPU configuration which reaches the
-highest percentage of peak is MATSBAIJMKL~\ref{pgfplots:sbaijmkl3} with $b=3$
+highest percentage of peak is \pk{MATSBAIJMKL}~\ref{pgfplots:sbaijmkl3} with $b=3$
 and $N=64$, rightmost middle plot in~\cref{fig:perfAXpar}, with approximately
-63 GFLOP/s. This format is not available yet in PETSc, but it is handled by the
+63 GFLOP/s. This format is not yet available in PETSc, but it is handled by the
 mini-app. Still, it is less than 1\% of peak for an Intel Xeon Gold 6248.

### 2020-07-20 MatProduct.c
Minor useless optimization which changes nothing

@@ -1,6 +1,6 @@
 /*
    Author(s): Pierre Jolivet <pierre.jolivet@enseeiht.fr>
-              Jose Roman <jroman@dsic.upv.es>
+              Jose E. Roman <jroman@dsic.upv.es>
               Stefano Zampini <stefano.zampini@kaust.edu.sa>
         Date: 2020-07-10
 
@@ -259,22 +259,19 @@ int main(int argc, char** argv) {
               const PetscScalar *c_ptr;
               PetscScalar       *d_ptr;
               ierr = MatCreateMAIJ(A, N[k], &E);CHKERRQ(ierr);
-              ierr = MatCreateVecs(E, &cC, &cD);CHKERRQ(ierr);
               ierr = MatTranspose(C, MAT_INPLACE_MATRIX, &C);CHKERRQ(ierr);
               ierr = MatTranspose(D, MAT_INPLACE_MATRIX, &D);CHKERRQ(ierr);
               ierr = MatDenseGetArrayRead(C, &c_ptr);CHKERRQ(ierr);
               ierr = MatDenseGetArrayWrite(D, &d_ptr);CHKERRQ(ierr);
-              ierr = VecPlaceArray(cC, c_ptr);CHKERRQ(ierr);
-              ierr = VecPlaceArray(cD, d_ptr);CHKERRQ(ierr);
+              ierr = VecCreateMPIWithArray(PETSC_COMM_WORLD, 1, AM * N[k], PETSC_DECIDE, c_ptr, &cC);CHKERRQ(ierr);
+              ierr = VecCreateMPIWithArray(PETSC_COMM_WORLD, 1, AM * N[k], PETSC_DECIDE, d_ptr, &cD);CHKERRQ(ierr);
               ierr = MatMult(E, cC, cD);CHKERRQ(ierr);
-              ierr = PetscPrintf(PETSC_COMM_WORLD, "Benchmarking MatMult: with A %s %Dx%D\n", MATMAIJ, AM, AN);
+              ierr = PetscPrintf(PETSC_COMM_WORLD, "Benchmarking MatMult: with A %s %Dx%D and B %s %Dx%D\n", MATMAIJ, AM, AN, VECMPI, AM * N[k], 1);
               ierr = PetscLogStagePush(stage);CHKERRQ(ierr);
               for (mm = 0; mm < trial; ++mm) {
                 ierr = MatMult(E, cC, cD);CHKERRQ(ierr);
               }
[…]

### 2020-07-21 article.tex
Proper reference

@@
 reusable libraries from which domain
-specialists can benefit \cite{Bartlett:2017:XFT}. In this paper, we focus on the interoperability of two such
+specialists can benefit~\cite{Bartlett:2017:XFT}. In this paper, we focus on the interoperability of two such
 libraries.
 
-On the one hand, PETSc \cite{balay1997cient}, the
+On the one hand, PETSc~\cite{petsc-web-page,petsc-user-ref}, the
 Portable and Extensible Toolkit for Scientific computation, is a
@@
 
-\bibitem{balay1997cient}
-{\sc S.~Balay, W.~D. Gropp, L.~Curfman~McInnes, and B.~F. Smith}, {\em
-  Efficient management of parallelism in object-oriented numerical software
-  libraries}, Modern Software Tools in Scientific Computing,  (1997),
-  pp.~163--202.
+\bibitem{petsc-user-ref}
+{\sc S.~Balay, S.~Abhyankar, M.~F. Adams, J.~Brown, P.~Brune, K.~Buschelman,
+  L.~Dalcin, A.~Dener, V.~Eijkhout, W.~D. Gropp, D.~Karpeyev, D.~Kaushik, M.~G.
+  Knepley, D.~A. May, L.~C. McInnes, R.~T. Mills, T.~Munson, K.~Rupp, P.~Sanan,
+  B.~F. Smith, S.~Zampini, H.~Zhang, and H.~Zhang}, {\em {PETS}c users manual},
+  Tech. Report ANL-95/11 - Revision 3.13, Argonne National Laboratory, 2020.
+
+\bibitem{petsc-web-page}
+{\sc S.~Balay, S.~Abhyankar, M.~F. Adams, J.~Brown, P.~Brune, K.~Buschelman,
+  L.~Dalcin, A.~Dener, V.~Eijkhout, W.~D. Gropp, D.~Karpeyev, D.~Kaushik, M.~G.
+  Knepley, D.~A. May, L.~C. McInnes, R.~T. Mills, T.~Munson, K.~Rupp, P.~Sanan,
+  B.~F. Smith, S.~Zampini, H.~Zhang, and H.~Zhang}, {\em {PETS}c web page},
+  2020, \url{http://www.mcs.anl.gov/petsc}.
 
@@
 {\sc NVIDIA}, {\em {cuSPARSE} web page}.
-\newblock \url{https://docs.nvidia.com/cuda/cusparse/}, 2020.
+\newblock \url{https://docs.nvidia.com/cuda/cusparse}, 2020.
 
@@
 \bibitem{saad2003iterative}
-{\sc Y.~Saad}, {\em {Iterative Methods for Sparse Linear Systems, Second
-  Edition}}, SIAM, 2003.
+{\sc Y.~Saad}, {\em {Iterative Methods for Sparse Linear Systems}}, SIAM,
+  second~ed., 2003.
 

### 2020-07-21 article.tex
Switch to Elsevier template

@@
-% SIAM Article Template
-\documentclass[review,onefignum,onetabnum]{siamart190516}
-
-% Information that is shared between the article and the supplement
-% (title and author information, macros, packages, etc.) goes into
-% ex_shared.tex. If there is no supplement, this file can be included
-% directly.
-
-\input{shared}
-
-% Optional PDF information
-\ifpdf
-\hypersetup{
-  pdftitle={KSPHPDDM and PCHPDDM: Extending PETSc with Robust Overlapping Schwarz Preconditioners and Advanced Krylov Methods},
-  pdfauthor={P. Jolivet, J. E. Roman, and S. Zampini}
-}
-\fi
-
-% The next statement enables references to information in the
-% supplement. See the xr-hyperref package for details.
-
-% \externaldocument{supplement}
-
-% FundRef data to be entered by SIAM
-%<funding-group specific-use="FundRef">
-%<award-group>
-%<funding-source>
-%<named-content content-type="funder-name">
-%</named-content>
-%<named-content content-type="funder-identifier">
-%</named-content>
-%</funding-source>
-%<award-id> </award-id>
-%</award-group>
-%</funding-group>
+\documentclass[3p,11pt]{elsarticle}
+
+\usepackage{lineno,hyperref}
+% \modulolinenumbers[5]
+
+\journal{Computers \& Mathematics with Applications}
 \newcommand\col{:}
@@
 \makeatother
-\usepackage{amsmath,etoolbox}
+\usepackage{amsmath,etoolbox,amssymb}
+\usepackage{cleveref}
 \AtBeginEnvironment{pmatrix}{\setlength{\arraycolsep}{2pt}}
@@
 
-\maketitle
+\begin{frontmatter}
+
+\title{KSPHPDDM and PCHPDDM: extending PETS\MakeLowercase{c} with advanced Krylov methods and robust multilevel Overlapping Schwarz preconditioners}
+
+\author[firstaddress]{Pierre Jolivet\corref{correspondingauthor}}
+\cortext[correspondingauthor]{Corresponding author}
+\ead{pierre.jolivet@enseeiht.fr}
+
+\author[secondaryaddress]{Jose E. Roman}
+\ead{jroman@dsic.upv.es}
+
+\author[thirdaddress]{Stefano Zampini}
+\ead{stefano.zampini@kaust.edu.sa}
+
+\address[firstaddress]{CNRS, IRIT--ENSEEIHT, Toulouse, France}
+\address[secondaryaddress]{Universitat Polit\`ecnica de Val\`encia, Val\`encia, Spain}
+\address[thirdaddress]{King Abdullah University of Science and Technology, Thuwal, Saudi Arabia}
 
@@
     shapes, and structures. The goal of this paper is to explain how two
-    libraries, PETSc and HPDDM, have been interfaced together in order to offer
+    libraries, PETSc and HPDDM, have been interfaced in order to offer
     end-users robust overlapping Schwarz preconditioners and advanced Krylov
-    methods featuring recycling and block solution techniques.
+    methods featuring recycling and the ability to deal with multiple right-hand sides.
     The flexibility of the implementation %, which for example may deal with various matrix formats,
@@
 
-% REQUIRED
-\begin{AMS}
-  35Q68, 65Y05
-\end{AMS}
+\end{frontmatter}
+
+\linenumbers
 
@@
 the way society solves many of its most urgent technological problems through predictive simulations.
-At the heart of simulations are robust and scalable solution algorithms, and the efficient
+At the heart of simulations are robust and scalable solution algorithms. The efficient
 production of applications requires a rich ecosystem of state-of-the-art
@@
 
-Fruits of the proposed enhancement are discussed for 
-applications from
-the Scalable Library for
-Eigenvalue Problem computations (SLEPc)~\cite{hernandez2005ssf}, 
-an add-on library that extends PETSc
-with several classes for linear and nonlinear eigenvalue problems, as well as with
-tools to facilitate the use of shift-and-invert spectral transformations.
+Fruits of the proposed enhancement are discussed for applications from
+SLEPc~\cite{hernandez2005ssf}, the Scalable Library for Eigenvalue Problem
+computations, an add-on library that extends PETSc with classes for
+linear and nonlinear eigenvalue problems, as well as with tools to facilitate
+the use of shift-and-invert spectral transformations.
 
@@
 (\pk{KSPHPDDM}) and discusses the readiness of the PETSc library with respect to block Krylov methods.
-Different applications from eigen-analysis are also provided.
+Different applications from eigenanalysis are also provided.
 \Cref{sec:PCHPDDM} introduces the suite of robust multilevel overlapping Schwarz methods (\pk{PCHPDDM})
@@
 convergence plateaus. Recycling techniques have been introduced to attenuate
-these effects. In PETSc, the Loose GMRES~\cite{baker2005technique}~(LGMRES) and the
-Deflated GMRES~\cite{erhel1996restarted,wakam2013158}~(DGMRES), implemented as
-\pk{KSPLGMRES} and \pk{KSPDGMRES}, are available. However, neither handle variable
-preconditioning, and \pk{KSPDGMRES} does not support complex arithmetic;
-moreover, extending the recycling capabilites of such methods to sequence of linear systems
-with smoothly varying coefficient matrices and right-hand sides is not trivial.
-On the other hand, HPDDM offers support for the
-generalized conjugate residual method with inner orthogonalization and deflated
-restarting~\cite{parks2006recycling}~(GCRODR), which does not suffer from the aforementioned
-limitations.
+these effects. On the one hand, PETSc implements the loose GMRES~\cite{baker2005technique}~(LGMRES) and the
+deflated GMRES~\cite{erhel1996restarted,wakam2013158}~(DGMRES) as
+\pk{KSPLGMRES} and \pk{KSPDGMRES}. However, neither handle variable
+preconditioning, unlike \pk{KSPFGMRES} or \pk{KSPGCR} which respectively
+implement the flexible GMRES~\cite{saad1993flexible}~(FGMRES) and the
+generalized conjugate residual method~\cite{eisenstat1983variational}~(GCR).
+Furtheremore, \pk{KSPDGMRES} does not support complex arithmetic.  Moreover,
+extending the recycling capabilities of such methods to sequence of linear
+systems with smoothly varying coefficient matrices and right-hand sides is not
+trivial.  On the other hand, HPDDM offers support for the generalized conjugate
+residual method with inner orthogonalization and deflated
+restarting~\cite{parks2006recycling}~(GCRODR), which does not suffer from the
+aforementioned limitations.
 
@@
 
-\pk{KSPHPDDM}, the interface between HPDDM Krylov methods and PETSc, is automatically
-registered since PETSc version 3.12 when configuring PETSc with the extra
-flag \pk{-{}-download-hpddm}. HPDDM linear solvers can be selected using the command
-line option \pk{-ksp\_type hpddm}, or programmatically via \pk{KSPSetType}(\pk{ksp},
-\pk{KSPHPDDM}), and the following Krylov methods can be accessed:
+\pk{KSPHPDDM}, the interface between HPDDM Krylov methods and PETSc, is
+automatically registered since PETSc version 3.12 when configuring PETSc with
+the extra flag \pk{-{}-download-hpddm}.  HPDDM linear solvers can be selected
+using the command line option \pk{-ksp\_type hpddm}, or programmatically via
+\pk{KSPSetType}(\pk{ksp}, \pk{KSPHPDDM}). Then, the following Krylov methods
+can be accessed:
 \begin{itemize}
@@
 While being mathematically equivalent to
-their ``standard'' counterparts, the pseudo-block variants try to fuse together multiple similar
+their ``standard'' counterparts, the pseudo-block variants fuse together multiple similar
 operations like matrix--vector
@@
 always applied on the right, except for the conjugate gradient methods BCG, 
-BFCG and BFBCG, that only handle left preconditioning.  It is also
+BFCG, and BFBCG, that only handle left preconditioning.  It is also
 possible to set the preconditioning side through the more common PETSc option
@@
 
-To fully support HPDDM solvers, from version 3.14.0 of PETSc we added interface routines
-for solving systems with multiple right-hand sides, \pk{KSPMatSolve}(\pk{ksp},\pk{X},\pk{Y}),
-and to apply preconditioning methods, i.e. \pk{PCMatApply}(\pk{pc},\pk{X},\pk{Y}).
+To fully support HPDDM solvers, in version 3.14.0 of PETSc, we added interface routines
+for solving systems with multiple right-hand sides, \pk{KSPMatSolve}(\pk{ksp}, \pk{X}, \pk{Y}),
+and to apply preconditioners, \pk{PCMatApply}(\pk{pc}, \pk{X}, \pk{Y}).
 Both input $X$ and output $Y$ matrices are currently limited to be dense tall-and-skinny matrices.
-With \pk{KSPHPDDM}, \mbox{(pseudo-)block} methods Krylov methods will be used; instead, when no specialized implementation
+With \pk{KSPHPDDM}, \mbox{(pseudo-)block} Krylov methods will be used. Instead, when no specialized implementation
 is available, PETSc will perform the solution phase
@@
 
-While solving the (block) linear system, HPDDM will repeatedly call the following PETSc routines:
+While solving linear systems, possibly with multiple right-hand sides, HPDDM will repeatedly call the following PETSc routines:
 \begin{itemize}
     \item \pk{MatMult}($A$, $x$, $y$) for $y = Ax$;
-    \item \pk{MatMatMult}($A$, $X$, {\small \pk{MAT\_REUSE\_MATRIX}}, {\small
-        \pk{PETSC\_DEFAULT}}, $Y$)\footnote{or \pk{MatProductNumeric}($Y$) with PETSc 3.14.0 and above} for $Y =
+    \item \pk{MatMatMult}($A$, $X$, \pk{MAT\_REUSE\_MATRIX}, 
+        \pk{PETSC\_DEFAULT}, $Y$)\footnote{or \pk{MatProductNumeric}($Y$) with PETSc 3.14.0 and above} for $Y =
         AX$;
-    \item \pk{PCApply}(M, $x$, $y$) to apply a preconditioner to $x$, i.e. for $y = M^{-1}x$;
+    \item \pk{PCApply}(M, $x$, $y$) to apply a preconditioner to $x$, i.e., for $y = M^{-1}x$;
     \item \pk{PCMatApply}(M, $X$, $Y$) for $Y = M^{-1}X$;
 \end{itemize}
-The rest of the operations are performed directly inside HPDDM.
-Specifically, within Krylov methods using the Arnoldi process, or when recycling is requested,
-one has to compute $QR$ factorizations of tall-and-skinny dense matrices to orthonormalize
-candidate basis vectors. Such operations are performed using the CholQR algorithm~\cite{stathopoulos2002block}, or
-via the (modified) Gram--Schmidt method.
-For Hessenberg matrices generated by the Arnoldi
-process, it is common to update their $QR$ decomposition using Givens
-rotations~\cite{saad2003iterative}. For block Hessenberg matrices, Householder reflectors are
-used instead~\cite{gutknecht2008updating}.
-These orthonormalization variants can be selected at runtime with the option \pk{-ksp\_hpddm\_qr (cholqr$|$mgs$|$cgs)}.
+The rest of the operations are performed directly inside HPDDM.  Specifically,
+within Krylov methods using the Arnoldi process, or when recycling is
+requested, one has to compute $QR$ factorizations of tall-and-skinny dense
+matrices to orthonormalize candidate basis vectors. Such operations are
+performed using the CholQR algorithm~\cite{stathopoulos2002block}, or via the
+(modified) Gram--Schmidt method.  These orthonormalization variants can be
+selected at runtime with the option \pk{-ksp\_hpddm\_qr (cholqr$|$mgs$|$cgs)}.
+For Hessenberg matrices generated by the Arnoldi process, it is common to
+update their $QR$ decomposition using Givens
+rotations~\cite{saad2003iterative}. For block Hessenberg matrices, Householder
+reflectors are used instead~\cite{gutknecht2008updating}.
 
 For matrix--matrix products, there are
-currently specialized implementations for the following \pk{MatType}:
+currently specialized implementations for the following \pk{MatType}s:
 \begin{itemize}
-    \item \pk{MATSEQAIJ}, standard sequential sparse matrix, based on compressed sparse row format;
-    \item \pk{MATSEQBAIJ}, block sparse matrix, based on block compressed sparse row format;
-    \item \pk{MATSEQSBAIJ}, symmetric block sparse matrix stored in upper triangular form.
-    \item \pk{MATMPIAIJ}, standard parallel sparse matrix;
-    \item \pk{MATSHELL}, user-defined matrix;
-    \item \pk{MATNEST}, block-defined matrix with nested submatrices;
-    \item \pk{MATAIJCUSPARSE}, sequential or parallel sparse matrix, offloaded to a GPU using cuSPARSE~\cite{cusparse-web-page}.
+    \item \pk{MATAIJ}: standard sequential or parallel sparse matrix, based on compressed sparse row format;
+    \item \pk{MATSEQBAIJ}: block sparse matrix, based on block compressed sparse row format;
+    \item \pk{MATSEQSBAIJ}: symmetric block sparse matrix stored in upper triangular form;
+    \item \pk{MATSHELL}: user-defined matrix;
+    \item \pk{MATNEST}: block-defined matrix with nested submatrices;
+    \item \pk{MATAIJCUSPARSE}: sequential or parallel sparse matrix, offloaded to a NVIDIA GPU using cuSPARSE~\cite{cusparse-web-page}.
 \end{itemize}
-
 The following preconditioners have specialized implementations for dealing
@@
     %\item \pk{PCNONE}: $P_{A'} = I$, identity matrix;
-    \item \pk{PCKSP}: preconditionining via an embedded Krylov;
-    \item \pk{PCMAT}: preconditioning via matrix multiplication;
+    \item \pk{PCKSP}: embedded Krylov method;
+    \item \pk{PCMAT}: matrix multiplication;
     %\item \pk{PCSHELL}: user-defined preconditioner, e.g., matrix-free;
-    % \item \pk{PCSPAI}: sparse approximate inverse~\cite{grote1997parallel};
-    \item \pk{PCHARA}~\cite{10.1145/3232850,ambartsumyan2020hierarchical}: hierarchical matrix preconditioners; 
+    %\item \pk{PCSPAI}: sparse approximate inverse~\cite{grote1997parallel};
+    \item \pk{PCHARA}: hierarchical matrices~\cite{10.1145/3232850,ambartsumyan2020hierarchical}; 
     \item \pk{PCHPDDM}: see~\cref{sec:PCHPDDM};\label{it:hpddm}
@@
 \end{enumerate}
-One appealing feature of domain decomposition-like preconditioners (i.e. 
-\cref{it:hpddm,it:asm,it:bjacobi}) is that they most often rely on exact or inexact
-factorizations~ (\cref{it:lu,it:ilu}) as subdomain solvers.
-In these cases, it is possible to access the so-called factored matrix $F$
-%via \pk{PCFactorGetMatrix} and then call \pk{MatMatSolve}($F$, $\tilde{Y}$, $\tilde{X}$).
-via \pk{PCFactorGetMatrix}, and then call \pk{MatMatSolve}($F$, $X$, $Y$) to take advantage
-of blocked forward eliminations and backward substitutions from the various factorization packages
-interfaced with PETSc. In the case of a sparse matrix,
-this strategy is possible with: MUMPS~\cite{amestoy2001fully},
-SuiteSparse~\cite{davis2004algorithm}, MKL
-PARDISO or CPARDISO~\cite{mkl-web-page}, and SuperLU~\cite{li2005overview} or
-SuperLU\_DIST~\cite{li2003superlu_dist}. For the case of a dense
-matrix: ScaLAPACK and Elemental~\cite{poulson2013elemental}. 
-Future work will consider extending the multigrid framework \pk{PCMG}, as well as other preconditioning classes to increase the arithmetic intensity of the
-block preconditioner application phase.
+One appealing feature of domain decomposition-like
+preconditioners~(\cref{it:hpddm,it:asm,it:bjacobi}) is that they most often
+rely on exact or inexact factorizations~(\cref{it:lu,it:ilu}) as subdomain
+solvers.  In these cases, it is possible to access the so-called factored
+matrix $F$
+%via \pk{PCFactorGetMatrix} and then call \pk{MatMatSolve}($F$, $\tilde{Y}$,
+%$\tilde{X}$).
+via \pk{PCFactorGetMatrix}, and then call \pk{MatMatSolve}($F$, $X$, $Y$) to
+take advantage of blocked forward eliminations and backward substitutions from
+the various factorization packages interfaced with PETSc. In the case of a
+sparse matrix, this strategy is possible with: MUMPS~\cite{amestoy2001fully},
+SuiteSparse~\cite{davis2004algorithm}, MKL PARDISO or
+CPARDISO~\cite{mkl-web-page}, and SuperLU~\cite{li2005overview} or
+SuperLU\_DIST~\cite{li2003superlu_dist}. For the case of a dense matrix:
+ScaLAPACK and Elemental~\cite{poulson2013elemental} are supported. Future work
+will consider extending the multigrid framework \pk{PCMG}, as well as other
+preconditioning classes to increase arithmetic intensity of the preconditioner
+application phase for blocks of right-hand sides.
 
@@
         GCRODR(40, 20), and Jacobi right-preconditioned GCRODR(40, 20);
-    \item in PETSc with unpreconditioned and unrestarted GMRES($\infty$) \pk{KSPGMRES};
+    \item in PETSc with no preconditioning and no restart GMRES($\infty$) \pk{KSPGMRES};
     \item in PETSc with \pk{KSPHPDDM}, all of the above.
@@
 The notation GCRODR($n$, $m$) indicates a restart after $n$ iterations and a recycling subspace of dimension $m$.
-In all the cases, convergence is declared when the initial unpreconditioned residual is reduced by 10 orders of magnitude.
+In all of the cases, convergence is declared when the initial unpreconditioned
+residual is reduced by 10 orders of magnitude.
 Except for the right-preconditioned GCRODR, these tests are the same as the
@@
 All PETSc tests are performed using four MPI
-processes and they can be launched using the following command lines: \\[-4pt]
+processes and can be launched using the following command lines. \\[-4pt]
 \begin{Verbatim}[fontsize=\footnotesize,frame=single,framerule=0.1mm,commandchars=&\[\]]
@@
 \end{tikzpicture}
-        \caption{\# of iterations needed to converge for various
+        \caption{Number of iterations needed to converge for various
         configurations as originally tested by Parks et
@@
         between \ref{pgfplots:r2} and \ref{pgfplots:r3}. In the other cases,
-        the numbers of iterations needed by \pk{KSPHPDDM} match with the respective
+        the number of iterations needed by \pk{KSPHPDDM} match with the respective
         MATLAB or PETSc reference implementation.\label{fig:comparison}}
@@
 
-In particular, we consider the first three matrix types from~\cref{sec:interface-ksp}, namely \pk{MATSEQAIJ}, \pk{MATSEQBAIJ}, and \pk{MATSEQSBAIJ}.
+In particular, we consider the first three sequential matrix types from~\cref{sec:interface-ksp}, namely \pk{MATSEQAIJ}, \pk{MATSEQBAIJ}, and \pk{MATSEQSBAIJ}.
 Their distributed memory counterparts, e.g., \pk{MATMPIAIJ}, rely on the sequential implementations, with each process
@@
         width=0.8\textwidth,
-        xlabel={\# of right-hand sides ($N$)},
+        xlabel={Number of right-hand sides ($N$)},
         ylabel={},
@@
 \begin{itemize}
-    \item with a block size of 1, top plot $b=1$ in~\cref{fig:perfAX}, using any type but \pk{MATAIJ}~\ref{pgfplots:aij1} is counterproductive when $N>1$, see~\ref{pgfplots:baij1} and~\ref{pgfplots:sbaij1}, even when using MKL, see~\ref{pgfplots:mklbaijseq1} and~\ref{pgfplots:mklsbaijseq1}.
-With 8 or more columns, the performance of \pk{MATAIJ} plateaus and the efficiency reaches approximately 200\%, while the performance of the MKL primitives stagnate at around 175\%, see~\ref{pgfplots:aij1};
-for the single column case, \pk{MATSBAIJ} and \pk{MATAIJMKL} deliver slightly better performances;
+    \item with a block size of 1, top plot $b=1$ in~\cref{fig:perfAX}, using any type but \pk{MATAIJ}~\ref{pgfplots:aij1} is counterproductive when $N>1$, see~\ref{pgfplots:baij1} and~\ref{pgfplots:sbaij1}, even when using the MKL, see~\ref{pgfplots:mklbaijseq1} and~\ref{pgfplots:mklsbaijseq1}.
+With 8 or more columns, the performance of \pk{MATAIJ} plateaus and the efficiency reaches approximately 200\%, while the performance of the MKL primitives stagnate at around 175\%, see~\ref{pgfplots:aij1}.
+For the single column case, \pk{MATSBAIJ} and \pk{MATAIJMKL} deliver slightly better performances;
     \item for block sizes lower or equal than 5, loops for matrix--vector and matrix--matrix multiplications are unrolled by hand for block formats.
@@
 For example, if the matrices were to be offloaded to a GPU, one could instead benchmark \pk{MATAIJcuSPARSE}.
-
-The mini-app is then used to benchmark intranode performance of matrix--matrix multiplications, either with multiple OpenMP threads, or by offloading the operation to a GPU and using cuSPARSE~\cite{cusparse-web-page} as interfaced in PETSc.
+The mini-app is then used to benchmark intranode performance of matrix--matrix
+multiplications, either with multiple OpenMP threads, or by offloading the
+operation to a NVIDIA GPU and using cuSPARSE~\cite{cusparse-web-page} as
+interfaced in PETSc.
 \begin{Verbatim}[fontsize=\footnotesize,frame=single,framerule=0.1mm,commandchars=&\[\]]
@@
 mini-app. Still, it is less than 1\% of peak for an Intel Xeon Gold 6248.
-For the GPU configuration, \pk{MATAIJCUSPARSE}~\ref{pgfplots:cusparse10} with $b=6$
+On the GPU, \pk{MATAIJCUSPARSE}~\ref{pgfplots:cusparse10} with $b=6$
 and $N=64$, rightmost bottom plot in~\cref{fig:perfAXpar}, performs at around
@@
 consider using interlaced layouts for storing the dense right-hand sides and extend the
-current PETSc functionality in order to maximize performances while maintaining interface flexibility
+current PETSc functionality in order to maximize performance while maintaining interface flexibility
 and the user friendliness of the library.
@@
               ytick={1,2,4,8,20},
+              ymajorgrids,
         yticklabel={
@@
           extra x tick labels = {1,2,4,8,16,\hspace*{0.3cm}20},
-          legend style={at={(0.02,0.98)}, legend cell align=left,anchor=north west},
+          legend style={at={(1.02,0.5)}, legend cell align=left,anchor=west},
       ]
@@
 evaluated using a given base flow $q_b = \begin{bmatrix}u_b \\ p_b\end{bmatrix}$ and
-$M$ the discretization of the mass matrix on the space of velocities.
+$M$ is the discretization of the mass matrix on the space of velocities.
 The mini-app employed for the numerical results is built on top of
@@
 as reported in~\cref{fig:comparison-slepc}, from 80 with \pk{KSPFGMRES} to 45.
-A similar speedup can be observed in terms of runtime, with the time spent for the solution of the eigenvalue problem being reduced from \SI[round-mode = places,round-precision=1,group-minimum-digits=3,group-separator = {,}]{1509}{\sec} to \SI[round-mode = places,round-precision=1,group-minimum-digits=3,group-separator = {,}]{1052}{\sec}.
-These timings, and all that follow, have been obtained on {Ir\`ene}, a system composed of \pgfmathprintnumber[assume math mode=true]{1656} nodes with two 24-core Intel Xeon Platinum 8168 clocked at \SI{2.7}{\giga\hertz}.
+A similar speedup can be observed in terms of runtime, with the time spent for
+the solution of the eigenvalue problem being reduced from \SI[round-mode = places,round-precision=1,group-minimum-digits=3,group-separator = {,}]{25.1}{\minute}
+to \SI[round-mode = places,round-precision=1,group-minimum-digits=3,group-separator = {,}]{17.5}{\minute}.
+These timings, and all that follow, have been obtained on
+Irène, a system composed of \pgfmathprintnumber[assume math mode=true]{1656}
+nodes with two 24-core Intel Xeon Platinum 8168 clocked at
+\SI{2.7}{\giga\hertz}.
 \pgfplotstableread{
@@
 \end{tikzpicture}
-        \caption{\# of inner iterations for systems~\cref{eq:ks} with~\ref{pgfplots:m0} or without~\ref{pgfplots:m1} recycling.\label{fig:comparison-slepc}}
+        \caption{Number of inner iterations for systems~\cref{eq:ks} with~\ref{pgfplots:m1} or without~\ref{pgfplots:m0} recycling.\label{fig:comparison-slepc}}
     \end{figure}
@@
 \subsubsection{Blocking inside SLEPc\label{sec:lobpcg}}
-The implementations of the Locally Optimal Block Preconditioned Conjugate
-Gradient~\cite{knyazev2001toward}~(LOBPCG) and the Contour Integral Spectrum
-Slicing method~\cite{sakurai2003projection}~(CISS) in SLEPc were previously not
+The implementations of the locally optimal block preconditioned conjugate
+gradient~\cite{knyazev2001toward}~(LOBPCG) and the contour integral spectrum
+slicing method~\cite{sakurai2003projection}~(CISS) in SLEPc were previously not
 using blocking when applying the preconditioner or solving linear systems. In
@@
 %implementation, the gains are striking, as reported in~\cref{sec:lobpcg}.
-
-The performance of LOBPCG~\cite{knyazev2001toward} as implemented in SLEPc is studied using three different preconditioners:
+The performance of LOBPCG implemented in SLEPc as \pk{EPSLOBPCG} is studied using three different preconditioners:
 \begin{itemize}
     \item \pk{PCASM}: one-level overlapping Schwarz method with one level of overlap and exact Cholesky factorizations in each subdomain;
-    \item \pk{PCHPDDM}: multilevel overlapping Schwarz method;
+    \item \pk{PCHPDDM}: multilevel overlapping Schwarz method, see~\cref{sec:PCHPDDM};
     \item \pk{PCGAMG}: algebraic multigrid method~\cite{adams2004ultra}.
 \end{itemize}
-More details on these solvers are given in the next section where focus is put on preconditioning rather than Krylov methods.
-We note here that only \pk{PCASM} and \pk{PCHPDDM} currently handle blocking (see the list~\cref{sec:interface-ksp}),
-and that \pk{PCHPDDM} and \pk{PCGAMG} have lower contraction factors than the cheaper alternative \pk{PCASM}.
+More details on these solvers are given in the next section where focus is put
+on preconditioning rather than Krylov methods.  We note here that only
+\pk{PCASM} and \pk{PCHPDDM} currently handle blocking, see the
+list~\cref{sec:interface-ksp}, and that \pk{PCHPDDM} and \pk{PCGAMG} have lower
+contraction factors than the cheaper alternative \pk{PCASM}.
 
@@
 elements using FreeFEM. The script \texttt{blocking-slepc.edp} available at
-\url{https://github.com/prj-/jolivet2020petsc} can be use for reproducibility.
+\url{https://github.com/prj-/jolivet2020petsc} can be used for reproducibility.
 The following timings and convergence histories have been obtained on $2{,}304$
-processes of Ir\`ene, for solving a problem of dimension
+processes of Irène, for solving a problem of dimension
 \pgfmathprintnumber{74618461}.
@@
 pseudo-block GMRES iterations with 40 vectors simultaneously, using \pk{MatProductNumeric}
-and \pk{PCMatApply}, instead of doing (at most) $40\times 5$ successive GMRES
+and \pk{PCMatApply}, instead of doing at most $40\times 5$ successive GMRES
 iterations, using \pk{MatMult} and \pk{PCApply}.
@@
 preconditioners have a similar setup time. The
-time spent in the full eigenvalue solver is reported in the last row. Clearly, having both a \pk{KSP}
+time spent in the full eigensolver is reported in the last row. Clearly, having both a \pk{KSP}
 and a \pk{PC} which can efficiently deal with multiple vectors is mandatory to
@@
 We finally note that blocking support is also available and tested in SLEPc for \pk{EPSSUBSPACE}
-and \pk{NEPCISS} since version 3.14.0.
-
-\section{PCHPDDM}\label{sec:PCHPDDM}
+since version 3.14.0.
+
+\section{Robust multilevel overlapping Schwarz preconditioners: PCHPDDM}\label{sec:PCHPDDM}
   \subsection{Related work}
@@
 in PETSc as \pk{PCBJACOBI} and \pk{PCASM} respectively, with some additional customization~\cite{cai1999restricted}
-available for \pk{PCASM}. However, none of these methods provide automatic support for a coarse level solver,
-which is mandatory to obtain algorithmic scalability with a large number of processes.
+available for \pk{PCASM}. However, none of these methods provide automatic support for a coarse-level solver,
+which is mandatory to obtain algorithmic scalability with large numbers of processes.
 %ASM stands for additive Schwarz
@@
 % preconditioner~\cite{dohrmann2008domain}. Its type is \pk{PCEXOTIC}.
-Furthermore, PETSc provides domain decomposition methods of the non-overlapping
-type. \pk{PCNN} implements a basic balancing Neumann--Neumann
+Furthermore, PETSc provides non-overlapping domain decomposition methods. \pk{PCNN} implements a basic balancing Neumann--Neumann
 method~\cite{mandel1993balancing}, while more advanced preconditioning
@@
 in PETSc with \pk{PCSetModifySubMatrices}, see~\cite{gander2019new}.  It also provides support for
-defining robust two-level methods using the Generalized Eigenvalue problem on
-the Overlap (GenEO) framework, for finite
+defining robust two-level methods using the generalized eigenvalue problem on
+the overlap (GenEO) framework, for finite
 elements~\cite{spillane2011robust,spillane2013abstract} and boundary
@@
   \subsection{Interfacing HPDDM overlapping Schwarz methods in PETSc\label{sec:interface-pc}}\
-The interface between HPDDM overlapping Schwarz methods and PETSc, \pk{PCHPDDM}, is automatically
-registered since PETSc version 3.12 when one configures PETSc with the extra
-flag \pk{-{}-download-hpddm -{}-download-slepc}. It is then
-possible to select the corresponding \pk{PC}
-using the command line option \pk{-pc\_type hpddm}, or the routine \pk{PCSetType}(\pk{pc},
-\pk{PCHPDDM}).
+\pk{PCHPDDM}, the interface between HPDDM overlapping Schwarz methods and
+PETSc, is automatically registered since PETSc version 3.12 when configuring
+PETSc with the extra flag \pk{-{}-download-hpddm -{}-download-slepc}.  It is
+then possible to select the corresponding \pk{PC} using the command line option
+\pk{-pc\_type hpddm}, or the routine \pk{PCSetType}(\pk{pc}, \pk{PCHPDDM}).
  % and its Fortran equivalent.
@@
 discretization of $\mathcal{L}$ on the extended local subdomain with overlap, endowed with natural boundary
-conditions. In the case of the 1D Poisson
-equation with homogeneous Dirichlet boundary conditions discretized using linear finite elements, an explicit representation of these  operators is given~\cref{fig:poisson} for a
+conditions. In the case of the one-dimensional Poisson
+equation with homogeneous Dirichlet boundary conditions discretized using trilinear finite elements, an explicit representation of these operators is given~\cref{fig:poisson} for a
 two-domain decomposition.
@@
     \end{minipage}
-\caption{Notations and basics of overlapping Schwarz methods.\label{fig:poisson}}
+\caption{Notation and basics of overlapping Schwarz methods.\label{fig:poisson}}
 \end{figure}
@@
         \begin{equation}\label{eq:gevp}
-            \mathring{A}_i y_i = \lambda_i \tilde{R}_i R_i^T (R_i A R_i^T) \tilde{R}_i R_i^T y_i ;
+            \mathring{A}_i y_i = \lambda_i \tilde{R}_i R_i^T (R_i A R_i^T) R_i \tilde{R}_i^T y_i ;
         \end{equation}
@@
 \end{equation*}
-Such a construction of a two-level method yields a preconditioner with which it is possible to bound the condition number of the preconditioned operator.
-Thus, the convergence speed of a preconditioned Krylov method such as the conjugate gradient can be guaranteed a priori.
+Such a construction of a two-level method yields a preconditioner with which it
+is possible to bound the condition number of the preconditioned operator. Thus,
+the convergence rate of a preconditioned Krylov method such as the conjugate
+gradient can be guaranteed a priori.
 
@@
 the size of the coarse operator $P^TAP$ is much lower
-than $n$, the size of $A$, and coarsening is much more aggressive than
+than the size of $A$, and coarsening is much more aggressive than
 when using algebraic or geometric multigrid methods.
@@
 Other \pk{DMType}s, including user-defined, may be supported by
-implementing the proper \pk{DMCreateNeumannOverlap} callback. Support for the structured grid infrastructure (\pk{DMDA})
-could be easily added upon request.
+implementing the proper \pk{DMCreateNeumannOverlap} callback. Support for the
+structured grid infrastructure (\pk{DMDA}) could be easily added.
 \begin{figure}[htbp]
@@
 With the local
-matrix pencils available, new options will be automatically registered such as
+matrix pencils available, new options are automatically registered such as
 \pk{-pc\_hpddm\_levels\_1\_eps\_nev} or \pk{-pc\_hpddm\_levels\_1\_eps\_threshold}, which
@@
 
-For performance reasons, it is better to compute the same number of
-eigenvectors on each process, since in this case the coarse operator will be
-assembled using the blocked matrix formats, with
-usually large block sizes, with consequent impact on memory, floating point and message passing performances.
+For performance, it is better to compute the same number of
+eigenvectors on each process, since in this case, the coarse operator will be
+assembled using symmetric or general blocked matrix formats. With
+usually large block sizes, this has consequent impacts on memory, floating-point, and message passing performances.
 % When applying the preconditioner,
@@
 In addition,
-standard eigensolvers such as \pk{EPSKRYLOVSCHUR} from SLEPc, cannot cheaply
+standard eigensolvers such as \pk{EPSKRYLOVSCHUR} from SLEPc, cannot inexpensively
 evaluate the exact number of eigenpairs in a given interval of $\mathbb{R}^+$.
-Instead, it is often more efficient to provide an upper bound of the number of
+Instead, it is often more efficient to provide an upper bound on the number of
 eigenpairs.
@@
 Some other parameters may be adjusted to define how the one-level preconditioner
-and the coarse grid operator interact. Besides the additive
+and the coarse-grid operator interact. Besides the additive
 correction shown in~\cref{eq:two}, end-users may select more numerically efficient
@@
 with $Q = P \left(P^T A P\right)^{-1} P^T$ and $M^{-1}$ defined in~\cref{eq:one}. This is parameterized using the
-options \pk{-pc\_hpddm\_coarse\_correction}
+option \pk{-pc\_hpddm\_coarse\_correction}
 \pk{(deflated$|$additive$|$balanced)} or the routine
@@
 incompressible elasticity or the system of Stokes. Indeed, for such equations,
-one needs to compute a lot of eigenvectors from the classical GenEO
+one needs to compute a large number of eigenvectors from the classical GenEO
 eigenproblem~\cref{eq:gevp} to generate a robust preconditioner. GenEO-2
 alleviates this phenomenon. In the case where the local operators
-$\{B_i\}_{i=1}^N$ are the discrete mass matrices on the subdomains interfaces,
+$\{B_i\}_{i=1}^N$ are the discrete mass matrices on the subdomain interfaces,
 the so-called Dirichlet-to-Neumann coarse operator is constructed. This has
@@
 fine level decomposition, switching to a multilevel scheme may alleviate the increasing cost
-of solving the coarse linear system. A multilevel extension of the GenEO framework has been recently
+of solving coarse linear systems. A multilevel extension of the GenEO framework has been recently
 proposed~\cite{aldaas2019multi}, with the advantage that the multilevel hierarchy can be
@@
 Because of the intrinsic nature of balancing domain
-decomposition methods, \pk{PCNN} and \pk{PCBDDC} must be supplied a \pk{MatIS}
+decomposition methods, \pk{PCNN} and \pk{PCBDDC} must be supplied with a \pk{MatIS}
 which stores local unassembled matrices and local to global mappings.
@@
 finite element context. Since the assembled form of these operators can be reconstructed,
-it is possible to get the Dirichlet operators
+it is possible to obtain the Dirichlet operators
 $\{R_i A R_i^T\}_{i=1}^N$ from~\cref{sec:interface-pc} as well. All tools needed by the
 GenEO framework are thus readily available. Note however that one-level
-overlapping Schwarz methods are known for converging slowly when there is in
-fact no overlap. The proposed methodology is not strictly equivalent to the
+overlapping Schwarz methods are known for converging slowly when there is
+no overlap. The proposed methodology is not strictly equivalent to the
 BDD-GenEO method~\cite{spillane2013automatic}, in which local Schur complements
@@
 \[
-	\varepsilon_{ij}(u) = \frac{1}{2} \left(\frac{\partial u_i}{\partial x_i} + \frac{\partial u_j}{\partial x_j} \right), \mu = \frac{E}{2(1+\nu)}, \text{ and }\lambda = \frac{E\nu}{1-2\nu}.
+    \varepsilon_{ij}(u) = \frac{1}{2} \left(\frac{\partial u_i}{\partial x_i} + \frac{\partial u_j}{\partial x_j}\right)\!, \text{ } \mu = \frac{E}{2(1+\nu)}, \text{ and }\lambda = \frac{E\nu}{1-2\nu}.
 \]
@@
 $\Gamma_N$ is defined as the complementary of $\Gamma_D$ with respect to the boundary of $\Omega$.
-\Cref{eq:elasticity} is discretized using piecewise linear
-finite elements resulting in $\pgfmathprintnumber{593} \times 10^6$ number of unknowns
-with approximately $\pgfmathprintnumber{45}$ nonzero coefficients per row.
+\Cref{eq:elasticity} is discretized using trilinear
+finite elements resulting in $\pgfmathprintnumber{593} \times 10^6$ unknowns
+with approximately $\pgfmathprintnumber{45}$ nonzero coefficients per row in the discrete coefficient matrix.
 The physical domain $\Omega$ is decomposed in $13{,}824$ subdomains using the automatic graph partitioner ParMETIS~\cite{karypis1998fast}.
@@
 
-Considering the parametrization used for the \pk{PCHPDDM} solvers, \cref{enum:exact} and \cref{enum:inexact} yield the same number of outer iterates.
+Considering the parametrization used for the \pk{PCHPDDM} solvers, \cref{enum:exact,enum:inexact} yield the same number of outer iterates.
 In~\cref{fig:gamg-geneo}, the convergence history of \pk{PCGAMG} and \pk{PCHPDDM} are reported.
 In~\cref{fig:inexact}, the number of inner coarse iterations are reported for various values of $M$.
-In the case of a two-level method,~\cref{enum:exact}, this number is equal to one and is thus not reported in the figure.
+In the case of a two-level method (\cref{enum:exact}) this number is equal to one and is thus not reported.
 \pgfplotstableread{
@@
     enlarge x limits=0.05,
-% yticklabel={
-% \pgfmathparse{int(\tick)}
-% \ifnum\pgfmathresult=0
-% $1$
-% \else
-% \pgfmathparse{int(\tick/ln(10))}{$10^{\pgfmathresult}$}
-% \fi
-% }
+yticklabel={
+\pgfmathparse{int(\tick)}
+\ifnum\pgfmathresult=0
+$1$
+\else
+\pgfmathparse{int(\tick/ln(10))}{$10^{\pgfmathresult}$}
+\fi
+}
 ]
@@
 }
-    \subfloat[\parbox{5cm}{\# of inner coarse iterations for various values of $M$, the number of second-level subdomains.}\label{fig:inexact}]{
+    \subfloat[\parbox{5cm}{Number of inner coarse iterations for various values of $M$, the number of second-level subdomains.}\label{fig:inexact}]{
         \begin{tikzpicture}[font=\scriptsize]
@@
 Note that in this test case, \pk{PCGAMG} is faster than all \pk{PCHPDDM}
-alternatives. For \pk{PCGAMG}, the setup phase takes 70 seconds, while the solution phase requires 19 seconds;
-on the other hand, the fastest \pk{PCHPDDM} configuration, which corresponds to
-configuration~\ref{pgfplots:p-256} with 256 second-level subdomains, requires 64 seconds to setup and 31 for the solution of the linear system.
+alternatives. On the one hand, for \pk{PCGAMG}, the setup phase takes 70 seconds, while the solution phase requires 19 seconds.
+On the other hand, the fastest \pk{PCHPDDM} configuration, which corresponds to the
+configuration with 256 second-level subdomains~\ref{pgfplots:p-256}, requires
+64 seconds to setup and 36 seconds for the solution of the linear system.
 Out of the 64 seconds needed for \pk{PCHPDDM} setup, 31 are spent in SLEPc
@@
 \begin{equation}
-    -\nabla \cdot (\kappa \nabla u) - 6.2 \exp^u = 0,\label{eq:bratu}
+    -\nabla \cdot (\kappa \nabla u) - 6.2 e^u = 0,\label{eq:bratu}
 \end{equation}
@@
 approximately $\pgfmathprintnumber{29}$ nonzero coefficients per row.
-It is solved using a \pk{SNES}, with \pk{PCHPDDM} being the preconditioner for
+It is solved using \pk{SNES}, with \pk{PCHPDDM} being the preconditioner for
 solving each linearized system, see~\cref{fig:u}.
@@
 \addplot[very thick,last] table [x expr=\thisrowno{0}+47, y expr=\thisrowno{4}/8.796157053238e-07] {\comparisonBratu};
-\node[font=\tiny,fill=white,inner sep=1pt] at (axis cs:0, 2) {12.0};
-\node[font=\tiny,fill=white,inner sep=1pt] at (axis cs:12, 2) {3.4E-4};
-\node[font=\tiny,fill=white,inner sep=1pt] at (axis cs:28, 2) {1.9E-5};
-\node[font=\tiny,fill=white,inner sep=1pt] at (axis cs:47, 2) {8.8E-7};
-\node[font=\tiny,fill=white,inner sep=1pt] at (axis cs:67, 2) {2.5E-9};
+      \contourlength{.7pt}
+\node[font=\tiny,inner sep=1pt] at (axis cs:0, 2) {\contour{white}{12.0}};
+\node[font=\tiny,inner sep=1pt] at (axis cs:12, 2) {\contour{white}{3.4E-4}};
+\node[font=\tiny,inner sep=1pt] at (axis cs:28, 2) {\contour{white}{1.9E-5}};
+\node[font=\tiny,inner sep=1pt] at (axis cs:47, 2) {\contour{white}{8.8E-7}};
+\node[font=\tiny,inner sep=1pt] at (axis cs:67, 2) {\contour{white}{2.5E-9}};
 \end{semilogyaxis}
@@
     \subsubsection{Using PCHPDDM as a coarse grid solver for PCBDDC\label{sec:bddc}}
-A last toy problem is here proposed to show how one may switch between a multilevel
+A last problem shows how one may switch between a multilevel
 BDDC solver to a two-level BDDC solver using \pk{PCHPDDM} for solving the
-coarse problem. The problem is generated using MFEM~\cite{anderson2019mfem}
+coarse problem. The problem is generated using the MFEM~\cite{anderson2019mfem}
 definite Maxwell example available at \url{https://github.com/mfem/mfem/blob/master/examples/petsc/ex3p.cpp}.
@@
 \end{align}
-It is discretized using first order N\'ed\'elec finite elements and \pk{PCBDDC}
+It is discretized using first order Nédélec finite elements and \pk{PCBDDC}
 is automatically customized to obtain a stable method with these
@@
 preconditioners can currently handle them efficiently, and additional
-storage formats may be considered to maximize performances.
+storage formats may be considered to maximize performance.
 
@@
 
-The interface between PETSc and HPDDM has still room for future improvements.
-First, HPDDM handles mixed-precision,
+The interface between PETSc and HPDDM has room for future improvements.  First,
+HPDDM handles mixed-precision,
 e.g., using single-precision scalars for assembling coarse operators while
@@
 handle GPU efficiently currently, but the interface with PETSc provides a very
-thorough testbed with \pk{Mat} and \pk{PC} implementations that are exploiting
-GPU.
+thorough testbed with \pk{Mat} and \pk{PC} implementations that do exploit
+GPUs.
 
 \section*{Acknowledgments}
-The authors would like to thank S.~Balay, J.~Brown, V.~Hapla, M.~Knepley,
-and B.~Smith for reviewing the successive merge requests in PETSc
-repository. This work was granted access to the GENCI-sponsored HPC resources of:
+The authors would like to thank S.~Balay, J.~Brown, V.~Hapla, M.~Knepley, and
+B.~Smith for reviewing the successive merge requests in PETSc repository and
+for their feedback on this manuscript. This work was granted access to the
+GENCI-sponsored HPC resources of:
 \begin{itemize}
@@
 
-\bibliographystyle{siamplain}
+\bibliographystyle{elsarticle-num}
 \begin{thebibliography}{10}

### 2020-07-21 article.tex
Fixing some formatting issues

@@
 
-\usepackage{lineno,hyperref}
 % \modulolinenumbers[5]
@@
 \usepackage[outline]{contour}
+\usepackage{enumitem}
 \usepackage{tikz}
@@
 \usepackage{amsmath,etoolbox,amssymb}
-\usepackage{cleveref}
 \AtBeginEnvironment{pmatrix}{\setlength{\arraycolsep}{2pt}}
@@
 \pgfmathsetmacro{\overlap}{1}
+\usepackage{lineno,hyperref}
+\usepackage[noabbrev]{cleveref}
+\hypersetup{
+  colorlinks=true
+}
 \begin{document}
+\hypersetup{
+  colorlinks=true,
+  linkcolor=red,
+  urlcolor=blue,
+  citecolor=green!50!black,
+  frenchlinks=false,
+  pdfborder={0 0 0},
+  naturalnames=false,
+  hypertexnames=false,
+  breaklinks
+}
 
@@
 
-\title{KSPHPDDM and PCHPDDM: extending PETS\MakeLowercase{c} with advanced Krylov methods and robust multilevel Overlapping Schwarz preconditioners}
+\title{KSPHPDDM and PCHPDDM: extending PETSc with advanced Krylov methods and robust multilevel overlapping Schwarz preconditioners}
 
@@
 % REQUIRED
-\begin{keywords}
+\begin{keyword}
   Krylov methods, domain decomposition preconditioners, distributed-memory parallel computing
-\end{keywords}
+\end{keyword}
 
@@
 \linenumbers
-
+\setlist{nosep}
 \section{Introduction}
@@
 options \pk{-ksp\_hpddm\_type}
-\pk{(gmres$|$cg$|$gcrodr$|$bgmres$|$bcg$|$bfbcg$|$bgcrodr$|$preonly)}, or programmatically 
-by using \pk{KSPHPDDMSetType}(\pk{ksp}, \pk{KSPHPDDMType}).
+\pk{(gmres$|$cg$|$gcrodr$|$\linebreak bgmres$|$bcg$|$bfbcg$|$bgcrodr$|$preonly)}, or programmatically 
+by using \pk{KSPHPDDMSetType}(\pk{ksp},\linebreak \pk{KSPHPDDMType}).
 Preconditioning variants can be specified via
-\pk{-ksp\_hpddm\_variant} \pk{(left$|$right$|$flexible)} to select whether the
+\pk{-ksp\_hpddm\_variant} \pk{(left$|$right$|$\linebreak flexible)} to select whether the
 preconditioner is applied on the left, on the right, or if it cannot be
@@
 \begin{Verbatim}[fontsize=\footnotesize,frame=single,framerule=0.1mm,commandchars=&\[\]]
-&fvtextcolor[mygreen][$] &fvtextcolor[myyellow][mpirun] -n 4 ./ex75 -ksp_converged_reason -pc_type none -ksp_rtol 1e-10             &fvtextcolor[myred][\]
-  -ksp_type hpddm -ksp_hpddm_type gcrodr -ksp_gmres_restart 40 -ksp_hpddm_recycle 20 &fvtextcolor[myred][\]
+&fvtextcolor[mygreen][$] &fvtextcolor[myyellow][mpirun] -n 4 ./ex75 -ksp_converged_reason -pc_type none -ksp_rtol 1e-10 -ksp_gmres_restart 40   &fvtextcolor[myred][\]
+  -ksp_type hpddm -ksp_hpddm_type gcrodr -ksp_hpddm_recycle 20                                   &fvtextcolor[myred][\]
   -load_dir &fvtextcolor[myyellow][${DATAFILESPATH}]/matrices/hpddm/GCRODR
@@
   -load_dir &fvtextcolor[myyellow][${DATAFILESPATH}]/matrices/hpddm/GCRODR -pc_type redundant
-&fvtextcolor[mygreen][$] &fvtextcolor[myyellow][mpirun] -n 4 ./ex75 -ksp_converged_reason -pc_type jacobi -ksp_rtol 1e-10           &fvtextcolor[myred][\]
-  -ksp_type hpddm -ksp_hpddm_type gcrodr -ksp_gmres_restart 40 -ksp_hpddm_recycle 20 &fvtextcolor[myred][\]
+&fvtextcolor[mygreen][$] &fvtextcolor[myyellow][mpirun] -n 4 ./ex75 -ksp_converged_reason -pc_type jacobi -ksp_rtol 1e-10 -ksp_gmres_restart 40 &fvtextcolor[myred][\]
+  -ksp_type hpddm -ksp_hpddm_type gcrodr -ksp_hpddm_recycle 20                                   &fvtextcolor[myred][\]
   -ksp_pc_side right -load_dir &fvtextcolor[myyellow][${DATAFILESPATH}]/matrices/hpddm/GCRODR
-&fvtextcolor[mygreen][$] &fvtextcolor[myyellow][mpirun] -n 4 ./ex75 -ksp_converged_reason -pc_type none -ksp_rtol 1e-10             &fvtextcolor[myred][\]
-  -ksp_type hpddm -ksp_gmres_restart 500                                             &fvtextcolor[myred][\]
-  -load_dir &fvtextcolor[myyellow][${DATAFILESPATH}]/matrices/hpddm/GCRODR
+&fvtextcolor[mygreen][$] &fvtextcolor[myyellow][mpirun] -n 4 ./ex75 -ksp_converged_reason -pc_type none -ksp_rtol 1e-10 -ksp_gmres_restart 500  &fvtextcolor[myred][\]
+  -ksp_type hpddm -load_dir &fvtextcolor[myyellow][${DATAFILESPATH}]/matrices/hpddm/GCRODR
 \end{Verbatim}
@@
 \begin{Verbatim}[fontsize=\footnotesize,frame=single,framerule=0.1mm,commandchars=&\[\]]
-&fvtextcolor[mygreen][$] &fvtextcolor[myyellow][mpicc] MatProduct.c -O3 -I&fvtextcolor[myyellow][${PETSC_DIR}]/&fvtextcolor[myyellow][${PETSC_ARCH}]/include                        &fvtextcolor[myred][\]
-  -I&fvtextcolor[myyellow][${PETSC_DIR}]/include -L&fvtextcolor[myyellow][${PETSC_DIR}]/&fvtextcolor[myyellow][${PETSC_ARCH}]/lib -lpetsc -o MatProduct
+&fvtextcolor[mygreen][$] &fvtextcolor[myyellow][mpicc] MatProduct.c -O3 -I&fvtextcolor[myyellow][${PETSC_DIR}]/&fvtextcolor[myyellow][${PETSC_ARCH}]/include -I&fvtextcolor[myyellow][${PETSC_DIR}]/include             &fvtextcolor[myred][\]
+  -L&fvtextcolor[myyellow][${PETSC_DIR}]/&fvtextcolor[myyellow][${PETSC_ARCH}]/lib -lpetsc -o MatProduct
 &fvtextcolor[mygreen][$] &fvtextcolor[myyellow][mpirun] -n 1 ./MatProduct -f binaryoutput -log_view                                             &fvtextcolor[myred][\]
@@
 
-\begin{figure}[tbp]
+\begin{figure}[!t]
 \pgfkeys{/pgf/fpu}
@@
           \path[shape=coordinate]
-(rel axis cs:-0.001,0.54) coordinate(b1) (rel axis cs: 1.001,0.54) coordinate(b2)
-(rel axis cs: 1.001,0.56) coordinate(b3) (rel axis cs:-0.001,0.56) coordinate(b4);
+(rel axis cs:-0.003,0.54) coordinate(b1) (rel axis cs: 1.003,0.54) coordinate(b2)
+(rel axis cs: 1.003,0.56) coordinate(b3) (rel axis cs:-0.003,0.56) coordinate(b4);
   \path[fill=white] (b2) -- (b3)
@@
 \pgfplotstablecreatecol[create col/expr={\refVI/\thisrow{aij6}}]{speedupVI}\loadedOMP
-\begin{figure}[tbp]
+\begin{figure}[!b]
   \centering
@@
 option \pk{-pc\_hpddm\_coarse\_correction}
-\pk{(deflated$|$additive$|$balanced)} or the routine
+\pk{(deflated$|$additive$|$balanced)} or the routine\linebreak
 \pk{PCHPDDMSetCoarseCorrectionType}. All of
@@
 }
-    \subfloat[\parbox{5cm}{\pk{SNES}, \pk{KSP}, and \pk{PC} options for
+    \subfloat[\parbox{4cm}{\pk{SNES}, \pk{KSP}, and \pk{PC} options for
     solving~\cref{eq:bratu}}\label{fig:bratu-options}]{
@@
 \begin{Verbatim}[fontsize=\footnotesize,frame=single,framerule=0.1mm,commandchars=&\[\]]
-&fvtextcolor[mygreen][$] &fvtextcolor[myyellow][mpirun] -n 512 ./ex3p -m ../../data/fichera.mesh -f 4 --petscopts rc_ex3p_bddc      &fvtextcolor[myred][\]
-  --nonoverlapping
+&fvtextcolor[mygreen][$] &fvtextcolor[myyellow][mpirun] -n 512 ./ex3p -m ../../data/fichera.mesh -f 4 --petscopts rc_ex3p_bddc --nonoverlapping
 \end{Verbatim}
@@
 \begin{thebibliography}{10}
-
-\bibitem{adams2004ultra}
-{\sc M.~Adams, H.~H. Bayraktar, T.~M. Keaveny, and P.~Papadopoulos}, {\em
-  {Ultrascalable Implicit Finite Element Analyses in Solid Mechanics with over
-  a Half a Billion Degrees of Freedom}}, in Proceedings of the 2004 ACM/IEEE
-  Conference on Supercomputing, SC04, IEEE Computer Society, 2004,
-  pp.~\mbox{34:1--34:15}.
-
-\bibitem{aldaas2019multi}
-{\sc H.~Al~Daas, L.~Grigori, P.~Jolivet, and P.-H. Tournier}, {\em {A
-  Multilevel Schwarz Preconditioner Based on a Hierarchy of Robust Coarse
-  Spaces}}, Journal of Scientific Computing,  (2019), p.~submitted for
-  publication, \url{https://github.com/prj-/aldaas2019multi}.
-
-\bibitem{ambartsumyan2020hierarchical}
-{\sc I.~Ambartsumyan, W.~Boukaram, T.~Bui-Thanh, O.~Ghattas, D.~Keyes,
-  G.~Stadler, G.~Turkiyyah, and S.~Zampini}, {\em {Hierarchical Matrix
-  Approximations of Hessians Arising in Inverse Problems Governed by PDEs}},
-  arXiv:2003.10173,  (2020).
-
-\bibitem{amestoy2001fully}
-{\sc P.~Amestoy, I.~Duff, J.-Y. L'Excellent, and J.~Koster}, {\em A fully
-  asynchronous multifrontal solver using distributed dynamic scheduling}, SIAM
-  Journal on Matrix Analysis and Applications, 23 (2001), pp.~15--41,
-  \url{http://mumps.enseeiht.fr}.
-
-\bibitem{anderson2019mfem}
-{\sc R.~Anderson, J.~Andrej, A.~Barker, J.~Bramwell, J.-S. Camier, J.~Cerveny,
-  V.~Dobrev, Y.~Dudouit, A.~Fisher, T.~Kolev, W.~Pazner, M.~Stowell, V.~Tomov,
-  I.~Akkerman, J.~Dahm, D.~Medina, and S.~Zampini}, {\em {MFEM: A modular
-  finite element methods library}}, Computers \& Mathematics with Applications,
-   (2020), \url{http://mfem.org}.
-
-\bibitem{badia2014highly}
-{\sc S.~Badia, A.~F. Mart{\'\i}n, and J.~Principe}, {\em A highly scalable
-  parallel implementation of balancing domain decomposition by constraints},
-  SIAM Journal on Scientific Computing, 36 (2014), pp.~C190--C218.
-
-\bibitem{baker2005technique}
-{\sc A.~H. Baker, E.~R. Jessup, and T.~Manteuffel}, {\em {A Technique for
-  Accelerating the Convergence of Restarted GMRES}}, SIAM Journal on Matrix
-  Analysis and Applications, 26 (2005), pp.~962--984.
-
-\bibitem{petsc-user-ref}
-{\sc S.~Balay, S.~Abhyankar, M.~F. Adams, J.~Brown, P.~Brune, K.~Buschelman,
+\expandafter\ifx\csname url\endcsname\relax
+  \def\url#1{\texttt{#1}}\fi
+\expandafter\ifx\csname urlprefix\endcsname\relax\def\urlprefix{URL }\fi
+\expandafter\ifx\csname href\endcsname\relax
+  \def\href#1#2{#2} \def\path#1{#1}\fi
+
+\bibitem{Bartlett:2017:XFT}
+R.~Bartlett, I.~Demeshko, T.~Gamblin, G.~Hammond, M.~A. Heroux, J.~Johnson,
+  A.~Klinvex, X.~S. Li, L.~C. McInnes, J.~D. Moulton, D.~Osei-Kuffuor,
+  J.~Sarich, B.~F. Smith, J.~Willenbring, U.~M. Yang,
+  \href{https://xsdk.info}{{xSDK} foundations: Toward an extreme-scale
+  scientific software development kit}, Supercomputing Frontiers and
+  Innovations 4~(1) (2017).
+\newline\urlprefix\url{https://xsdk.info}
+
+\bibitem{petsc-web-page}
+S.~Balay, S.~Abhyankar, M.~F. Adams, J.~Brown, P.~Brune, K.~Buschelman,
   L.~Dalcin, A.~Dener, V.~Eijkhout, W.~D. Gropp, D.~Karpeyev, D.~Kaushik, M.~G.
   Knepley, D.~A. May, L.~C. McInnes, R.~T. Mills, T.~Munson, K.~Rupp, P.~Sanan,
-  B.~F. Smith, S.~Zampini, H.~Zhang, and H.~Zhang}, {\em {PETS}c users manual},
-  Tech. Report ANL-95/11 - Revision 3.13, Argonne National Laboratory, 2020.
-
-\bibitem{petsc-web-page}
-{\sc S.~Balay, S.~Abhyankar, M.~F. Adams, J.~Brown, P.~Brune, K.~Buschelman,
+  B.~F. Smith, S.~Zampini, H.~Zhang, H.~Zhang,
+  \href{http://www.mcs.anl.gov/petsc}{{PETS}c web page} (2020).
+\newline\urlprefix\url{http://www.mcs.anl.gov/petsc}
+
+\bibitem{petsc-user-ref}
+S.~Balay, S.~Abhyankar, M.~F. Adams, J.~Brown, P.~Brune, K.~Buschelman,
   L.~Dalcin, A.~Dener, V.~Eijkhout, W.~D. Gropp, D.~Karpeyev, D.~Kaushik, M.~G.
   Knepley, D.~A. May, L.~C. McInnes, R.~T. Mills, T.~Munson, K.~Rupp, P.~Sanan,
-  B.~F. Smith, S.~Zampini, H.~Zhang, and H.~Zhang}, {\em {PETS}c web page},
-  2020, \url{http://www.mcs.anl.gov/petsc}.
+  B.~F. Smith, S.~Zampini, H.~Zhang, H.~Zhang, {PETS}c users manual, Tech. Rep.
+  ANL-95/11 - Revision 3.13, Argonne National Laboratory (2020).
+
+\bibitem{falgout2002hypre}
+R.~Falgout, U.~M. Yang, \href{https://www.llnl.gov/casc/hypre}{\emph{hypre}: A
+  library of high performance preconditioners}, Computational Science---ICCS
+  2002 (2002) 632--641.
+\newline\urlprefix\url{https://www.llnl.gov/casc/hypre}
+
+\bibitem{si2013tet}
+H.~Si, \href{http://wias-berlin.de/software/tetgen}{{TetGen}: A quality
+  tetrahedral mesh generator and {3D D}elaunay triangulator}, Tech. Rep.~13
+  (2013).
+\newline\urlprefix\url{http://wias-berlin.de/software/tetgen}
+
+\bibitem{BursteddeWilcoxGhattas11}
+C.~Burstedde, L.~C. Wilcox, O.~Ghattas,
+  \href{http://www.p4est.org}{{\texttt{p4est}}: Scalable algorithms for
+  parallel adaptive mesh refinement on forests of octrees}, SIAM Journal on
+  Scientific Computing 33~(3) (2011) 1103--1133.
+\newline\urlprefix\url{http://www.p4est.org}
+
+\bibitem{jolivet2013scalable}
+P.~Jolivet, F.~Hecht, F.~Nataf, C.~Prud'homme, Scalable domain decomposition
+  preconditioners for heterogeneous elliptic problems, in: Proceedings of the
+  International Conference on High Performance Computing, Networking, Storage
+  and Analysis, SC13, ACM, 2013.
+
+\bibitem{jolivet2016block}
+P.~Jolivet, P.-H. Tournier, Block iterative methods and recycling for improved
+  scalability of linear solvers, in: Proceedings of the 2016 International
+  Conference for High Performance Computing, Networking, Storage and Analysis,
+  SC16, IEEE, 2016.
+
+\bibitem{hoefler2015scientific}
+T.~Hoefler, R.~Belli, Scientific benchmarking of parallel computing systems:
+  Twelve ways to tell the masses when reporting performance results, in:
+  Proceedings of the 2015 International Conference for High Performance
+  Computing, Networking, Storage and Analysis, SC15, 2015.
+
+\bibitem{hernandez2005ssf}
+V.~Hernandez, J.~E. Roman, V.~Vidal, \href{https://slepc.upv.es}{{SLEP}c: A
+  scalable and flexible toolkit for the solution of eigenvalue problems}, ACM
+  Transactions on Mathematical Software 31~(3) (2005) 351--362.
+\newline\urlprefix\url{https://slepc.upv.es}
 
 \bibitem{barrett1994templates}
-{\sc R.~Barrett, M.~Berry, T.~F. Chan, J.~Demmel, J.~Donato, J.~Dongarra,
-  V.~Eijkhout, R.~Pozo, C.~Romine, and H.~Van~der Vorst}, {\em Templates for
-  the solution of linear systems: building blocks for iterative methods}, SIAM,
-  1994.
-
-\bibitem{Bartlett:2017:XFT}
-{\sc R.~Bartlett, I.~Demeshko, T.~Gamblin, G.~Hammond, M.~A. Heroux,
-  J.~Johnson, A.~Klinvex, X.~Li, L.~C. McInnes, J.~D. Moulton, D.~Osei-Kuffuor,
-  J.~Sarich, B.~Smith, J.~Willenbring, and U.~M. Yang}, {\em {xSDK Foundations:
-  Toward an Extreme-scale Scientific Software Development Kit}}, Supercomputing
-  Frontiers and Innovations, 4 (2017).
+R.~Barrett, M.~Berry, T.~F. Chan, J.~Demmel, J.~Donato, J.~Dongarra,
+  V.~Eijkhout, R.~Pozo, C.~Romine, H.~Van~der Vorst, Templates for the Solution
+  of Linear Systems: Building Blocks for Iterative Methods, SIAM, 1994.
+
+\bibitem{saad2003iterative}
+Y.~Saad, Iterative Methods for Sparse Linear Systems, 2nd Edition, SIAM, 2003.
+
+\bibitem{saad1986gmres}
+Y.~Saad, M.~H. Schultz, {GMRES}: A generalized minimal residual algorithm for
+  solving nonsymmetric linear systems, SIAM Journal on Scientific and
+  Statistical Computing 7~(3) (1986) 856--869.
+
+\bibitem{hestenes1952methods}
+M.~R. Hestenes, E.~Stiefel, Methods of conjugate gradients for solving linear
+  systems, Journal of Research of the National Bureau of Standards 49~(6)
+  (1952) 409--436.
+
+\bibitem{van1992bi}
+H.~A. Van~der Vorst, {Bi-CGSTAB}: A fast and smoothly converging variant of
+  {Bi-CG} for the solution of nonsymmetric linear systems, SIAM Journal on
+  Scientific and Statistical Computing 13~(2) (1992) 631--644.
+
+\bibitem{baker2005technique}
+A.~H. Baker, E.~R. Jessup, T.~Manteuffel, A technique for accelerating the
+  convergence of restarted {GMRES}, SIAM Journal on Matrix Analysis and
+  Applications 26~(4) (2005) 962--984.
+
+\bibitem{erhel1996restarted}
+J.~Erhel, K.~Burrage, B.~Pohl, Restarted {GMRES} preconditioned by deflation,
+  Journal of Computational and Applied Mathematics 69~(2) (1996) 303--318.
+
+\bibitem{wakam2013158}
+D.~N. Wakam, F.~Pacull, Memory efficient hybrid algebraic solvers for linear
+  systems arising from compressible flows, Computers \& Fluids 80 (2013)
+  158--167.
+
+\bibitem{saad1993flexible}
+Y.~Saad, A flexible inner--outer preconditioned {GMRES} algorithm, SIAM Journal
+  on Scientific Computing 14~(2) (1993) 461--469.
+
+\bibitem{eisenstat1983variational}
+S.~C. Eisenstat, H.~C. Elman, M.~H. Schultz, Variational iterative methods for
+  nonsymmetric systems of linear equations, SIAM Journal on Numerical Analysis
+  20~(2) (1983) 345--357.
+
+\bibitem{parks2006recycling}
+M.~L. Parks, E.~de~Sturler, G.~Mackey, D.~D. Johnson, S.~Maiti, Recycling
+  {K}rylov subspaces for sequences of linear systems, SIAM Journal on
+  Scientific Computing 28~(5) (2006) 1651--1674.
+
+\bibitem{gutknecht2006block}
+M.~H. Gutknecht, Block {K}rylov space methods for linear systems with multiple
+  right-hand sides: An introduction, in: A.~Siddiqui, I.~Duff, O.~Christensen
+  (Eds.), Modern Mathematical Models, Methods and Algorithms for Real World
+  Systems, 2006, pp. 420--447.
+
+\bibitem{frommer2017block}
+A.~Frommer, K.~Lund, D.~B. Szyld, Block {K}rylov subspace methods for functions
+  of matrices, Electronic Transactions on Numerical Analysis 47 (2017)
+  100--126.
+
+\bibitem{tournier2016micro}
+P.-H. Tournier, I.~Aliferis, M.~Bonazzoli, M.~de~Buhan, M.~Darbas, V.~Dolean,
+  F.~Hecht, P.~Jolivet, I.~El~Kanfoud, C.~Migliaccio, F.~Nataf, C.~Pichot,
+  S.~Semenov, Microwave tomographic imaging of cerebrovascular accidents by
+  using high-performance computing, Parallel Computing 85 (2019) 88--97.
+
+\bibitem{KALANTZIS2018136}
+V.~Kalantzis, A.~C.~I. Malossi, C.~Bekas, A.~Curioni, E.~Gallopoulos, Y.~Saad,
+  A scalable iterative dense linear system solver for multiple right-hand sides
+  in data analytics, Parallel Computing 74 (2018) 136--153.
+
+\bibitem{knyazev2001toward}
+A.~V. Knyazev, Toward the optimal preconditioned eigensolver: Locally optimal
+  block preconditioned conjugate gradient method, SIAM Journal on Scientific
+  Computing 23~(2) (2001) 517--541.
+
+\bibitem{calandra2012flexible}
+H.~Calandra, S.~Gratton, J.~Langou, X.~Pinel, X.~Vasseur, Flexible variants of
+  block restarted {GMRES} methods with application to geophysics, SIAM Journal
+  on Scientific Computing 34~(2) (2012) A714--A736.
+
+\bibitem{sakurai2010application}
+T.~Sakurai, H.~Tadano, Y.~Kuramashi, Application of block {K}rylov subspace
+  algorithms to the {W}ilson--{D}irac equation with multiple right-hand sides
+  in lattice {QCD}, Computer Physics Communications 181~(1) (2010) 113--117.
+
+\bibitem{heroux2005overview}
+M.~A. Heroux, R.~A. Bartlett, V.~E. Howle, R.~J. Hoekstra, J.~J. Hu, T.~G.
+  Kolda, R.~B. Lehoucq, K.~R. Long, R.~P. Pawlowski, E.~T. Phipps, et~al.,
+  \href{https://trilinos.github.io}{An overview of the {T}rilinos project}, ACM
+  Transactions on Mathematical Software (TOMS) 31~(3) (2005) 397--423.
+\newline\urlprefix\url{https://trilinos.github.io}
+
+\bibitem{bavier2012amesos2}
+E.~Bavier, M.~Hoemmen, S.~Rajamanickam, H.~Thornquist, {Amesos2} and {Belos}:
+  Direct and iterative solvers for large sparse linear systems, Scientific
+  Programming 20~(3) (2012) 241--255.
+
+\bibitem{notay2000flexible}
+Y.~Notay, Flexible conjugate gradients, SIAM Journal on Scientific Computing
+  22~(4) (2000) 1444--1460.
+
+\bibitem{carvalho2011flexible}
+L.~M. Carvalho, S.~Gratton, R.~Lago, X.~Vasseur, A flexible generalized
+  conjugate residual method with inner orthogonalization and deflated
+  restarting, SIAM Journal on Matrix Analysis and Applications 32~(4) (2011)
+  1212--1235.
+
+\bibitem{o1980block}
+D.~P. O'Leary, The block conjugate gradient algorithm and related methods,
+  Linear Algebra and its Applications 29 (1980) 293--322.
+
+\bibitem{ji2017breakdown}
+H.~Ji, Y.~Li, A breakdown-free block conjugate gradient method, BIT Numerical
+  Mathematics 57~(2) (2017) 379--403.
+
+\bibitem{saad2000deflated}
+Y.~Saad, M.~Yeung, J.~Erhel, F.~Guyomarc'h, A deflated version of the conjugate
+  gradient algorithm, SIAM Journal on Scientific Computing 21~(5) (2000)
+  1909--1926.
+
+\bibitem{stathopoulos2009deflation}
+A.~Stathopoulos, A.~Abdel-Rehim, K.~Orginos, Deflation for inversion with
+  multiple right-hand sides in {QCD}, in: Journal of Physics: Conference
+  Series, Vol. 180, IOP Publishing, 2009.
+
+\bibitem{soodhalter2014krylov}
+K.~M. Soodhalter, D.~B. Szyld, F.~Xue, Krylov subspace recycling for sequences
+  of shifted linear systems, Applied Numerical Mathematics 81 (2014) 105--118.
+
+\bibitem{amestoy2001fully}
+P.~Amestoy, I.~Duff, J.-Y. L'Excellent, J.~Koster,
+  \href{http://mumps.enseeiht.fr}{A fully asynchronous multifrontal solver
+  using distributed dynamic scheduling}, SIAM Journal on Matrix Analysis and
+  Applications 23~(1) (2001) 15--41.
+\newline\urlprefix\url{http://mumps.enseeiht.fr}
+
+\bibitem{stathopoulos2002block}
+A.~Stathopoulos, K.~Wu, A block orthogonalization procedure with constant
+  synchronization requirements, SIAM Journal on Scientific Computing 23~(6)
+  (2002) 2165--2182.
+
+\bibitem{gutknecht2008updating}
+M.~H. Gutknecht, T.~Schmelzer, Updating the {$QR$} decomposition of block
+  tridiagonal and block {H}essenberg matrices, {Applied Numerical Mathematics}
+  58~(6) (2008) 871--883.
+
+\bibitem{cusparse-web-page}
+NVIDIA, {cuSPARSE} web page, \url{https://docs.nvidia.com/cuda/cusparse}
+  (2020).
+
+\bibitem{10.1145/3232850}
+W.~Boukaram, G.~Turkiyyah, D.~Keyes, Hierarchical matrix operations on {GPU}s:
+  Matrix--vector multiplication and compression, ACM Transactions on
+  Mathematical Software 45 (2019).
+
+\bibitem{ambartsumyan2020hierarchical}
+I.~Ambartsumyan, W.~Boukaram, T.~Bui-Thanh, O.~Ghattas, D.~Keyes, G.~Stadler,
+  G.~Turkiyyah, S.~Zampini, Hierarchical matrix approximations of {H}essians
+  arising in inverse problems governed by {PDE}s, arXiv:2003.10173 (2020).
+
+\bibitem{davis2004algorithm}
+T.~A. Davis, Algorithm 832: {UMFPACK}—an unsymmetric-pattern multifrontal
+  method, ACM Transactions on Mathematical Software 30~(2) (2004) 196--199.
+
+\bibitem{mkl-web-page}
+Intel, {MKL} web page,
+  \url{https://software.intel.com/content/www/us/en/develop/tools/math-kernel-library.html}
+  (2020).
+
+\bibitem{li2005overview}
+X.~S. Li, An overview of {S}uper{LU}: Algorithms, implementation, and user
+  interface, ACM Transactions on Mathematical Software 31~(3) (2005) 302--325.
+
+\bibitem{li2003superlu_dist}
+X.~S. Li, J.~Demmel, {S}uper{LU\_DIST}: A scalable distributed-memory sparse
+  direct solver for unsymmetric linear systems, ACM Transactions on
+  Mathematical Software 29~(2) (2003) 110--140.
+
+\bibitem{poulson2013elemental}
+J.~Poulson, B.~Marker, R.~A. Van~de Geijn, J.~R. Hammond, N.~A. Romero,
+  Elemental: A new framework for distributed memory dense matrix computations,
+  ACM Transactions on Mathematical Software 39~(2) (2013).
+
+\bibitem{heinecke2016libxsmm}
+A.~Heinecke, G.~Henry, M.~Hutchinson, H.~Pabst,
+  \href{https://github.com/hfp/libxsmm}{{LIBXSMM}: Accelerating small matrix
+  multiplications by runtime code generation}, in: Proceedings of the 2016
+  International Conference for High Performance Computing, Networking, Storage
+  and Analysis, SC16, IEEE, 2016.
+\newline\urlprefix\url{https://github.com/hfp/libxsmm}
+
+\bibitem{hecht2012new}
+F.~Hecht, \href{http://freefem.org}{New development in {FreeFem++}}, Journal of
+  Numerical Mathematics 20~(3-4) (2012) 251--266.
+\newline\urlprefix\url{http://freefem.org}
+
+\bibitem{moulin2018al}
+J.~Moulin, P.~Jolivet, O.~Marquet,
+  \href{https://github.com/prj-/moulin2019al}{Augmented {L}agrangian
+  preconditioner for large-scale hydrodynamic stability analysis}, Computer
+  Methods in Applied Mechanics and Engineering 351 (2019) 718--743.
+\newline\urlprefix\url{https://github.com/prj-/moulin2019al}
+
+\bibitem{Stewart2002}
+G.~W. Stewart, A {K}rylov--{S}chur algorithm for large eigenproblems, SIAM
+  Journal on Matrix Analysis and Applications 23~(3) (2002) 601--614.
+
+\bibitem{Benzi2011a}
+M.~Benzi, M.~A. Olshanskii, Z.~Wang, Modified augmented {L}agrangian
+  preconditioners for the incompressible {N}avier--{S}tokes equations,
+  International Journal for Numerical Methods in Fluids 66~(4) (2011) 486--508.
+
+\bibitem{sakurai2003projection}
+T.~Sakurai, H.~Sugiura, A projection method for generalized eigenvalue problems
+  using numerical integration, Journal of Computational and Applied Mathematics
+  159~(1) (2003) 119--128.
+
+\bibitem{adams2004ultra}
+M.~Adams, H.~H. Bayraktar, T.~M. Keaveny, P.~Papadopoulos, Ultrascalable
+  implicit finite element analyses in solid mechanics with over a half a
+  billion degrees of freedom, in: Proceedings of the 2004 ACM/IEEE Conference
+  on Supercomputing, SC04, IEEE Computer Society, 2004, pp. \mbox{34:1--34:15}.
+
+\bibitem{smith2004domain}
+B.~F. Smith, P.~Bj{\o}rstad, W.~D. Gropp, Domain Decomposition: Parallel
+  Multilevel Methods for Elliptic Partial Differential Equations, Cambridge
+  University Press, 2004.
+
+\bibitem{dolean2014ddm}
+V.~Dolean, P.~Jolivet, F.~Nataf, An Introduction to Domain Decomposition
+  Methods\col\ Algorithms, Theory and Parallel Implementation, SIAM, 2015.
+
+\bibitem{toselli2005domain}
+A.~Toselli, O.~B. Widlund, Domain decomposition methods: algorithms and theory,
+  Vol.~34 of Series in Computational Mathematics, Springer, 2005.
+
+\bibitem{pechstein2012finite}
+C.~Pechstein, Finite and Boundary Element Tearing and Interconnecting Solvers
+  for Multiscale Problems, Vol.~90 of Lecture Notes in Computational Science
+  and Engineering, Springer, 2012.
+
+\bibitem{quarteroni1999domain}
+A.~Quarteroni, A.~Valli, Domain Decomposition Methods for Partial Differential
+  Equations, Vol.~10, Clarendon Press, 1999.
+
+\bibitem{mathew2008domain}
+T.~Mathew, Domain Decomposition Methods for the Numerical Solution of Partial
+  Differential Equations, Vol.~61, Springer Science \& Business Media, 2008.
+
+\bibitem{cai1999restricted}
+X.-C. Cai, M.~Sarkis, A restricted additive {S}chwarz preconditioner for
+  general sparse linear systems, SIAM Journal on Scientific Computing 21~(2)
+  (1999) 792--797.
+
+\bibitem{mandel1993balancing}
+J.~Mandel, Balancing domain decomposition, Communications in Numerical Methods
+  in Engineering 9~(3) (1993) 233--241.
+
+\bibitem{zampini2016pcbddc}
+S.~Zampini,
+  \href{https://www.mcs.anl.gov/petsc/petsc-master/docs/manualpages/PC/PCBDDC.html}{{PCBDDC}:
+  A class of robust dual--primal methods in {PETS}c}, SIAM Journal on
+  Scientific Computing 38~(5) (2016) S282--S306.
+\newline\urlprefix\url{https://www.mcs.anl.gov/petsc/petsc-master/docs/manualpages/PC/PCBDDC.html}
+
+\bibitem{pechstein2017unified}
+C.~Pechstein, C.~R. Dohrmann, A unified framework for adaptive {BDDC},
+  Electronic Transactions on Numerical Analysis 46 (2017) 273--336.
+
+\bibitem{zampini2017balancing}
+S.~Zampini, P.~Vassilevski, V.~Dobrev, T.~Kolev, Balancing domain decomposition
+  by constraints algorithms for curl-conforming spaces of arbitrary order, in:
+  International Conference on Domain Decomposition Methods, Springer, 2017, pp.
+  103--116.
+
+\bibitem{oh2018bddc}
+D.-S. Oh, O.~B. Widlund, S.~Zampini, C.~Dohrmann, {BDDC} algorithms with deluxe
+  scaling and adaptive selection of primal constraints for {R}aviart--{T}homas
+  vector fields, Mathematics of Computation 87~(310) (2018) 659--692.
+
+\bibitem{da2014isogeometric}
+L.~B. Da~Veiga, L.~F. Pavarino, S.~Scacchi, O.~B. Widlund, S.~Zampini,
+  Isogeometric {BDDC} preconditioners with deluxe scaling, SIAM Journal on
+  Scientific Computing 36~(3) (2014) A1118--A1139.
+
+\bibitem{gee2006ml}
+M.~W. Gee, C.~M. Siefert, J.~J. Hu, R.~S. Tuminaro, M.~G. Sala,
+  \href{https://trilinos.github.io/ml.html}{{ML} 5.0 smoothed aggregation
+  user's guide}, Tech. Rep. SAND2006-2649, Sandia National Laboratories (2006).
+\newline\urlprefix\url{https://trilinos.github.io/ml.html}
+
+\bibitem{Heinlein:2016:PIT}
+A.~Heinlein, A.~Klawonn, O.~Rheinbach, A parallel implementation of a two-level
+  overlapping {S}chwarz method with energy-minimizing coarse space based on
+  {T}rilinos, SIAM Journal on Scientific Computing 38~(6) (2016) C713--C747.
+
+\bibitem{vsistek2013parallel}
+J.~{\v{S}}{\'\i}stek, J.~Mandel, B.~Soused{\'\i}k, P.~Burda,
+  \href{http://users.math.cas.cz/~sistek/software/bddcml.html}{Parallel
+  implementation of multilevel {BDDC}}, in: Numerical Mathematics and Advanced
+  Applications 2011, Springer, 2013, pp. 681--689.
+\newline\urlprefix\url{http://users.math.cas.cz/~sistek/software/bddcml.html}
+
+\bibitem{badia2014highly}
+S.~Badia, A.~F. Mart{\'\i}n, J.~Principe, A highly scalable parallel
+  implementation of balancing domain decomposition by constraints, SIAM Journal
+  on Scientific Computing 36~(2) (2014) C190--C218.
+
+\bibitem{gander2006optimized}
+M.~J. Gander, Optimized {S}chwarz {M}ethods, SIAM Journal on Numerical Analysis
+  44~(2) (2006) 699--731.
+
+\bibitem{gander2019new}
+M.~J. Gander, S.~Van~Criekingen, New coarse corrections for optimized
+  restricted additive {S}chwarz using {PETS}c, in: International Conference on
+  Domain Decomposition Methods, Springer, 2019.
+
+\bibitem{spillane2011robust}
+N.~Spillane, V.~Dolean, P.~Hauret, F.~Nataf, C.~Pechstein, R.~Scheichl, A
+  robust two-level domain decomposition preconditioner for systems of {PDE}s,
+  Comptes Rendus Mathématique 349~(23) (2011) 1255--1259.
+
+\bibitem{spillane2013abstract}
+N.~Spillane, V.~Dolean, P.~Hauret, F.~Nataf, C.~Pechstein, R.~Scheichl,
+  Abstract robust coarse spaces for systems of {PDEs} via generalized
+  eigenproblems in the overlaps, Numerische Mathematik 126~(4) (2013) 741--700.
+
+\bibitem{marchand2019two}
+P.~Marchand, X.~Claeys, P.~Jolivet, F.~Nataf, P.-H. Tournier, Two-level
+  preconditioning for $h$-version boundary element approximation of
+  hypersingular operator with {GenEO}, Numerische Mathematik (2019) submitted
+  for publication.
+
+\bibitem{brown2012composable}
+J.~Brown, M.~G. Knepley, D.~A. May, L.~Curfman~McInnes, B.~F. Smith, Composable
+  linear solvers for multiphysics, in: 2012 11th International Symposium on
+  Parallel and Distributed Computing, IEEE, 2012, pp. 55--62.
+
+\bibitem{butler2020high}
+R.~Butler, T.~Dodwell, A.~Reinarz, A.~Sandhu, R.~Scheichl, L.~Seelinger,
+  High-performance \textsc{Dune} modules for solving large-scale, strongly
+  anisotropic elliptic problems with applications to aerospace composites,
+  Computer Physics Communications 249 (2020).
+
+\bibitem{may2016extreme}
+D.~A. May, P.~Sanan, K.~Rupp, M.~G. Knepley, B.~F. Smith, Extreme-scale
+  multigrid components within {PETS}c, in: Proceedings of the Platform for
+  Advanced Scientific Computing Conference, 2016.
+
+\bibitem{knepley2009mesh}
+M.~G. Knepley, D.~A. Karpeev, Mesh algorithms for {PDE} with {S}ieve {I}: Mesh
+  distribution, Scientific Programming 17~(3) (2009) 215--230.
+
+\bibitem{tang2009comparison}
+J.~Tang, R.~Nabben, C.~Vuik, Y.~Erlangga, Comparison of two-level
+  preconditioners derived from deflation, domain decomposition and multigrid
+  methods, Journal of Scientific Computing 39~(3) (2009) 340--370.
 
 \bibitem{bastian1998additive}
-{\sc P.~Bastian, G.~Wittum, and W.~Hackbusch}, {\em Additive and multiplicative
-  multi-grid—a comparison}, Computing, 60 (1998), pp.~345--364.
-
-\bibitem{bavier2012amesos2}
-{\sc E.~Bavier, M.~Hoemmen, S.~Rajamanickam, and H.~Thornquist}, {\em {Amesos2
-  and Belos: Direct and Iterative Solvers for Large Sparse Linear Systems}},
-  Scientific Programming, 20 (2012), pp.~241--255.
-
-\bibitem{Benzi2011a}
-{\sc M.~Benzi, M.~A. Olshanskii, and Z.~Wang}, {\em {Modified augmented
-  Lagrangian preconditioners for the incompressible Navier--Stokes equations}},
-  International Journal for Numerical Methods in Fluids, 66 (2011),
-  pp.~486--508.
-
-\bibitem{10.1145/3232850}
-{\sc W.~Boukaram, G.~Turkiyyah, and D.~Keyes}, {\em {Hierarchical Matrix
-  Operations on GPUs: Matrix-Vector Multiplication and Compression}}, ACM
-  Transactions on Mathematical Software, 45 (2019).
-
-\bibitem{brown2012composable}
-{\sc J.~Brown, M.~G. Knepley, D.~A. May, L.~Curfman~McInnes, and B.~F. Smith},
-  {\em Composable linear solvers for multiphysics}, in 2012 11th International
-  Symposium on Parallel and Distributed Computing, IEEE, 2012, pp.~55--62.
-
-\bibitem{BursteddeWilcoxGhattas11}
-{\sc C.~Burstedde, L.~C. Wilcox, and O.~Ghattas}, {\em {\texttt{p4est}}:
-  Scalable algorithms for parallel adaptive mesh refinement on forests of
-  octrees}, SIAM Journal on Scientific Computing, 33 (2011), pp.~1103--1133,
-  \url{http://www.p4est.org}.
-
-\bibitem{butler2020high}
-{\sc R.~Butler, T.~Dodwell, A.~Reinarz, A.~Sandhu, R.~Scheichl, and
-  L.~Seelinger}, {\em High-performance dune modules for solving large-scale,
-  strongly anisotropic elliptic problems with applications to aerospace
-  composites}, Computer Physics Communications, 249 (2020).
-
-\bibitem{cai1999restricted}
-{\sc X.-C. Cai and M.~Sarkis}, {\em {A restricted additive Schwarz
-  preconditioner for general sparse linear systems}}, SIAM Journal on
-  Scientific Computing, 21 (1999), pp.~792--797.
-
-\bibitem{calandra2012flexible}
-{\sc H.~Calandra, S.~Gratton, J.~Langou, X.~Pinel, and X.~Vasseur}, {\em
-  {Flexible Variants of Block Restarted GMRES Methods with Application to
-  Geophysics}}, SIAM Journal on Scientific Computing, 34 (2012),
-  pp.~A714--A736.
-
-\bibitem{carvalho2011flexible}
-{\sc L.~M. Carvalho, S.~Gratton, R.~Lago, and X.~Vasseur}, {\em {A Flexible
-  Generalized Conjugate Residual Method With Inner Orthogonalization and
-  Deflated Restarting}}, SIAM Journal on Matrix Analysis and Applications, 32
-  (2011), pp.~1212--1235.
+P.~Bastian, G.~Wittum, W.~Hackbusch, Additive and multiplicative multi-grid—a
+  comparison, Computing 60~(4) (1998) 345--364.
+
+\bibitem{haferssas2017soras}
+R.~Haferssas, P.~Jolivet, F.~Nataf, An additive {S}chwarz method type theory
+  for {L}ions's algorithm and a symmetrized optimized restricted additive
+  {S}chwarz method, Journal on Scientific Computing 39~(4) (2017) A1345--A1365.
 
 \bibitem{conen2014coarse}
-{\sc L.~Conen, V.~Dolean, R.~Krause, and F.~Nataf}, {\em {A coarse space for
-  heterogeneous Helmholtz problems based on the Dirichlet-to-Neumann
-  operator}}, Journal of Computational and Applied Mathematics, 271 (2014),
-  pp.~83--99.
-
-\bibitem{da2014isogeometric}
-{\sc L.~B. Da~Veiga, L.~F. Pavarino, S.~Scacchi, O.~B. Widlund, and
-  S.~Zampini}, {\em {Isogeometric BDDC preconditioners with deluxe scaling}},
-  SIAM Journal on Scientific Computing, 36 (2014), pp.~A1118--A1139.
-
-\bibitem{davis2004algorithm}
-{\sc T.~A. Davis}, {\em Algorithm 832: {UMFPACK}—an unsymmetric-pattern
-  multifrontal method}, ACM Transactions on Mathematical Software, 30 (2004),
-  pp.~196--199.
-
-\bibitem{dolean2014ddm}
-{\sc V.~Dolean, P.~Jolivet, and F.~Nataf}, {\em An Introduction to Domain
-  Decomposition Methods\col\ Algorithms, Theory and Parallel Implementation},
-  SIAM, 2015.
-
-\bibitem{erhel1996restarted}
-{\sc J.~Erhel, K.~Burrage, and B.~Pohl}, {\em {Restarted GMRES preconditioned
-  by deflation}}, Journal of Computational and Applied Mathematics, 69 (1996),
-  pp.~303--318.
-
-\bibitem{falgout2002hypre}
-{\sc R.~Falgout and U.~M. Yang}, {\em \emph{hypre}: a library of high
-  performance preconditioners}, Computational Science---ICCS 2002,  (2002),
-  pp.~632--641, \url{https://www.llnl.gov/casc/hypre}.
-
-\bibitem{frommer2017block}
-{\sc A.~Frommer, K.~Lund, and D.~B. Szyld}, {\em {Block Krylov Subspace Methods
-  for Functions of Matrices}}, Electronic Transactions on Numerical Analysis,
-  47 (2017), pp.~100--126.
-
-\bibitem{gander2006optimized}
-{\sc M.~J. Gander}, {\em Optimized {S}chwarz {M}ethods}, SIAM Journal on
-  Numerical Analysis, 44 (2006), pp.~699--731.
-
-\bibitem{gander2019new}
-{\sc M.~J. Gander and S.~Van~Criekingen}, {\em {New Coarse Corrections for
-  Optimized Restricted Additive Schwarz Using PETSc}}, in International
-  Conference on Domain Decomposition Methods, Springer, 2019.
-
-\bibitem{gee2006ml}
-{\sc M.~W. Gee, C.~M. Siefert, J.~J. Hu, R.~S. Tuminaro, and M.~G. Sala}, {\em
-  {ML} 5.0 smoothed aggregation user's guide}, Tech. Report SAND2006-2649,
-  Sandia National Laboratories, 2006, \url{https://trilinos.github.io/ml.html}.
-
-\bibitem{gutknecht2006block}
-{\sc M.~H. Gutknecht}, {\em Block {K}rylov space methods for linear systems
-  with multiple right-hand sides: an introduction}, in Modern Mathematical
-  Models, Methods and Algorithms for Real World Systems, A.~Siddiqui, I.~Duff,
-  and O.~Christensen, eds., 2006, pp.~420--447.
-
-\bibitem{gutknecht2008updating}
-{\sc M.~H. Gutknecht and T.~Schmelzer}, {\em {Updating the $QR$ decomposition
-  of block tridiagonal and block Hessenberg matrices}}, {Applied Numerical
-  Mathematics}, 58 (2008), pp.~871--883.
-
-\bibitem{haferssas2017soras}
-{\sc R.~Haferssas, P.~Jolivet, and F.~Nataf}, {\em {An additive Schwarz method
-  type theory for Lions's algorithm and a Symmetrized Optimized Restricted
-  Additive Schwarz Method}}, Journal on Scientific Computing, 39 (2017),
-  pp.~A1345--A1365.
-
-\bibitem{hecht2012new}
-{\sc F.~Hecht}, {\em New development in {FreeFem++}}, Journal of Numerical
-  Mathematics, 20 (2012), pp.~251--266, \url{http://freefem.org}.
-
-\bibitem{heinecke2016libxsmm}
-{\sc A.~Heinecke, G.~Henry, M.~Hutchinson, and H.~Pabst}, {\em {LIBXSMM:
-  accelerating small matrix multiplications by runtime code generation}}, in
-  Proceedings of the 2016 International Conference for High Performance
-  Computing, Networking, Storage and Analysis, SC16, IEEE, 2016,
-  \url{https://github.com/hfp/libxsmm}.
-
-\bibitem{Heinlein:2016:PIT}
-{\sc A.~Heinlein, A.~Klawonn, and O.~Rheinbach}, {\em A parallel implementation
-  of a two-level overlapping {S}chwarz method with energy-minimizing coarse
-  space based on {T}rilinos}, SIAM Journal on Scientific Computing, 38 (2016),
-  pp.~C713--C747.
-
-\bibitem{hernandez2005ssf}
-{\sc V.~Hernandez, J.~E. Roman, and V.~Vidal}, {\em {SLEP}c: A scalable and
-  flexible toolkit for the solution of eigenvalue problems}, ACM Transactions
-  on Mathematical Software, 31 (2005), pp.~351--362,
-  \url{https://slepc.upv.es}.
-
-\bibitem{heroux2005overview}
-{\sc M.~A. Heroux, R.~A. Bartlett, V.~E. Howle, R.~J. Hoekstra, J.~J. Hu, T.~G.
-  Kolda, R.~B. Lehoucq, K.~R. Long, R.~P. Pawlowski, E.~T. Phipps, et~al.},
-  {\em {An overview of the Trilinos project}}, ACM Transactions on Mathematical
-  Software (TOMS), 31 (2005), pp.~397--423, \url{https://trilinos.github.io}.
-
-\bibitem{hestenes1952methods}
-{\sc M.~R. Hestenes and E.~Stiefel}, {\em {Methods of Conjugate Gradients for
-  Solving Linear Systems}}, Journal of Research of the National Bureau of
-  Standards, 49 (1952), pp.~409--436.
-
-\bibitem{hoefler2015scientific}
-{\sc T.~Hoefler and R.~Belli}, {\em Scientific benchmarking of parallel
-  computing systems: twelve ways to tell the masses when reporting performance
-  results}, in Proceedings of the 2015 International Conference for High
-  Performance Computing, Networking, Storage and Analysis, SC15, 2015.
-
-\bibitem{mkl-web-page}
-{\sc Intel}, {\em {MKL} web page}.
-\newblock
-  \url{https://software.intel.com/content/www/us/en/develop/tools/math-kernel-library.html},
-  2020.
-
-\bibitem{ji2017breakdown}
-{\sc H.~Ji and Y.~Li}, {\em A breakdown-free block conjugate gradient method},
-  BIT Numerical Mathematics, 57 (2017), pp.~379--403.
-
-\bibitem{jolivet2013scalable}
-{\sc P.~Jolivet, F.~Hecht, F.~Nataf, and C.~Prud'homme}, {\em {Scalable Domain
-  Decomposition Preconditioners for Heterogeneous Elliptic Problems}}, in
-  Proceedings of the International Conference on High Performance Computing,
-  Networking, Storage and Analysis, SC13, ACM, 2013.
-
-\bibitem{jolivet2016block}
-{\sc P.~Jolivet and P.-H. Tournier}, {\em {Block Iterative Methods and
-  Recycling for Improved Scalability of Linear Solvers}}, in Proceedings of the
-  2016 International Conference for High Performance Computing, Networking,
-  Storage and Analysis, SC16, IEEE, 2016.
-
-\bibitem{KALANTZIS2018136}
-{\sc V.~Kalantzis, A.~C.~I. Malossi, C.~Bekas, A.~Curioni, E.~Gallopoulos, and
-  Y.~Saad}, {\em A scalable iterative dense linear system solver for multiple
-  right-hand sides in data analytics}, Parallel Computing, 74 (2018),
-  pp.~136--153.
+L.~Conen, V.~Dolean, R.~Krause, F.~Nataf, A coarse space for heterogeneous
+  {H}elmholtz problems based on the {D}irichlet-to-{N}eumann operator, Journal
+  of Computational and Applied Mathematics 271 (2014) 83--99.
+
+\bibitem{aldaas2019multi}
+H.~Al~Daas, L.~Grigori, P.~Jolivet, P.-H. Tournier,
+  \href{https://github.com/prj-/aldaas2019multi}{A multilevel {S}chwarz
+  preconditioner based on a hierarchy of robust coarse spaces}, Journal of
+  Scientific Computing (2019) submitted for publication.
+\newline\urlprefix\url{https://github.com/prj-/aldaas2019multi}
+
+\bibitem{spillane2013automatic}
+N.~Spillane, D.~J. Rixen, Automatic spectral coarse spaces for robust finite
+  element tearing and interconnecting and balanced domain decomposition
+  algorithms, International Journal for Numerical Methods in Engineering
+  95~(11) (2013) 953--990.
 
 \bibitem{karypis1998fast}
-{\sc G.~Karypis and V.~Kumar}, {\em A fast and high quality multilevel scheme
-  for partitioning irregular graphs}, SIAM Journal on Scientific Computing, 20
-  (1998), pp.~359--392,
-  \url{http://glaros.dtc.umn.edu/gkhome/metis/parmetis/overview}.
-
-\bibitem{knepley2009mesh}
-{\sc M.~G. Knepley and D.~A. Karpeev}, {\em {Mesh algorithms for PDE with Sieve
-  I: Mesh distribution}}, Scientific Programming, 17 (2009), pp.~215--230.
-
-\bibitem{knyazev2001toward}
-{\sc A.~V. Knyazev}, {\em {Toward the optimal preconditioned eigensolver:
-  Locally optimal block preconditioned conjugate gradient method}}, SIAM
-  journal on scientific computing, 23 (2001), pp.~517--541.
-
-\bibitem{li2005overview}
-{\sc X.~S. Li}, {\em {An overview of SuperLU: Algorithms, implementation, and
-  user interface}}, ACM Transactions on Mathematical Software, 31 (2005),
-  pp.~302--325.
-
-\bibitem{li2003superlu_dist}
-{\sc X.~S. Li and J.~Demmel}, {\em {SuperLU\_DIST: A scalable
-  distributed-memory sparse direct solver for unsymmetric linear systems}}, ACM
-  Transactions on Mathematical Software, 29 (2003), pp.~110--140.
-
-\bibitem{mandel1993balancing}
-{\sc J.~Mandel}, {\em Balancing domain decomposition}, Communications in
-  Numerical Methods in Engineering, 9 (1993), pp.~233--241.
-
-\bibitem{marchand2019two}
-{\sc P.~Marchand, X.~Claeys, P.~Jolivet, F.~Nataf, and P.-H. Tournier}, {\em
-  {Two-level preconditioning for $h$-version boundary element approximation of
-  hypersingular operator with GenEO}}, Numerische Mathematik,  (2019),
-  p.~submitted for publication.
-
-\bibitem{mathew2008domain}
-{\sc T.~Mathew}, {\em Domain decomposition methods for the numerical solution
-  of partial differential equations}, vol.~61, Springer Science \& Business
-  Media, 2008.
-
-\bibitem{may2016extreme}
-{\sc D.~A. May, P.~Sanan, K.~Rupp, M.~G. Knepley, and B.~F. Smith}, {\em
-  {Extreme-scale multigrid components within PETSc}}, in Proceedings of the
-  Platform for Advanced Scientific Computing Conference, 2016.
-
-\bibitem{moulin2018al}
-{\sc J.~Moulin, P.~Jolivet, and O.~Marquet}, {\em {Augmented Lagrangian
-  Preconditioner for Large-Scale Hydrodynamic Stability Analysis}}, Computer
-  Methods in Applied Mechanics and Engineering, 351 (2019), pp.~718--743,
-  \url{https://github.com/prj-/moulin2019al}.
-
-\bibitem{notay2000flexible}
-{\sc Y.~Notay}, {\em Flexible conjugate gradients}, SIAM Journal on Scientific
-  Computing, 22 (2000), pp.~1444--1460.
-
-\bibitem{cusparse-web-page}
-{\sc NVIDIA}, {\em {cuSPARSE} web page}.
-\newblock \url{https://docs.nvidia.com/cuda/cusparse}, 2020.
-
-\bibitem{oh2018bddc}
-{\sc D.-S. Oh, O.~B. Widlund, S.~Zampini, and C.~Dohrmann}, {\em {BDDC
-  algorithms with deluxe scaling and adaptive selection of primal constraints
-  for Raviart--Thomas vector fields}}, Mathematics of Computation, 87 (2018),
-  pp.~659--692.
-
-\bibitem{o1980block}
-{\sc D.~P. O'Leary}, {\em {The Block Conjugate Gradient Algorithm and Related
-  Methods}}, Linear Algebra and its Applications, 29 (1980), pp.~293--322.
-
-\bibitem{parks2006recycling}
-{\sc M.~L. Parks, E.~de~Sturler, G.~Mackey, D.~D. Johnson, and S.~Maiti}, {\em
-  {Recycling Krylov Subspaces for Sequences of Linear Systems}}, SIAM Journal
-  on Scientific Computing, 28 (2006), pp.~1651--1674.
-
-\bibitem{pechstein2012finite}
-{\sc C.~Pechstein}, {\em {Finite and Boundary Element Tearing and
-  Interconnecting Solvers for Multiscale Problems}}, vol.~90 of Lecture Notes
-  in Computational Science and Engineering, Springer, 2012.
-
-\bibitem{pechstein2017unified}
-{\sc C.~Pechstein and C.~R. Dohrmann}, {\em {A unified framework for adaptive
-  BDDC}}, Electronic Transactions on Numerical Analysis, 46 (2017),
-  pp.~273--336.
-
-\bibitem{poulson2013elemental}
-{\sc J.~Poulson, B.~Marker, R.~A. Van~de Geijn, J.~R. Hammond, and N.~A.
-  Romero}, {\em {Elemental: A new framework for distributed memory dense matrix
-  computations}}, ACM Transactions on Mathematical Software, 39 (2013).
-
-\bibitem{quarteroni1999domain}
-{\sc A.~Quarteroni and A.~Valli}, {\em Domain decomposition methods for partial
-  differential equations}, vol.~10, Clarendon Press, 1999.
-
-\bibitem{saad1993flexible}
-{\sc Y.~Saad}, {\em {A Flexible Inner--Outer Preconditioned GMRES Algorithm}},
-  SIAM Journal on Scientific Computing, 14 (1993), pp.~461--469.
-
-\bibitem{saad2003iterative}
-{\sc Y.~Saad}, {\em {Iterative Methods for Sparse Linear Systems}}, SIAM,
-  second~ed., 2003.
-
-\bibitem{saad1986gmres}
-{\sc Y.~Saad and M.~H. Schultz}, {\em {GMRES: a generalized minimal residual
-  algorithm for solving nonsymmetric linear systems}}, SIAM Journal on
-  Scientific and Statistical Computing, 7 (1986), pp.~856--869.
-
-\bibitem{saad2000deflated}
-{\sc Y.~Saad, M.~Yeung, J.~Erhel, and F.~Guyomarc'h}, {\em A deflated version
-  of the conjugate gradient algorithm}, SIAM Journal on Scientific Computing,
-  21 (2000), pp.~1909--1926.
-
-\bibitem{sakurai2003projection}
-{\sc T.~Sakurai and H.~Sugiura}, {\em A projection method for generalized
-  eigenvalue problems using numerical integration}, Journal of Computational
-  and Applied Mathematics, 159 (2003), pp.~119--128.
-
-\bibitem{sakurai2010application}
-{\sc T.~Sakurai, H.~Tadano, and Y.~Kuramashi}, {\em {Application of block
-  Krylov subspace algorithms to the Wilson--Dirac equation with multiple
-  right-hand sides in lattice QCD}}, Computer Physics Communications, 181
-  (2010), pp.~113--117.
-
-\bibitem{si2013tet}
-{\sc H.~Si}, {\em {TetGen: A Quality Tetrahedral Mesh Generator and 3D Delaunay
-  Triangulator}}, Tech. Report~13, 2013,
-  \url{http://wias-berlin.de/software/tetgen}.
-
-\bibitem{vsistek2013parallel}
-{\sc J.~{\v{S}}{\'\i}stek, J.~Mandel, B.~Soused{\'\i}k, and P.~Burda}, {\em
-  Parallel implementation of multilevel {BDDC}}, in Numerical Mathematics and
-  Advanced Applications 2011, Springer, 2013, pp.~681--689,
-  \url{http://users.math.cas.cz/~sistek/software/bddcml.html}.
-
-\bibitem{smith2004domain}
-{\sc B.~F. Smith, P.~Bj{\o}rstad, and W.~D. Gropp}, {\em Domain decomposition:
-  parallel multilevel methods for elliptic partial differential equations},
-  Cambridge University Press, 2004.
-
-\bibitem{soodhalter2014krylov}
-{\sc K.~M. Soodhalter, D.~B. Szyld, and F.~Xue}, {\em Krylov subspace recycling
-  for sequences of shifted linear systems}, Applied Numerical Mathematics, 81
-  (2014), pp.~105--118.
-
-\bibitem{spillane2011robust}
-{\sc N.~Spillane, V.~Dolean, P.~Hauret, F.~Nataf, C.~Pechstein, and
-  R.~Scheichl}, {\em A robust two-level domain decomposition preconditioner for
-  systems of {PDE}s}, Comptes Rendus Mathématique, 349 (2011), pp.~1255--1259.
-
-\bibitem{spillane2013abstract}
-{\sc N.~Spillane, V.~Dolean, P.~Hauret, F.~Nataf, C.~Pechstein, and
-  R.~Scheichl}, {\em Abstract robust coarse spaces for systems of {PDEs} via
-  generalized eigenproblems in the overlaps}, Numerische Mathematik, 126
-  (2013), pp.~741--700.
-
-\bibitem{spillane2013automatic}
-{\sc N.~Spillane and D.~J. Rixen}, {\em Automatic spectral coarse spaces for
-  robust finite element tearing and interconnecting and balanced domain
-  decomposition algorithms}, International Journal for Numerical Methods in
-  Engineering, 95 (2013), pp.~953--990.
-
-\bibitem{stathopoulos2009deflation}
-{\sc A.~Stathopoulos, A.~Abdel-Rehim, and K.~Orginos}, {\em {Deflation for
-  inversion with multiple right-hand sides in QCD}}, in Journal of Physics:
-  Conference Series, vol.~180, IOP Publishing, 2009.
-
-\bibitem{stathopoulos2002block}
-{\sc A.~Stathopoulos and K.~Wu}, {\em {A Block Orthogonalization Procedure with
-  Constant Synchronization Requirements}}, SIAM Journal on Scientific
-  Computing, 23 (2002), pp.~2165--2182.
-
-\bibitem{Stewart2002}
-{\sc G.~W. Stewart}, {\em {A Krylov--Schur Algorithm for Large Eigenproblems}},
-  SIAM Journal on Matrix Analysis and Applications, 23 (2002), pp.~601--614.
-
-\bibitem{tang2009comparison}
-{\sc J.~Tang, R.~Nabben, C.~Vuik, and Y.~Erlangga}, {\em Comparison of
-  two-level preconditioners derived from deflation, domain decomposition and
-  multigrid methods}, Journal of Scientific Computing, 39 (2009), pp.~340--370.
-
-\bibitem{toselli2005domain}
-{\sc A.~Toselli and O.~B. Widlund}, {\em Domain decomposition methods:
-  algorithms and theory}, vol.~34 of Series in Computational Mathematics,
-  Springer, 2005.
-
-\bibitem{tournier2016micro}
-{\sc P.-H. Tournier, I.~Aliferis, M.~Bonazzoli, M.~de~Buhan, M.~Darbas,
-  V.~Dolean, F.~Hecht, P.~Jolivet, I.~El~Kanfoud, C.~Migliaccio, F.~Nataf,
-  C.~Pichot, and S.~Semenov}, {\em {Microwave Tomographic Imaging of
-  Cerebrovascular Accidents by Using High-Performance Computing}}, Parallel
-  Computing, 85 (2019), pp.~88--97.
-
-\bibitem{van1992bi}
-{\sc H.~A. Van~der Vorst}, {\em {Bi-CGSTAB: a fast and smoothly converging
-  variant of Bi-CG for the solution of nonsymmetric linear systems}}, SIAM
-  Journal on Scientific and Statistical Computing, 13 (1992), pp.~631--644.
-
-\bibitem{wakam2013158}
-{\sc D.~N. Wakam and F.~Pacull}, {\em Memory efficient hybrid algebraic solvers
-  for linear systems arising from compressible flows}, Computers \& Fluids, 80
-  (2013), pp.~158--167.
-
-\bibitem{zampini2016pcbddc}
-{\sc S.~Zampini}, {\em {PCBDDC: a class of robust dual--primal methods in
-  PETSc}}, SIAM Journal on Scientific Computing, 38 (2016), pp.~S282--S306,
-  \url{https://www.mcs.anl.gov/petsc/petsc-master/docs/manualpages/PC/PCBDDC.html}.
-
-\bibitem{zampini2017balancing}
-{\sc S.~Zampini, P.~Vassilevski, V.~Dobrev, and T.~Kolev}, {\em {Balancing
-  Domain Decomposition by Constraints Algorithms for Curl-Conforming Spaces of
-  Arbitrary Order}}, in International Conference on Domain Decomposition
-  Methods, Springer, 2017, pp.~103--116.
+G.~Karypis, V.~Kumar,
+  \href{http://glaros.dtc.umn.edu/gkhome/metis/parmetis/overview}{A fast and
+  high quality multilevel scheme for partitioning irregular graphs}, SIAM
+  Journal on Scientific Computing 20~(1) (1998) 359--392.
+\newline\urlprefix\url{http://glaros.dtc.umn.edu/gkhome/metis/parmetis/overview}
+
+\bibitem{anderson2019mfem}
+R.~Anderson, J.~Andrej, A.~Barker, J.~Bramwell, J.-S. Camier, J.~Cerveny,
+  V.~Dobrev, Y.~Dudouit, A.~Fisher, T.~Kolev, W.~Pazner, M.~Stowell, V.~Tomov,
+  I.~Akkerman, J.~Dahm, D.~Medina, S.~Zampini, \href{http://mfem.org}{{MFEM}: A
+  modular finite element methods library}, Computers \& Mathematics with
+  Applications (2020).
+\newline\urlprefix\url{http://mfem.org}
 
 \bibitem{zhou2008block}
-{\sc Y.~Zhou and Y.~Saad}, {\em {Block Krylov--Schur method for large symmetric
-  eigenvalue problems}}, Numerical Algorithms, 47 (2008), pp.~341--359.
+Y.~Zhou, Y.~Saad, Block {K}rylov--{S}chur method for large symmetric eigenvalue
+  problems, Numerical Algorithms 47~(4) (2008) 341--359.
 

### 2020-07-24 article.tex
Fix labelling in figure caption
Style fixes

--- 2020-07-21 article.tex
@@
-\documentclass[3p,11pt]{elsarticle}
-
-\usepackage{lineno,hyperref}
-% \modulolinenumbers[5]
-
-\journal{Computers \& Mathematics with Applications}
+% SIAM Article Template
+\documentclass[review,onefignum,onetabnum]{siamart190516}
+
+% Information that is shared between the article and the supplement
+% (title and author information, macros, packages, etc.) goes into
+% ex_shared.tex. If there is no supplement, this file can be included
+% directly.
+
+\input{shared}
+
+% Optional PDF information
+\ifpdf
+\hypersetup{
+  pdftitle={KSPHPDDM and PCHPDDM: Extending PETSc with Robust Overlapping Schwarz Preconditioners and Advanced Krylov Methods},
+  pdfauthor={P. Jolivet, J. E. Roman, and S. Zampini}
+}
+\fi
+
+% The next statement enables references to information in the
+% supplement. See the xr-hyperref package for details.
+
+% \externaldocument{supplement}
+
+% FundRef data to be entered by SIAM
+%<funding-group specific-use="FundRef">
+%<award-group>
+%<funding-source>
+%<named-content content-type="funder-name">
+%</named-content>
+%<named-content content-type="funder-identifier">
+%</named-content>
+%</funding-source>
+%<award-id> </award-id>
+%</award-group>
+%</funding-group>
 \newcommand\col{:}
@@
 \makeatother
-\usepackage{amsmath,etoolbox,amssymb}
-\usepackage{cleveref}
+\usepackage{amsmath,etoolbox}
 \AtBeginEnvironment{pmatrix}{\setlength{\arraycolsep}{2pt}}
@@
 
-\begin{frontmatter}
-
-\title{KSPHPDDM and PCHPDDM: extending PETS\MakeLowercase{c} with advanced Krylov methods and robust multilevel Overlapping Schwarz preconditioners}
-
-\author[firstaddress]{Pierre Jolivet\corref{correspondingauthor}}
-\cortext[correspondingauthor]{Corresponding author}
-\ead{pierre.jolivet@enseeiht.fr}
-
-\author[secondaryaddress]{Jose E. Roman}
-\ead{jroman@dsic.upv.es}
-
-\author[thirdaddress]{Stefano Zampini}
-\ead{stefano.zampini@kaust.edu.sa}
-
-\address[firstaddress]{CNRS, IRIT--ENSEEIHT, Toulouse, France}
-\address[secondaryaddress]{Universitat Polit\`ecnica de Val\`encia, Val\`encia, Spain}
-\address[thirdaddress]{King Abdullah University of Science and Technology, Thuwal, Saudi Arabia}
+\maketitle
 
@@
     shapes, and structures. The goal of this paper is to explain how two
-    libraries, PETSc and HPDDM, have been interfaced in order to offer
+    libraries, PETSc and HPDDM, have been interfaced together in order to offer
     end-users robust overlapping Schwarz preconditioners and advanced Krylov
@@
 
-\end{frontmatter}
-
-\linenumbers
+% REQUIRED
+\begin{AMS}
+  35Q68, 65Y05
+\end{AMS}
 
@@
 SLEPc~\cite{hernandez2005ssf}, the Scalable Library for Eigenvalue Problem
-computations, an add-on library that extends PETSc with classes for
+computations, an add-on library that extends PETSc with several classes for
 linear and nonlinear eigenvalue problems, as well as with tools to facilitate
@@
 convergence plateaus. Recycling techniques have been introduced to attenuate
-these effects. On the one hand, PETSc implements the loose GMRES~\cite{baker2005technique}~(LGMRES) and the
-deflated GMRES~\cite{erhel1996restarted,wakam2013158}~(DGMRES) as
+these effects. On the one hand, PETSc implements the Loose GMRES~\cite{baker2005technique}~(LGMRES) and the
+Deflated GMRES~\cite{erhel1996restarted,wakam2013158}~(DGMRES) as
 \pk{KSPLGMRES} and \pk{KSPDGMRES}. However, neither handle variable
-preconditioning, unlike \pk{KSPFGMRES} or \pk{KSPGCR} which respectively
-implement the flexible GMRES~\cite{saad1993flexible}~(FGMRES) and the
-generalized conjugate residual method~\cite{eisenstat1983variational}~(GCR).
-Furtheremore, \pk{KSPDGMRES} does not support complex arithmetic.  Moreover,
-extending the recycling capabilities of such methods to sequence of linear
-systems with smoothly varying coefficient matrices and right-hand sides is not
-trivial.  On the other hand, HPDDM offers support for the generalized conjugate
-residual method with inner orthogonalization and deflated
-restarting~\cite{parks2006recycling}~(GCRODR), which does not suffer from the
-aforementioned limitations.
+preconditioning, and \pk{KSPDGMRES} does not support complex arithmetic.
+Moreover, extending the recycling capabilities of such methods to sequence of linear systems
+with smoothly varying coefficient matrices and right-hand sides is not trivial.
+On the other hand, HPDDM offers support for the
+generalized conjugate residual method with inner orthogonalization and deflated
+restarting~\cite{parks2006recycling}~(GCRODR), which does not suffer from the aforementioned
+limitations.
 
@@
     \item \pk{MATNEST}: block-defined matrix with nested submatrices;
-    \item \pk{MATAIJCUSPARSE}: sequential or parallel sparse matrix, offloaded to a NVIDIA GPU using cuSPARSE~\cite{cusparse-web-page}.
+    \item \pk{MATAIJCUSPARSE}: sequential or parallel sparse matrix, offloaded to a GPU using cuSPARSE~\cite{cusparse-web-page}.
 \end{itemize}
@@
 SuperLU\_DIST~\cite{li2003superlu_dist}. For the case of a dense matrix:
-ScaLAPACK and Elemental~\cite{poulson2013elemental} are supported. Future work
-will consider extending the multigrid framework \pk{PCMG}, as well as other
-preconditioning classes to increase arithmetic intensity of the preconditioner
-application phase for blocks of right-hand sides.
+ScaLAPACK and Elemental~\cite{poulson2013elemental}.  Future work will consider
+extending the multigrid framework \pk{PCMG}, as well as other preconditioning
+classes to increase arithmetic intensity of the preconditioner application
+phase for blocks of right-hand sides.
 
@@
         GCRODR(40, 20), and Jacobi right-preconditioned GCRODR(40, 20);
-    \item in PETSc with no preconditioning and no restart GMRES($\infty$) \pk{KSPGMRES};
+    \item in PETSc with unpreconditioned and unrestarted GMRES($\infty$) \pk{KSPGMRES};
     \item in PETSc with \pk{KSPHPDDM}, all of the above.
@@
 The notation GCRODR($n$, $m$) indicates a restart after $n$ iterations and a recycling subspace of dimension $m$.
-In all of the cases, convergence is declared when the initial unpreconditioned
-residual is reduced by 10 orders of magnitude.
+In all the cases, convergence is declared when the initial unpreconditioned residual is reduced by 10 orders of magnitude.
 Except for the right-preconditioned GCRODR, these tests are the same as the
@@
 All PETSc tests are performed using four MPI
-processes and can be launched using the following command lines. \\[-4pt]
+processes and they can be launched using the following command lines. \\[-4pt]
 \begin{Verbatim}[fontsize=\footnotesize,frame=single,framerule=0.1mm,commandchars=&\[\]]
@@
 \end{tikzpicture}
-        \caption{Number of iterations needed to converge for various
+        \caption{\# of iterations needed to converge for various
         configurations as originally tested by Parks et
@@
         between \ref{pgfplots:r2} and \ref{pgfplots:r3}. In the other cases,
-        the number of iterations needed by \pk{KSPHPDDM} match with the respective
+        the numbers of iterations needed by \pk{KSPHPDDM} match with the respective
         MATLAB or PETSc reference implementation.\label{fig:comparison}}
@@
         width=0.8\textwidth,
-        xlabel={Number of right-hand sides ($N$)},
+        xlabel={\# of right-hand sides ($N$)},
         ylabel={},
@@
 For example, if the matrices were to be offloaded to a GPU, one could instead benchmark \pk{MATAIJcuSPARSE}.
-The mini-app is then used to benchmark intranode performance of matrix--matrix
-multiplications, either with multiple OpenMP threads, or by offloading the
-operation to a NVIDIA GPU and using cuSPARSE~\cite{cusparse-web-page} as
-interfaced in PETSc.
+The mini-app is then used to benchmark intranode performance of matrix--matrix multiplications, either with multiple OpenMP threads, or by offloading the operation to a GPU and using cuSPARSE~\cite{cusparse-web-page} as interfaced in PETSc.
 \begin{Verbatim}[fontsize=\footnotesize,frame=single,framerule=0.1mm,commandchars=&\[\]]
@@
 evaluated using a given base flow $q_b = \begin{bmatrix}u_b \\ p_b\end{bmatrix}$ and
-$M$ is the discretization of the mass matrix on the space of velocities.
+$M$ the discretization of the mass matrix on the space of velocities.
 The mini-app employed for the numerical results is built on top of
@@
 as reported in~\cref{fig:comparison-slepc}, from 80 with \pk{KSPFGMRES} to 45.
-A similar speedup can be observed in terms of runtime, with the time spent for
-the solution of the eigenvalue problem being reduced from \SI[round-mode = places,round-precision=1,group-minimum-digits=3,group-separator = {,}]{25.1}{\minute}
-to \SI[round-mode = places,round-precision=1,group-minimum-digits=3,group-separator = {,}]{17.5}{\minute}.
-These timings, and all that follow, have been obtained on
-Irène, a system composed of \pgfmathprintnumber[assume math mode=true]{1656}
-nodes with two 24-core Intel Xeon Platinum 8168 clocked at
-\SI{2.7}{\giga\hertz}.
+A similar speedup can be observed in terms of runtime, with the time spent for the solution of the eigenvalue problem being reduced from \SI[round-mode = places,round-precision=1,group-minimum-digits=3,group-separator = {,}]{1509}{\sec} to \SI[round-mode = places,round-precision=1,group-minimum-digits=3,group-separator = {,}]{1052}{\sec}.
+These timings, and all that follow, have been obtained on {Irène}, a system composed of \pgfmathprintnumber[assume math mode=true]{1656} nodes with two 24-core Intel Xeon Platinum 8168 clocked at \SI{2.7}{\giga\hertz}.
 \pgfplotstableread{
@@
 \end{tikzpicture}
-        \caption{Number of inner iterations for systems~\cref{eq:ks} with~\ref{pgfplots:m1} or without~\ref{pgfplots:m0} recycling.\label{fig:comparison-slepc}}
+        \caption{\# of inner iterations for systems~\cref{eq:ks} with~\ref{pgfplots:m1} or without~\ref{pgfplots:m0} recycling.\label{fig:comparison-slepc}}
     \end{figure}
@@
 \subsubsection{Blocking inside SLEPc\label{sec:lobpcg}}
-The implementations of the locally optimal block preconditioned conjugate
-gradient~\cite{knyazev2001toward}~(LOBPCG) and the contour integral spectrum
-slicing method~\cite{sakurai2003projection}~(CISS) in SLEPc were previously not
+The implementations of the Locally Optimal Block Preconditioned Conjugate
+Gradient~\cite{knyazev2001toward}~(LOBPCG) and the Contour Integral Spectrum
+Slicing method~\cite{sakurai2003projection}~(CISS) in SLEPc were previously not
 using blocking when applying the preconditioner or solving linear systems. In
@@
 We finally note that blocking support is also available and tested in SLEPc for \pk{EPSSUBSPACE}
-since version 3.14.0.
+and \pk{NEPCISS} since version 3.14.0.
 
@@
 in PETSc with \pk{PCSetModifySubMatrices}, see~\cite{gander2019new}.  It also provides support for
-defining robust two-level methods using the generalized eigenvalue problem on
-the overlap (GenEO) framework, for finite
+defining robust two-level methods using the Generalized Eigenvalue problem on
+the Overlap (GenEO) framework, for finite
 elements~\cite{spillane2011robust,spillane2013abstract} and boundary
@@
 discretization of $\mathcal{L}$ on the extended local subdomain with overlap, endowed with natural boundary
-conditions. In the case of the one-dimensional Poisson
+conditions. In the case of the 1D Poisson
 equation with homogeneous Dirichlet boundary conditions discretized using trilinear finite elements, an explicit representation of these operators is given~\cref{fig:poisson} for a
@@
     \end{minipage}
-\caption{Notation and basics of overlapping Schwarz methods.\label{fig:poisson}}
+\caption{Notations and basics of overlapping Schwarz methods.\label{fig:poisson}}
 \end{figure}
@@
         \begin{equation}\label{eq:gevp}
-            \mathring{A}_i y_i = \lambda_i \tilde{R}_i R_i^T (R_i A R_i^T) R_i \tilde{R}_i^T y_i ;
+            \mathring{A}_i y_i = \lambda_i \tilde{R}_i R_i^T (R_i A R_i^T) \tilde{R}_i R_i^T y_i ;
         \end{equation}
@@
 \end{equation*}
-Such a construction of a two-level method yields a preconditioner with which it
-is possible to bound the condition number of the preconditioned operator. Thus,
-the convergence rate of a preconditioned Krylov method such as the conjugate
-gradient can be guaranteed a priori.
+Such a construction of a two-level method yields a preconditioner with which it is possible to bound the condition number of the preconditioned operator.
+Thus, the convergence speed of a preconditioned Krylov method such as the conjugate gradient can be guaranteed a priori.
 
@@
 the size of the coarse operator $P^TAP$ is much lower
-than the size of $A$, and coarsening is much more aggressive than
+than $n$, the size of $A$, and coarsening is much more aggressive than
 when using algebraic or geometric multigrid methods.
@@
 Other \pk{DMType}s, including user-defined, may be supported by
-implementing the proper \pk{DMCreateNeumannOverlap} callback. Support for the
-structured grid infrastructure (\pk{DMDA}) could be easily added.
+implementing the proper \pk{DMCreateNeumannOverlap} callback. Support for the structured grid infrastructure (\pk{DMDA})
+could be easily added upon request.
 \begin{figure}[htbp]
@@
 
-For performance, it is better to compute the same number of
+For performance reasons, it is better to compute the same number of
 eigenvectors on each process, since in this case, the coarse operator will be
@@
 In addition,
-standard eigensolvers such as \pk{EPSKRYLOVSCHUR} from SLEPc, cannot inexpensively
+standard eigensolvers such as \pk{EPSKRYLOVSCHUR} from SLEPc, cannot cheaply
 evaluate the exact number of eigenpairs in a given interval of $\mathbb{R}^+$.
-Instead, it is often more efficient to provide an upper bound on the number of
+Instead, it is often more efficient to provide an upper bound of the number of
 eigenpairs.
@@
 incompressible elasticity or the system of Stokes. Indeed, for such equations,
-one needs to compute a large number of eigenvectors from the classical GenEO
+one needs to compute a lot of eigenvectors from the classical GenEO
 eigenproblem~\cref{eq:gevp} to generate a robust preconditioner. GenEO-2
@@
 Because of the intrinsic nature of balancing domain
-decomposition methods, \pk{PCNN} and \pk{PCBDDC} must be supplied with a \pk{MatIS}
+decomposition methods, \pk{PCNN} and \pk{PCBDDC} must be supplied a \pk{MatIS}
 which stores local unassembled matrices and local to global mappings.
@@
 finite element context. Since the assembled form of these operators can be reconstructed,
-it is possible to obtain the Dirichlet operators
+it is possible to get the Dirichlet operators
 $\{R_i A R_i^T\}_{i=1}^N$ from~\cref{sec:interface-pc} as well. All tools needed by the
 GenEO framework are thus readily available. Note however that one-level
-overlapping Schwarz methods are known for converging slowly when there is
-no overlap. The proposed methodology is not strictly equivalent to the
+overlapping Schwarz methods are known for converging slowly when there is in
+fact no overlap. The proposed methodology is not strictly equivalent to the
 BDD-GenEO method~\cite{spillane2013automatic}, in which local Schur complements
@@
 \[
-    \varepsilon_{ij}(u) = \frac{1}{2} \left(\frac{\partial u_i}{\partial x_i} + \frac{\partial u_j}{\partial x_j}\right)\!, \text{ } \mu = \frac{E}{2(1+\nu)}, \text{ and }\lambda = \frac{E\nu}{1-2\nu}.
+	\varepsilon_{ij}(u) = \frac{1}{2} \left(\frac{\partial u_i}{\partial x_i} + \frac{\partial u_j}{\partial x_j} \right), \mu = \frac{E}{2(1+\nu)}, \text{ and }\lambda = \frac{E\nu}{1-2\nu}.
 \]
@@
 In~\cref{fig:inexact}, the number of inner coarse iterations are reported for various values of $M$.
-In the case of a two-level method (\cref{enum:exact}) this number is equal to one and is thus not reported.
+In the case of a two-level method (\cref{enum:exact}) this number is equal to one and is thus not reported in the figure.
 \pgfplotstableread{
@@
 }
-    \subfloat[\parbox{5cm}{Number of inner coarse iterations for various values of $M$, the number of second-level subdomains.}\label{fig:inexact}]{
+    \subfloat[\parbox{5cm}{\# of inner coarse iterations for various values of $M$, the number of second-level subdomains.}\label{fig:inexact}]{
         \begin{tikzpicture}[font=\scriptsize]
@@
 On the other hand, the fastest \pk{PCHPDDM} configuration, which corresponds to the
-configuration with 256 second-level subdomains~\ref{pgfplots:p-256}, requires
-64 seconds to setup and 36 seconds for the solution of the linear system.
+configuration with 256 second-level subdomains~\ref{pgfplots:p-256}, requires 64 seconds to setup and 31 for the solution of the linear system.
 Out of the 64 seconds needed for \pk{PCHPDDM} setup, 31 are spent in SLEPc
@@
 \begin{equation}
-    -\nabla \cdot (\kappa \nabla u) - 6.2 e^u = 0,\label{eq:bratu}
+    -\nabla \cdot (\kappa \nabla u) - 6.2 \exp^u = 0,\label{eq:bratu}
 \end{equation}
@@
 approximately $\pgfmathprintnumber{29}$ nonzero coefficients per row.
-It is solved using \pk{SNES}, with \pk{PCHPDDM} being the preconditioner for
+It is solved using a \pk{SNES}, with \pk{PCHPDDM} being the preconditioner for
 solving each linearized system, see~\cref{fig:u}.
@@
     \subsubsection{Using PCHPDDM as a coarse grid solver for PCBDDC\label{sec:bddc}}
-A last problem shows how one may switch between a multilevel
+A last toy problem is here proposed to show how one may switch between a multilevel
 BDDC solver to a two-level BDDC solver using \pk{PCHPDDM} for solving the
-coarse problem. The problem is generated using the MFEM~\cite{anderson2019mfem}
+coarse problem. The problem is generated using MFEM~\cite{anderson2019mfem}
 definite Maxwell example available at \url{https://github.com/mfem/mfem/blob/master/examples/petsc/ex3p.cpp}.
@@
 
-The interface between PETSc and HPDDM has room for future improvements.  First,
-HPDDM handles mixed-precision,
+The interface between PETSc and HPDDM has still room for future improvements.
+First, HPDDM handles mixed-precision,
 e.g., using single-precision scalars for assembling coarse operators while
@@
 handle GPU efficiently currently, but the interface with PETSc provides a very
-thorough testbed with \pk{Mat} and \pk{PC} implementations that do exploit
-GPUs.
+thorough testbed with \pk{Mat} and \pk{PC} implementations that are exploiting
+GPU.
 
 \section*{Acknowledgments}
-The authors would like to thank S.~Balay, J.~Brown, V.~Hapla, M.~Knepley, and
-B.~Smith for reviewing the successive merge requests in PETSc repository and
-for their feedback on this manuscript. This work was granted access to the
-GENCI-sponsored HPC resources of:
+The authors would like to thank S.~Balay, J.~Brown, V.~Hapla, M.~Knepley,
+and B.~Smith for reviewing the successive merge requests in PETSc
+repository. This work was granted access to the GENCI-sponsored HPC resources of:
 \begin{itemize}
@@
 
-\bibliographystyle{elsarticle-num}
+\bibliographystyle{siamplain}
 \begin{thebibliography}{10}

### 2020-07-27 article.tex
Remove NEPCISS
Comments from Barry (not for the bibliography)

--- 2020-07-21 article.tex
@@
 \section*{Acknowledgments}
-The authors would like to thank S.~Balay, J.~Brown, V.~Hapla, M.~Knepley, and
-B.~Smith for reviewing the successive merge requests in PETSc repository and
-for their feedback on this manuscript. This work was granted access to the
-GENCI-sponsored HPC resources of:
+The authors would like to thank S.~Balay, J.~Brown, V.~Hapla, M.~Knepley,
+and B.~Smith for reviewing the successive merge requests in PETSc
+repository. This work was granted access to the GENCI-sponsored HPC resources of:
 \begin{itemize}

### 2020-07-28 article.tex
Some last minor fixes

@@
 as reported in~\cref{fig:comparison-slepc}, from 80 with \pk{KSPFGMRES} to 45.
-A similar speedup can be observed in terms of runtime, with the time spent for the solution of the eigenvalue problem being reduced from \SI[round-mode = places,round-precision=1,group-minimum-digits=3,group-separator = {,}]{1509}{\sec} to \SI[round-mode = places,round-precision=1,group-minimum-digits=3,group-separator = {,}]{1052}{\sec}.
-These timings, and all that follow, have been obtained on {Irène}, a system composed of \pgfmathprintnumber[assume math mode=true]{1656} nodes with two 24-core Intel Xeon Platinum 8168 clocked at \SI{2.7}{\giga\hertz}.
+A similar speedup can be observed in terms of runtime, with the time spent for
+the solution of the eigenvalue problem being reduced from \SI[round-mode = places,round-precision=1,group-minimum-digits=3,group-separator = {,}]{25.1}{\minute}
+to \SI[round-mode = places,round-precision=1,group-minimum-digits=3,group-separator = {,}]{17.5}{\minute}.
+These timings, and all that follow, have been obtained on
+Irène, a system composed of \pgfmathprintnumber[assume math mode=true]{1656}
+nodes with two 24-core Intel Xeon Platinum 8168 clocked at
+\SI{2.7}{\giga\hertz}.
 \pgfplotstableread{
@@
 \section*{Acknowledgments}
-The authors would like to thank S.~Balay, J.~Brown, V.~Hapla, M.~Knepley,
-and B.~Smith for reviewing the successive merge requests in PETSc
-repository. This work was granted access to the GENCI-sponsored HPC resources of:
+The authors would like to thank S.~Balay, J.~Brown, V.~Hapla, M.~Knepley, and
+B.~Smith for reviewing the successive merge requests in PETSc repository and
+for their feedback on this manuscript. This work was granted access to the
+GENCI-sponsored HPC resources of:
 \begin{itemize}

### 2020-07-28 MatProduct.c
Cleaning up MatProduct.c a bit

@@ -27,7 +27,7 @@
     PetscStackPush(#func);                                     \
     __ierr = func args;                                        \
     PetscStackPop;                                             \
-    if (__ierr != SPARSE_STATUS_SUCCESS) SETERRQ2(PETSC_COMM_SELF, PETSC_ERR_LIB, "Error in %s(): error code %d",#func,(int)__ierr); \
+    if (__ierr != SPARSE_STATUS_SUCCESS) SETERRQ2(PETSC_COMM_SELF, PETSC_ERR_LIB, "Error in %s(): error code %d", #func, (int)__ierr); \
   } while (0)
 #else
 #define PetscStackCallMKLSparse(func, args) do {               \
@@ -69,9 +69,9 @@ int main(int argc, char** argv) {
     ntype = 1;
     ierr = PetscStrallocpy(MATSEQAIJ, &type[0]);CHKERRQ(ierr);
   }
-  ierr = PetscOptionsGetBool(NULL, NULL, "-check_results", &check, NULL);CHKERRQ(ierr);
-  ierr = PetscOptionsGetBool(NULL, NULL, "-test_trans", &trans, NULL);CHKERRQ(ierr);
-  for (int j = 0; j < nbs; ++j) {
+  ierr = PetscOptionsGetBool(NULL, NULL, "-check", &check, NULL);CHKERRQ(ierr);
+  ierr = PetscOptionsGetBool(NULL, NULL, "-trans", &trans, NULL);CHKERRQ(ierr);
+  for (PetscInt j = 0; j < nbs; ++j) {
     ierr = PetscViewerBinaryOpen(PETSC_COMM_WORLD, file, FILE_MODE_READ, &viewer);CHKERRQ(ierr);
     ierr = MatCreate(PETSC_COMM_WORLD, &A);CHKERRQ(ierr);
     ierr = MatSetFromOptions(A);CHKERRQ(ierr);
@@ -126,9 +126,9 @@ int main(int argc, char** argv) {
       ierr = MatConvert(A, MATSEQAIJ, MAT_INPLACE_MATRIX, &A);CHKERRQ(ierr);
       ierr = MatSetOption(A, MAT_SYMMETRIC, PETSC_TRUE);CHKERRQ(ierr);
     }
-    for (int i = 0; i < ntype; ++i) {
+    for (PetscInt i = 0; i < ntype; ++i) {
       char        *tmp;
-      PetscInt    *ia_ptr, *ja_ptr;
[…]
